### 2020-03-06 main.tex
Initial Overleaf Import

%\documentclass{article}
%\usepackage[a4paper, total={6in, 8in}]{geometry}
\documentclass[reprint,amsmath,amssymb,rmp,onecolumn,notitlepage,11pt]{revtex4-1}
\usepackage[utf8]{inputenc}
%\usepackage{authblk}
%\usepackage{natbib}
\usepackage[normalem]{ulem}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\newcommand{\red}[1]{\textcolor{red!80!black}{#1}}
\newcommand{\blue}[1]{\textcolor{blue!80!black}{#1}}
\newcommand{\green}[1]{\textcolor{green!70!black}{#1}}
\usepackage{mathtools}
\DeclarePairedDelimiter{\evdel}{\langle}{\rangle}

\begin{document}
\title{Linklength distribution in Proteins and our model}
\title{The role of geometry in the chain separation distribution of links in protein contact networks}
%\title{}
%\title{}
%\title{}
\author{Antonia Mey}
\affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
\author{Steffen Mühle}
\affiliation{Uni Göttingen, 37077 Göttingen, Germany}
\author{Nora Molkenthin}
\email[Electronic Address: ]{n.molkenthin@gmx.de}
\affiliation{Potsdam Institut für Klimafolgenforschung, Potsdam, Germany}
\affiliation{Chair for Network Dynamics, Institute for Theoretical Physics and Center for Advancing Electronics Dresden (cfaed), Technical University of Dresden, 01069 Dresden}
%\author{Marc Timme}
% \affiliation{Network Dynamics, Max Planck Institute for Dynamics and Self-Organization (MPIDS), 37077 Göttingen, Germany}
% \affiliation{Chair for Network Dynamics, Institute for Theoretical Physics and Center for Advancing Electronics Dresden (cfaed), Technical University of Dresden, 01069 Dresden}

\begin{abstract}
The distribution of chain separations of interacting amino acids in proteins roughly follows a power law. Here we show, analytically and in simulations, that a geometrical stochastic model of folding chains can explain this behaviour. %Separating the data in $\alpha$-dominated, $\beta$-dominated and intrinsically disordered proteins (IDP's) shows characteristic structures in the distribution due to the secondary structure.
\end{abstract}
\maketitle

\section*{Introduction}
\begin{enumerate}
    \item Proteins are very important \cite{a bunch of stuff}
    \item Proteins can be modeled as networks of amino acids and their connections \cite{Vendruscolo2002,DiPaola2013,Estrada2011}.
    \item They postulate/observe that proteins have a sequence separation distribution, that is 1/l in \cite{bartoli2008effect}.
    \item Here we explain this from geometric constraints alone using a model introduced in \cite{Molkenthin2016,molkenthin2020self}
\end{enumerate}
TO DO: Write text
\section*{Model}
\begin{enumerate}
    \item Model in General
    \item Simplified 2D version
\end{enumerate}
TO DO: Write text
\section*{Analytic approximation}
Starting from a closed chain of $N$ unit discs, as introduced for the analytic calculations in \cite{molkenthin2016scaling}.
Let us start by introducing the auxiliary variable $F(s)$, defined to be the number of possible links with a sequence separation of $s$. Before any links are added all sequence separations are equally likely, as there are $N$ possibilities of making a link of each separation $s$ with $2\leq s < N/2$.

As we add more links, not only are links taken out of this pool, because they have already been realized, each existing link can also geometrically prohibit other connections.
Let us call the poos of available links after $i$ links have been added $F_i(s)$.

-------------- Add Figure explaining which links are geometrically excluded --------------

The expected distribution of possible links after one step is given by the average over available link pools taken over all possible lengths of the initial link
\begin{equation}
    F_1(s)= \frac{1}{N/2-2} \sum_{s_1=2}^{N/2} { \begin{cases}
    N-2(s-1) \text{ , for } s<s_1\\
    N-2(s_1 -1)\text{ , for } s\geq s_1
    \end{cases}}.
\end{equation}

In subsequent steps we get the same reduction but starting from the pool of the step before rather than the full $N$ links for each length, leading to an additional probability factor 
\begin{equation}
    P_k(s)=\frac{F_k(s)}{C_k},
    \label{eq.Pk}
\end{equation}
with $C_k=\sum_{s=2}^{N/2}F^k(s)$, for each segment separation $s$ to occur and thus
\begin{equation}
    F_k(s)= \frac{1}{N/2-2} \sum_{s_k=2}^{N/2} {\begin{cases}
     F_{k-1}(s)-2(s-1) P^k(s) \text{ , for } s<s_k\\
     F_{k-1}(s)-2(s_k -1)P^k(s)\text{ , for } s\geq s_k
    \end{cases}}.
\end{equation}
This can be simplified to 
\begin{align}
   F_k(s)&= \frac{1}{N/2-2} \frac{F_{k-1}(s)}{C_{k-1}}\left( \sum_{s_k=2}^{N/2}C_{k-1} - \sum_{s_k=2}^{s} 2(s-1) - \sum_{s_k=s+1}^{N/2} 2(s_k -1) \right) \nonumber \\
   &= \frac{1}{N/2-2}\frac{F_{k-1}(s)}{C_{k-1}}\left((N/2-2)C_{k-1}- 2(N/2-(s+1))(s-1) - ((s-1)(s+1))\right)  \nonumber \\
   &=F_{k-1}(s)\left(1-\frac{f(s)}{(N/2-2)C_{k-1}} \right)
   \label{eq.Fk_rec}
\end{align}
Where we defined $f(s)=2(N/2-(s+1))(s-1) + ((s-1)(s+1))$ for notational brevity. \red{Maybe we can try to simplify this expression some more.}

We can now use the recursive expression Eq.~\ref{eq.Fk_rec} to write down a closed expression for $F_k(s)$

\begin{equation}
    F_k(s)=N\prod_{i}^k\left(1-\frac{f(s)}{(N/2-2)C_{k-1}} \right)
\end{equation}

The probability distribution of the realized sequence separations of all added links is then given by the average over the available pools at each link addition step, leading to
\begin{align}
    P(s)&=\frac{1}{N-3}\sum_{k=0}^{N-3} P_k(s) \nonumber \\
    &= \frac{N}{N-3}\sum_{k=0}^{N-3} \frac{1}{C_k}\prod_{i}^k\left(1-\frac{f(s)}{(N/2-2)C_{k-1}} \right),
\end{align}
which makes use of Eq.~\ref{eq.Pk}. The only unknown in this expression is $C_k$, which we can not compute exactly. However, since $C_k$ monotonically decreases with $k$, $N>C_k>C_{N-2}$ always holds and we can obtain an upper and lower bound by replacing all $C_k$ in the expression with either $N$ (underestimates all probabilities) or $C_{N-2}$ (overestimates all probabilities). Each time we get an expression of the form 
\begin{align}
     P(s)&\approx\frac{N}{N-3}\frac{1}{C}\sum_{k=0}^{N-3} \prod_{i}^k\left(1-\frac{f(s)}{(N/2-2)C} \right)\nonumber \\
     &=\frac{N}{N-3}\frac{1}{C} \frac{1}{\alpha s}.
\end{align}
Here we have made use of the geometric series and made the substitution $1-\frac{f(s)}{(N/2-2)C} = 1- \alpha s $.

\red{ah, I forgot to copy down why $f(s)$ is linear in s, which is kind of crucial for this argument, it's in my notes somewhere.}

\red{is that plausible? I guess if the difference of $N$ and $C_{N-2}$ is large that leaves quite a large range...}

\section*{Data and simulations}
\begin{enumerate}
    \item Show plots of distribution for data, simulation and analytics
    \item Maybe comment on chin length dependence
    \item Look into the dip we saw there once
\end{enumerate}
\begin{figure}[h]
        \centering
	\includegraphics[width=0.45\textwidth]{both_dist_200.pdf}
	\includegraphics[width=0.45\textwidth]{both_dist_400.pdf}
        \caption{New figure is averaged over more structures (no alpha beta selection), need to explain why the rule only works for link lengths between 20 and N/2. Or why there is a bump, it seems more like a bump than a dip now...
        }
        \label{fig:time_scaling}
\end{figure}
\section*{Conclusion}
\begin{enumerate}
    \item We can now explain the sequence separation from geometry and even do it analytically.
    \item Maybe something about helixes and sheets and the dip.
\end{enumerate}
TO DO: Leave until the rest is done.
\section*{Declarations}
\subsection{Availability of data and materials}
All data generated or analysed during this study are included in this published article [and its supplementary information files].
\subsection{Competing interests}
The authors declare no competing interests.
\subsection{Funding}
This research was supported by ...
\subsection{Authors' contributions}

\subsection{Acknowledgements}
We thank ... for fruitful discussions.

\bibliographystyle{unsrt}
\bibliography{proteins}

\appendix
\section{Supplemental Material Collection}
\section{Random ideas}
\begin{itemize}
    \item Why is the beginning different for simulations?
    \item Does the dip disappear for IDPs
    \item Can we use all NMR structures for IDPs
    \item Would it help using more simulated structures than just the Crystal structure. 
\end{itemize}
TODO: @Toni: Collect data and plotting scripts in github
TODO: @Toni: Connect overleaf to github

\end{document}

### 2020-03-06 pdb_linklength.py
Add files via upload

@@ -1,0 +1,222 @@
+# -*- coding: utf-8 -*-
+from __future__ import print_function
+import networkx as nx
+import scipy as sp
+import scipy.stats
+from scipy.optimize import curve_fit
+import matplotlib.pyplot as plt
+import matplotlib.colors as colors
+import numpy as np
+from Bio.PDB import *
+from MDAnalysis import *
+import urllib
+import os
+import seaborn as sns
+import glob
+
+sns.set(font_scale=1.5)  # crazy big
+sns.set_style("ticks")
+sns.despine()
+
+def func(x, a,b):
+     return a * x**(-b)
+
+def func2(x, a):
+     return a / x
+
+names_xray = glob.glob("pdb_data2/all_xray/*.pdb")
+names_xray = glob.glob("pdb_data2/xray_data_short/*.pdb")
+names_xray = glob.glob("pdb_data2/xray_data/*.pdb")
+
+names_nmr = glob.glob("pdb_data2/nmr_data/*.pdb")
+names_nmr = glob.glob("pdb_data2/nmr_data_short/*.pdb")
+
+names_both = glob.glob("pdb_data2/both/*.pdb")
+
+names_used=names_both
+
+num=28
+lengths=[3,4,5,6,7,8,9,10,11,12,13,14,16,17,19,21,23,25,28,30,33,36,40,44,48,52,58,63,69,76,83,91,100,110,120,132,145,158,174,191,209,229,251,275,
+    302,331,363,398]#[0:45]
+print(len(lengths))
+
+numnum=20
+logscale=np.logspace(np.log10(3), np.log10(1000), num=numnum)
+print('logscale',logscale)
+results=[]
+binnedresults=[]
+
+if __name__ == '__main__':
+    directories = ['pdb_data2/both']
+    nameslen=len(names_used)
+    #nameslen=50
+    path = os.getcwd()
+    meanies=[]
+    linklengths=[]
+    for d in directories:
+        print(d)
+        output = os.path.join(path,d)
+        count=0
+        n=1000
+        t=6.5
+        counter=0
+        for counter in range(nameslen):
+            #pdb_file = os.path.join(output,pdb.strip())+'.pdb'
+            pdb_file = names_used[counter]
+            print(pdb_file)
+            u = Universe(pdb_file)
+            #calphas = u.select_atoms("name CA and segid A")
[…]

### 2020-04-13 paper/main.tex
restructuring repository
Adding small changes to intro

--- 2020-03-06 main.tex
@@
 \section*{Introduction}
+Proteins, the molecular machines of every living organism perform vital tasks required for life to persists ranging from transport (e.g.
+hemoglobin), signal transduction (e.g.
+rhodopsin), immune responses (e.g.
+antibodies), and hormonal regulation (e.g.
+insulin)~\cite{}.
+All natural proteins are made up of 20 different amino acids which dictate the three dimensional conformations the proteins adopt in order to function~\cite{}.
+One way of looking at the functional forms of proteins and classifying structure properties of proteins is by looking at protein residue or contact networks~\cite{Vendruscolo2002,DiPaola2013,Estrada2011}.
+Protein Contact Networks (PCN),
+
+Since amino acids take up space and are inherently geometrical objects, any plausible network ensemble for describing them, would have to be spatially embedded [5, 10].
+Geometrical models for protein folding include models that derive characteristics of the secondary structure from constraints on bond and torsion angles [2, 7] and models focused on the formation mechanisms of the tertiary structure [8].
+Based on such considerations, we have constructed a network formation process, in which random connections are added to a chain of unit disks, without causing them to intersect [8].
+Here we analyze the link length distribution in those models, where the length of a link is defined as the distance of the two ends along the original chain.
+This is then compared to the link length distribution ex- tracted from protein residue networks (PRN) from Pro- tein Data Bank (PDB) data.
+We find that the frequency of links drops off with their length as
+f(l) ∝ l−γ, (1)
+both for the model and the PRNs (see Fig.
+2), although the PRNs show a dip in the distribution for small lengths.
+Here f(l) denotes the frequency of links of length l.
+Interestingly in [https://iopscience.iop.org/article/10.1088/1478-3975/4/4/L01/fulltext/], such a link length distribution (with γ = 1) was imposed on a model (without further explanation), reproducing protein-like average shortest path lengths and clustering coefficient.
 \begin{enumerate}
@@
 \section*{Analytic approximation}
-Starting from a closed chain of $N$ unit discs, as introduced for the analytic calculations in \cite{molkenthin2016scaling}.
+Starting from a closed chain of $N$ unit discs, as introduced for the analytic calculations in \cite{molkenthin2016scaling}. \red{briefly explain that here too.}
 Let us start by introducing the auxiliary variable $F(s)$, defined to be the number of possible links with a sequence separation of $s$.
@@
 As we add more links, not only are links taken out of this pool, because they have already been realized, each existing link can also geometrically prohibit other connections.
-Let us call the poos of available links after $i$ links have been added $F_i(s)$.
+Let us call the pools of available links after $i$ links have been added $F_i(s)$.
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{prn_vis.pdf}
+    \caption{Caption}
+    \label{fig:fig_prn_vis}
+\end{figure}
 
@@
 
-In subsequent steps we get the same reduction but starting from the pool of the step before rather than the full $N$ links for each length, leading to an additional probability factor 
+In subsequent steps we get the same reduction but starting from the pool of the step before rather than the full $N$ links for each length, leading to an additional probability factor \red{explain this in more detail, I already forgot exactly why that is...}
 \begin{equation}
@@
     F_k(s)= \frac{1}{N/2-2} \sum_{s_k=2}^{N/2} {\begin{cases}
-     F_{k-1}(s)-2(s-1) P^k(s) \text{ , for } s<s_k\\
-     F_{k-1}(s)-2(s_k -1)P^k(s)\text{ , for } s\geq s_k
+     F_{k-1}(s)-2(s-1) P_k(s) \text{ , for } s<s_k\\
+     F_{k-1}(s)-2(s_k -1)P_k(s)\text{ , for } s\geq s_k
     \end{cases}}.
@@
 \begin{align}
-   F_k(s)&= \frac{1}{N/2-2} \frac{F_{k-1}(s)}{C_{k-1}}\left( \sum_{s_k=2}^{N/2}C_{k-1} - \sum_{s_k=2}^{s} 2(s-1) - \sum_{s_k=s+1}^{N/2} 2(s_k -1) \right) \nonumber \\
-   &= \frac{1}{N/2-2}\frac{F_{k-1}(s)}{C_{k-1}}\left((N/2-2)C_{k-1}- 2(N/2-(s+1))(s-1) - ((s-1)(s+1))\right)  \nonumber \\
-   &=F_{k-1}(s)\left(1-\frac{f(s)}{(N/2-2)C_{k-1}} \right)
+   F_k(s)&= \frac{1}{\frac{N}{2}-2} \frac{F_{k-1}(s)}{C_{k-1}}\left( \sum_{s_k=2}^{N/2}C_{k-1} - \sum_{s_k=2}^{s} 2(s_k-1) - \sum_{s_k=s+1}^{N/2} 2(s -1) \right) \nonumber \\
+   &= \frac{1}{\frac{N}{2}-2}\frac{F_{k-1}(s)}{C_{k-1}}\left(\left(\frac{N}{2}-2\right)C_{k-1} -(s^2-s)-2\left(\frac{N}{2}-(s+1)\right)(s-1)\right)  \nonumber \\
+   &= F_{k-1}(s)\left(1-\frac{-s^2 +(N-1)s-(N+1)}{(\frac{N}{2}-2)C_{k-1}} \right)\nonumber \\
+   &=F_{k-1}(s)\left(1-\frac{f(s)}{(\frac{N}{2}-2)C_{k-1}} \right)
    \label{eq.Fk_rec}
 \end{align}
-Where we defined $f(s)=2(N/2-(s+1))(s-1) + ((s-1)(s+1))$ for notational brevity. \red{Maybe we can try to simplify this expression some more.}
+Where we defined $f(s)=-s^2 +(N-1)s-(N+1) \approx N s - s^2 - N$ for notational brevity.
 
@@
 \begin{equation}
-    F_k(s)=N\prod_{i}^k\left(1-\frac{f(s)}{(N/2-2)C_{k-1}} \right)
+    F_k(s)=N\prod_{i}^k\left(1-\frac{f(s)}{(\frac{N}{2}-2)C_{i-1}} \right)
 \end{equation}
@@
     P(s)&=\frac{1}{N-3}\sum_{k=0}^{N-3} P_k(s) \nonumber \\
-    &= \frac{N}{N-3}\sum_{k=0}^{N-3} \frac{1}{C_k}\prod_{i}^k\left(1-\frac{f(s)}{(N/2-2)C_{k-1}} \right),
+    &= \frac{N}{N-3}\sum_{k=0}^{N-3} \frac{1}{C_k}\prod_{i}^k\left(1-\frac{f(s)}{(\frac{N}{2}-2)C_{i-1}} \right),
 \end{align}
@@
 \begin{align}
-     P(s)&\approx\frac{N}{N-3}\frac{1}{C}\sum_{k=0}^{N-3} \prod_{i}^k\left(1-\frac{f(s)}{(N/2-2)C} \right)\nonumber \\
-     &=\frac{N}{N-3}\frac{1}{C} \frac{1}{\alpha s}.
+     P(s)&\approx\frac{N}{N-3}\sum_{k=0}^{N-3} \frac{1}{C_k}\prod_{i}^k\left(1-\frac{f(s)}{(\frac{N}{2}-2)C} \right)\nonumber \\
+     &=\frac{N}{N-3} \sum_{k=0}^{N-3}\frac{1}{C_k}\left(1-\frac{f(s)}{(\frac{N}{2}-2)C} \right)^k \nonumber \\
+     &\approx \frac{A}{f(s)}
 \end{align}
-Here we have made use of the geometric series and made the substitution $1-\frac{f(s)}{(N/2-2)C} = 1- \alpha s $.
+for $1<s<\frac{N}{2}$,
+where $A$ is an unknown constant, containing an approximation of the various normalization constants $C_i$, as well as the direct effects of the chain length. 
+Since for long chains, $f(s)$ is largely dominated by the linear term, the distribution is well approximated by a power law with an exponent of $-1$.
+We have made use of the geometric series to approximately find a closed solution.
 
-\red{ah, I forgot to copy down why $f(s)$ is linear in s, which is kind of crucial for this argument, it's in my notes somewhere.}
+In the last step of this approximation, we take the $C_k$ to all be approximately equal in order to apply the geometric series.
+However, in reality we know $1/C_k$ to increase with $k$, giving large $k$ summands relatively more weight than in a geometric series.
+This would lead to a correction in the exponent.
 
-\red{is that plausible? I guess if the difference of $N$ and $C_{N-2}$ is large that leaves quite a large range...}
+\begin{equation}
+    P_{corrected}(s)\approx A f(s)^{-1+\alpha},
+\end{equation}
+where $0<\alpha<1$ \red{why cant $\alpha$ be >1?}.
 
 \section*{Data and simulations}
-\begin{enumerate}
-    \item Show plots of distribution for data, simulation and analytics
-    \item Maybe comment on chin length dependence
-    \item Look into the dip we saw there once
-\end{enumerate}
+ For large values of the chain length $N$ we thus find an approximate power law probability distribution for the sequence separation with an exponent of -1 \red{include note of correction}.
+Here we use simulations, as introduced in \cite{molkenthin2020self} and networks extracted from measured protein data from the PDB \cite{PDB} to compare this behaviour in realistic 3D settings and show that the basic principle persists, indicating that the sequence distance separation distribution is largely determined by simple geometric constraints.
+
+The complex interaction pattern between the amino acids in a protein can be very naturally expressed as a network, in which each amino acid is represented by a node and spatial proximity is encoded as a link.
+Whenever two central $C_\alpha$ atoms are closer together than a threshold $d_c$, they are connected.
+These connections are encoded in the adjacency matrix, which is the binary matrix
+\begin{equation}
+  A^{\textsf{PDB}}_{ij}=
+  \begin{cases}
+   0, & \text{ if } d_{i,j}>d_c \text{ or } i=j\\
+      1, & \text{ if } d_{i,j}\leq d_c .
+      \end{cases}
+    \label{eq:aij}
+\end{equation}
+
 \begin{figure}[h]
         \centering
-	\includegraphics[width=0.45\textwidth]{both_dist_200.pdf}
-	\includegraphics[width=0.45\textwidth]{both_dist_400.pdf}
-        \caption{New figure is averaged over more structures (no alpha beta selection), need to explain why the rule only works for link lengths between 20 and N/2.
-Or why there is a bump, it seems more like a bump than a dip now...
+	\includegraphics[width=0.45\textwidth]{figures/idp_200.pdf}
+	\includegraphics[width=0.45\textwidth]{figures/idp_400.pdf}
+        \caption{Sequence distance distributions for real and simulated folds for a) Chain lengths around 200 b) chain lengths around 400.
+The theoretical curve for 2D rings works okay for 3D open chains (because the rings are closed sequence separations of more than N/2 are a bit meaningless and thus not accounted for) The parameters used in this are $A=80$ and $\alpha=0.2$ in a) and $0.3$ in b).
+I thought in b) it makes sense for $\alpha$ to be larger as there are even larger exponents $k$ coming into relevance.
+I'm not sure we want to use that whole $\alpha$ argument too much though as it's a but ad-hoc and would be nicer to just fit $A$ and be done with it, but then it may look a bit more off.
         }
@@
 TODO: @Toni: Collect data and plotting scripts in github
-TODO: @Toni: Connect overleaf to github
+TODO: @Toni: Connect overleaf to github \green{(check)}
 

### 2020-04-13 paper/main.tex
First pass at introduction

@@
 \DeclarePairedDelimiter{\evdel}{\langle}{\rangle}
+\usepackage[colorinlistoftodos]{todonotes}
 
 \begin{document}
-\title{Linklength distribution in Proteins and our model}
 \title{The role of geometry in the chain separation distribution of links in protein contact networks}
@@
 %\title{}
-\author{Antonia Mey}
-\affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
-\author{Steffen Mühle}
-\affiliation{Uni Göttingen, 37077 Göttingen, Germany}
 \author{Nora Molkenthin}
@@
 \affiliation{Chair for Network Dynamics, Institute for Theoretical Physics and Center for Advancing Electronics Dresden (cfaed), Technical University of Dresden, 01069 Dresden}
+\author{Antonia S J S Mey}
+\affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
+\author{Steffen Mühle}
+\affiliation{Uni Göttingen, 37077 Göttingen, Germany}
+
 %\author{Marc Timme}
@@
 One way of looking at the functional forms of proteins and classifying structure properties of proteins is by looking at protein residue or contact networks~\cite{Vendruscolo2002,DiPaola2013,Estrada2011}.
-Protein Contact Networks (PCN),
+Protein Contact Networks (PCN), have shown promise in understanding protein folding patterns as well as revealing allosteric communication pathways~\cite{https://pubs.acs.org/doi/10.1021/acs.jcim.9b00320, and others}.
+Prom a physicist perspective it is interesting to understand the emergent behaviour in protein structure from PCNs, by investigating common network measures and their behaviour for PCNs.
+For example Bartoli et al.
+observe that proteins have a sequence separation distribution that follows $\frac{1}{L}$, where $L$ is the length of the protein sequence, however do not provide an explanation for this observations~\cite{bartoli2008effect}.
+Here, we will show how simple geometrical model can capture this behaviour well, even providing a simple analytical approximation.
+The analytical approximation and geometrical model build on the idea that inherently geometrical objects, such as amino acids, that any resulting PCN network ensembles for such geometrical objects has to be spatially embedded~\cite{Grabow2010,Wu2015}.
+Different approaches have been used in the past for geometrical models describing protein folding, some of which derived characteristics of the secondary structure from constraints on bond and torsion angles~\cite{Danielsson2010,Molkenthin2011} and others focused on the formation mechanisms of the tertiary structure~\cite{molkenthin2016scaling, molkenthin2020self}.
+Based on such considerations, we have constructed a network formation process, in which random connections are added to a chain of unit disks, without causing them to intersect~\cite{molkenthin2016scaling}.
+Such a model cannow be used to investigate the emergenct behaviour of sequence separation deistributions as observed by Bartoli et al.~\cite{bartoli2008effect}.
+To this end, we analyze the link length distribution that emerge from the geometric model, where the length of a link is defined as the distance of the two ends along the original chain.
+This is then compared to the link length distribution extracted from protein residue networks (PRN) generated from protein data bank data.
 
-Since amino acids take up space and are inherently geometrical objects, any plausible network ensemble for describing them, would have to be spatially embedded [5, 10].
-Geometrical models for protein folding include models that derive characteristics of the secondary structure from constraints on bond and torsion angles [2, 7] and models focused on the formation mechanisms of the tertiary structure [8].
-Based on such considerations, we have constructed a network formation process, in which random connections are added to a chain of unit disks, without causing them to intersect [8].
-Here we analyze the link length distribution in those models, where the length of a link is defined as the distance of the two ends along the original chain.
-This is then compared to the link length distribution ex- tracted from protein residue networks (PRN) from Pro- tein Data Bank (PDB) data.
-We find that the frequency of links drops off with their length as
-f(l) ∝ l−γ, (1)
-both for the model and the PRNs (see Fig.
-2), although the PRNs show a dip in the distribution for small lengths.
-Here f(l) denotes the frequency of links of length l.
-Interestingly in [https://iopscience.iop.org/article/10.1088/1478-3975/4/4/L01/fulltext/], such a link length distribution (with γ = 1) was imposed on a model (without further explanation), reproducing protein-like average shortest path lengths and clustering coefficient.
-\begin{enumerate}
-    \item Proteins are very important \cite{a bunch of stuff}
-    \item Proteins can be modeled as networks of amino acids and their connections \cite{Vendruscolo2002,DiPaola2013,Estrada2011}.
-    \item They postulate/observe that proteins have a sequence separation distribution, that is 1/l in \cite{bartoli2008effect}.
-    \item Here we explain this from geometric constraints alone using a model introduced in \cite{Molkenthin2016,molkenthin2020self}
-\end{enumerate}
-TO DO: Write text
+\todo[inline, color={green!40}]{@Nora: Summary of our findings and why this is cool.}
+
 \section*{Model}
@@
 \end{enumerate}
-TO DO: Write text
+\todo[inline, color={green!40}]{@Nora: Add model information}
 \section*{Analytic approximation}
@@
     \centering
-    \includegraphics[width=\columnwidth]{prn_vis.pdf}
+    \includegraphics[width=0.5\columnwidth]{figures/prn_vis.pdf}
     \caption{Caption}
@@
 \end{itemize}
-TODO: @Toni: Collect data and plotting scripts in github
-TODO: @Toni: Connect overleaf to github \green{(check)}
 

### 2020-04-20 paper/main.tex
updates from Nora

@@
 \section*{Introduction}
-Proteins, the molecular machines of every living organism perform vital tasks required for life to persists ranging from transport (e.g.
+Proteins, the molecular machines of every living organism perform vital tasks required for life to persist, ranging from transport (e.g.
 hemoglobin), signal transduction (e.g.
@@
 All natural proteins are made up of 20 different amino acids which dictate the three dimensional conformations the proteins adopt in order to function~\cite{}.
-One way of looking at the functional forms of proteins and classifying structure properties of proteins is by looking at protein residue or contact networks~\cite{Vendruscolo2002,DiPaola2013,Estrada2011}.
+One way of looking at the functional forms of proteins and classifying their structural properties is by looking at protein residue or contact networks~\cite{Vendruscolo2002,DiPaola2013,Estrada2011}.
 Protein Contact Networks (PCN), have shown promise in understanding protein folding patterns as well as revealing allosteric communication pathways~\cite{https://pubs.acs.org/doi/10.1021/acs.jcim.9b00320, and others}.
@@
 For example Bartoli et al.
-observe that proteins have a sequence separation distribution that follows $\frac{1}{L}$, where $L$ is the length of the protein sequence, however do not provide an explanation for this observations~\cite{bartoli2008effect}.
-Here, we will show how simple geometrical model can capture this behaviour well, even providing a simple analytical approximation.
-The analytical approximation and geometrical model build on the idea that inherently geometrical objects, such as amino acids, that any resulting PCN network ensembles for such geometrical objects has to be spatially embedded~\cite{Grabow2010,Wu2015}.
+observe that proteins have a sequence separation distribution that follows $\frac{1}{\ell}$, where $\ell$ is the separation of two connected amino acids along the backbone chain.
+However, they do not provide an explanation for this observation~\cite{bartoli2008effect}.
+Here, we will show how a simple geometrical model can capture this behaviour well, even providing a simple analytical approximation.
+The analytical approximation and geometrical model build on the idea that inherently geometrical objects, such as amino acids, imply that any resulting PCN network ensemble modelling them has to be spatially embedded~\cite{}.
 Different approaches have been used in the past for geometrical models describing protein folding, some of which derived characteristics of the secondary structure from constraints on bond and torsion angles~\cite{Danielsson2010,Molkenthin2011} and others focused on the formation mechanisms of the tertiary structure~\cite{molkenthin2016scaling, molkenthin2020self}.
 Based on such considerations, we have constructed a network formation process, in which random connections are added to a chain of unit disks, without causing them to intersect~\cite{molkenthin2016scaling}.
-Such a model cannow be used to investigate the emergenct behaviour of sequence separation deistributions as observed by Bartoli et al.~\cite{bartoli2008effect}.
-To this end, we analyze the link length distribution that emerge from the geometric model, where the length of a link is defined as the distance of the two ends along the original chain.
-This is then compared to the link length distribution extracted from protein residue networks (PRN) generated from protein data bank data.
+Such a model can now be used to investigate the emergenct behaviour of sequence separation distributions as observed by Bartoli et al.~\cite{bartoli2008effect}.
+To this end, we analyze the sequence separation distribution that emerge from the geometric model, where the length $\ell$ of a link is defined as the distance of the two ends along the original chain.
+This is then compared to the sequence separation distribution extracted from protein residue networks (PRN) generated from protein data bank data.
 
@@
 
-\section*{Model}
-\begin{enumerate}
-    \item Model in General
-    \item Simplified 2D version
-\end{enumerate}
-\todo[inline, color={green!40}]{@Nora: Add model information}
+\section*{Protein Residue Networks and their synthetic counterparts}
+The complex interaction pattern between the amino acids in a protein can be very naturally expressed as a network, in which each amino acid is represented by a node and spatial proximity is encoded as a link.
+Whenever two central $C_\alpha$ atoms are closer together than a threshold $d_c$, they are connected.
+These connections are encoded in the adjacency matrix of the folded protein, which is the binary matrix
+\begin{equation}
+  A^{\textsf{PDB}}_{ij}=
+  \begin{cases}
+   0, & \text{ if } d_{i,j}>d_c \text{ or } i=j\\
+      1, & \text{ if } d_{i,j}\leq d_c .
+      \end{cases}
+    \label{eq:aij}
+\end{equation}
+Adjacency matrices are commonly used to describe networks and in this case represent the PRN.
+
+The underlying formation mechanisms of individual protein folds are incredibly complex and depend on the surrounding solvent as well as the specific amino acid sequence and their quantum-mechanical interactions.
+Here we want to study the ensemble of all proteins, however, and will thus "average out" such specifics.
+
+Based on the basic assumption that amino acids are objects in space, that can not overlap indefinitely, we introduced a model in \cite{molkenthin2020self} starting from a chain of identical spheres, each in contact only with the neighbours it is connected to.
+Then additional connections form by randomly selecting two spheres and moving them towards each other until they touch, the new links formed that way can not be broken in later steps and function as constraints on subsequent links.
+If contact of the two selected spheres is geometrically impossible without breaking previously made connections or leading to overlaps of any spheres, the link is not made and taken out of the pool of possible connections.
+This process is repeated until no more links can be formed without violating the geometric constraints.
+
+In \cite{molkenthin2016scaling}, we have introduced a simplified, two-dimensional version of this model, which starts from a closed chain of $N$ unit discs and subsequently adds links, such that connected discs touch, yet no discs intersect.
+The advantage of this model is, that it can be approximated by a purely topological simplification, which can be treated analytically.
+
+The topological constraints in the resulting network model are:
+(a) New links always form between two units that are part of the same face of the graph (region enclosed by a cycle in the network).
+This prevents the overlapping of discs.
+(b) No links form across the outer face.
+This prevents the enclosure of a unit by less than six other units (which is geometrically impossible) such that.
+(c) The maximum degree of each unit is six, as six is the maximum number of unit discs, one central unit disc can touch.
+(d) Once connected by a link, pairs of units do not disconnect.
+
 \section*{Analytic approximation}
-Starting from a closed chain of $N$ unit discs, as introduced for the analytic calculations in \cite{molkenthin2016scaling}. \red{briefly explain that here too.}
+Here we want to analytically approximate the sequence separation distribution for the simplified 2D model, that is the separation along the backbone chain of two nodes connected by a link, as illustrated in Fig~\ref{fig:fig_prn_vis}.
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.3\columnwidth]{figures/prn_vis.pdf}
+    \caption{The sequence separation distribution is defined as the separation along the chain (blue) of two connected units (red).}
+    \label{fig:fig_prn_vis}
+\end{figure}
+
 Let us start by introducing the auxiliary variable $F(s)$, defined to be the number of possible links with a sequence separation of $s$.
@@
 Let us call the pools of available links after $i$ links have been added $F_i(s)$.
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\columnwidth]{figures/prn_vis.pdf}
-    \caption{Caption}
-    \label{fig:fig_prn_vis}
-\end{figure}
+\\
 
--------------- Add Figure explaining which links are geometrically excluded --------------
+\red{----------- Add Figure explaining which links are geometrically excluded, maybe include in fig 1 -----------}
+\\
 
@@
 
-The complex interaction pattern between the amino acids in a protein can be very naturally expressed as a network, in which each amino acid is represented by a node and spatial proximity is encoded as a link.
-Whenever two central $C_\alpha$ atoms are closer together than a threshold $d_c$, they are connected.
-These connections are encoded in the adjacency matrix, which is the binary matrix
-\begin{equation}
-  A^{\textsf{PDB}}_{ij}=
-  \begin{cases}
-   0, & \text{ if } d_{i,j}>d_c \text{ or } i=j\\
-      1, & \text{ if } d_{i,j}\leq d_c .
-      \end{cases}
-    \label{eq:aij}
-\end{equation}
+\begin{figure}[h]
+        \centering
+	\includegraphics[width=0.45\textwidth]{figures/both_200.pdf}
+	\includegraphics[width=0.45\textwidth]{figures/both_400.pdf}
+        \caption{Sequence distance distributions for real and simulated folds for a) Chain lengths around 200 b) chain lengths around 400.
+The theoretical curve for 2D rings works okay for 3D open chains (because the rings are closed sequence separations of more than N/2 are a bit meaningless and thus not accounted for) The parameters used in this are $A=80$ and $\alpha=0.2$ in a) and $0.3$ in b).
+I thought in b) it makes sense for $\alpha$ to be larger as there are even larger exponents $k$ coming into relevance.
+I'm not sure we want to use that whole $\alpha$ argument too much though as it's a but ad-hoc and would be nicer to just fit $A$ and be done with it, but then it may look a bit more off.
+        }
+        \label{fig:time_scaling}
+\end{figure}
+\red{We observe that, while the general shape seems to match, in real proteins sequence separations of 20 to 60 amino acids are slightly over represented.
+We need to comment on that somehow.
+Also links of length 2 and 3 are over represented, we need to discuss that too.}
 
+\red{To check if the bump has something to do with the secondary structure that is present in real proteins but not in the model, we have generated a sample of intrinsically disordered proteins, that also do not contain secondary structures.
+The bump does indeed disappear.
+The over representation of 2 and 3 does not though and there os probably a simple explanation for that and maybe we can simply take those points out and redo the mean degree matching and get to lines that are not just parallel but actually on top of one another that way.}
 \begin{figure}[h]
@@
 	\includegraphics[width=0.45\textwidth]{figures/idp_400.pdf}
-        \caption{Sequence distance distributions for real and simulated folds for a) Chain lengths around 200 b) chain lengths around 400.
+        \caption{Sequence distance distributions for intrinsically disordered proteins and simulated folds for a) Chain lengths around 200 b) chain lengths around 400.
 The theoretical curve for 2D rings works okay for 3D open chains (because the rings are closed sequence separations of more than N/2 are a bit meaningless and thus not accounted for) The parameters used in this are $A=80$ and $\alpha=0.2$ in a) and $0.3$ in b).

### 2020-04-20 pdb_linklength.py

@@ -24,6 +24,12 @@ def func(x, a,b):
 def func2(x, a):
      return a / x
 
+def theofunc(x, C,N,ex):
+     return (C)* np.power(N/(2*(-x**2+N*x-N)),ex)
+
+#def theofunc(x, C,N):
+#     return (C)* N/(2*(N*x-N))
+
 names_xray = glob.glob("pdb_data2/all_xray/*.pdb")
 names_xray = glob.glob("pdb_data2/xray_data_short/*.pdb")
 names_xray = glob.glob("pdb_data2/xray_data/*.pdb")
@@ -32,6 +38,8 @@ names_nmr = glob.glob("pdb_data2/nmr_data/*.pdb")
 names_nmr = glob.glob("pdb_data2/nmr_data_short/*.pdb")
 
 names_both = glob.glob("pdb_data2/both/*.pdb")
+
+#names_both = glob.glob("idp/*.pdb")
 
 names_used=names_both
 
@@ -48,6 +56,7 @@ binnedresults=[]
 
 if __name__ == '__main__':
     directories = ['pdb_data2/both']
+    #directories = ['idp']
     nameslen=len(names_used)
     #nameslen=50
     path = os.getcwd()
@@ -58,7 +67,7 @@ if __name__ == '__main__':
         output = os.path.join(path,d)
         count=0
         n=1000
-        t=6.5
+        t=8.0
         counter=0
         for counter in range(nameslen):
             #pdb_file = os.path.join(output,pdb.strip())+'.pdb'
@@ -128,7 +137,7 @@ for i in range(len(results)):
     me=results[i]
     bin_means, bin_edges, binnumber = sp.stats.binned_statistic(me[1][1][:-1],me[1][0], statistic=np.mean, bins=logscale)
     bin_centers = bin_edges[1:] - (bin_edges[1:] - bin_edges[0:numnum-1])/2
-    if me[0]<230 and me[0]>170:
+    if me[0]<260 and me[0]>130:
         sum=sum+me[1][0]
         bsum=bsum+bin_means
         countbin+=1
@@ -148,25 +157,28 @@ for i in range(len(sim_results)):
     me=sim_results[i]
     bin_means, bin_edges, binnumber = sp.stats.binned_statistic(me[1][1][:-1],me[1][0], statistic=np.mean, bins=logscale)
[…]

### 2020-05-25 data/pdb_data2/snippet.py
tidying code and adding data for plotting

@@ -1,0 +1,17 @@
+
+import glob
+from MDAnalysis import *
+
+names_xray = glob.glob("/xray_data/*.pdb")
+names_nmr = glob.glob("nmr_data/*.pdb")
+for i in range(len(names_xray)):
+    pdb_file = names[i]
+    u = Universe(pdb_file)
+    calphas = u.select_atoms("name CA and segid "+u.residues.segments.segids[0])
+    #Then do the stuff you did before
+
+for i in range(len(names_nmr)):
+    pdb_file = names[i]
+    u = Universe(pdb_file)
+    calphas = u.select_atoms("name CA and segid "+u.residues.segments.segids[0])
+    #Then do the stuff you did before

### 2020-05-25 pdb_linklength.py

@@ -1,245 +1,0 @@
-# -*- coding: utf-8 -*-
-from __future__ import print_function
-
-import glob
-import os
-
-import matplotlib.pyplot as plt
-import networkx as nx
-import numpy as np
-import scipy as sp
-import scipy.stats
-import seaborn as sns
-from Bio.PDB import *
-from MDAnalysis import *
-from scipy.optimize import curve_fit
-
-sns.set(font_scale=1.5)  # crazy big
-sns.set_style("ticks")
-sns.despine()
-
-
-def func(x, a, b):
-    return a * x ** (-b)
-
-
-def func2(x, a):
-    return a / x
-
-
-def theofunc(x, C, N, ex):
-    return (C) * np.power(N / (2 * (-x ** 2 + N * x - N)), ex)
-
-
-# def theofunc(x, C,N):
-#     return (C)* N/(2*(N*x-N))
-
-names_xray = glob.glob("pdb_data2/all_xray/*.pdb")
-names_xray = glob.glob("pdb_data2/xray_data_short/*.pdb")
-names_xray = glob.glob("pdb_data2/xray_data/*.pdb")
-
-names_nmr = glob.glob("pdb_data2/nmr_data/*.pdb")
-names_nmr = glob.glob("pdb_data2/nmr_data_short/*.pdb")
-
-names_both = glob.glob("pdb_data2/both/*.pdb")
-
-# names_both = glob.glob("idp/*.pdb")
-
-names_used = names_both
-
-num = 28
-lengths = [3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 16, 17, 19, 21, 23, 25, 28, 30, 33, 36, 40, 44, 48, 52, 58, 63, 69,
-           76, 83, 91, 100, 110, 120, 132, 145, 158, 174, 191, 209, 229, 251, 275,
-           302, 331, 363, 398]  # [0:45]
-print(len(lengths))
-
-numnum = 20
-logscale = np.logspace(np.log10(3), np.log10(1000), num=numnum)
-print('logscale', logscale)
-results = []
-binnedresults = []
-
-if __name__ == '__main__':
-    directories = ['pdb_data2/both']
-    # directories = ['idp']
-    nameslen = len(names_used)
-    # nameslen=50
-    path = os.getcwd()
-    meanies = []
-    linklengths = []
-    for d in directories:
-        print(d)
-        output = os.path.join(path, d)
-        count = 0
-        n = 1000
-        t = 8.0
[…]

### 2020-05-25 paper/main.tex
Updates from Overleaf

@@
 \title{The role of geometry in the chain separation distribution of links in protein contact networks}
-%\title{}
+\title{\red{We need a better title, the old one is too long}}
 %\title{}
@@
 \author{Nora Molkenthin}
-\email[Electronic Address: ]{n.molkenthin@gmx.de}
-\affiliation{Potsdam Institut für Klimafolgenforschung, Potsdam, Germany}
-\affiliation{Chair for Network Dynamics, Institute for Theoretical Physics and Center for Advancing Electronics Dresden (cfaed), Technical University of Dresden, 01069 Dresden}
-\author{Antonia S J S Mey}
-\affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
+\affiliation{Potsdam Institut für Klimafolgenforschung, Potsdam}
+\affiliation{Center for Advancing Electronics Dresden (cfaed), Technical University of Dresden, 01069 Dresden}
 \author{Steffen Mühle}
 \affiliation{Uni Göttingen, 37077 Göttingen, Germany}
-
+\author{Antonia S J S Mey}
+\email[Electronic Address: ]{antonia.mey@ed.ac.uk}
+\affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
 %\author{Marc Timme}
@@
 Protein Contact Networks (PCN), have shown promise in understanding protein folding patterns as well as revealing allosteric communication pathways~\cite{https://pubs.acs.org/doi/10.1021/acs.jcim.9b00320, and others}.
-Prom a physicist perspective it is interesting to understand the emergent behaviour in protein structure from PCNs, by investigating common network measures and their behaviour for PCNs.
-For example Bartoli et al.
-observe that proteins have a sequence separation distribution that follows $\frac{1}{\ell}$, where $\ell$ is the separation of two connected amino acids along the backbone chain.
-However, they do not provide an explanation for this observation~\cite{bartoli2008effect}.
+From a physicist perspective it is interesting to understand the emergent behaviour in protein structure from PCNs, by investigating common network measures and their behaviour for PCNs.
+For example Bartoli et al.~\cite{bartoli2008effect} assume that proteins have a sequence separation distribution that follows $\frac{1}{\ell}$, where $\ell$ is the separation of two connected amino acids along the backbone chain as the simplest implementation of a connection probability, that decreases with the distance along the sequence.
+However, they do not provide a quantitative explanation for this assumption.
+
 Here, we will show how a simple geometrical model can capture this behaviour well, even providing a simple analytical approximation.
@@
 \section*{Analytic approximation}
-Here we want to analytically approximate the sequence separation distribution for the simplified 2D model, that is the separation along the backbone chain of two nodes connected by a link, as illustrated in Fig~\ref{fig:fig_prn_vis}.
+Here we want to analytically approximate the sequence separation distribution for the simplified 2D model, that is the separation along the backbone chain of two nodes connected by a link, as illustrated in Fig~\ref{fig:schematic}~a).
+
 \begin{figure}[h]
     \centering
-    \includegraphics[width=0.3\columnwidth]{figures/prn_vis.pdf}
-    \caption{The sequence separation distribution is defined as the separation along the chain (blue) of two connected units (red).}
-    \label{fig:fig_prn_vis}
+    \includegraphics[width=0.4\columnwidth]{figures/Fig1.pdf}
+    \caption{a) The sequence separation distribution is defined as the separation along the chain (blue) of two connected units (red).
+b) The solid blue link prohibits some connections from node $A$ (dotted red), while others are still permitted (dotted blue).
+Counting all newly excluded links, we find that a link of length $k$ prohibits $2(l-1)$ links of length $l$ up to a maximum of $k$ and $2(k-1)$ links of length $k$ or more.}
+    \label{fig:schematic}
 \end{figure}
@@
 As we add more links, not only are links taken out of this pool, because they have already been realized, each existing link can also geometrically prohibit other connections.
-Let us call the pools of available links after $i$ links have been added $F_i(s)$.
-\\
+Adding a link of length $s_1$ prohibits $2(s-1)$ links of length $s<s_1$ and $2(s_1-1)$ links of length $s>s_1$ (see Fig.~\ref{fig:schematic}~b).
+To find the expected number of links still available in subsequent stops, we average over all possible sequence separations $s_1$ of the first link added.
 
-\red{----------- Add Figure explaining which links are geometrically excluded, maybe include in fig 1 -----------}
-\\
-
-The expected distribution of possible links after one step is given by the average over available link pools taken over all possible lengths of the initial link
+The expected distribution of possible links after one step is given thus by the average over available link pools taken over all possible lengths of the initial link
 \begin{equation}
@@
 \end{equation}
+Here we denote the size of the pool of available links after $i$ links have been added $F_i(s)$.
 
-In subsequent steps we get the same reduction but starting from the pool of the step before rather than the full $N$ links for each length, leading to an additional probability factor \red{explain this in more detail, I already forgot exactly why that is...}
+In subsequent steps we get the same reduction but starting from the pool $F_{k-1}(s)$ of the step before rather than the full $N$ links for each length.
+The reduction also needs to be multiplied by a factor
 \begin{equation}
@@
 \end{equation}
-with $C_k=\sum_{s=2}^{N/2}F^k(s)$, for each segment separation $s$ to occur and thus
+with $C_k=\sum_{s=2}^{N/2}F^k(s)$, to only remove links that are still in the pool. This leads to the recursion
 \begin{equation}
     F_k(s)= \frac{1}{N/2-2} \sum_{s_k=2}^{N/2} {\begin{cases}
-     F_{k-1}(s)-2(s-1) P_k(s) \text{ , for } s<s_k\\
-     F_{k-1}(s)-2(s_k -1)P_k(s)\text{ , for } s\geq s_k
-    \end{cases}}.
+     F_{k-1}(s)-2(s-1) P_{k-1}(s) \text{ , for } s<s_k\\
+     F_{k-1}(s)-2(s_k -1)P_{k-1}(s)\text{ , for } s\geq s_k
+    \end{cases}},
 \end{equation}
-This can be simplified to
+which can be simplified to
 \begin{align}
@@
 \end{equation}
-where $0<\alpha<1$ \red{why cant $\alpha$ be >1?}.
+where $0<\alpha<<1$ \red{, I think this is because we want $\alpha$ to be a small correction, so in fact we'd want it to be much smaller than 1}.
 
 \section*{Data and simulations}
- For large values of the chain length $N$ we thus find an approximate power law probability distribution for the sequence separation with an exponent of -1 \red{include note of correction}.
+ For large values of the chain length $N$ we thus find an approximate power law probability distribution for the sequence separation with an exponent of $-1+\alpha$.
 Here we use simulations, as introduced in \cite{molkenthin2020self} and networks extracted from measured protein data from the PDB \cite{PDB} to compare this behaviour in realistic 3D settings and show that the basic principle persists, indicating that the sequence distance separation distribution is largely determined by simple geometric constraints.
+
+ \red{write about how we picked out the specific structures to analyze, (i think we established that 200 to 400 is a very common range of lengths for proteins and i guess we will select them randomly, drawing from "typical" proteins or something), mention how many structures were analyzed and all that stuff.}
 
@@
 I'm not sure we want to use that whole $\alpha$ argument too much though as it's a but ad-hoc and would be nicer to just fit $A$ and be done with it, but then it may look a bit more off.
+\red{Should I remove the fits?
+(grey line and blue line) and only leave in the theory line?
+Also axes labels and stuff.}
         }
-        \label{fig:time_scaling}
+        \label{fig:sdd}
 \end{figure}
-\red{We observe that, while the general shape seems to match, in real proteins sequence separations of 20 to 60 amino acids are slightly over represented.
-We need to comment on that somehow.
-Also links of length 2 and 3 are over represented, we need to discuss that too.}
+We observe that, while the general shape seems to match, in real proteins sequence separations of 2 to 4 as well as 20 to 50 amino acids are slightly over-represented (see Fig.\ref{fig:sdd}).
 
-\red{To check if the bump has something to do with the secondary structure that is present in real proteins but not in the model, we have generated a sample of intrinsically disordered proteins, that also do not contain secondary structures.
-The bump does indeed disappear.
-The over representation of 2 and 3 does not though and there os probably a simple explanation for that and maybe we can simply take those points out and redo the mean degree matching and get to lines that are not just parallel but actually on top of one another that way.}
+The over-representation of very short sequence separations is an artefact of the different network construction for simulated networks and PRNs.
+While in the simulation all links are equally long and links are made until no longer possible, the parameter, determining the connectivity in PRNs is the cutoff distance $d_c$, which is chosen such that the number of links in the PRN is comparable to those in a simulated network of the same $N$.
+This leads to a larger range of bond angles randomly resulting in connections for small $s$ and thus an over-representation of those links.
+
+We expect the mid-separation bump to be the result of secondary structure effects.
+While in the simulation all links are made with equal probability apart from geometric constraints, real protein chains have preferred bond angles leading to the energetic preference for $\alpha-$helix and $\beta-$sheet structures, within which connections only short connections can occur.
+Longer connections thus tend to be between different helices.
+
+\red{With helices ranging in length from a minimum of about 2 to 3 turns or 7-11 residues up to beyond 30 residues \cite{manning1988circular}, with the (rarer) longer helices accounting relatively for more connections, sequence separations of 20 to 50 seem likely to result from connections between subsequent helices.
+(I guess we could test this somehow, though i'm not sure how I would go about interpreting the results.
+I think the PDB files contain information on which residues belong to which structure, so we could do an analysis of how many connections of $s>4$ there are a) within one helix b) between a helix and the following loop) c) between 2 subsequent helices etc.
+Where I would count helices and sheets the same...)}
+\begin{figure}[h]
+        \centering
+	\includegraphics[width=0.35\textwidth]{paper/figures/neighbouring_helices.pdf}
+        \caption{The average length of a link between two subsequent helices is the average length of a helix plus the average length of a loop.
+\red{Does this suffice or do we want a fancy 3d image of this?
+If so, do you have the software that makes those?}
+        }
+        \label{fig:bumpexplain}
+\end{figure}
+
+For a more direct comparison between simulation results and PRN's, we have generated a sample of intrinsically disordered proteins \red{cite something}, that also do not contain secondary structures.
+The results are shown in Fig.~\ref{fig:sdd_idp}.
+\red{Maybe write how we found those samples and how many there are and all that.}
+We find that the over representation of connections with $20<s<50$ disappears, further confirming our expectation.
 \begin{figure}[h]
@@
         }
-        \label{fig:time_scaling}
+        \label{fig:sdd_idp}
 \end{figure}
+\red{Anything else we want to say to that?
+Maybe write down more fit parameters?
+Somehow i'm reluctant to do that because we wouldnt learn anything new from that, that i can think of and it doesnt look great number wise i guess.}
 \section*{Conclusion}
-\begin{enumerate}
-    \item We can now explain the sequence separation from geometry and even do it analytically.
-    \item Maybe something about helixes and sheets and the dip.
-\end{enumerate}
-TO DO: Leave until the rest is done.
+We have demonstrated in analytical calculations and direct simulations how a random linking models with geometric constraints, such as the ones introduced in \cite{molkenthin2016scaling, molkenthin2020self} generate sequence distance distributions, that resemble a power-law, similar to the one found in folded proteins.
+This feature of protein network structure has been used to generate protein-like structures in \cite{bartoli2008effect}.
+
+For proteins with a high percentage of $\alpha$-helix structures, however, we find an over-representation of links with a sequence separation of 20 to 50 residues, implying a higher number of connections between neighbouring helices.
+We show that this is indeed the case by analyzing the sequence separation distributions of intrinsically disordered proteins.
+As helix structures arise from local bond angle preferences \cite{Danielsson2010}, rather than geometric constraints, these are not replicated in the model and thus the bump is absent in our theoretical calculations as well as simulations.
+
+\red{This makes the sequence separation distribution an indicator for the disorderedness of a protein structure.
+If such a thing is interesting we could try to formalize this measure and see how good it is at identifying idp's, but i'm guessing they are quite easy to identify already.}
+
+% \begin{enumerate}
+%     \item Bartoli et al. made a random network model based on a power-law sequence separation distribution \cite{bartoli2008effect}
+%     \item Here we show how such a distribution can be explained directly from geometric constraints
+%     \item In the analysis of realistic sequence separation distributions from PRN's we find that the secondary structure leads to an over representation of $20<s<50$, which we can explain as connections between subsequent helices.
+%     \item The bump is absent for idp's
+%     \item This makes the sequence separation distribution an indicator for the disorderedness of a protein structure.
+If such a thing is interesting we could try to formalize this measure and see how good it is at identifying idp's, but i'm guessing they are quite easy to identify already.
+% \end{enumerate}
 \section*{Declarations}

### 2020-05-26 pdb_linklength.py
Updates from Overleaf

@@ -1,0 +1,237 @@
+# -*- coding: utf-8 -*-
+from __future__ import print_function
+import networkx as nx
+import scipy as sp
+import scipy.stats
+from scipy.optimize import curve_fit
+import matplotlib.pyplot as plt
+import matplotlib.colors as colors
+import numpy as np
+from Bio.PDB import *
+from MDAnalysis import *
+import urllib
+import os
+import seaborn as sns
+import glob
+
+sns.set(font_scale=1.5)  # crazy big
+sns.set_style("ticks")
+sns.despine()
+
+def func(x, a,b):
+     return a * x**(-b)
+
+def func2(x, a):
+     return a / x
+
+def theofunc(x, C,N,ex):
+     return (C)* np.power(N/(2*(-x**2+N*x-N)),ex)
+
+#def theofunc(x, C,N):
+#     return (C)* N/(2*(N*x-N))
+
+names_xray = glob.glob("pdb_data2/all_xray/*.pdb")
+names_xray = glob.glob("pdb_data2/xray_data_short/*.pdb")
+names_xray = glob.glob("pdb_data2/xray_data/*.pdb")
+
+names_nmr = glob.glob("pdb_data2/nmr_data/*.pdb")
+names_nmr = glob.glob("pdb_data2/nmr_data_short/*.pdb")
+
+names_both = glob.glob("pdb_data2/both/*.pdb")
+
+#names_both = glob.glob("idp/*.pdb")
+
+names_used=names_both
+
+num=28
+lengths=[3,4,5,6,7,8,9,10,11,12,13,14,16,17,19,21,23,25,28,30,33,36,40,44,48,52,58,63,69,76,83,91,100,110,120,132,145,158,174,191,209,229,251,275,
+    302,331,363,398]#[0:45]
+print(len(lengths))
+
+numnum=20
+logscale=np.logspace(np.log10(3), np.log10(1000), num=numnum)
+print('logscale',logscale)
+results=[]
+binnedresults=[]
+
+if __name__ == '__main__':
+    directories = ['pdb_data2/both']
+    #directories = ['idp']
+    nameslen=len(names_used)
+    #nameslen=50
+    path = os.getcwd()
+    meanies=[]
+    linklengths=[]
+    for d in directories:
+        print(d)
+        output = os.path.join(path,d)
+        count=0
+        n=1000
+        t=8.0
+        counter=0
+        for counter in range(nameslen):
+            #pdb_file = os.path.join(output,pdb.strip())+'.pdb'
[…]

### 2020-05-26 src/pdb_linklength.py

@@ -1,254 +1,0 @@
-# -*- coding: utf-8 -*-
-from __future__ import print_function
-
-import glob
-import os
-
-import matplotlib.pyplot as plt
-import networkx as nx
-import numpy as np
-import scipy as sp
-import scipy.stats
-import seaborn as sns
-from Bio.PDB import *
-from MDAnalysis import *
-from scipy.optimize import curve_fit
-import pickle
-
-sns.set(font_scale=1.5)  # crazy big
-sns.set_style("ticks")
-sns.despine()
-
-
-def func(x, a, b):
-    return a * x ** (-b)
-
-
-def func2(x, a):
-    return a / x
-
-
-def theofunc(x, C, N, ex):
-    return (C) * np.power(N / (2 * (-x ** 2 + N * x - N)), ex)
-
-
-def connected_component_subgraphs(G):
-    for c in nx.connected_components(G):
-        yield G.subgraph(c)
-
-
-# def theofunc(x, C,N):
-#     return (C)* N/(2*(N*x-N))
-
-names_xray = glob.glob("../data/pdb_data2/all_xray/*.pdb")
-names_xray = glob.glob("../data/pdb_data2/xray_data_short/*.pdb")
-names_xray = glob.glob("../data/pdb_data2/xray_data/*.pdb")
-
-names_nmr = glob.glob("../datapdb_data2/nmr_data/*.pdb")
-names_nmr = glob.glob("../data/pdb_data2/nmr_data_short/*.pdb")
-
-names_both = glob.glob("../data/pdb_data2/both/*.pdb")
-
-# names_both = glob.glob("idp/*.pdb")
-
-names_used = names_both
-
-num = 5
-lengths = [3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 16, 17, 19, 21, 23, 25, 28, 30, 33, 36, 40, 44, 48, 52, 58, 63, 69,
-           76, 83, 91, 100, 110, 120, 132, 145, 158, 174, 191, 209, 229, 251, 275,
-           302]#, 331, 363, 398]  # [0:45]
-print(len(lengths))
-
-numnum = 20
-logscale = np.logspace(np.log10(3), np.log10(1000), num=numnum)
-print('logscale', logscale)
-results = []
-binnedresults = []
-
-if __name__ == '__main__':
-    directories = ['../data/pdb_data2/both']
-    # directories = ['idp']
-    nameslen = len(names_used)
-    # nameslen=50
-    path = os.getcwd()
[…]

### 2020-05-26 paper/main.tex

@@
 If so, do you have the software that makes those?}
+        \green{I can make a prettier thing provided we can actually make this argument based on the pdb data. I still struggle make the pdb data comply.}
         }
@@
 
-For a more direct comparison between simulation results and PRN's, we have generated a sample of intrinsically disordered proteins \red{cite something}, that also do not contain secondary structures.
+For a more direct comparison between simulation results and PRNs, we have generated a sample of intrinsically disordered proteins \red{cite something}, that also do not contain secondary structures.
 The results are shown in Fig.~\ref{fig:sdd_idp}.
+In order to compare the 3D protein model simulations to PRNs, we selected protein structure lengths based on frequency and comparable length to the simulations.
+Fig.~\ref{fig:pdb_stats} shows that most proteins in the PDB have chain lengths between 100-400 protein residues.
+Therefore using example lengths of 200 and 400 residues seemed appropriate.
+From each of the over 20000 structures around the 200 and 400 residue length, we made the following choices for selecting different types of structures:
+\red{now fill in what we actually did once we finalise this}.
+
+\begin{figure}[h]
+        \centering
+	\includegraphics[width=0.45\textwidth]{figures/pdb_statistics.pdf}
+        \caption{Statistics on all PDB entries based on the number of amino acid residues, with most entries found in the interval of 100-400 residues as of May 26th 2020. 
+        }
+        \label{fig:pdb_stats}
+\end{figure}
 \red{Maybe write how we found those samples and how many there are and all that.}

### 2020-08-04 paper/main.tex

@@
 \newcommand{\green}[1]{\textcolor{green!70!black}{#1}}
+\newcommand{\gray}[1]{\textcolor{gray!80!black}{#1}}
 \usepackage{mathtools}
@@
 \title{The role of geometry in the chain separation distribution of links in protein contact networks}
-\title{\red{We need a better title, the old one is too long}}
-%\title{}
+\title{\red{We need a better title, the old one is too long} \green{Yes! But no better ideas at the moment}}
+\title{\red{Chain separation distribution in protein contact networks}}
 %\title{}
 \author{Nora Molkenthin}
-\affiliation{Potsdam Institut für Klimafolgenforschung, Potsdam}
-\affiliation{Center for Advancing Electronics Dresden (cfaed), Technical University of Dresden, 01069 Dresden}
+\affiliation{Potsdam Institut für Klimafolgenforschung}
+\affiliation{Center for Advancing Electronics Dresden (cfaed), Technical University of Dresden}
 \author{Steffen Mühle}
@@
 Here we show, analytically and in simulations, that a geometrical stochastic model of folding chains can explain this behaviour.
-%Separating the data in $\alpha$-dominated, $\beta$-dominated and intrinsically disordered proteins (IDP's) shows characteristic structures in the distribution due to the secondary structure.
 \end{abstract}
@@
 antibodies), and hormonal regulation (e.g.
-insulin)~\cite{}.
-All natural proteins are made up of 20 different amino acids which dictate the three dimensional conformations the proteins adopt in order to function~\cite{}.
+insulin)~\cite{something}.
+All natural proteins are made up of 20 different amino acids which dictate the three dimensional conformations the proteins adopt in order to function~\cite{stuff}.
 One way of looking at the functional forms of proteins and classifying their structural properties is by looking at protein residue or contact networks~\cite{Vendruscolo2002,DiPaola2013,Estrada2011}.
@@
 From a physicist perspective it is interesting to understand the emergent behaviour in protein structure from PCNs, by investigating common network measures and their behaviour for PCNs.
-For example Bartoli et al.~\cite{bartoli2008effect} assume that proteins have a sequence separation distribution that follows $\frac{1}{\ell}$, where $\ell$ is the separation of two connected amino acids along the backbone chain as the simplest implementation of a connection probability, that decreases with the distance along the sequence.
+For example Bartoli et al.~\cite{bartoli2008effect} build a model in which they assume that proteins have a sequence separation distribution that follows $\frac{1}{\ell}$, where $\ell$ is the separation of two connected amino acids along the backbone chain as the simplest implementation of a connection probability, that decreases with the distance along the sequence.
 However, they do not provide a quantitative explanation for this assumption.
 
-Here, we will show how a simple geometrical model can capture this behaviour well, even providing a simple analytical approximation.
-The analytical approximation and geometrical model build on the idea that inherently geometrical objects, such as amino acids, imply that any resulting PCN network ensemble modelling them has to be spatially embedded~\cite{}.
-Different approaches have been used in the past for geometrical models describing protein folding, some of which derived characteristics of the secondary structure from constraints on bond and torsion angles~\cite{Danielsson2010,Molkenthin2011} and others focused on the formation mechanisms of the tertiary structure~\cite{molkenthin2016scaling, molkenthin2020self}.
-Based on such considerations, we have constructed a network formation process, in which random connections are added to a chain of unit disks, without causing them to intersect~\cite{molkenthin2016scaling}.
-Such a model can now be used to investigate the emergenct behaviour of sequence separation distributions as observed by Bartoli et al.~\cite{bartoli2008effect}.
-To this end, we analyze the sequence separation distribution that emerge from the geometric model, where the length $\ell$ of a link is defined as the distance of the two ends along the original chain.
+Here, we show how a simple geometrical model can reproduce this heuristically found behaviour well, even providing a simple analytical approximation.
+The analytical approximation and geometrical model build on the idea that inherently geometrical objects, such as amino acids, imply that any resulting PCN network ensemble modelling them has to be spatially embedded~\cite{molkenthin2016scaling, molkenthin2020self}.
+Different approaches have been used in the past for geometrical models describing protein folding, some of which derived characteristics of the secondary structure from constraints on bond and torsion angles~\cite{maybe cite Flory stuff here as well?,Danielsson2010,Molkenthin2011} and others focused on the formation mechanisms of the tertiary structure~\cite{molkenthin2016scaling, molkenthin2020self}.
+
+Based on such considerations, the network formation process, repeatedly \red{(this may not be the adverb i'm looking for, but subsequently also doesnt quite work)} connects random pairs of units, without causing any intersections along the chain ~\cite{molkenthin2016scaling}.
+Such a model can now be used to investigate the emergent behaviour of sequence separation distributions as observed by Bartoli et al.~\cite{bartoli2008effect}.
+To this end, we analyze the sequence separation distributions that emerge from the geometric model, where the \emph{sequence separation} $\ell$ of a link is defined as the distance of the two end units along the original chain.
 This is then compared to the sequence separation distribution extracted from protein residue networks (PRN) generated from protein data bank data.
 
-\todo[inline, color={green!40}]{@Nora: Summary of our findings and why this is cool.}
+\red{We find that, despite the highly simplified model, the generated sequence distance distribution is in agreement with the one observed in networks generated from PDB data and roughly consistent with the heuristic $1/\ell$ finding.
+The analytical approximation provides a correction term on top of the $1/\ell$ heuristic.}
 
@@
 I'm not sure we want to use that whole $\alpha$ argument too much though as it's a but ad-hoc and would be nicer to just fit $A$ and be done with it, but then it may look a bit more off.
-\red{Should I remove the fits?
-(grey line and blue line) and only leave in the theory line?
-Also axes labels and stuff.}
+\red{Redo figure fancier, with more data, axis labels and possibly different lines (remove fit lines, maybe add just one 1/l heuristic line or something).}
         }
@@
 \end{figure}
-We observe that, while the general shape seems to match, in real proteins sequence separations of 2 to 4 as well as 20 to 50 amino acids are slightly over-represented (see Fig.\ref{fig:sdd}).
+We observe that, while the general shape seems to match, in real proteins sequence separations of 2 to 4 as well as 20 to 50 amino acids are slightly over-represented (see Fig.\ref{fig:sdd}) \red{I have a vague memory of figures with more data where the hump was less pronounced?
+Depending on this rephrase the above.}.
 
@@
 
-We expect the mid-separation bump to be the result of secondary structure effects.
+\gray{We expect the mid-separation bump to be the result of secondary structure effects.
 While in the simulation all links are made with equal probability apart from geometric constraints, real protein chains have preferred bond angles leading to the energetic preference for $\alpha-$helix and $\beta-$sheet structures, within which connections only short connections can occur.
 Longer connections thus tend to be between different helices.
-
-\red{With helices ranging in length from a minimum of about 2 to 3 turns or 7-11 residues up to beyond 30 residues \cite{manning1988circular}, with the (rarer) longer helices accounting relatively for more connections, sequence separations of 20 to 50 seem likely to result from connections between subsequent helices.
+With helices ranging in length from a minimum of about 2 to 3 turns or 7-11 residues up to beyond 30 residues \cite{manning1988circular}, with the (rarer) longer helices accounting relatively for more connections, sequence separations of 20 to 50 seem likely to result from connections between subsequent helices.
 (I guess we could test this somehow, though i'm not sure how I would go about interpreting the results.
@@
 	\includegraphics[width=0.35\textwidth]{paper/figures/neighbouring_helices.pdf}
-        \caption{The average length of a link between two subsequent helices is the average length of a helix plus the average length of a loop.
+        \caption{\gray{The average length of a link between two subsequent helices is the average length of a helix plus the average length of a loop.
 \red{Does this suffice or do we want a fancy 3d image of this?
 If so, do you have the software that makes those?}
-        \green{I can make a prettier thing provided we can actually make this argument based on the pdb data. I still struggle make the pdb data comply.}
+        \green{I can make a prettier thing provided we can actually make this argument based on the pdb data. I still struggle make the pdb data comply.}}
         }
@@
 
-For a more direct comparison between simulation results and PRNs, we have generated a sample of intrinsically disordered proteins \red{cite something}, that also do not contain secondary structures.
+\gray{For a more direct comparison between simulation results and PRNs, we have generated a sample of intrinsically disordered proteins \cite{something}, that also do not contain secondary structures.
 The results are shown in Fig.~\ref{fig:sdd_idp}.
@@
 From each of the over 20000 structures around the 200 and 400 residue length, we made the following choices for selecting different types of structures:
-\red{now fill in what we actually did once we finalise this}.
+...}
 
@@
         \caption{Statistics on all PDB entries based on the number of amino acid residues, with most entries found in the interval of 100-400 residues as of May 26th 2020.
+\red{should this figure be before the current fig 2?
+Because it's the reason we do 200 +- and 400 +-}
         }
@@
 \end{figure}
-\red{Maybe write how we found those samples and how many there are and all that.}
-We find that the over representation of connections with $20<s<50$ disappears, further confirming our expectation.
+
+\gray{We find that the over representation of connections with $20<s<50$ disappears, further confirming our expectation.}
 \begin{figure}[h]
@@
 	\includegraphics[width=0.45\textwidth]{figures/idp_400.pdf}
-        \caption{Sequence distance distributions for intrinsically disordered proteins and simulated folds for a) Chain lengths around 200 b) chain lengths around 400.
+        \caption{\gray{Sequence distance distributions for intrinsically disordered proteins and simulated folds for a) Chain lengths around 200 b) chain lengths around 400.
 The theoretical curve for 2D rings works okay for 3D open chains (because the rings are closed sequence separations of more than N/2 are a bit meaningless and thus not accounted for) The parameters used in this are $A=80$ and $\alpha=0.2$ in a) and $0.3$ in b).
 I thought in b) it makes sense for $\alpha$ to be larger as there are even larger exponents $k$ coming into relevance.
-I'm not sure we want to use that whole $\alpha$ argument too much though as it's a but ad-hoc and would be nicer to just fit $A$ and be done with it, but then it may look a bit more off.
+I'm not sure we want to use that whole $\alpha$ argument too much though as it's a but ad-hoc and would be nicer to just fit $A$ and be done with it, but then it may look a bit more off.}
         }
@@
 \end{figure}
-\red{Anything else we want to say to that?
-Maybe write down more fit parameters?
-Somehow i'm reluctant to do that because we wouldnt learn anything new from that, that i can think of and it doesnt look great number wise i guess.}
+\red{Anything else we want to say to that? This is a bit short now}
 \section*{Conclusion}
@@
 
-For proteins with a high percentage of $\alpha$-helix structures, however, we find an over-representation of links with a sequence separation of 20 to 50 residues, implying a higher number of connections between neighbouring helices.
+\red{I remember we agreed on just deleting all the stuff about idp's beause the bump wasn't very pronounced once we used more data and thus the whole narrative of helix lengths and stuff wasn't particularly convincing.
+Additionally this means less work and thus sooner publication, so we were actually quite happy about it.
+But then we need to come up with something to write in the conclusion, so far it's a bit short.}
+
+\gray{For proteins with a high percentage of $\alpha$-helix structures, however, we find an over-representation of links with a sequence separation of 20 to 50 residues, implying a higher number of connections between neighbouring helices.
 We show that this is indeed the case by analyzing the sequence separation distributions of intrinsically disordered proteins.
-As helix structures arise from local bond angle preferences \cite{Danielsson2010}, rather than geometric constraints, these are not replicated in the model and thus the bump is absent in our theoretical calculations as well as simulations.
+As helix structures arise from local bond angle preferences \cite{Danielsson2010}, rather than geometric constraints, these are not replicated in the model and thus the bump is absent in our theoretical calculations as well as simulations.}
 
-\red{This makes the sequence separation distribution an indicator for the disorderedness of a protein structure.
+\gray{This makes the sequence separation distribution an indicator for the disorderedness of a protein structure.
 If such a thing is interesting we could try to formalize this measure and see how good it is at identifying idp's, but i'm guessing they are quite easy to identify already.}

### 2020-08-04 paper/main.tex
Updates from Overleaf

@@
 %\usepackage[a4paper, total={6in, 8in}]{geometry}
-\documentclass[reprint,amsmath,amssymb,rmp,onecolumn,notitlepage,11pt]{revtex4-1}
+%\documentclass[reprint,amsmath,amssymb,rmp,onecolumn,notitlepage,11pt]{revtex4-1}
+\documentclass[
+reprint,
+twocolumn,
+%preprint
+%onecolumn,
+amsmath,amssymb,superscriptaddress,aps,
+pre]{revtex4-1}
 \usepackage[utf8]{inputenc}
@@
 \begin{document}
-\title{The role of geometry in the chain separation distribution of links in protein contact networks}
-\title{\red{We need a better title, the old one is too long} \green{Yes! But no better ideas at the moment}}
-\title{\red{Chain separation distribution in protein contact networks}}
+\title{Chain separation distribution in protein contact networks}
+\title{Theoretical approximation of chain separation distribution in protein contact networks}
+\title{Theory of chain separation distributions gives insights in real protein contact networks}
 %\title{}
@@
 To this end, we analyze the sequence separation distributions that emerge from the geometric model, where the \emph{sequence separation} $\ell$ of a link is defined as the distance of the two end units along the original chain.
-This is then compared to the sequence separation distribution extracted from protein residue networks (PRN) generated from protein data bank data.
+This is then compared to the sequence separation distribution extracted from protein contact networks generated from protein data bank data.
 
@@
 \end{equation}
-Adjacency matrices are commonly used to describe networks and in this case represent the PRN.
+Adjacency matrices are commonly used to describe networks and in this case represent the PCN.
 
 The underlying formation mechanisms of individual protein folds are incredibly complex and depend on the surrounding solvent as well as the specific amino acid sequence and their quantum-mechanical interactions.
-Here we want to study the ensemble of all proteins, however, and will thus "average out" such specifics.
+Here we want to study the ensemble of all proteins, however, and will thus "average" over such specifics.
 
@@
     \centering
-    \includegraphics[width=0.4\columnwidth]{figures/Fig1.pdf}
+    \includegraphics[width=0.7\columnwidth]{figures/Fig1.pdf}
     \caption{a) The sequence separation distribution is defined as the separation along the chain (blue) of two connected units (red).
@@
 with $C_k=\sum_{s=2}^{N/2}F^k(s)$, to only remove links that are still in the pool. This leads to the recursion
-\begin{equation}
-    F_k(s)= \frac{1}{N/2-2} \sum_{s_k=2}^{N/2} {\begin{cases}
+\begin{align}
+    F_k(s)&= \frac{1}{N/2-2} \nonumber \\
+    &\sum_{s_k=2}^{N/2} {\begin{cases}
      F_{k-1}(s)-2(s-1) P_{k-1}(s) \text{ , for } s<s_k\\
      F_{k-1}(s)-2(s_k -1)P_{k-1}(s)\text{ , for } s\geq s_k
-    \end{cases}},
-\end{equation}
+    \end{cases}}
+\end{align}
 which can be simplified to
 \begin{align}
-   F_k(s)&= \frac{1}{\frac{N}{2}-2} \frac{F_{k-1}(s)}{C_{k-1}}\left( \sum_{s_k=2}^{N/2}C_{k-1} - \sum_{s_k=2}^{s} 2(s_k-1) - \sum_{s_k=s+1}^{N/2} 2(s -1) \right) \nonumber \\
-   &= \frac{1}{\frac{N}{2}-2}\frac{F_{k-1}(s)}{C_{k-1}}\left(\left(\frac{N}{2}-2\right)C_{k-1} -(s^2-s)-2\left(\frac{N}{2}-(s+1)\right)(s-1)\right)  \nonumber \\
+   F_k(s)&= \frac{1}{\frac{N}{2}-2} \frac{F_{k-1}(s)}{C_{k-1}}\nonumber \\
+   &\left( \sum_{s_k=2}^{N/2}C_{k-1} - \sum_{s_k=2}^{s} 2(s_k-1) - \sum_{s_k=s+1}^{N/2} 2(s -1) \right) \nonumber \\
+   &= \frac{1}{\frac{N}{2}-2}\frac{F_{k-1}(s)}{C_{k-1}}\nonumber \\
+   &\left(\left(\frac{N}{2}-2\right)C_{k-1} -(s^2-s)
+   -2\left(\frac{N}{2}-(s+1)\right)(s-1)\right)  \nonumber \\
    &= F_{k-1}(s)\left(1-\frac{-s^2 +(N-1)s-(N+1)}{(\frac{N}{2}-2)C_{k-1}} \right)\nonumber \\
@@
 However, in reality we know $1/C_k$ to increase with $k$, giving large $k$ summands relatively more weight than in a geometric series.
-This would lead to a correction in the exponent.
+This can be expressed as a small correction $\alpha$ in the exponent.
 
@@
 \end{equation}
-where $0<\alpha<<1$ \red{, I think this is because we want $\alpha$ to be a small correction, so in fact we'd want it to be much smaller than 1}.
+where $0<\alpha \ll 1$.
 
@@
  For large values of the chain length $N$ we thus find an approximate power law probability distribution for the sequence separation with an exponent of $-1+\alpha$.
-Here we use simulations, as introduced in \cite{molkenthin2020self} and networks extracted from measured protein data from the PDB \cite{PDB} to compare this behaviour in realistic 3D settings and show that the basic principle persists, indicating that the sequence distance separation distribution is largely determined by simple geometric constraints.
+Here we use simulations, as introduced in \cite{molkenthin2020self} and networks extracted from measured protein data from the PDB \cite{PDB} to compare this behaviour in realistic 3D settings and show that the basic principle persists, indicating that much of the sequence distance separation distribution can be explained by simple geometric constraints.
+ \begin{figure}[h]
+        \centering
+	\includegraphics[width=0.45\textwidth]{figures/pdb_statistics.pdf}
+        \caption{Statistics on all PDB entries based on the number of amino acid residues, with most entries found in the interval of 100-400 residues as of May 26th 2020.}
+        \label{fig:pdb_stats}
+\end{figure}
 
@@
 
-\begin{figure}[h]
+\begin{figure*}[t]
         \centering
-	\includegraphics[width=0.45\textwidth]{figures/both_200.pdf}
-	\includegraphics[width=0.45\textwidth]{figures/both_400.pdf}
-        \caption{Sequence distance distributions for real and simulated folds for a) Chain lengths around 200 b) chain lengths around 400.
-The theoretical curve for 2D rings works okay for 3D open chains (because the rings are closed sequence separations of more than N/2 are a bit meaningless and thus not accounted for) The parameters used in this are $A=80$ and $\alpha=0.2$ in a) and $0.3$ in b).
-I thought in b) it makes sense for $\alpha$ to be larger as there are even larger exponents $k$ coming into relevance.
-I'm not sure we want to use that whole $\alpha$ argument too much though as it's a but ad-hoc and would be nicer to just fit $A$ and be done with it, but then it may look a bit more off.
-\red{Redo figure fancier, with more data, axis labels and possibly different lines (remove fit lines, maybe add just one 1/l heuristic line or something).}
+	\includegraphics[width=\textwidth]{paper/figures/resultsfig_old.pdf}
+        \caption{Sequence distance distributions for real and simulated folds for a) Chain lengths around 200 b) chain lengths around 300 c) chain lengths around 400.
         }
         \label{fig:sdd}
-\end{figure}
+\end{figure*}
 We observe that, while the general shape seems to match, in real proteins sequence separations of 2 to 4 as well as 20 to 50 amino acids are slightly over-represented (see Fig.\ref{fig:sdd}) \red{I have a vague memory of figures with more data where the hump was less pronounced?
@@
 
-\gray{We expect the mid-separation bump to be the result of secondary structure effects.
-While in the simulation all links are made with equal probability apart from geometric constraints, real protein chains have preferred bond angles leading to the energetic preference for $\alpha-$helix and $\beta-$sheet structures, within which connections only short connections can occur.
-Longer connections thus tend to be between different helices.
-With helices ranging in length from a minimum of about 2 to 3 turns or 7-11 residues up to beyond 30 residues \cite{manning1988circular}, with the (rarer) longer helices accounting relatively for more connections, sequence separations of 20 to 50 seem likely to result from connections between subsequent helices.
-(I guess we could test this somehow, though i'm not sure how I would go about interpreting the results.
-I think the PDB files contain information on which residues belong to which structure, so we could do an analysis of how many connections of $s>4$ there are a) within one helix b) between a helix and the following loop) c) between 2 subsequent helices etc.
-Where I would count helices and sheets the same...)}
-\begin{figure}[h]
-        \centering
-	\includegraphics[width=0.35\textwidth]{paper/figures/neighbouring_helices.pdf}
-        \caption{\gray{The average length of a link between two subsequent helices is the average length of a helix plus the average length of a loop.
-\red{Does this suffice or do we want a fancy 3d image of this?
-If so, do you have the software that makes those?}
-        \green{I can make a prettier thing provided we can actually make this argument based on the pdb data. I still struggle make the pdb data comply.}}
-        }
-        \label{fig:bumpexplain}
-\end{figure}
-
-\gray{For a more direct comparison between simulation results and PRNs, we have generated a sample of intrinsically disordered proteins \cite{something}, that also do not contain secondary structures.
-The results are shown in Fig.~\ref{fig:sdd_idp}.
-In order to compare the 3D protein model simulations to PRNs, we selected protein structure lengths based on frequency and comparable length to the simulations.
-Fig.~\ref{fig:pdb_stats} shows that most proteins in the PDB have chain lengths between 100-400 protein residues.
-Therefore using example lengths of 200 and 400 residues seemed appropriate.
-From each of the over 20000 structures around the 200 and 400 residue length, we made the following choices for selecting different types of structures:
-...}
-
-\begin{figure}[h]
-        \centering
-	\includegraphics[width=0.45\textwidth]{figures/pdb_statistics.pdf}
-        \caption{Statistics on all PDB entries based on the number of amino acid residues, with most entries found in the interval of 100-400 residues as of May 26th 2020.
-\red{should this figure be before the current fig 2?
-Because it's the reason we do 200 +- and 400 +-}
-        }
-        \label{fig:pdb_stats}
-\end{figure}
-
-\gray{We find that the over representation of connections with $20<s<50$ disappears, further confirming our expectation.}
-\begin{figure}[h]
-        \centering
-	\includegraphics[width=0.45\textwidth]{figures/idp_200.pdf}
-	\includegraphics[width=0.45\textwidth]{figures/idp_400.pdf}
-        \caption{\gray{Sequence distance distributions for intrinsically disordered proteins and simulated folds for a) Chain lengths around 200 b) chain lengths around 400.
-The theoretical curve for 2D rings works okay for 3D open chains (because the rings are closed sequence separations of more than N/2 are a bit meaningless and thus not accounted for) The parameters used in this are $A=80$ and $\alpha=0.2$ in a) and $0.3$ in b).
-I thought in b) it makes sense for $\alpha$ to be larger as there are even larger exponents $k$ coming into relevance.
-I'm not sure we want to use that whole $\alpha$ argument too much though as it's a but ad-hoc and would be nicer to just fit $A$ and be done with it, but then it may look a bit more off.}
-        }
-        \label{fig:sdd_idp}
-\end{figure}
-\red{Anything else we want to say to that? This is a bit short now}
 \section*{Conclusion}
@@
 
-\gray{For proteins with a high percentage of $\alpha$-helix structures, however, we find an over-representation of links with a sequence separation of 20 to 50 residues, implying a higher number of connections between neighbouring helices.
-We show that this is indeed the case by analyzing the sequence separation distributions of intrinsically disordered proteins.
-As helix structures arise from local bond angle preferences \cite{Danielsson2010}, rather than geometric constraints, these are not replicated in the model and thus the bump is absent in our theoretical calculations as well as simulations.}
-
-\gray{This makes the sequence separation distribution an indicator for the disorderedness of a protein structure.
-If such a thing is interesting we could try to formalize this measure and see how good it is at identifying idp's, but i'm guessing they are quite easy to identify already.}
-
-% \begin{enumerate}
-%     \item Bartoli et al. made a random network model based on a power-law sequence separation distribution \cite{bartoli2008effect}
-%     \item Here we show how such a distribution can be explained directly from geometric constraints
-%     \item In the analysis of realistic sequence separation distributions from PRN's we find that the secondary structure leads to an over representation of $20<s<50$, which we can explain as connections between subsequent helices.
-%     \item The bump is absent for idp's
-%     \item This makes the sequence separation distribution an indicator for the disorderedness of a protein structure.
-If such a thing is interesting we could try to formalize this measure and see how good it is at identifying idp's, but i'm guessing they are quite easy to identify already.
-% \end{enumerate}
 \section*{Declarations}
@@
 \subsection{Acknowledgements}
-We thank ... for fruitful discussions.
+We thank Marc Timme for fruitful discussions.
 

### 2020-08-04 paper/main.tex
Updates from Overleaf

@@
     \centering
-    \includegraphics[width=0.7\columnwidth]{figures/Fig1.pdf}
+    \includegraphics[width=0.7\columnwidth]{figures/FigFig1.pdf}
     \caption{a) The sequence separation distribution is defined as the separation along the chain (blue) of two connected units (red).
@@
 
+Protein chain lengths in the PDB vary from small fragments of less than 10 amino acids to large agglomerates of over 2000 residues.
+The most frequently occurring chain lengths, however are between 100 and 400 residues as shown in Fig.~\ref{fig:pdb_stats}.
+For the analysis of our theoretical results, we have thus chosen ............
+structures of around 200, ........
+structures of around 300 and ..........
+structures of around 400 residues.
+
  \red{write about how we picked out the specific structures to analyze, (i think we established that 200 to 400 is a very common range of lengths for proteins and i guess we will select them randomly, drawing from "typical" proteins or something), mention how many structures were analyzed and all that stuff.}

### 2020-08-04 paper/main.tex

@@
 structures of around 400 residues.
-
- \red{write about how we picked out the specific structures to analyze, (i think we established that 200 to 400 is a very common range of lengths for proteins and i guess we will select them randomly, drawing from "typical" proteins or something), mention how many structures were analyzed and all that stuff.}
+Those have been converted to PCN's
 

### 2020-08-06 paper/main.tex
nora changes

@@
 \begin{abstract}
-The distribution of chain separations of interacting amino acids in proteins roughly follows a power law.
+The mechanisms by which the folded structures of proteins are determined has long been one of the key mysteries of Biophysics.
+Here we approach the problem from a network ensemble perspective, analyzing the relative frequencies with which residues at different distances along the chain come into contact.
+We define and analyze the residue distance distribution for measured protein structures and structures resulting from a geometrically constrained random interaction model and find that the distributions are compatible.
+We furthermore derive an analytic approximation for the two dimensional analog of the model and find this approximation to also describe the distribution.
+We therefore propose the conclusion that the approximate power-law behaviour of the residue distance distribution is largely caused by geometric constraints of chains rather than specifics of protein sequences.
+%The distribution of chain separations of interacting amino acids in proteins roughly follows a power law.
 Here we show, analytically and in simulations, that a geometrical stochastic model of folding chains can explain this behaviour.
@@
 From a physicist perspective it is interesting to understand the emergent behaviour in protein structure from PCNs, by investigating common network measures and their behaviour for PCNs.
-For example Bartoli et al.~\cite{bartoli2008effect} build a model in which they assume that proteins have a sequence separation distribution that follows $\frac{1}{\ell}$, where $\ell$ is the separation of two connected amino acids along the backbone chain as the simplest implementation of a connection probability, that decreases with the distance along the sequence.
+For example Bartoli et al.~\cite{bartoli2008effect} build a model in which they assume that proteins have a residue distance distribution that follows $\frac{1}{\ell}$, where $\ell$ is the separation of two connected amino acids along the backbone chain as the simplest implementation of a connection probability, that decreases with the distance along the sequence.
 However, they do not provide a quantitative explanation for this assumption.
@@
 
-Based on such considerations, the network formation process, repeatedly \red{(this may not be the adverb i'm looking for, but subsequently also doesnt quite work)} connects random pairs of units, without causing any intersections along the chain ~\cite{molkenthin2016scaling}.
-Such a model can now be used to investigate the emergent behaviour of sequence separation distributions as observed by Bartoli et al.~\cite{bartoli2008effect}.
-To this end, we analyze the sequence separation distributions that emerge from the geometric model, where the \emph{sequence separation} $\ell$ of a link is defined as the distance of the two end units along the original chain.
-This is then compared to the sequence separation distribution extracted from protein contact networks generated from protein data bank data.
+Based on such considerations, the network formation process, repeatedly connects random pairs of units, without causing any intersections along the chain ~\cite{molkenthin2016scaling}.
+Such a model can now be used to investigate the emergent behaviour of residue distance distributions as used heuristically by Bartoli et al.~\cite{bartoli2008effect}.
+To this end, we define and analyze the residue distance distributions $P(s)$ that emerge from the geometric model, where the \emph{residue distance} $s$ of a link is defined as the distance of the two end units along the original chain.
+This is then compared to the residue distance distribution extracted from protein contact networks generated from protein data bank data.
 
-\red{We find that, despite the highly simplified model, the generated sequence distance distribution is in agreement with the one observed in networks generated from PDB data and roughly consistent with the heuristic $1/\ell$ finding.
-The analytical approximation provides a correction term on top of the $1/\ell$ heuristic.}
+We find that, despite the highly simplified model, the generated residue distance distribution is in agreement with the one observed in networks generated from PDB data and roughly consistent with the heuristic $1/s$ finding.
+The analytical approximation provides a correction term on top of the $1/s$ heuristic.
 
@@
 \section*{Analytic approximation}
-Here we want to analytically approximate the sequence separation distribution for the simplified 2D model, that is the separation along the backbone chain of two nodes connected by a link, as illustrated in Fig~\ref{fig:schematic}~a).
+Here we want to analytically approximate the residue distance distribution for the simplified 2D model, that is the separation along the backbone chain of two nodes connected by a link, as illustrated in Fig~\ref{fig:schematic}~a).
 
@@
     \centering
-    \includegraphics[width=0.7\columnwidth]{figures/FigFig1.pdf}
-    \caption{a) The sequence separation distribution is defined as the separation along the chain (blue) of two connected units (red).
+    \includegraphics[width=0.7\columnwidth]{figures/Fig1/Fig1.pdf}
+    \caption{a) The residue distance distribution is defined as the separation along the chain (blue) of two connected units (red).
 b) The solid blue link prohibits some connections from node $A$ (dotted red), while others are still permitted (dotted blue).
@@
 
-Let us start by introducing the auxiliary variable $F(s)$, defined to be the number of possible links with a sequence separation of $s$.
-Before any links are added all sequence separations are equally likely, as there are $N$ possibilities of making a link of each separation $s$ with $2\leq s < N/2$.
+Let us start by introducing the auxiliary variable $F(s)$, defined to be the number of possible links with a residue distance of $s$.
+Before any links are added all residue distances are equally likely, as there are $N$ possibilities of making a link of each separation $s$ with $2\leq s < N/2$.
 
@@
 Adding a link of length $s_1$ prohibits $2(s-1)$ links of length $s<s_1$ and $2(s_1-1)$ links of length $s>s_1$ (see Fig.~\ref{fig:schematic}~b).
-To find the expected number of links still available in subsequent stops, we average over all possible sequence separations $s_1$ of the first link added.
+To find the expected number of links still available in subsequent stops, we average over all possible residue distances $s_1$ of the first link added.
 
@@
 
-The probability distribution of the realized sequence separations of all added links is then given by the average over the available pools at each link addition step, leading to
+The probability distribution of the realized residue distances of all added links is then given by the average over the available pools at each link addition step, leading to
 \begin{align}
@@
 \section*{Data and simulations}
- For large values of the chain length $N$ we thus find an approximate power law probability distribution for the sequence separation with an exponent of $-1+\alpha$.
-Here we use simulations, as introduced in \cite{molkenthin2020self} and networks extracted from measured protein data from the PDB \cite{PDB} to compare this behaviour in realistic 3D settings and show that the basic principle persists, indicating that much of the sequence distance separation distribution can be explained by simple geometric constraints.
+For large values of the chain length $N$ we thus find an approximate power law probability distribution for the residue distance with an exponent of $-1+\alpha$.
+Here we use simulations, as introduced in~\cite{molkenthin2020self} and networks extracted from measured protein data from the PDB \cite{PDB} to compare this behaviour in realistic 3D settings and show that the basic principle persists, indicating that much of the residue distance distribution can be explained by simple geometric constraints.
  \begin{figure}[h]
         \centering
-	\includegraphics[width=0.45\textwidth]{figures/pdb_statistics.pdf}
+	\includegraphics[width=0.45\textwidth]{paper/figures/Fig2/pdb_statistics.pdf}
         \caption{Statistics on all PDB entries based on the number of amino acid residues, with most entries found in the interval of 100-400 residues as of May 26th 2020.}
@@
 The most frequently occurring chain lengths, however are between 100 and 400 residues as shown in Fig.~\ref{fig:pdb_stats}.
-For the analysis of our theoretical results, we have thus chosen ............
+\red{For our analysis, we have thus chosen ............
 structures of around 200, ........
 structures of around 300 and ..........
-structures of around 400 residues.
-Those have been converted to PCN's
+structures of around 400 residues from the PDB.} Those have been converted to PCN's based on the measured locations of their residues to analyze the distribution of residue distances between connected amino acids.
+The resulting distribution is compared to those generated in simulations in~\cite{molkenthin2020self} as well as the above derived theoretical expression.
+The latter, however, is scaled with a factor to account for the different dimensionality, as well as approximations made in the derivation.
 
-\begin{figure*}[t]
+The results are shown logarithmically scaled in Fig.\ref{fig:sdd}.
+\red{For all three lengths, the distributions show an approximate power law decay with slopes of $...\pm ...$, which is consistent with both, the simulated results ans the theoretical approximation.}
+
+\begin{figure}[t]
         \centering
-	\includegraphics[width=\textwidth]{paper/figures/resultsfig_old.pdf}
-        \caption{Sequence distance distributions for real and simulated folds for a) Chain lengths around 200 b) chain lengths around 300 c) chain lengths around 400.
+	\includegraphics[width=0.45\textwidth]{paper/figures/Fig3/Fig3.pdf}
+        \caption{residue distance distributions for real and simulated folds for a) Chain lengths around 200 b) chain lengths around 300 c) chain lengths around 400.
+\red{Depending on what this looks like when we have all the data, perhaps adding error bars could be nice.
+Also, since i'm mentioning the overrepresentation of separations 2 to 4, maybe those should be shown.
+If it fits better we can take them out for the fits and stuff.}
         }
         \label{fig:sdd}
-\end{figure*}
-We observe that, while the general shape seems to match, in real proteins sequence separations of 2 to 4 as well as 20 to 50 amino acids are slightly over-represented (see Fig.\ref{fig:sdd}) \red{I have a vague memory of figures with more data where the hump was less pronounced?
-Depending on this rephrase the above.}.
+\end{figure}
 
-The over-representation of very short sequence separations is an artefact of the different network construction for simulated networks and PRNs.
-While in the simulation all links are equally long and links are made until no longer possible, the parameter, determining the connectivity in PRNs is the cutoff distance $d_c$, which is chosen such that the number of links in the PRN is comparable to those in a simulated network of the same $N$.
+We observe that, while the general shape seems to match, in real proteins residue distances of 2 to 4 amino acids are slightly over-represented (see Fig.\ref{fig:sdd}).
+The over-representation of very short residue distances is an artefact of the different network construction for simulated (and theoretically approximated) networks and PCNs.
+While in the simulation all links are equally long and links are made until no longer possible, the parameter, determining the connectivity in PCNs is the cutoff distance $d_c$, which is chosen such that the number of links in the PCN is comparable to those in a simulated network of the same $N$.
 This leads to a larger range of bond angles randomly resulting in connections for small $s$ and thus an over-representation of those links.
@@
 \section*{Conclusion}
-We have demonstrated in analytical calculations and direct simulations how a random linking models with geometric constraints, such as the ones introduced in \cite{molkenthin2016scaling, molkenthin2020self} generate sequence distance distributions, that resemble a power-law, similar to the one found in folded proteins.
-This feature of protein network structure has been used to generate protein-like structures in \cite{bartoli2008effect}.
+The tertiary structures of folded proteins have long been one of the most important mysteries of Biophysics.
+While for individual structures, detailed molecular dynamics or statistical models are essential in approaching these questions, general statements for the ensemble of folded proteins can be made with simple models.
 
-\red{I remember we agreed on just deleting all the stuff about idp's beause the bump wasn't very pronounced once we used more data and thus the whole narrative of helix lengths and stuff wasn't particularly convincing.
-Additionally this means less work and thus sooner publication, so we were actually quite happy about it.
-But then we need to come up with something to write in the conclusion, so far it's a bit short.}
+Here we have introduced and analyzed the residue distance $s$ and its probability distribution $P(s)$ in measured and simulated protein structures.
+For the equivalent model in two dimensions we have used a mean-field approach to derive an analytic approximation for the residue distance distribution, which appears to agree with the simulated and measured distributions.
+We have thus demonstrated in analytical calculations as well as direct simulations how a random linking model with geometric constraints, such as the ones introduced in \cite{molkenthin2016scaling, molkenthin2020self} generates a residue distance distribution, that resembles a power-law, similar to the one found in folded proteins.
+This feature of protein network structure has previously been used in the heuristic model to generate protein-like structures in \cite{bartoli2008effect}.
+
+Gaining a better understanding of the ensemble of folded protein structures can help guide the way to a better understanding of the constraints within which structures may occur.
+Together with an understanding of secondary structure principles, such as those in
+\cite{Danielsson2010, Molkenthin2011}, this can help to narrow down the complex energy landscapes and find paths through them more effectively in the future.
 

### 2020-08-06 src/get_link_lengths.py
adding long script

@@ -1,0 +1,112 @@
+
+
+# Imports
+import numpy as np
+import networkx as nx
+import pickle
+import scipy as sp
+import scipy.stats
+from scipy.optimize import curve_fit
+import glob
+import os
+from Bio.PDB import *
+from MDAnalysis import *
+import warnings
+import urllib
+import pandas as pd
+warnings.filterwarnings('ignore')
+
+
+
+def connected_component_subgraphs(G):
+    for c in nx.connected_components(G):
+        yield G.subgraph(c)
+
+# Get data from the pdb
+def do_analysis(calphas, link_length_list, t=8.0):
+    linklengths = []
+    r = calphas.positions
+    maxdist = 0
+    G = nx.empty_graph(len(r))
+    avd = 0
+    for i in range(len(r)):
+        for j in range(i):
+            dist = np.linalg.norm(r[i] - r[j])
+            maxdist = max(dist, maxdist)
+            avd += dist ** 2
+            if dist < t:
+                G.add_edge(i, j)
+                linklengths.append(dist)
+    avd = np.sqrt(avd) / (len(r))
+    # take largest component
+    if len(G.edges()) > 0:
+        G = list(connected_component_subgraphs(G))[0]
+        lena = len(G.nodes())
+        Li = list(set(G.edges()) - set(nx.cycle_graph(lena).edges()))
+        for link in Li:
+            #return(abs(link[0]-link[1]))
+            link_length_list.append(abs(link[0] - link[1]))
+
+if __name__ == "__main__":
+    # Reading the pdb IDs
+    query = pd.read_csv('../data/mix.txt')
+    hundred = []
+    twohundered = []
+    threehundred = []
+    fourhundred = []
+    ll_100 = []
+    ll_200 = []
+    ll_300 = []
+    ll_400 = []
+    for pdb in query.columns:
+        #print (pdb)
+        download = 'https://files.rcsb.org/download/%s.pdb' %pdb
+        #print(download)
+        try:
+            file_name = '../data/temp/'+pdb+'.pdb'
+            urllib.request.urlretrieve(download, file_name)
+        except:
+            continue
+        u = Universe(file_name)
+        calphas = u.select_atoms("name CA and segid " + u.residues.segments.segids[0])
+        chain_len = len(calphas)
+        print(chain_len)
+        if chain_len in range(85,115):
+            hundred.append(pdb)
[…]

### 2020-08-13 src/get_link_lengths.py
trying to fix things

@@ -69,7 +69,7 @@ if __name__ == "__main__":
     counter = 0
     n_pdbs = len(query.columns)
     with open("log.txt", "w") as log:
-        for pdb in query.columns:
+        for pdb in query.columns[:100]:
             if counter % 500 == 0:
                 print("We are at entry %d/%d!" % (counter, n_pdbs), flush=True)
                 save_current_data('100', ll_100, hundred)
@@ -79,8 +79,10 @@ if __name__ == "__main__":
             download = 'https://files.rcsb.org/download/%s.pdb' % pdb
             try:
                 file_name = '../data/temp/'+pdb+'.pdb'
-                urllib.request.urlretrieve(download, file_name)
+                bio.retrieve_pdb_file(pdb,pdir='../data/tmep/', file_format='pbd')
+                #urllib.request.urlretrieve(download, file_name)
             except Exception:
+                print("failed at %s" % pdb)
                 traceback.print_exc(file=log)
                 continue
             if os.path.isfile(file_name):

### 2020-08-13 src/get_link_lengths.py
adding final results

@@ -79,10 +79,10 @@ if __name__ == "__main__":
             download = 'https://files.rcsb.org/download/%s.pdb' % pdb
             try:
                 file_name = '../data/temp/'+pdb+'.pdb'
-                bio.retrieve_pdb_file(pdb,pdir='../data/tmep/', file_format='pbd')
-                #urllib.request.urlretrieve(download, file_name)
+                #bio.retrieve_pdb_file(pdb,pdir='../data/tmep/', file_format='pbd')
+                urllib.request.urlretrieve(download, file_name)
             except Exception:
-                print("failed at %s" % pdb)
+                print("failed at %s" % pdb, flush=True)
                 traceback.print_exc(file=log)
                 continue
             if os.path.isfile(file_name):

### 2020-08-13 src/get_link_lengths.py
small fix

@@ -56,7 +56,7 @@ def save_current_data(length, ll_list, id_list):
 
 if __name__ == "__main__":
     # Reading the pdb IDs
-    socket.setdefaulttimeout(15)
+    socket.setdefaulttimeout(20)
     query = pd.read_csv('../data/combined_ids.txt')
     hundred = []
     twohundered = []
@@ -69,7 +69,7 @@ if __name__ == "__main__":
     counter = 0
     n_pdbs = len(query.columns)
     with open("log.txt", "w") as log:
-        for pdb in query.columns[:100]:
+        for pdb in query.columns:
             if counter % 500 == 0:
                 print("We are at entry %d/%d!" % (counter, n_pdbs), flush=True)
                 save_current_data('100', ll_100, hundred)

### 2021-12-02 paper/main.tex
Updates from Overleaf

@@
 \affiliation{Center for Advancing Electronics Dresden (cfaed), Technical University of Dresden}
+\author{J. J.  Güven}
+\affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
 \author{Steffen Mühle}
@@
 \affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
-%\author{Marc Timme}
-% \affiliation{Network Dynamics, Max Planck Institute for Dynamics and Self-Organization (MPIDS), 37077 Göttingen, Germany}
-% \affiliation{Chair for Network Dynamics, Institute for Theoretical Physics and Center for Advancing Electronics Dresden (cfaed), Technical University of Dresden, 01069 Dresden}
 
 \begin{abstract}
-The mechanisms by which the folded structures of proteins are determined has long been one of the key mysteries of Biophysics.
-Here we approach the problem from a network ensemble perspective, analyzing the relative frequencies with which residues at different distances along the chain come into contact.
+The mechanisms by which the folded structures of proteins are determined have long been one of the key mysteries of Biophysics.
+Here we approach this problem from a network ensemble perspective, analyzing the relative frequencies with which residues at different distances along the chain come into contact.
 We define and analyze the residue distance distribution for measured protein structures and structures resulting from a geometrically constrained random interaction model and find that the distributions are compatible.
-We furthermore derive an analytic approximation for the two dimensional analog of the model and find this approximation to also describe the distribution.
+We furthermore derive an analytic approximation for the two dimensional analog of the model and find this approximation to also describe the same distribution.
 We therefore propose the conclusion that the approximate power-law behaviour of the residue distance distribution is largely caused by geometric constraints of chains rather than specifics of protein sequences.
@@
 antibodies), and hormonal regulation (e.g.
-insulin)~\cite{something}.
-All natural proteins are made up of 20 different amino acids which dictate the three dimensional conformations the proteins adopt in order to function~\cite{stuff}.
+insulin)~\cite{dill2008protein,Dill1042} \red{(Actually in the last paper we focused more on modelling of proteins rather than why they are so important so i didn't find much)}.
+All natural proteins are made up of 20 different amino acids which dictate the three dimensional conformations the proteins adopt in order to function~\cite{scheraga2007protein}.
 One way of looking at the functional forms of proteins and classifying their structural properties is by looking at protein residue or contact networks~\cite{Vendruscolo2002,DiPaola2013,Estrada2011}.
-Protein Contact Networks (PCN), have shown promise in understanding protein folding patterns as well as revealing allosteric communication pathways~\cite{https://pubs.acs.org/doi/10.1021/acs.jcim.9b00320, and others}.
-From a physicist perspective it is interesting to understand the emergent behaviour in protein structure from PCNs, by investigating common network measures and their behaviour for PCNs.
-For example Bartoli et al.~\cite{bartoli2008effect} build a model in which they assume that proteins have a residue distance distribution that follows $\frac{1}{\ell}$, where $\ell$ is the separation of two connected amino acids along the backbone chain as the simplest implementation of a connection probability, that decreases with the distance along the sequence.
+Protein Contact Networks (PCN) have shown promise in understanding protein folding patterns as well as revealing allosteric communication pathways~\cite{yao2019establishing, menichetti2016network,dokholyan2002topological} \red{(i found some papers that use PCN's but not necessarily for allosteric communication)}.
+From a physicist's perspective it is interesting to understand the emergent behaviour in protein structure from PCNs, by investigating common network measures and their behaviour for PCNs.
+For example Bartoli et al.~\cite{bartoli2008effect} build a model in which they assume that proteins have a residue distance distribution that follows $\frac{1}{s}$, where $s$ is the separation of two connected amino acids along the backbone chain (in the following called \emph{residue distance}) as the simplest implementation of a connection probability, thus decreasing with the distance along the sequence.
 However, they do not provide a quantitative explanation for this assumption.
 
-Here, we show how a simple geometrical model can reproduce this heuristically found behaviour well, even providing a simple analytical approximation.
-The analytical approximation and geometrical model build on the idea that inherently geometrical objects, such as amino acids, imply that any resulting PCN network ensemble modelling them has to be spatially embedded~\cite{molkenthin2016scaling, molkenthin2020self}.
-Different approaches have been used in the past for geometrical models describing protein folding, some of which derived characteristics of the secondary structure from constraints on bond and torsion angles~\cite{maybe cite Flory stuff here as well?,Danielsson2010,Molkenthin2011} and others focused on the formation mechanisms of the tertiary structure~\cite{molkenthin2016scaling, molkenthin2020self}.
+Here, we show how a simple geometrical model can reproduce this heuristically found behaviour well, and even provide a simple analytical approximation.
+The three-dimensional geometrical model and the analytical approximation of its two-dimensional analogue both build on the idea that inherently geometrical objects, such as amino acids, imply that any resulting PCN network ensemble modelling them has to be spatially embedded~\cite{molkenthin2016scaling, molkenthin2020self}.
+Different approaches have been used in the past for geometrical models describing protein folding, some of which derived characteristics of the secondary structure from constraints on bond and torsion angles~\cite{bhattacharjee2013flory,Danielsson2010,Molkenthin2011} and others focused on the formation mechanisms of the tertiary structure~\cite{molkenthin2016scaling, molkenthin2020self}.
 
-Based on such considerations, the network formation process, repeatedly connects random pairs of units, without causing any intersections along the chain ~\cite{molkenthin2016scaling}.
+Based on such considerations, the network formation process repeatedly connects random pairs of units, without causing any intersections along the chain ~\cite{molkenthin2016scaling}.
 Such a model can now be used to investigate the emergent behaviour of residue distance distributions as used heuristically by Bartoli et al.~\cite{bartoli2008effect}.
 To this end, we define and analyze the residue distance distributions $P(s)$ that emerge from the geometric model, where the \emph{residue distance} $s$ of a link is defined as the distance of the two end units along the original chain.
-This is then compared to the residue distance distribution extracted from protein contact networks generated from protein data bank data.
+This distribution is then compared to the residue distance distribution extracted from protein contact networks generated from protein data bank data.
 
@@
 The underlying formation mechanisms of individual protein folds are incredibly complex and depend on the surrounding solvent as well as the specific amino acid sequence and their quantum-mechanical interactions.
-Here we want to study the ensemble of all proteins, however, and will thus "average" over such specifics.
+Here we want to study the ensemble of all folded proteins, however, and will thus "average" over such specifics.
 
-Based on the basic assumption that amino acids are objects in space, that can not overlap indefinitely, we introduced a model in \cite{molkenthin2020self} starting from a chain of identical spheres, each in contact only with the neighbours it is connected to.
+Based on the basic observation that amino acids are objects in space which can not overlap indefinitely, we introduced a model in \cite{molkenthin2020self} starting from a chain of identical spheres, each in contact only with the neighbours it is connected to.
 Then additional connections form by randomly selecting two spheres and moving them towards each other until they touch, the new links formed that way can not be broken in later steps and function as constraints on subsequent links.
@@
 
-In \cite{molkenthin2016scaling}, we have introduced a simplified, two-dimensional version of this model, which starts from a closed chain of $N$ unit discs and subsequently adds links, such that connected discs touch, yet no discs intersect.
-The advantage of this model is, that it can be approximated by a purely topological simplification, which can be treated analytically.
+In \cite{molkenthin2016scaling}, we have introduced a simplified, two-dimensional version of this model, which starts from a closed chain (i.e.
+a ring) of $N$ unit discs and subsequently adds links, such that connected discs touch, yet no discs intersect.
+The advantage of this model is that it can be approximated by a purely topological simplification, which can be treated analytically.
 
@@
 (b) No links form across the outer face.
-This prevents the enclosure of a unit by less than six other units (which is geometrically impossible) such that.
-(c) The maximum degree of each unit is six, as six is the maximum number of unit discs, one central unit disc can touch.
+This prevents the enclosure of a unit by less than six other units (which is geometrically impossible) such that (c) the maximum degree of each unit is six, as six is the maximum number of unit discs one central unit disc can touch.
 (d) Once connected by a link, pairs of units do not disconnect.
@@
     \includegraphics[width=0.7\columnwidth]{figures/Fig1/Fig1.pdf}
-    \caption{a) The residue distance distribution is defined as the separation along the chain (blue) of two connected units (red).
+    \caption{a) The residue distance $s$ is defined as the separation along the chain (blue) of two connected units (red).
 b) The solid blue link prohibits some connections from node $A$ (dotted red), while others are still permitted (dotted blue).
-Counting all newly excluded links, we find that a link of length $k$ prohibits $2(l-1)$ links of length $l$ up to a maximum of $k$ and $2(k-1)$ links of length $k$ or more.}
+Counting all newly excluded links, we find that a link of length $s$ prohibits $2(l-1)$ links of length $l$ up to a maximum of $s$ and $2(s-1)$ links of length $s$ or more.}
     \label{fig:schematic}
@@
 
-Let us start by introducing the auxiliary variable $F(s)$, defined to be the number of possible links with a residue distance of $s$.
-Before any links are added all residue distances are equally likely, as there are $N$ possibilities of making a link of each separation $s$ with $2\leq s < N/2$.
+Let us start by introducing the auxiliary variable $F_k(s)$, defined to be the number of possible links with a residue distance of $s$ if $k$ links have been added before.
+Before any links are added all residue distances are equally likely, as there are $F_0(s)=N$ possibilities of making a link of each separation $s$ with $2\leq s < N/2$.
 
@@
 The expected distribution of possible links after one step is given thus by the average over available link pools taken over all possible lengths of the initial link
-\begin{equation}
-    F_1(s)= \frac{1}{N/2-2} \sum_{s_1=2}^{N/2} { \begin{cases}
+\begin{align}
+    F_1(s)&=\frac{1}{N/2-1} \sum_{s_1=2}^{N/2}{ \begin{cases}
+    F_0(s)-2(s-1) \text{ , for } s<s_1\\
+    F_0(s)-2(s_1 -1)\text{ , for } s\geq s_1
+    \end{cases}}
+    \nonumber \\
+    &\approx\frac{2}{N} \sum_{s_1=2}^{N/2} { \begin{cases}
     N-2(s-1) \text{ , for } s<s_1\\
     N-2(s_1 -1)\text{ , for } s\geq s_1
-    \end{cases}}.
+    \end{cases}}
+\end{align}
+
+In subsequent steps we take the weighted average over new links being added at step $k-1$, where the probability that a link of length $s$ can still be made is
+\begin{equation}
+    P_{in}^k(s)=F_k(s)/C_k,
 \end{equation}
-Here we denote the size of the pool of available links after $i$ links have been added $F_i(s)$.
-
-In subsequent steps we get the same reduction but starting from the pool $F_{k-1}(s)$ of the step before rather than the full $N$ links for each length.
-The reduction also needs to be multiplied by a factor
+where $C_k=\sum_{s=2}^{N/2}F_k(s)$ and get the same reduction (compare Fig.~\ref{fig:schematic}) but starting from the pool $F_{k-1}(s)$ of the step before rather than the full $N$ links for each length.
+To account for possible overlaps between newly removed links and links that were removed previously, we multiply by
 \begin{equation}
-    P_k(s)=\frac{F_k(s)}{C_k},
+    \frac{F_k(s)}{F_0(s)} \approx F_k(s)\frac{1}{N}.
     \label{eq.Pk}
 \end{equation}
-with $C_k=\sum_{s=2}^{N/2}F^k(s)$, to only remove links that are still in the pool. This leads to the recursion
+% \begin{equation}
+%     P_k(s)=\frac{F_k(s)}{C_k},
+%     \label{eq.Pk}
+% \end{equation}
+%with $C_k=\sum_{s=2}^{N/2}F_k(s)$, 
+This ignores any bias in which links are removed, however, in small, manageable examples this assumption seemed to hold well. 
+This leads to the recursion
 \begin{align}
-    F_k(s)&= \frac{1}{N/2-2} \nonumber \\
-    &\sum_{s_k=2}^{N/2} {\begin{cases}
-     F_{k-1}(s)-2(s-1) P_{k-1}(s) \text{ , for } s<s_k\\
-     F_{k-1}(s)-2(s_k -1)P_{k-1}(s)\text{ , for } s\geq s_k
-    \end{cases}}
+    F_k(s)&= \sum_{s_k=2}^{N/2} P_{in}^k(s_k)\nonumber \\
+    & (F_{k-1}(s)-  F_k(s)\frac{1}{N}
+    {\begin{cases}
+     2(s-1) \text{ , for } s<s_k\\
+     2(s_k -1))\text{ , for } s\geq s_k
+    \end{cases}} )
 \end{align}
-which can be simplified to
+which simplifies to
 \begin{align}
-   F_k(s)&= \frac{1}{\frac{N}{2}-2} \frac{F_{k-1}(s)}{C_{k-1}}\nonumber \\
-   &\left( \sum_{s_k=2}^{N/2}C_{k-1} - \sum_{s_k=2}^{s} 2(s_k-1) - \sum_{s_k=s+1}^{N/2} 2(s -1) \right) \nonumber \\
-   &= \frac{1}{\frac{N}{2}-2}\frac{F_{k-1}(s)}{C_{k-1}}\nonumber \\
-   &\left(\left(\frac{N}{2}-2\right)C_{k-1} -(s^2-s)
-   -2\left(\frac{N}{2}-(s+1)\right)(s-1)\right)  \nonumber \\
-   &= F_{k-1}(s)\left(1-\frac{-s^2 +(N-1)s-(N+1)}{(\frac{N}{2}-2)C_{k-1}} \right)\nonumber \\
-   &=F_{k-1}(s)\left(1-\frac{f(s)}{(\frac{N}{2}-2)C_{k-1}} \right)
+   F_k(s)&=F_{k-1}(s)(1-\frac{1}{N} \sum_{s_k=2}^{s} (2 (s_k-1) P_{in}^k(s_k))\nonumber \\
+   &-\frac{1}{N} \sum_{s_k=s+1}^{N/2} (2 (s-1) P_{in}^k(s_k)))\nonumber \\
+   &=F_{k-1}(s)(1-\frac{1}{N}
+   \sum_{s_k=2}^{s} (2 (s_k-1) P_{in}^k(s_k))\nonumber \\
+   &-\frac{2 (s-1)}{N} (\sum_{s_k=2}^{N/2} P_{in}^k(s_k) - \sum_{s_k=2}^{s} P_{in}^k(s_k)))\nonumber \\
+   &=F_{k-1}(s)(1-\frac{2}{N} (s-1) 
+   +\frac{2}{N} s
+   \sum_{s_k=2}^{s}P_{in}^k(s_k)\nonumber \\
+   &-\frac{2}{N}
+   \sum_{s_k=2}^{s} s_k P_{in}^k(s_k))
    \label{eq.Fk_rec}
 \end{align}
-Where we defined $f(s)=-s^2 +(N-1)s-(N+1) \approx N s - s^2 - N$ for notational brevity.
 
-We can now use the recursive expression Eq.~\ref{eq.Fk_rec} to write down a closed expression for $F_k(s)$
+To approximately evaluate this we make a guess for $P_{in}^k(s)$ at low $k$ and at high $k$, namely
+\begin{equation}
+    P_{in}^{k\ll N/2}(s)\approx P_{in}^{k=0}(s)=2/N.
+    \label{eq.P_small}
+\end{equation}
+For later steps $k$, the probability of a link to still be in the pool decreases with $s$, we thus make the ansatz, that it drops off as $1/s$, leading to
+\begin{equation}
+    P_{in}^{k\gg 1}(s)\approx \frac{s^{-1}}{ (H_{\frac{N}{2}}-1)},
+    \label{eq.P_large}
+\end{equation}
+where $H_{\frac{N}{2}}$ is the harmonic number. 
+We suspect that the approximation for low $k$ only holds for a couple of steps while Eq.~\ref{eq.P_large} holds, at least on average for larger $k$.
+We have no approximation for the intermediate steps but assume that the errors roughly cancel each other out.
+Inserting Eq.\ref{eq.P_small} and Eq.\ref{eq.P_large} into Eq.\ref{eq.Fk_rec} results in the following two recursions for the early and late evolution of the sequence distance distribution respectively.
+\begin{align}
+   F_{k\ll N/2}(s)&\approx 
+   F_{k-1}(s)(1-\frac{2}{N} (s-1) 
+   +\frac{4}{N^2} s (s-1)\nonumber \\
+   &-\frac{2}{N^2}(s^2+s-2)\nonumber \\
+   &=F_{k-1}(s)(1-f_{low}(s)),
+   \label{eq.Fk_rec_low}
+\end{align}
+where 
+\[f_{low}(s)=-\frac{2}{N^2}s^2+(\frac{2}{N}+\frac{6}{N^2})s-(\frac{2}{N}+\frac{4}{N^2}),\]
+and
+\begin{align}
+   F_{k\gg1}(s)&\approx 
+   F_{k-1}(s)(1-\frac{2}{N} (s-1) 
+   +\frac{2}{N} s \sum_{s_k=2}^{s}\frac{1}{s_k(H_{\frac{N}{2}}-1)}\nonumber \\
+   &-\frac{2}{N}
+   \sum_{s_k=2}^{s} s_k \frac{1}{s_k(H_{\frac{N}{2}}-1)})\nonumber \\
+   &=
+   F_{k-1}(s)(1 -(\frac{2}{N}-\frac{2}{N}\frac{H_{s}-1}{H_{\frac{N}{2}}-1}+\frac{2}{N(H_{\frac{N}{2}}-1)})s \nonumber \\ 
+   &+ \frac{2}{N(H_{\frac{N}{2}}-1)} + \frac{2}{N})\nonumber \\
+   &=F_{k-1}(s)(1-f_{high}(s)).
+   \label{eq.Fk_rec_high}
+\end{align}
+where 
+\[f_{high}(s)=\frac{2}{N}(\frac{H_{\frac{N}{2}}-H_{s}+1}{H_{\frac{N}{2}}-1}s-\frac{H_{\frac{N}{2}}}{H_{\frac{N}{2}}-1}).\]
 
-\begin{equation}
-    F_k(s)=N\prod_{i}^k\left(1-\frac{f(s)}{(\frac{N}{2}-2)C_{i-1}} \right)
-\end{equation}
+We can now use the recursive expressions to write down a closed expression for $F_k(s)$
 
+\begin{align}
+    F_k(s)&=N\prod_{i_1}^a\left(1-f_{low}(s)\right)\prod_{i=a+1}^k\left(1-f_{high}(s)\right) \nonumber \\
+    &= N\left(\frac{1-f_{low}(s)}{1-f_{high}(s)}\right)^a\left(1-f_{high}(s)\right)^{k}
+\end{align}
+
+In 2D this is repeated for $N-3$ steps until no more links can be added \cite{molkenthin2016scaling}.
 The probability distribution of the realized residue distances of all added links is then given by the average over the available pools at each link addition step, leading to
 \begin{align}
-    P(s)&=\frac{1}{N-3}\sum_{k=0}^{N-3} P_k(s) \nonumber \\
-    &= \frac{N}{N-3}\sum_{k=0}^{N-3} \frac{1}{C_k}\prod_{i}^k\left(1-\frac{f(s)}{(\frac{N}{2}-2)C_{i-1}} \right),
+    P(s)&=\frac{1}{N-3}\sum_{k=0}^{N-4} \frac{F_k(s)}{C_k} \nonumber \\
+    &= \frac{N}{N-3}\left(\frac{1-f_{low}(s)}{1-f_{high}(s)}\right)^a
+\sum_{k=0}^{N-4} \frac{1}{C_k} \left(1-f_{high}(s)\right)^{k} \nonumber \\
+    &\approx \left(\frac{1-f_{low}(s)}{1-f_{high}(s)}\right)^a\frac{A}{f_{high}(s)},
+    \label{eq.Ps}
 \end{align}
-which makes use of Eq.~\ref{eq.Pk}.
-The only unknown in this expression is $C_k$, which we can not compute exactly.
+
+\red{The assumption that the sum is basically a harmonic series seems like a bit of a stretch I shall check it again but the results look good}
+
+for $1<s<\frac{N}{2}$.
+The approximation sets all $C_k=C$ to allow the use of the geometric series for solving the equation.
+We can do this because $C_k$ monotonically decreases with $k$ and $N>C_k>C_{N-2}$ meaning we can obtain an upper and lower bound by replacing all $C_k$ in the expression with either $N$ (underestimates all probabilities) or $C_{N-2}$ (overestimates all probabilities), however, as neither of those depend on $s$, these bounds are merely in the constant factor $A$ and do not affect the scaling as such.
+%which makes use of Eq.~\ref{eq.Pk}. 
+%The only unknown in this expression is $C_k$, \red{which we can not compute exactly...?}.
 However, since $C_k$ monotonically decreases with $k$, $N>C_k>C_{N-2}$ always holds and we can obtain an upper and lower bound by replacing all $C_k$ in the expression with either $N$ (underestimates all probabilities) or $C_{N-2}$ (overestimates all probabilities).
 Each time we get an expression of the form
-\begin{align}
-     P(s)&\approx\frac{N}{N-3}\sum_{k=0}^{N-3} \frac{1}{C_k}\prod_{i}^k\left(1-\frac{f(s)}{(\frac{N}{2}-2)C} \right)\nonumber \\
-     &=\frac{N}{N-3} \sum_{k=0}^{N-3}\frac{1}{C_k}\left(1-\frac{f(s)}{(\frac{N}{2}-2)C} \right)^k \nonumber \\
-     &\approx \frac{A}{f(s)}
-\end{align}
-for $1<s<\frac{N}{2}$,
-where $A$ is an unknown constant, containing an approximation of the various normalization constants $C_i$, as well as the direct effects of the chain length. 
-Since for long chains, $f(s)$ is largely dominated by the linear term, the distribution is well approximated by a power law with an exponent of $-1$.
-We have made use of the geometric series to approximately find a closed solution.
+% \begin{align}
+%      P(s)&\approx\frac{N}{N-3}\sum_{k=0}^{N-3} \frac{1}{C_k}\prod_{i}^k\left(1-\frac{f(s)}{(\frac{N}{2}-2)C} \right)\nonumber \\
+%      &=\frac{N}{N-3} \sum_{k=0}^{N-3}\frac{1}{C_k}\left(1-\frac{f(s)}{(\frac{N}{2}-2)C} \right)^k \nonumber \\
+%      &\approx \frac{A}{f(s)}
+% \end{align}
+% for $1<s<\frac{N}{2}$,
+% where $A$ is an unknown constant, containing an approximation of the various normalization constants $C_i$, as well as the direct effects of the chain length. 
+Since for long chains and $s \ll N/2$, $f(s)$ is dominated by the linear term, the distribution is well approximated by a power law with an exponent of $-1$.
 
-In the last step of this approximation, we take the $C_k$ to all be approximately equal in order to apply the geometric series.
+%In the last step of this approximation, we took the $C_k$ to all be approximately equal in order to apply the geometric series.
 However, in reality we know $1/C_k$ to increase with $k$, giving large $k$ summands relatively more weight than in a geometric series.
@@
 
-\begin{equation}
-    P_{corrected}(s)\approx A f(s)^{-1+\alpha},
-\end{equation}
-where $0<\alpha \ll 1$.
+%\begin{equation}
+ %   P_{corrected}(s)\approx A f(s)^{-1+\alpha},
+%\end{equation}
+%where $0<\alpha \ll 1$.
+\blue{The parameters $A$ and $\alpha$ in this approximation are treated as free parameters and used for fitting the theoretical result to data and simulations.
+While they could be estimated by evaluating the middle line of Eq.~\ref{eq.Ps}, this estimate is for the 2D model, in which the exclusion of connections is faster, leading to a steeper decrease of $C_k$, as well as a different factor $A$.}
+
+\red{We could add a figure with simulated 2D results (which i should have...
+somewhere) and the exact evaluation with the sums and all that.
+I'm not super motivated to do that but i guess it might be a nice addition and the only place for the exact formula really.}
 
@@
 For large values of the chain length $N$ we thus find an approximate power law probability distribution for the residue distance with an exponent of $-1+\alpha$.
-Here we use simulations, as introduced in~\cite{molkenthin2020self} and networks extracted from measured protein data from the PDB \cite{PDB} to compare this behaviour in realistic 3D settings and show that the basic principle persists, indicating that much of the residue distance distribution can be explained by simple geometric constraints.
+Here we use simulations, as introduced in~\cite{molkenthin2020self}, and networks extracted from measured protein data from the PDB \cite{PDB} to compare this behaviour with the more realistic 3D settings and show that the basic principle persists, indicating that much of the residue distance distribution can be explained by generic geometric constraints.
  \begin{figure}[h]
@@
 Protein chain lengths in the PDB vary from small fragments of less than 10 amino acids to large agglomerates of over 2000 residues.
-The most frequently occurring chain lengths, however are between 100 and 400 residues as shown in Fig.~\ref{fig:pdb_stats}.
+The most frequently occurring chain lengths, however, are between 100 and 400 residues as shown in Fig.~\ref{fig:pdb_stats}.
 \red{For our analysis, we have thus chosen ............
@@
 structures of around 300 and ..........
-structures of around 400 residues from the PDB.} Those have been converted to PCN's based on the measured locations of their residues to analyze the distribution of residue distances between connected amino acids.
-The resulting distribution is compared to those generated in simulations in~\cite{molkenthin2020self} as well as the above derived theoretical expression.
-The latter, however, is scaled with a factor to account for the different dimensionality, as well as approximations made in the derivation.
+structures of around 400 residues from the PDB.} Those have been converted into PCN's based on the measured locations of their residues to analyze the distribution of residue distances between connected amino acids.
+The resulting distribution is compared to those generated in simulations in~\cite{molkenthin2020self} as well as the above derived theoretical expression scaled with a factor to account for the different dimensionality, as well as approximations made in the derivation.
 
@@
         \centering
-	\includegraphics[width=0.45\textwidth]{paper/figures/Fig3/Fig3.pdf}
+	\includegraphics[width=\columnwidth]{paper/figures/Fig3/Fig3.pdf}
         \caption{residue distance distributions for real and simulated folds for a) Chain lengths around 200 b) chain lengths around 300 c) chain lengths around 400.
@@
 The over-representation of very short residue distances is an artefact of the different network construction for simulated (and theoretically approximated) networks and PCNs.
-While in the simulation all links are equally long and links are made until no longer possible, the parameter, determining the connectivity in PCNs is the cutoff distance $d_c$, which is chosen such that the number of links in the PCN is comparable to those in a simulated network of the same $N$.
+While in the simulation all links are equally long and links are made until no longer possible, the parameter, determining the connectivity in PCNs is the cutoff distance $d_c$\blue{(mention used value?)}, which is chosen such that the number of links in the PCN is comparable to those in a simulated network of the same $N$.
 This leads to a larger range of bond angles randomly resulting in connections for small $s$ and thus an over-representation of those links.
@@
 For the equivalent model in two dimensions we have used a mean-field approach to derive an analytic approximation for the residue distance distribution, which appears to agree with the simulated and measured distributions.
-We have thus demonstrated in analytical calculations as well as direct simulations how a random linking model with geometric constraints, such as the ones introduced in \cite{molkenthin2016scaling, molkenthin2020self} generates a residue distance distribution, that resembles a power-law, similar to the one found in folded proteins.
-This feature of protein network structure has previously been used in the heuristic model to generate protein-like structures in \cite{bartoli2008effect}.
+We have thus demonstrated in analytical calculations as well as direct simulations how a random linking model with geometric constraints, such as the ones introduced in \cite{molkenthin2016scaling, molkenthin2020self}, generates a residue distance distribution that resembles a power-law, similar to the one found in folded proteins.
+This feature of protein network structures has previously been used in a heuristic manner to model and generate protein-like structures in \cite{bartoli2008effect}.
 
@@
 \section{Supplemental Material Collection}
+\red{Do we have anything that needs to go here or can we just leave this?}
 \section{Random ideas}

### 2021-12-02 src/PCN.py
first draft of classes and plotting

@@ -1,0 +1,138 @@
+'''
+Base class for PCNs. 
+
+'''
+import networkx as nx
+import MDAnalysis as mda
+import numpy as np
+import random
+
+class PCN():
+	'''
+	Base class for PCNs. 
+	
+	Takes a PDB file and creates a PCN from the alpha carbons.
+
+	'''
+	# Define a threshold value for defining a link between two atoms
+	link_length_threshold = 8.0 # Å
+	
+	def __init__(self, PDB_file):
+		self.universe = mda.Universe(PDB_file)
+
+	def get_C_alphas(self):
+		'''
+		Take a universe, get segment A and return the C-alphas for the PDB. 
+
+		Parameters
+		----------
+		universe: MDAnalysis.Universe(PDB_file)
+
+		Return
+		------
+		C_alphas: MDAnalysis AtomGroup
+			contains the C-alphas from the PDB
+		'''
+		"------------------------MOVE-------------------------------"
+		segments = self.universe.residues.segments
+			
+		number_of_segments = len(segments)
+
+		for index in range(number_of_segments):
+			segment_id = self.universe.residues.segments.segids[index]
+
+			try:
+				C_alphas = self.universe.select_atoms(f'name CA and segid {segment_id}')
+				print(segment_id)
+				break
+			
+			except:
+				print('Error in first segment id. Trying the next one.')
+				continue
+
+		"-----------------------------------------------------------"
+		return C_alphas
+
+	def get_chain_length(self,C_alphas):
+		'''
+		Take C-alphas and use that to return the length of the chain.
+
+		Parameters
+		----------
+		C_alphas: MDAnalysis AtomGroup
+			contains the C-alphas from the PDB
+
+		Return
+		------
+		chain_length: int
+			length of the chain
+		'''
+		chain_length = len(C_alphas)
+		
+		return chain_length
+
+	def create_connected_component_subgraphs(self,protein_graph):
+		'''
+		Take a protein graph and create connected component subgraphs.
+		
+		Parameters
+		----------
+		protein_graph: nx.graph()
+			protein graph containing the links
+
+		Return
+		------
+		cc_subgraph: nx.subgraph() generator
+			subgraph of connected components in the protein graph
+		'''
+		for components in nx.connected_components(protein_graph):
+			yield protein_graph.subgraph(components)
+
+	def get_link_lengths(self, C_alphas):
+		'''
[…]

### 2021-12-02 src/get_ll_stats_tidy.py

@@ -46,20 +46,20 @@ def get_link_lengths(CA):
 	graph = nx.empty_graph(len(r))
 
 	# Define max and average distances
-	max_dist = 0
-	avg_dist = 0
+	max_dist = 0 # could be commented out
+	avg_dist = 0 # could be commented out
 	'''-------------------------------???-------------------------------'''
 	# Create the adjacency matrix
 	for i in range(len(r)):
 		for j in range(i):
 			# Get the norm, i.e. distnces between adjacent atoms
 			distance = np.linalg.norm(r[i] - r[j])
-			max_dist = max(distance, max_dist)
-			avg_dist += distance ** 2
+			max_dist = max(distance, max_dist) # could be commented out
+			avg_dist += distance ** 2 # could be commented out
 			# If closer than threshold, add link
 			if distance < threshold:
 				graph.add_edge(i,j)
-				linklengths.append(distance)
+				linklengths.append(distance) # could be commented out
 	'''-------------------------------???-------------------------------'''
 	# Get the average distance
 	avg_dist = np.sqrt(avg_dist) / (len(r))
@@ -79,30 +79,35 @@ def get_link_lengths(CA):
 	return link_length_list
 
 
-def bootsrap(ll_list, link_length_id_list):
-	'''Bootstrap LL data.'''
-	# Specify a sample size to use for bootstrapping
-	sample_size = 1000
-	data_histogram_list = []
-	mean_list = []
-	std_list = []
-	# Loop over number of samples
-	for n in range(sample_size):
-		# Randomly choose a list of observations from the sample
-		bts_sample = [random.choice(ll_list) for _ in ll_list]
-		# Sample with replacement
-		bts_sample = [item for items in bts_sample for item in items]
-		# Make a histogram
-		data_histogram, data_histogram_edges = np.histogram(bts_sample, bins = range(3, 1000), density = True)
-		# Normalise the histogram? But doesn't density already do this?
-		data_histogram = data_histogram / len(link_length_id_list)
-		data_histogram_list.append(data_histogram)
-	data_histogram_list = np.array(data_histogram_list)
-	mean_list = data_histogram_list.mean(axis = 0)
-	std_list = data_histogram_list.std(axis = 0)
-
-
-
+def bootsrap(length, ll_list, link_length_id_list):
[…]

### 2021-12-02 src/remove_duplicates.py

@@ -1,0 +1,33 @@
+import pandas as pd
+import numpy as np
+
+DF = pd.read_csv('../data/duplicates.csv')
+
+#DF_T = DF.transpose()
+
+string_list = DF['DOWNLOADED'].astype(str).to_numpy()
+
+down_list = [x.replace('.pdb', '').replace('.gz', '') for x in string_list]
+all_list = DF['ALL'].astype(str).tolist()
+
+duplicates = []
+
+for i in range(len(all_list)):
+	print(f'We are at entry: {i}')
+	duplicates.append(down_list[i])
+	duplicates.append(all_list[i])
+
+
+duplicate_dict = {'duplicates':duplicates}
+print(f'Length of duplicates list is {len(duplicates)}')
+
+df = pd.DataFrame()
+df = df.from_dict(duplicate_dict)
+
+no_duplicates = df.drop_duplicates(inplace=False)
+
+no_dupes_list = no_duplicates['duplicates'].to_numpy()
+print(f'No duplicates list length is {len(no_dupes_list)}')
+
+np.savetxt('no_duplicates.csv',no_dupes_list,delimiter=',')
+#print(no_duplicates)

### 2021-12-02 src/test_methods.py

@@ -1,11 +1,48 @@
 import random
 import numpy as np
 import pandas as pd
+import PCN
+import os
+import MDAnalysis as mda
+PDB_IDs = ['5EUB']
 
-sample_size = 1000
+for PDB_ID in PDB_IDs:
+	PDB_ID_lower = str(PDB_ID).lower()
+	PDB_file = f'/Volumes/Seagate_Extension_Plus/PDBs/pdb{PDB_ID_lower}.ent'
+	if os.path.isfile(PDB_file):
+		u = mda.Universe(PDB_file)
 
-sample = np.arange(0,21,0.1)
 
-DF = pd.read_csv('../data/ids/id_file_0.txt', header = None)
-DF = DF.transpose()
-print(DF)
+		# for seg_index in range(number_of_segments):
+
+		segments = u.universe.residues.segments
+			
+		number_of_segments = len(segments)
+
+		for index in range(number_of_segments):
+			segment_id = u.universe.residues.segments.segids[index]
+
+			try:
+				C_alphas = u.universe.select_atoms(f'name CA and segid {segment_id}')
+				
+				print(len(C_alphas))
+				break
+			
+			except:
+				print('Error in first segment id. Trying the next one.')
+				continue
+
+
+
+
+		# if segment_id:
+		# 	print(segment_id)
+		# segments = self.universe.residues.segments
+		# number_of_segments = len(segments)
+		# for seg_index in range(number_of_segments):
+		# 	str_seg_ids = str(segments[seg_index])
+			
+		# 	if '>' != str_seg_ids[9]:
+		# 	 	segment_id = segments[seg_index]
+
+

### 2021-12-03 data/Lunaplot.py
untrack files in .gitignore

@@ -1,135 +1,0 @@
-import networkx as nx
-import numpy as np
-import scipy as sp
-import scipy.stats
-import matplotlib.pyplot as plt
-import matplotlib
-import math as m
-import random
-import time
-import bisect
-from scipy.optimize import curve_fit
-#from compiler.ast import flatten
-from operator import add
-
-def func(x, a, b):
-     return a * x**(-b)
-
-def func1(x, a):
-     return a/x
-
-def expfit(x, a, b):
-     return a * np.exp(x*(-b))
-
-n=1000
-Luna=np.zeros((10,498))
-Luna[1]=np.loadtxt('luna_noencircle1')
-Luna[2]=np.loadtxt('luna_noencircle2')
-Luna[3]=np.loadtxt('luna_noencircle3')
-Luna[4]=np.loadtxt('luna_noencircle4')
-Luna[5]=np.loadtxt('luna_noencircle5')
-Luna[6]=np.loadtxt('luna_noencircle6')
-Luna[7]=np.loadtxt('luna_noencircle7')
-Luna[8]=np.loadtxt('luna_noencircle8')
-Luna[9]=np.loadtxt('luna_noencircle9')
-Luna[0]=np.loadtxt('luna_noencircle10')
-
-#n=2000
-#Luna[1]=np.loadtxt('luna1')
-#Luna[2]=np.loadtxt('luna2')
-#Luna[3]=np.loadtxt('luna3')
-#Luna[4]=np.loadtxt('luna4')
-#Luna[5]=np.loadtxt('luna5')
-#Luna[6]=np.loadtxt('luna6')
-#Luna[7]=np.loadtxt('luna7')
-#Luna[8]=np.loadtxt('luna8')
-#Luna[9]=np.loadtxt('luna9')
-#Luna[10]=np.loadtxt('luna10')
-#Luna[11]=np.loadtxt('luna11')
-#Luna[12]=np.loadtxt('luna12')
-#Luna[13]=np.loadtxt('luna13')
-#Luna[14]=np.loadtxt('luna14')
-#Luna[15]=np.loadtxt('luna15')
-#Luna[16]=np.loadtxt('luna16')
-#Luna[17]=np.loadtxt('luna17')
-#Luna[18]=np.loadtxt('luna18')
-#Luna[19]=np.loadtxt('luna19')
-#Luna[0]=np.loadtxt('luna10')
-
-Lunatot=np.mean(Luna,axis=0)
[…]

### 2021-12-03 pdb_linklength.py

@@ -1,237 +1,0 @@
-# -*- coding: utf-8 -*-
-from __future__ import print_function
-import networkx as nx
-import scipy as sp
-import scipy.stats
-from scipy.optimize import curve_fit
-import matplotlib.pyplot as plt
-import matplotlib.colors as colors
-import numpy as np
-from Bio.PDB import *
-from MDAnalysis import *
-import urllib
-import os
-import seaborn as sns
-import glob
-
-sns.set(font_scale=1.5)  # crazy big
-sns.set_style("ticks")
-sns.despine()
-
-def func(x, a,b):
-     return a * x**(-b)
-
-def func2(x, a):
-     return a / x
-
-def theofunc(x, C,N,ex):
-     return (C)* np.power(N/(2*(-x**2+N*x-N)),ex)
-
-#def theofunc(x, C,N):
-#     return (C)* N/(2*(N*x-N))
-
-names_xray = glob.glob("pdb_data2/all_xray/*.pdb")
-names_xray = glob.glob("pdb_data2/xray_data_short/*.pdb")
-names_xray = glob.glob("pdb_data2/xray_data/*.pdb")
-
-names_nmr = glob.glob("pdb_data2/nmr_data/*.pdb")
-names_nmr = glob.glob("pdb_data2/nmr_data_short/*.pdb")
-
-names_both = glob.glob("pdb_data2/both/*.pdb")
-
-#names_both = glob.glob("idp/*.pdb")
-
-names_used=names_both
-
-num=28
-lengths=[3,4,5,6,7,8,9,10,11,12,13,14,16,17,19,21,23,25,28,30,33,36,40,44,48,52,58,63,69,76,83,91,100,110,120,132,145,158,174,191,209,229,251,275,
-    302,331,363,398]#[0:45]
-print(len(lengths))
-
-numnum=20
-logscale=np.logspace(np.log10(3), np.log10(1000), num=numnum)
-print('logscale',logscale)
-results=[]
-binnedresults=[]
-
-if __name__ == '__main__':
-    directories = ['pdb_data2/both']
-    #directories = ['idp']
-    nameslen=len(names_used)
-    #nameslen=50
-    path = os.getcwd()
-    meanies=[]
-    linklengths=[]
-    for d in directories:
-        print(d)
-        output = os.path.join(path,d)
-        count=0
-        n=1000
-        t=8.0
-        counter=0
-        for counter in range(nameslen):
-            #pdb_file = os.path.join(output,pdb.strip())+'.pdb'
[…]

### 2021-12-03 src/functions.py

@@ -1,224 +1,0 @@
-'''
-Functions to be used in main.py. 
-
-'''
-import os
-import pandas as pd
-import Bio.PDB.PDBList as biopy
-import PCN
-import traceback
-import tarfile
-import random
-import numpy as np
-
-def concat_ID_files():
-	'''
-	Opens the chunked PDB id files as a DataFrame. Returns an array with of all the IDs.
-	
-	Parameters
-	----------
-	None
-
-	Return
-	------
-	PDB_ID_array: numpy array
-		array containing all the PDB IDs
-	'''
-	frames = []
-	for filename in os.listdir('../data/ids/'):
-		if filename.endswith('.txt'):
-			temp_df = pd.read_csv('../data/ids/'+str(filename), header = None)
-			frames.append(temp_df)
-
-	concatenated_DF = pd.concat(frames).reset_index()
-	concatenated_DF = concatenated_DF.drop(columns = 'index')
-	PDB_ID_array = concatenated_DF[0].to_numpy()
-	
-	return PDB_ID_array
-
-def bioretrieve_PDBs(PDB_ID_array, save_directory):
-	'''
-	Take the array of PDB IDs and use BioPython to retrieve PDB files as .ent files.
-
-	Parameters
-	----------
-	PDB_ID_array: numpy array
-		array containing all the PDB IDs
-	save_directory: string 
-		path to the directory where the file will be saved
-
-	Return
-	------
-	None
-	'''
-	pdblist = biopy.PDBList()
-	PDB_IDs = concat_ID_files()
-	counter = 1
-	for PDB_ID in PDB_IDs:
-		print(f'At entry {counter} out of {len(PDB_IDs)} entries.')
-		pdblist.retrieve_pdb_file(pdb_code = PDB_ID, file_format = 'pdb', pdir = save_directory)
-		counter += 1
-
-def bootstrap(length, link_lengths, length_range):
-	'''
-	Take the link lengths and bootstrap the data. 
-
-	Saves the mean and median of the bootstrapped data and the PDB ids belonging to
-	that length range. Files are saved in .npy format. 
-
-	Parameters
-	----------
-	length: string
-		defines the length range as '100', '200' or '300'
-
-	link_lengths: list
-		list containing lengths of links
-
-	length_range: list
-		list containing the PDB ids that fall in the length range given
-
-	Return
-	------
-	None
-	'''
-	if len(link_lengths) > 0:
-		sample_size = 1000
-		histogram_list = []
-		means = []
-
-		for n in range(sample_size):
[…]

### 2021-12-03 src/main.py

@@ -1,8 +1,0 @@
-'''
-Main 
-
-'''
-import functions
-
-if __name__ == '__main__':
-	functions.main()

### 2021-12-03 src/read_ids.py

@@ -1,14 +1,0 @@
-import pandas as pd
-import numpy as np
-
-filename = '../data/combined_ids_tr.csv'
-chunksize = 1000
-filename_counter = 0
-
-for chunk in pd.read_csv(filename, chunksize=chunksize, sep='\n', header=None):
-
-	np_array = chunk.to_numpy()
-	np.savetxt('../data/ids/id_file_'+str(filename_counter)+'.txt', np_array, fmt='%s')
-
-	filename_counter += 1
-

### 2021-12-03 src/remove_duplicates.py

@@ -1,33 +1,0 @@
-import pandas as pd
-import numpy as np
-
-DF = pd.read_csv('../data/duplicates.csv')
-
-#DF_T = DF.transpose()
-
-string_list = DF['DOWNLOADED'].astype(str).to_numpy()
-
-down_list = [x.replace('.pdb', '').replace('.gz', '') for x in string_list]
-all_list = DF['ALL'].astype(str).tolist()
-
-duplicates = []
-
-for i in range(len(all_list)):
-	print(f'We are at entry: {i}')
-	duplicates.append(down_list[i])
-	duplicates.append(all_list[i])
-
-
-duplicate_dict = {'duplicates':duplicates}
-print(f'Length of duplicates list is {len(duplicates)}')
-
-df = pd.DataFrame()
-df = df.from_dict(duplicate_dict)
-
-no_duplicates = df.drop_duplicates(inplace=False)
-
-no_dupes_list = no_duplicates['duplicates'].to_numpy()
-print(f'No duplicates list length is {len(no_dupes_list)}')
-
-np.savetxt('no_duplicates.csv',no_dupes_list,delimiter=',')
-#print(no_duplicates)

### 2021-12-03 src/test_methods.py

@@ -1,48 +1,0 @@
-import random
-import numpy as np
-import pandas as pd
-import PCN
-import os
-import MDAnalysis as mda
-PDB_IDs = ['5EUB']
-
-for PDB_ID in PDB_IDs:
-	PDB_ID_lower = str(PDB_ID).lower()
-	PDB_file = f'/Volumes/Seagate_Extension_Plus/PDBs/pdb{PDB_ID_lower}.ent'
-	if os.path.isfile(PDB_file):
-		u = mda.Universe(PDB_file)
-
-
-		# for seg_index in range(number_of_segments):
-
-		segments = u.universe.residues.segments
-			
-		number_of_segments = len(segments)
-
-		for index in range(number_of_segments):
-			segment_id = u.universe.residues.segments.segids[index]
-
-			try:
-				C_alphas = u.universe.select_atoms(f'name CA and segid {segment_id}')
-				
-				print(len(C_alphas))
-				break
-			
-			except:
-				print('Error in first segment id. Trying the next one.')
-				continue
-
-
-
-
-		# if segment_id:
-		# 	print(segment_id)
-		# segments = self.universe.residues.segments
-		# number_of_segments = len(segments)
-		# for seg_index in range(number_of_segments):
-		# 	str_seg_ids = str(segments[seg_index])
-			
-		# 	if '>' != str_seg_ids[9]:
-		# 	 	segment_id = segments[seg_index]
-
-

### 2021-12-03 src/PCN.py
add new src files

@@ -1,0 +1,138 @@
+'''
+Base class for PCNs. 
+
+'''
+import networkx as nx
+import MDAnalysis as mda
+import numpy as np
+import random
+
+class PCN():
+	'''
+	Base class for PCNs. 
+	
+	Takes a PDB file and creates a PCN from the alpha carbons.
+
+	'''
+	# Define a threshold value for defining a link between two atoms
+	link_length_threshold = 8.0 # Å
+	
+	def __init__(self, PDB_file):
+		self.universe = mda.Universe(PDB_file)
+
+	def get_C_alphas(self):
+		'''
+		Take a universe, get segment A and return the C-alphas for the PDB. 
+
+		Parameters
+		----------
+		universe: MDAnalysis.Universe(PDB_file)
+
+		Return
+		------
+		C_alphas: MDAnalysis AtomGroup
+			contains the C-alphas from the PDB
+		'''
+		"------------------------MOVE-------------------------------"
+		segments = self.universe.residues.segments
+			
+		number_of_segments = len(segments)
+
+		for index in range(number_of_segments):
+			segment_id = self.universe.residues.segments.segids[index]
+
+			try:
+				C_alphas = self.universe.select_atoms(f'name CA and segid {segment_id}')
+				#print(segment_id)
+				break
+			
+			except:
+				print('Error in first segment id. Trying the next one.')
+				continue
+
+		"-----------------------------------------------------------"
+		return C_alphas
+
+	def get_chain_length(self,C_alphas):
+		'''
+		Take C-alphas and use that to return the length of the chain.
+
+		Parameters
+		----------
+		C_alphas: MDAnalysis AtomGroup
+			contains the C-alphas from the PDB
+
+		Return
+		------
+		chain_length: int
+			length of the chain
+		'''
+		chain_length = len(C_alphas)
+		
+		return chain_length
+
+	def create_connected_component_subgraphs(self,protein_graph):
+		'''
+		Take a protein graph and create connected component subgraphs.
+		
+		Parameters
+		----------
+		protein_graph: nx.graph()
+			protein graph containing the links
+
+		Return
+		------
+		cc_subgraph: nx.subgraph() generator
+			subgraph of connected components in the protein graph
+		'''
+		for components in nx.connected_components(protein_graph):
+			yield protein_graph.subgraph(components)
+
+	def get_link_lengths(self, C_alphas):
+		'''
[…]

### 2021-12-03 src/bio_retrieve_pdbs.py

@@ -1,0 +1,33 @@
+from Bio.PDB.PDBList import PDBList
+import os
+import pandas as pd
+
+
+pdblist = PDBList()
+
+def concat_ID_files():
+	''' Opens id_file_N.txt as a DF and concats them all into one DF '''
+	frames = []
+	for filename in os.listdir('../data/ids/'):
+		if filename.endswith(".txt"):
+			temp_df = pd.read_csv('../data/ids/'+str(filename), header=None)
+			frames.append(temp_df)
+
+	DF = pd.concat(frames).reset_index()
+	DF = DF.drop(columns='index')
+	PDB_list = DF[0].to_numpy()
+	return PDB_list
+
+PDBs = concat_ID_files()
+counter = 1
+for pdb in PDBs:
+	print(f'At entry {counter}')
+	pdblist.retrieve_pdb_file(pdb_code = pdb, 
+						      file_format = "pdb", 
+						      pdir = '/Volumes/Seagate_Extension_Plus/PDBs/')
+	counter += 1
+
+# pdb = '1AKI'
+# pdblist.retrieve_pdb_file(pdb_code = pdb, 
+# 						  file_format = "bundle", 
+# 						  pdir = '../data/bioretrieve_PDBs/')

### 2021-12-03 src/ff1.py

@@ -1,0 +1,264 @@
+import networkx as nx
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import math as m
+import random
+import time
+import bisect
+from scipy.optimize import curve_fit
+# from compiler.ast import flatten
+from operator import add
+import collections
+
+def flatten(iterable):
+    for el in iterable:
+        if isinstance(el, collections.Iterable) and not isinstance(el, str): 
+            yield from flatten(el)
+        else:
+            yield el
+            
+def func(x, a, b):
+     return a * x**b
+def func2(x, a):
+     #return a * x**meanp
+     return a * x**0.60228
+def pi(x):
+    #return 1.0/3
+    if (x<n/2.0): return float(4*x)/(3*(n+2*x)) +1.0/3
+    else: return float(2)/3
+    #return float(3*x+4+n)/(3*(n+x))
+def pimin(x):
+    return 1.0/3
+
+
+def diameter_size(G):
+    shortest_path_pairwise=nx.shortest_path(G)
+    all_shortest_paths=[]
+
+    for shortest_paths_from_src in shortest_path_pairwise.values():
+        all_shortest_paths+=shortest_paths_from_src.values()
+
+    diameter=max([len(path) for path in all_shortest_paths])
+    paths_forming_diameter=[path for path in all_shortest_paths if len(path)==diameter]
+    return len(set([node for path in paths_forming_diameter for node in path]))
+#flattening the list and using set() to get rid of duplicate nodes that are in multiple diameters
+
+print(  '---------------------------------------')
+starttime = time.time()
+s=1
+maxi=30
+#fitmin=8
+#mini=8
+#meandias=np.zeros((maxi-mini,2))
+#meandias2=np.zeros((maxi-mini,2))
+#meanvol=np.zeros((maxi-mini,2))
+#meanp=np.zeros((maxi-mini,2))
+meandias=np.zeros((maxi,2))
+meandias2=np.zeros((maxi,2))
[…]

### 2021-12-03 src/functions.py

@@ -1,0 +1,224 @@
+'''
+Functions to be used in main.py. 
+
+'''
+import os
+import pandas as pd
+import Bio.PDB.PDBList as biopy
+import PCN
+import traceback
+import tarfile
+import random
+import numpy as np
+
+def concat_ID_files():
+	'''
+	Opens the chunked PDB id files as a DataFrame. Returns an array with of all the IDs.
+	
+	Parameters
+	----------
+	None
+
+	Return
+	------
+	PDB_ID_array: numpy array
+		array containing all the PDB IDs
+	'''
+	frames = []
+	for filename in os.listdir('../data/ids/'):
+		if filename.endswith('.txt'):
+			temp_df = pd.read_csv('../data/ids/'+str(filename), header = None)
+			frames.append(temp_df)
+
+	concatenated_DF = pd.concat(frames).reset_index()
+	concatenated_DF = concatenated_DF.drop(columns = 'index')
+	PDB_ID_array = concatenated_DF[0].to_numpy()
+	
+	return PDB_ID_array
+
+def bioretrieve_PDBs(PDB_ID_array, save_directory):
+	'''
+	Take the array of PDB IDs and use BioPython to retrieve PDB files as .ent files.
+
+	Parameters
+	----------
+	PDB_ID_array: numpy array
+		array containing all the PDB IDs
+	save_directory: string 
+		path to the directory where the file will be saved
+
+	Return
+	------
+	None
+	'''
+	pdblist = biopy.PDBList()
+	PDB_IDs = concat_ID_files()
+	counter = 1
+	for PDB_ID in PDB_IDs:
+		print(f'At entry {counter} out of {len(PDB_IDs)} entries.')
+		pdblist.retrieve_pdb_file(pdb_code = PDB_ID, file_format = 'pdb', pdir = save_directory)
+		counter += 1
+
+def bootstrap(length, link_lengths, length_range):
+	'''
+	Take the link lengths and bootstrap the data. 
+
+	Saves the mean and median of the bootstrapped data and the PDB ids belonging to
+	that length range. Files are saved in .npy format. 
+
+	Parameters
+	----------
+	length: string
+		defines the length range as '100', '200' or '300'
+
+	link_lengths: list
+		list containing lengths of links
+
+	length_range: list
+		list containing the PDB ids that fall in the length range given
+
+	Return
+	------
+	None
+	'''
+	if len(link_lengths) > 0:
+		sample_size = 1000
+		histogram_list = []
+		means = []
+
+		for n in range(sample_size):
[…]

### 2021-12-03 src/get_link_lengths.py

@@ -1,0 +1,134 @@
+
+
+# Imports
+import numpy as np
+import networkx as nx
+import traceback
+import os
+from Bio.PDB import *
+from MDAnalysis import *
+import warnings
+import urllib
+import pandas as pd
+import socket
+warnings.filterwarnings('ignore')
+
+
+
+def connected_component_subgraphs(G):
+    for c in nx.connected_components(G):
+        yield G.subgraph(c)
+
+# Get data from the pdb
+def do_analysis(calphas, link_length_list, t=8.0):
+    linklengths = []
+    r = calphas.positions
+    maxdist = 0
+    G = nx.empty_graph(len(r))
+    avd = 0
+    for i in range(len(r)):
+        for j in range(i):
+            dist = np.linalg.norm(r[i] - r[j])
+            maxdist = max(dist, maxdist)
+            avd += dist ** 2
+            if dist < t:
+                G.add_edge(i, j)
+                linklengths.append(dist)
+    avd = np.sqrt(avd) / (len(r))
+    # take largest component
+    if len(G.edges()) > 0:
+        G = list(connected_component_subgraphs(G))[0]
+        lena = len(G.nodes())
+        Li = list(set(G.edges()) - set(nx.cycle_graph(lena).edges()))
+        for link in Li:
+            #return(abs(link[0]-link[1]))
+            link_length_list.append(abs(link[0] - link[1]))
+
+def save_current_data(length, ll_list, id_list):
+    if len(ll_list)>0:
+        sim_data_hist, sim_data_edges = np.histogram(ll_list, bins=range(3, 1000), normed=True)
+        sim_data_hist = sim_data_hist/len(id_list)
+
+        fname1 = 'simd_data_hist_%s_curr.npy' % length
+        fname2 = 'Ids_%s_curr.npy' % length
+        np.save(fname2, id_list)
+        np.save(fname1, sim_data_hist)
+
+if __name__ == "__main__":
+    # Reading the pdb IDs
+    socket.setdefaulttimeout(20)
+    query = pd.read_csv('../data/combined_ids.txt')
+    hundred = []
+    twohundered = []
+    threehundred = []
+    fourhundred = []
+    ll_100 = []
+    ll_200 = []
+    ll_300 = []
+    ll_400 = []
+    counter = 0
+    n_pdbs = len(query.columns)
+    with open("log.txt", "w") as log:
+        for pdb in query.columns:
+            if counter % 500 == 0:
[…]

### 2021-12-03 src/get_link_lengths_with_stats.py

@@ -1,0 +1,172 @@
+
+
+# Imports
+import numpy as np
+import networkx as nx
+import traceback
+import os
+from Bio.PDB import *
+from MDAnalysis import *
+import warnings
+import urllib
+import pandas as pd
+import socket
+import random
+import operator
+warnings.filterwarnings('ignore')
+
+def connected_component_subgraphs(G):
+    for c in nx.connected_components(G):
+        yield G.subgraph(c)
+
+# Get data from the pdb
+def do_analysis(calphas, t=8.0):
+    link_length_list = []
+    linklengths = []
+    r = calphas.positions
+    maxdist = 0
+    G = nx.empty_graph(len(r))
+    avd = 0
+    for i in range(len(r)):
+        for j in range(i):
+            dist = np.linalg.norm(r[i] - r[j])
+            maxdist = max(dist, maxdist)
+            avd += dist ** 2
+            if dist < t:
+                G.add_edge(i, j)
+                linklengths.append(dist)
+    avd = np.sqrt(avd) / (len(r))
+    # take largest component
+    if len(G.edges()) > 0:
+        G = list(connected_component_subgraphs(G))[0]
+        lena = len(G.nodes())
+        Li = list(set(G.edges()) - set(nx.cycle_graph(lena).edges()))
+        for link in Li:
+            #return(abs(link[0]-link[1]))
+            link_length_list.append(abs(link[0] - link[1]))
+    return link_length_list
+
+def save_current_data(length, ll_list, id_list):
+    if len(ll_list)>0:
+        sim_data_hist, sim_data_edges = np.histogram(ll_list, bins=range(3, 1000), normed=True)
+        sim_data_hist = sim_data_hist/len(id_list)
+
+        fname1 = 'simd_data_hist_%s_curr.npy' % length
+        fname2 = 'Ids_%s_curr.npy' % length
+        np.save(fname2, id_list)
+        np.save(fname1, sim_data_hist)
+
+def save_bootstrapped_current_data(length,ll_list,id_list, n_samples = 1000):
+    print (np.shape(ll_list), flush=True)
+    if len(ll_list)>0:
+
+        sim_data_hist_list = []
+        sim_data_edges = None
+        # Now we need to boostrap the list a bunch of times
+        for n in range(n_samples):
+            bts_ll_list = [random.choice(ll_list) for _ in ll_list]
+            #bts_ll_list = np.array(bts_ll_list)
[…]

### 2021-12-03 src/main.py

@@ -1,0 +1,8 @@
+'''
+Main 
+
+'''
+import functions
+
+if __name__ == '__main__':
+	functions.main()

### 2021-12-03 src/pdb_linklength.py

@@ -1,0 +1,254 @@
+# -*- coding: utf-8 -*-
+from __future__ import print_function
+
+import glob
+import os
+
+import matplotlib.pyplot as plt
+import networkx as nx
+import numpy as np
+import scipy as sp
+import scipy.stats
+import seaborn as sns
+from Bio.PDB import *
+from MDAnalysis import *
+from scipy.optimize import curve_fit
+import pickle
+
+sns.set(font_scale=1.5)  # crazy big
+sns.set_style("ticks")
+sns.despine()
+
+
+def func(x, a, b):
+    return a * x ** (-b)
+
+
+def func2(x, a):
+    return a / x
+
+
+def theofunc(x, C, N, ex):
+    return (C) * np.power(N / (2 * (-x ** 2 + N * x - N)), ex)
+
+
+def connected_component_subgraphs(G):
+    for c in nx.connected_components(G):
+        yield G.subgraph(c)
+
+
+# def theofunc(x, C,N):
+#     return (C)* N/(2*(N*x-N))
+
+names_xray = glob.glob("../data/pdb_data2/all_xray/*.pdb")
+names_xray = glob.glob("../data/pdb_data2/xray_data_short/*.pdb")
+names_xray = glob.glob("../data/pdb_data2/xray_data/*.pdb")
+
+names_nmr = glob.glob("../datapdb_data2/nmr_data/*.pdb")
+names_nmr = glob.glob("../data/pdb_data2/nmr_data_short/*.pdb")
+
+names_both = glob.glob("../data/pdb_data2/both/*.pdb")
+
+# names_both = glob.glob("idp/*.pdb")
+
+names_used = names_both
+
+num = 5
+lengths = [3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 16, 17, 19, 21, 23, 25, 28, 30, 33, 36, 40, 44, 48, 52, 58, 63, 69,
+           76, 83, 91, 100, 110, 120, 132, 145, 158, 174, 191, 209, 229, 251, 275,
+           302]#, 331, 363, 398]  # [0:45]
+print(len(lengths))
+
+numnum = 20
+logscale = np.logspace(np.log10(3), np.log10(1000), num=numnum)
+print('logscale', logscale)
+results = []
+binnedresults = []
+
+if __name__ == '__main__':
+    directories = ['../data/pdb_data2/both']
+    # directories = ['idp']
+    nameslen = len(names_used)
+    # nameslen=50
+    path = os.getcwd()
[…]

### 2021-12-03 paper/main.tex
fixing folder

@@
 \affiliation{Center for Advancing Electronics Dresden (cfaed), Technical University of Dresden}
-\author{J. J.  Güven}
-\affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
 \author{Steffen Mühle}
@@
 \affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
+%\author{Marc Timme}
+% \affiliation{Network Dynamics, Max Planck Institute for Dynamics and Self-Organization (MPIDS), 37077 Göttingen, Germany}
+% \affiliation{Chair for Network Dynamics, Institute for Theoretical Physics and Center for Advancing Electronics Dresden (cfaed), Technical University of Dresden, 01069 Dresden}
 
@@
 Here, we show how a simple geometrical model can reproduce this heuristically found behaviour well, and even provide a simple analytical approximation.
-The three-dimensional geometrical model and the analytical approximation of its two-dimensional analogue both build on the idea that inherently geometrical objects, such as amino acids, imply that any resulting PCN network ensemble modelling them has to be spatially embedded~\cite{molkenthin2016scaling, molkenthin2020self}.
+\blue{The three-dimensional geometrical model and the analytical approximation of its two-dimensional analogue both} build on the idea that inherently geometrical objects, such as amino acids, imply that any resulting PCN network ensemble modelling them has to be spatially embedded~\cite{molkenthin2016scaling, molkenthin2020self}.
 Different approaches have been used in the past for geometrical models describing protein folding, some of which derived characteristics of the secondary structure from constraints on bond and torsion angles~\cite{bhattacharjee2013flory,Danielsson2010,Molkenthin2011} and others focused on the formation mechanisms of the tertiary structure~\cite{molkenthin2016scaling, molkenthin2020self}.
@@
 The underlying formation mechanisms of individual protein folds are incredibly complex and depend on the surrounding solvent as well as the specific amino acid sequence and their quantum-mechanical interactions.
-Here we want to study the ensemble of all folded proteins, however, and will thus "average" over such specifics.
+Here we want to study the ensemble of all \blue{folded(?)} proteins, however, and will thus "average" over such specifics.
 
-Based on the basic observation that amino acids are objects in space which can not overlap indefinitely, we introduced a model in \cite{molkenthin2020self} starting from a chain of identical spheres, each in contact only with the neighbours it is connected to.
+Based on the basic \blue{observation} that amino acids are objects in space which can not overlap indefinitely, we introduced a model in \cite{molkenthin2020self} starting from a chain of identical spheres, each in contact only with the neighbours it is connected to.
 Then additional connections form by randomly selecting two spheres and moving them towards each other until they touch, the new links formed that way can not be broken in later steps and function as constraints on subsequent links.
@@
 
-In \cite{molkenthin2016scaling}, we have introduced a simplified, two-dimensional version of this model, which starts from a closed chain (i.e.
-a ring) of $N$ unit discs and subsequently adds links, such that connected discs touch, yet no discs intersect.
+In \cite{molkenthin2016scaling}, we have introduced a simplified, two-dimensional version of this model, which starts from a closed chain \blue{(i.e.
+a ring)} of $N$ unit discs and subsequently adds links, such that connected discs touch, yet no discs intersect.
 The advantage of this model is that it can be approximated by a purely topological simplification, which can be treated analytically.
@@
 The expected distribution of possible links after one step is given thus by the average over available link pools taken over all possible lengths of the initial link
-\begin{align}
+\blue{\begin{align}
     F_1(s)&=\frac{1}{N/2-1} \sum_{s_1=2}^{N/2}{ \begin{cases}
@@
     \end{cases}}
-\end{align}
+\end{align}}
+Here we denote the size of the pool of available links after $i$ links have been added $F_i(s)$.
 
-In subsequent steps we take the weighted average over new links being added at step $k-1$, where the probability that a link of length $s$ can still be made is
-\begin{equation}
-    P_{in}^k(s)=F_k(s)/C_k,
-\end{equation}
-where $C_k=\sum_{s=2}^{N/2}F_k(s)$ and get the same reduction (compare Fig.~\ref{fig:schematic}) but starting from the pool $F_{k-1}(s)$ of the step before rather than the full $N$ links for each length.
-To account for possible overlaps between newly removed links and links that were removed previously, we multiply by
-\begin{equation}
-    \frac{F_k(s)}{F_0(s)} \approx F_k(s)\frac{1}{N}.
+In subsequent steps we \blue{still average over $N/2-1$ possible link lengths made at step $k-1$, each occurring with probability $F_k(s)/C_k$, where $C_k=\sum_{s=2}^{N/2}F_k(s)$ and} get the same reduction \blue{(compare Fig.~\ref{fig:schematic})} but starting from the pool $F_{k-1}(s)$ of the step before rather than the full $N$ links for each length.
+The reduction also needs to be multiplied by a factor \blue{accounting for already removed links}
+\blue{\begin{equation}
+    P_k(s)=\frac{F_k(s)}{N/2-1} \approx F_k(s)\frac{2}{N},
     \label{eq.Pk}
-\end{equation}
+\end{equation}}
 % \begin{equation}
@@
 %with $C_k=\sum_{s=2}^{N/2}F_k(s)$, 
-This ignores any bias in which links are removed, however, in small, manageable examples this assumption seemed to hold well. 
-This leads to the recursion
-\begin{align}
-    F_k(s)&= \sum_{s_k=2}^{N/2} P_{in}^k(s_k)\nonumber \\
-    & (F_{k-1}(s)-  F_k(s)\frac{1}{N}
-    {\begin{cases}
-     2(s-1) \text{ , for } s<s_k\\
-     2(s_k -1))\text{ , for } s\geq s_k
-    \end{cases}} )
-\end{align}
-which simplifies to
-\begin{align}
-   F_k(s)&=F_{k-1}(s)(1-\frac{1}{N} \sum_{s_k=2}^{s} (2 (s_k-1) P_{in}^k(s_k))\nonumber \\
-   &-\frac{1}{N} \sum_{s_k=s+1}^{N/2} (2 (s-1) P_{in}^k(s_k)))\nonumber \\
-   &=F_{k-1}(s)(1-\frac{1}{N}
-   \sum_{s_k=2}^{s} (2 (s_k-1) P_{in}^k(s_k))\nonumber \\
-   &-\frac{2 (s-1)}{N} (\sum_{s_k=2}^{N/2} P_{in}^k(s_k) - \sum_{s_k=2}^{s} P_{in}^k(s_k)))\nonumber \\
-   &=F_{k-1}(s)(1-\frac{2}{N} (s-1) 
-   +\frac{2}{N} s
-   \sum_{s_k=2}^{s}P_{in}^k(s_k)\nonumber \\
-   &-\frac{2}{N}
-   \sum_{s_k=2}^{s} s_k P_{in}^k(s_k))
+to only remove links that are still in the pool. This leads to the recursion
+\blue{\begin{align}
+    F_k(s)&= \sum_{s_k=2}^{N/2} F_k(s)/C_k\nonumber \\
+    &{\begin{cases}
+     F_{k-1}(s)-2(s-1) P_{k-1}(s) \text{ , for } s<s_k\\
+     F_{k-1}(s)-2(s_k -1)P_{k-1}(s)\text{ , for } s\geq s_k
+    \end{cases}}\nonumber \\
+    &\approx \frac{1}{N/2} \sum_{s_k=2}^{N/2} \nonumber \\
+    &{\begin{cases}
+     F_{k-1}(s)-2(s-1) P_{k-1}(s) \text{ , for } s<s_k\\
+     F_{k-1}(s)-2(s_k -1)P_{k-1}(s)\text{ , for } s\geq s_k
+    \end{cases}},
+    \label{eq.Fk_def}
+\end{align}}
+\blue{where we made the approximation that the probabilities $F_k(s)/C_k\approx 2/N$ remain approximately uniform.
+This approximation holds best for small $s$ and large $N$.} \red{Toni and Steffen, does either of you have access to a Mathematica license?
+I could start bugging my admin about getting one but i think that would probably take longer than we want to spend on this.
+If we find a license we could probably take the first line of eq.
+4 and get a plot out using Rsolve or something...
+And plot the result with the approximation.}
+
+\blue{We further simplify Eq.~\ref{eq.Fk_def}.}
+
+\blue{\begin{align}
+   F_k(s)&\approx \frac{4}{N^2} F_{k-1}\nonumber \\
+   &\left( \sum_{s_k=2}^{N/2}\frac{N}{2} - \sum_{s_k=2}^{s} 2(s_k-1) - \sum_{s_k=s+1}^{N/2} 2(s -1) \right) \nonumber \\
+   &= \frac{4}{N^2} F_{k-1}\nonumber \\
+   &\left(\frac{N^2}{4} -(s^2-s)
+   -2\left(\frac{N}{2}-(s+1)\right)(s-1)\right)  \nonumber \\
+   &= F_{k-1}(s)\left(1-\frac{-s^2 +(N-1)s-(N+2)}{\frac{N^2}{4}} \right)\nonumber \\
+   &=F_{k-1}(s)\left(1-\frac{f(s)}{\frac{N^2}{4}} \right)
    \label{eq.Fk_rec}
-\end{align}
+\end{align}}
+% \begin{align}
+%   F_k(s)&= \frac{1}{\frac{N}{2}-2} \frac{F_{k-1}(s)}{C_{k-1}}\nonumber \\
+%   &\left( \sum_{s_k=2}^{N/2}C_{k-1} - \sum_{s_k=2}^{s} 2(s_k-1) - \sum_{s_k=s+1}^{N/2} 2(s -1) \right) \nonumber \\
+%   &= \frac{1}{\frac{N}{2}-2}\frac{F_{k-1}(s)}{C_{k-1}}\nonumber \\
+%   &\left(\left(\frac{N}{2}-2\right)C_{k-1} -(s^2-s)
+%   -2\left(\frac{N}{2}-(s+1)\right)(s-1)\right)  \nonumber \\
+%   &= F_{k-1}(s)\left(1-\frac{-s^2 +(N-1)s-(N+1)}{(\frac{N}{2}-2)C_{k-1}} \right)\nonumber \\
+%   &=F_{k-1}(s)\left(1-\frac{f(s)}{(\frac{N}{2}-2)C_{k-1}} \right)
+%   \label{eq.Fk_rec}
+% \end{align}
+Where we defined $f(s)=-s^2 +(N-1)s-(N+2) \approx N s - s^2 - N$ for notational brevity.
 
-To approximately evaluate this we make a guess for $P_{in}^k(s)$ at low $k$ and at high $k$, namely
-\begin{equation}
-    P_{in}^{k\ll N/2}(s)\approx P_{in}^{k=0}(s)=2/N.
-    \label{eq.P_small}
-\end{equation}
-For later steps $k$, the probability of a link to still be in the pool decreases with $s$, we thus make the ansatz, that it drops off as $1/s$, leading to
-\begin{equation}
-    P_{in}^{k\gg 1}(s)\approx \frac{s^{-1}}{ (H_{\frac{N}{2}}-1)},
-    \label{eq.P_large}
-\end{equation}
-where $H_{\frac{N}{2}}$ is the harmonic number. 
-We suspect that the approximation for low $k$ only holds for a couple of steps while Eq.~\ref{eq.P_large} holds, at least on average for larger $k$.
-We have no approximation for the intermediate steps but assume that the errors roughly cancel each other out.
-Inserting Eq.\ref{eq.P_small} and Eq.\ref{eq.P_large} into Eq.\ref{eq.Fk_rec} results in the following two recursions for the early and late evolution of the sequence distance distribution respectively.
-\begin{align}
-   F_{k\ll N/2}(s)&\approx 
-   F_{k-1}(s)(1-\frac{2}{N} (s-1) 
-   +\frac{4}{N^2} s (s-1)\nonumber \\
-   &-\frac{2}{N^2}(s^2+s-2)\nonumber \\
-   &=F_{k-1}(s)(1-f_{low}(s)),
-   \label{eq.Fk_rec_low}
-\end{align}
-where 
-\[f_{low}(s)=-\frac{2}{N^2}s^2+(\frac{2}{N}+\frac{6}{N^2})s-(\frac{2}{N}+\frac{4}{N^2}),\]
-and
-\begin{align}
-   F_{k\gg1}(s)&\approx 
-   F_{k-1}(s)(1-\frac{2}{N} (s-1) 
-   +\frac{2}{N} s \sum_{s_k=2}^{s}\frac{1}{s_k(H_{\frac{N}{2}}-1)}\nonumber \\
-   &-\frac{2}{N}
-   \sum_{s_k=2}^{s} s_k \frac{1}{s_k(H_{\frac{N}{2}}-1)})\nonumber \\
-   &=
-   F_{k-1}(s)(1 -(\frac{2}{N}-\frac{2}{N}\frac{H_{s}-1}{H_{\frac{N}{2}}-1}+\frac{2}{N(H_{\frac{N}{2}}-1)})s \nonumber \\ 
-   &+ \frac{2}{N(H_{\frac{N}{2}}-1)} + \frac{2}{N})\nonumber \\
-   &=F_{k-1}(s)(1-f_{high}(s)).
-   \label{eq.Fk_rec_high}
-\end{align}
-where 
-\[f_{high}(s)=\frac{2}{N}(\frac{H_{\frac{N}{2}}-H_{s}+1}{H_{\frac{N}{2}}-1}s-\frac{H_{\frac{N}{2}}}{H_{\frac{N}{2}}-1}).\]
+We can now use the recursive expression Eq.~\ref{eq.Fk_rec} to write down a closed expression for $F_k(s)$
 
-We can now use the recursive expressions to write down a closed expression for $F_k(s)$
+\blue{\begin{equation}
+    F_k(s)=N\prod_{i}^k\left(1-\frac{f(s)}{\frac{N^2}{4}} \right) = N \left(1-\frac{f(s)}{\frac{N^2}{4}} \right)^k
+\end{equation}}
+% \begin{equation}
+%     F_k(s)=N\prod_{i}^k\left(1-\frac{f(s)}{(\frac{N}{2}-2)C_{i-1}} \right)
+% \end{equation}
 
-\begin{align}
-    F_k(s)&=N\prod_{i_1}^a\left(1-f_{low}(s)\right)\prod_{i=a+1}^k\left(1-f_{high}(s)\right) \nonumber \\
-    &= N\left(\frac{1-f_{low}(s)}{1-f_{high}(s)}\right)^a\left(1-f_{high}(s)\right)^{k}
-\end{align}
-
-In 2D this is repeated for $N-3$ steps until no more links can be added \cite{molkenthin2016scaling}.
+\blue{In 2D this is repeated for $N-3$ steps until no more links can be added \cite{molkenthin2016scaling}.}
 The probability distribution of the realized residue distances of all added links is then given by the average over the available pools at each link addition step, leading to
-\begin{align}
+\blue{\begin{align}
     P(s)&=\frac{1}{N-3}\sum_{k=0}^{N-4} \frac{F_k(s)}{C_k} \nonumber \\
-    &= \frac{N}{N-3}\left(\frac{1-f_{low}(s)}{1-f_{high}(s)}\right)^a
-\sum_{k=0}^{N-4} \frac{1}{C_k} \left(1-f_{high}(s)\right)^{k} \nonumber \\
-    &\approx \left(\frac{1-f_{low}(s)}{1-f_{high}(s)}\right)^a\frac{A}{f_{high}(s)},
+    &= \frac{N}{N-3}\sum_{k=0}^{N-4} \frac{1}{C_k} \left(1-\frac{f(s)}{\frac{N^2}{4}} \right)^k \nonumber \\
+    &\approx \frac{A}{f(s)},
     \label{eq.Ps}
-\end{align}
-
-\red{The assumption that the sum is basically a harmonic series seems like a bit of a stretch I shall check it again but the results look good}
-
-for $1<s<\frac{N}{2}$.
+\end{align}}
+% \begin{align}
+%     P(s)&=\frac{1}{N-3}\sum_{k=0}^{N-3} P_k(s) \nonumber \\
+%     &= \frac{N}{N-3}\sum_{k=0}^{N-3} \frac{1}{C_k}\prod_{i}^k\left(1-\frac{f(s)}{(\frac{N}{2}-2)C_{i-1}} \right),
+% \end{align}
+\blue{for $1<s<\frac{N}{2}$.
 The approximation sets all $C_k=C$ to allow the use of the geometric series for solving the equation.
-We can do this because $C_k$ monotonically decreases with $k$ and $N>C_k>C_{N-2}$ meaning we can obtain an upper and lower bound by replacing all $C_k$ in the expression with either $N$ (underestimates all probabilities) or $C_{N-2}$ (overestimates all probabilities), however, as neither of those depend on $s$, these bounds are merely in the constant factor $A$ and do not affect the scaling as such.
+We can do this because $C_k$ monotonically decreases with $k$ and $N>C_k>C_{N-2}$ meaning we can obtain an upper and lower bound by replacing all $C_k$ in the expression with either $N$ (underestimates all probabilities) or $C_{N-2}$ (overestimates all probabilities), however, as neither of those depend on $s$, these bounds are merely in the constant factor $A$ and do not affect the scaling as such.}
 %which makes use of Eq.~\ref{eq.Pk}. 
@@
 % where $A$ is an unknown constant, containing an approximation of the various normalization constants $C_i$, as well as the direct effects of the chain length. 
-Since for long chains and $s \ll N/2$, $f(s)$ is dominated by the linear term, the distribution is well approximated by a power law with an exponent of $-1$.
+\blue{Since for long chains and $s \ll N/2$, $f(s)$ is dominated by the linear term, the distribution is well approximated by a power law with an exponent of $-1$.}
 
-%In the last step of this approximation, we took the $C_k$ to all be approximately equal in order to apply the geometric series.
+In the last step of this approximation, we took the $C_k$ to all be approximately equal in order to apply the geometric series.
 However, in reality we know $1/C_k$ to increase with $k$, giving large $k$ summands relatively more weight than in a geometric series.
@@
 
-%\begin{equation}
- %   P_{corrected}(s)\approx A f(s)^{-1+\alpha},
-%\end{equation}
-%where $0<\alpha \ll 1$.
+\begin{equation}
+    P_{corrected}(s)\approx A f(s)^{-1+\alpha},
+\end{equation}
+where $0<\alpha \ll 1$.
 \blue{The parameters $A$ and $\alpha$ in this approximation are treated as free parameters and used for fitting the theoretical result to data and simulations.
@@
 For large values of the chain length $N$ we thus find an approximate power law probability distribution for the residue distance with an exponent of $-1+\alpha$.
-Here we use simulations, as introduced in~\cite{molkenthin2020self}, and networks extracted from measured protein data from the PDB \cite{PDB} to compare this behaviour with the more realistic 3D settings and show that the basic principle persists, indicating that much of the residue distance distribution can be explained by generic geometric constraints.
+Here we use simulations, as introduced in~\cite{molkenthin2020self}, and networks extracted from measured protein data from the PDB \cite{PDB} to compare this behaviour \blue{with the more} realistic 3D settings and show that the basic principle persists, indicating that much of the residue distance distribution can be explained by simple\blue{/generic?} geometric constraints.
  \begin{figure}[h]
@@
 structures of around 400 residues from the PDB.} Those have been converted into PCN's based on the measured locations of their residues to analyze the distribution of residue distances between connected amino acids.
-The resulting distribution is compared to those generated in simulations in~\cite{molkenthin2020self} as well as the above derived theoretical expression scaled with a factor to account for the different dimensionality, as well as approximations made in the derivation.
+The resulting distribution is compared to those generated in simulations in~\cite{molkenthin2020self} \blue{as well as the above derived theoretical expression scaled with a factor to account for the different dimensionality, as well as approximations made in the derivation.}
 

### 2021-12-03 data/Lunaplot.py

@@ -1,0 +1,135 @@
+import networkx as nx
+import numpy as np
+import scipy as sp
+import scipy.stats
+import matplotlib.pyplot as plt
+import matplotlib
+import math as m
+import random
+import time
+import bisect
+from scipy.optimize import curve_fit
+#from compiler.ast import flatten
+from operator import add
+
+def func(x, a, b):
+     return a * x**(-b)
+
+def func1(x, a):
+     return a/x
+
+def expfit(x, a, b):
+     return a * np.exp(x*(-b))
+
+n=1000
+Luna=np.zeros((10,498))
+Luna[1]=np.loadtxt('luna_noencircle1')
+Luna[2]=np.loadtxt('luna_noencircle2')
+Luna[3]=np.loadtxt('luna_noencircle3')
+Luna[4]=np.loadtxt('luna_noencircle4')
+Luna[5]=np.loadtxt('luna_noencircle5')
+Luna[6]=np.loadtxt('luna_noencircle6')
+Luna[7]=np.loadtxt('luna_noencircle7')
+Luna[8]=np.loadtxt('luna_noencircle8')
+Luna[9]=np.loadtxt('luna_noencircle9')
+Luna[0]=np.loadtxt('luna_noencircle10')
+
+#n=2000
+#Luna[1]=np.loadtxt('luna1')
+#Luna[2]=np.loadtxt('luna2')
+#Luna[3]=np.loadtxt('luna3')
+#Luna[4]=np.loadtxt('luna4')
+#Luna[5]=np.loadtxt('luna5')
+#Luna[6]=np.loadtxt('luna6')
+#Luna[7]=np.loadtxt('luna7')
+#Luna[8]=np.loadtxt('luna8')
+#Luna[9]=np.loadtxt('luna9')
+#Luna[10]=np.loadtxt('luna10')
+#Luna[11]=np.loadtxt('luna11')
+#Luna[12]=np.loadtxt('luna12')
+#Luna[13]=np.loadtxt('luna13')
+#Luna[14]=np.loadtxt('luna14')
+#Luna[15]=np.loadtxt('luna15')
+#Luna[16]=np.loadtxt('luna16')
+#Luna[17]=np.loadtxt('luna17')
+#Luna[18]=np.loadtxt('luna18')
+#Luna[19]=np.loadtxt('luna19')
+#Luna[0]=np.loadtxt('luna10')
+
+Lunatot=np.mean(Luna,axis=0)
[…]

### 2022-01-14 src/PCN.py
shameful commit dump: WIP alphafold stuff

@@ -6,6 +6,9 @@ import networkx as nx
 import MDAnalysis as mda
 import numpy as np
 import random
+import warnings
+warnings.filterwarnings("ignore")
+import pandas as pd
 
 class PCN():
     """
@@ -90,8 +93,8 @@ class PCN():
         None
         """
         raise AttributeError('The threshold value cannot be deleted. You can set it to 0.')
-
-    def get_C_alphas(self):
+        
+    def get_alpha_carbons(self):
         
         """
         Take a universe, get segment A and return the C-alphas for the PDB. 
@@ -102,20 +105,40 @@ class PCN():
 
         Return
         ------
-        C_alphas: MDAnalysis AtomGroup
+        alpha_carbons: MDAnalysis AtomGroup
             contains the C-alphas from the PDB
         """
         "------------------------MOVE-------------------------------"
         segments = self.universe.residues.segments
             
-        number_of_segments = len(segments)
+        n_segments = len(segments)
 
-        for index in range(number_of_segments):
-            segment_id = self.universe.residues.segments.segids[index]
+        for i in range(n_segments):
+            segment_id = self.universe.residues.segments.segids[i]
 
             try:
-                C_alphas = self.universe.select_atoms(f'name CA and segid {segment_id}')
-                #print(segment_id)
+                protein_atoms = self.universe.select_atoms('protein')
+                MDA_altlocs = protein_atoms.altLocs
+                alternative_locations = []
+                for loc in MDA_altlocs:
+                    if loc != '':
+                        alternative_locations.append(loc)
+                # Remove duplicates and sort alphabetically
+                alternative_locations = np.sort(list(set(alternative_locations)))
+                n_altlocs = len(alternative_locations)
+
+                if n_altlocs > 1:
+                    first_altloc = alternative_locations[0]
+                    residue_ids = protein_atoms.residues.resids
+                    alpha_carbons = self.universe.select_atoms(f'resid {residue_ids[0]}:{residue_ids[-1]} and name CA and segid {segment_id}')
+
+                    for i in range(1, n_altlocs):
+                        exclude_atoms = self.universe.select_atoms(f'resid {residue_ids[0]}:{residue_ids[-1]} and name CA and segid {segment_id} and altLoc {alternative_locations[i]}')
+                        alpha_carbons -= exclude_atoms
+                else:
+                    residue_ids = protein_atoms.residues.resids
[…]

### 2022-01-14 src/PDBs_from_uniprot.py

@@ -1,0 +1,23 @@
+
+import functions
+import urllib
+
+sequence_lengths = ['100', '200', '300']
+
+for sequence_length in sequence_lengths:
+    counter = 0
+
+    uniprot_IDs = functions.get_uniprot_IDs(sequence_length)
+    for uniprot_ID in uniprot_IDs:
+        print(f'At entry {counter}/{len(uniprot_IDs)}')
+        download = f'https://alphafold.ebi.ac.uk/files/AF-{uniprot_ID}-F1-model_v2.pdb'
+
+        try:
+            file_name = f'../data/alphafold_data/AF_PDBs/AF-{uniprot_ID}-F1-model_v2.pdb'
+            #bio.retrieve_pdb_file(pdb,pdir='../data/tmep/', file_format='pbd')
+            urllib.request.urlretrieve(download, file_name)
+        except Exception:
+            
+            'error'
+        counter += 1
+         

### 2022-01-14 src/bio_retrieve_pdbs.py

@@ -6,28 +6,28 @@ import pandas as pd
 pdblist = PDBList()
 
 def concat_ID_files():
-	''' Opens id_file_N.txt as a DF and concats them all into one DF '''
-	frames = []
-	for filename in os.listdir('../data/ids/'):
-		if filename.endswith(".txt"):
-			temp_df = pd.read_csv('../data/ids/'+str(filename), header=None)
-			frames.append(temp_df)
+    ''' Opens id_file_N.txt as a DF and concats them all into one DF '''
+    frames = []
+    for filename in os.listdir('../data/ids/'):
+        if filename.endswith(".txt"):
+            temp_df = pd.read_csv('../data/ids/'+str(filename), header=None)
+            frames.append(temp_df)
 
-	DF = pd.concat(frames).reset_index()
-	DF = DF.drop(columns='index')
-	PDB_list = DF[0].to_numpy()
-	return PDB_list
+    DF = pd.concat(frames).reset_index()
+    DF = DF.drop(columns='index')
+    PDB_list = DF[0].to_numpy()
+    return PDB_list
 
 PDBs = concat_ID_files()
 counter = 1
 for pdb in PDBs:
-	print(f'At entry {counter}')
-	pdblist.retrieve_pdb_file(pdb_code = pdb, 
-						      file_format = "pdb", 
-						      pdir = '/Volumes/Seagate_Extension_Plus/PDBs/')
-	counter += 1
+    print(f'At entry {counter}')
+    pdblist.retrieve_pdb_file(pdb_code = pdb, 
+                              file_format = "pdb", 
+                              pdir = '/Volumes/Seagate_Extension_Plus/PDBs/')
+    counter += 1
 
 # pdb = '1AKI'
 # pdblist.retrieve_pdb_file(pdb_code = pdb, 
-# 						  file_format = "bundle", 
-# 						  pdir = '../data/bioretrieve_PDBs/')
+#                           file_format = "bundle", 
+#                           pdir = '../data/bioretrieve_PDBs/')

### 2022-01-14 src/exclude.py

@@ -10,29 +10,29 @@ import pandas as pd
 import os
 
 def concat_ID_files():
-	'''
-	Opens the chunked PDB id files as a DataFrame. Returns an array with of all the IDs.
-	
-	Parameters
-	----------
-	None
+    '''
+    Opens the chunked PDB id files as a DataFrame. Returns an array with of all the IDs.
+    
+    Parameters
+    ----------
+    None
 
-	Return
-	------
-	PDB_ID_array: numpy array
-		array containing all the PDB IDs
-	'''
-	frames = []
-	for filename in os.listdir('/Users/jasminguven/Documents/GitHub/sequence_distance_distribution/data/ids/'):
-		if filename.endswith('.txt'):
-			temp_df = pd.read_csv('/Users/jasminguven/Documents/GitHub/sequence_distance_distribution/data/ids/'+str(filename), header = None)
-			frames.append(temp_df)
+    Return
+    ------
+    PDB_ID_array: numpy array
+        array containing all the PDB IDs
+    '''
+    frames = []
+    for filename in os.listdir('/Users/jasminguven/Documents/GitHub/sequence_distance_distribution/data/ids/'):
+        if filename.endswith('.txt'):
+            temp_df = pd.read_csv('/Users/jasminguven/Documents/GitHub/sequence_distance_distribution/data/ids/'+str(filename), header = None)
+            frames.append(temp_df)
 
-	concatenated_DF = pd.concat(frames).reset_index()
-	concatenated_DF = concatenated_DF.drop(columns = 'index')
-	PDB_ID_array = concatenated_DF[0].to_numpy()
-	
-	return PDB_ID_array
+    concatenated_DF = pd.concat(frames).reset_index()
+    concatenated_DF = concatenated_DF.drop(columns = 'index')
+    PDB_ID_array = concatenated_DF[0].to_numpy()
+    
+    return PDB_ID_array
 
 
 PDB_IDs = concat_ID_files()
@@ -45,48 +45,48 @@ range_300_PDB = []
 exclude_ids = []
 
 for PDB_ID in PDB_IDs:
-	print('--------------------------------------------------------------------------------------------------')
-	print(f'At entry {counter+1} out of {number_PDBs+1} entries.')
-	print(f'The PDB ID is {PDB_ID}')
-	# This will be replaced with untarring of pdb files
-	#-------------------------------------------------------------------
-	try:
[…]

### 2022-01-14 src/functions.py

@@ -8,8 +8,10 @@ import Bio.PDB.PDBList as biopy
 import PCN
 import traceback
 import tarfile
-import random
 import numpy as np
+
+import warnings
+warnings.filterwarnings("ignore")
 
 def concat_ID_files():
     """
@@ -36,6 +38,93 @@ def concat_ID_files():
     
     return PDB_ID_array
 
+def is_int(input_string, input_var): # NOT WORKING
+    """
+    Check if user input is an integer.
+
+    Parameters
+    ----------
+    input_string: str
+        string to be printed out to get user input
+    input_var: int
+        variable to hold user input, should be integer otherwise ask for new input
+    
+    Return
+    ------ 
+    integer_input: int
+        user input converted to an integer
+    
+    """
+    while True:
+
+        try:
+            integer_input = int(input_var)
+        
+        except ValueError:
+
+            print('Did not understand the command. Try again.')
+            continue
+        
+        else:
+            break
+    
+    return integer_input
+
+
+def is_file(filename):
+    """
+    Check if given filename exists. 
+    
+    Parameters
+    ----------
+    filename: str
+        filename given by user
+
+    Return
+    ------
+    checked_filename: str
+        correct filename
+
+    """
+    is_file = os.path.isfile(str(filename))
+    while is_file == False:
+        checked_filename = input(f'Invalid filename. Please enter again: ')
+        is_file = os.path.isfile(str(checked_filename))
+    return checked_filename
+
+def get_uniprot_IDs(sequence_length):
+    """
+    Take the sequence length and return an array of Uniprot IDs in that range.
+    Parameters
+    ----------
+    sequence_length: str
+        sequence length of PDBs, either 100, 200 or 300
+    
+    Return
+    ------
+    uniprot_IDs: numpy.array
+        array containing all the uniprot IDs of the given sequence length
+    """
+    uniprot_IDs_df = pd.read_excel(f'../data/alphafold_data/uniprot_{sequence_length}s.xlsx')
+    uniprot_IDs = uniprot_IDs_df['Entry'].to_numpy()
+
+    return uniprot_IDs
+
+def get_alphafold_PDBs(uniprot_IDs):
+    """
+    Parameters
+    ----------
+    uniprot_IDs: numpy.array
+        uniprot IDs of proteins in a given sequence length range
+    
+    Return
+    ------
+
+    """
+    sequence_length = '100'
[…]

### 2022-01-14 src/get_ll_stats_tidy.py

@@ -14,186 +14,186 @@ import operator
 warnings.filterwarnings('ignore')
 
 def concat_ID_files():
-	''' Opens id_file_N.txt as a DF and concats them all into one DF '''
-	frames = []
-	for filename in os.listdir('../data/ids/'):
-		if filename.endswith(".txt"):
-			temp_df = pd.read_csv('../data/ids/'+str(filename), header=None)
-			frames.append(temp_df)
+    ''' Opens id_file_N.txt as a DF and concats them all into one DF '''
+    frames = []
+    for filename in os.listdir('../data/ids/'):
+        if filename.endswith(".txt"):
+            temp_df = pd.read_csv('../data/ids/'+str(filename), header=None)
+            frames.append(temp_df)
 
-	DF = pd.concat(frames).reset_index()
-	DF = DF.drop(columns='index')
-	PDB_list = DF[0].to_numpy()
-	return PDB_list
+    DF = pd.concat(frames).reset_index()
+    DF = DF.drop(columns='index')
+    PDB_list = DF[0].to_numpy()
+    return PDB_list
 
 def make_connected_component_subgraphs(graph):
-	'''Take a graph and get subgraphs.'''
-	for components in nx.connected_components(graph):
-		yield graph.subgraph(components)
+    '''Take a graph and get subgraphs.'''
+    for components in nx.connected_components(graph):
+        yield graph.subgraph(components)
 
 
 def get_link_lengths(CA):
-	'''Get link lengths and return a list.'''
-	# Define the threshold for connections
-	threshold = 8.0 #Å
+    '''Get link lengths and return a list.'''
+    # Define the threshold for connections
+    threshold = 8.0 #Å
 
-	link_length_list = []
-	linklengths = []
+    link_length_list = []
+    linklengths = []
 
-	# Get positions of C alphas
-	r = CA.positions
-	# Create an empty graph
-	graph = nx.empty_graph(len(r))
+    # Get positions of C alphas
+    r = CA.positions
+    # Create an empty graph
+    graph = nx.empty_graph(len(r))
 
-	# Define max and average distances
-	max_dist = 0 # could be commented out
-	avg_dist = 0 # could be commented out
-	'''-------------------------------???-------------------------------'''
-	# Create the adjacency matrix
-	for i in range(len(r)):
-		for j in range(i):
-			# Get the norm, i.e. distnces between adjacent atoms
-			distance = np.linalg.norm(r[i] - r[j])
[…]

### 2022-01-14 src/get_subgraphs.py

@@ -1,0 +1,124 @@
+import PCN
+import functions
+import numpy as np
+import networkx as nx
+import os 
+
+
+PDB_IDs = functions.concat_ID_files()
+number_PDBs = len(PDB_IDs)
+
+counter = 0
+exclude_counter = 0
+
+exclude_IDs = np.load('../data/exclude.npy')
+
+subgraphs_100 = []
+subgraphs_200 = []
+subgraphs_300 = []
+ids_100 = []
+ids_200 = []
+ids_300 = []
+for i in range(number_PDBs):
+
+    if PDB_IDs[i] not in exclude_IDs:
+        print('--------------------------------------------------------------------------------------------------')
+        print(f'At entry {counter+1} out of {number_PDBs+1} entries.')
+        print(f'The PDB ID is {PDB_IDs[i]}')
+
+        try:
+            # PDB_tarfile = tarfile.open(f'../data/PDBs.tar.gz', 'r:gz')
+            PDB_IDs_lower = str(PDB_IDs[i]).lower()
+            # PDB_file = PDB_tarfile.extractfile(f'../data/PDBs/pdb{PDB_IDs_lower}.ent')
+            PDB_file = f'/Volumes/Seagate_Extension_Plus/PDBs/pdb{PDB_IDs_lower}.ent'
+            PDB_check = open(PDB_file, 'r')
+            print('Successfully opened file.')
+        except:
+            print('Could not open file.') # Log the exception
+            continue # Jump back to the top of the loop and try another PDB
+
+        if os.path.isfile(PDB_file):
+            # Create the PCN instance
+            protein_contact_network = PCN.PCN(PDB_file)
+            alpha_carbons = protein_contact_network.get_alpha_carbons()
+            chain_length = protein_contact_network.get_chain_length(alpha_carbons)
+            # Split into chain length ranges
+            if chain_length in range(85,116):
+                print('This PDB is in the length range 85-115.')
+                alpha_carbon_positions = alpha_carbons.positions
+                n_alpha_carbons = len(alpha_carbons)
+                parent_graph = nx.empty_graph(n_alpha_carbons)
+
+                for i in range(n_alpha_carbons):
+                    for j in range(i):
+                        # Get the distance between two adjacent atoms
+                        distance = np.linalg.norm(alpha_carbon_positions[i] - alpha_carbon_positions[j])
+                        if distance < protein_contact_network.threshold:
+                            parent_graph.add_edge(i,j)
[…]

### 2022-01-14 src/main.py

@@ -2,10 +2,15 @@
 Main 
 
 """
-import functions
+import temp_af_ll
 import PCN
 
+import sys
+import warnings
+warnings.filterwarnings("ignore")
 
 if __name__ == '__main__':
+    # args = sys.argv[1:]
 
-    functions.main()
+   # functions.main()
+   temp_af_ll.main()

### 2022-01-14 src/plot_link_lengths.py

@@ -12,7 +12,9 @@ import numpy as np
 import warnings
 warnings.filterwarnings('ignore')
 import matplotlib.pyplot as plt
-
+from sklearn.metrics import r2_score
+import scipy.stats as stats
+import pylab
 def nth_harmonic(n):
     '''
     Compute the nth harmonic number, i.e. the sum of the reciprocals of the first n natural numbers.
@@ -183,7 +185,7 @@ def F_k(f_low, f_high, chain_length, a, k):
     F_k = chain_length * (((1 - f_low)/(1 - f_high))**a) * (1 - f_high)**k
     return F_k
 
-def powerlaw(link_length_array, chain_length):
+def powerlaw(scale_factor, link_length_array, chain_length):
     '''
     Compute the power law relationship: N * (1/x^1.2)
     
@@ -200,7 +202,7 @@ def powerlaw(link_length_array, chain_length):
     pwr: float
         Power law relationship N * (1/x^1.2)
     '''
-    pwr = N * (1/pow(link_length_array, 1.2))
+    pwr =  N * (1/pow(link_length_array * scale_factor, 1.2))
     return pwr
 
 
@@ -264,76 +266,158 @@ def normalise_data(data_array, bin_width):
 
 
 link_length_ranges = ['100', '200', '300']
-constant_A = [0.001, 0.001, 0.001] # check these
+constant_A = [0.005, 0.001, 0.001] # check these
 constant_a = [3, 3, 3] # check these
-#shift_x = [9, 8.5, 8.5] 
-#shift_y = [500, 1000, 1000]
-for ll_index in range(len(link_length_ranges)):
-    
-    # Load PDB data
-    pdb_ids = np.load(f'../data/data_for_plotting/ids_{link_length_ranges[ll_index]}.npy')
-    pdb_mean_data = np.load(f'../data/data_for_plotting/means_{link_length_ranges[ll_index]}.npy')
-    
-    # Get PDB bin widths and centres
-    pdb_edges = np.load(f'../data/data_for_plotting/edges_{link_length_ranges[ll_index]}.npy')[0]
-    pdb_bin_centres = pdb_edges[:-1] + np.diff(pdb_edges)/2
-    pdb_bin_width = pdb_edges[1] - pdb_edges[0]
-    
-    # Normalise PDB data
-    pdb_mean_data = normalise_data(pdb_mean_data, pdb_bin_width)
[…]

### 2022-01-14 src/read_ids.py

@@ -7,8 +7,8 @@ filename_counter = 0
 
 for chunk in pd.read_csv(filename, chunksize=chunksize, sep='\n', header=None):
 
-	np_array = chunk.to_numpy()
-	np.savetxt('../data/ids/id_file_'+str(filename_counter)+'.txt', np_array, fmt='%s')
+    np_array = chunk.to_numpy()
+    np.savetxt('../data/ids/id_file_'+str(filename_counter)+'.txt', np_array, fmt='%s')
 
-	filename_counter += 1
+    filename_counter += 1
 

### 2022-01-14 src/remove_duplicates.py

@@ -13,9 +13,9 @@ all_list = DF['ALL'].astype(str).tolist()
 duplicates = []
 
 for i in range(len(all_list)):
-	print(f'We are at entry: {i}')
-	duplicates.append(down_list[i])
-	duplicates.append(all_list[i])
+    print(f'We are at entry: {i}')
+    duplicates.append(down_list[i])
+    duplicates.append(all_list[i])
 
 
 duplicate_dict = {'duplicates':duplicates}

### 2022-01-14 src/temp_af_ll.py

@@ -1,0 +1,338 @@
+# temp_af_ll.py
+
+"""
+Functions to be used in main.py. 
+
+"""
+import os
+import pandas as pd
+import Bio.PDB.PDBList as biopy
+import PCN
+import traceback
+import tarfile
+import numpy as np
+
+import warnings
+warnings.filterwarnings("ignore")
+
+
+def is_file(filename):
+    """
+    Check if given filename exists. 
+    
+    Parameters
+    ----------
+    filename: str
+        filename given by user
+
+    Return
+    ------
+    checked_filename: str
+        correct filename
+
+    """
+    is_file = os.path.isfile(str(filename))
+    while is_file == False:
+        checked_filename = input(f'Invalid filename. Please enter again: ')
+        is_file = os.path.isfile(str(checked_filename))
+    return checked_filename
+
+def get_uniprot_IDs(sequence_length):
+    """
+    Take the sequence length and return an array of Uniprot IDs in that range.
+    Parameters
+    ----------
+    sequence_length: str
+        sequence length of PDBs, either 100, 200 or 300
+    
+    Return
+    ------
+    uniprot_IDs: numpy.array
+        array containing all the uniprot IDs of the given sequence length
+    """
+    uniprot_IDs_df = pd.read_excel(f'../data/alphafold_data/uniprot_{sequence_length}s.xlsx')
+    uniprot_IDs = uniprot_IDs_df['Entry'].to_numpy()
+
+    return uniprot_IDs
+
+def get_alphafold_PDBs(uniprot_IDs):
+    """
+    Parameters
+    ----------
+    uniprot_IDs: numpy.array
+        uniprot IDs of proteins in a given sequence length range
+    
+    Return
+    ------
+
+    """
+    sequence_length = '100'
+    uniprot_IDs = get_uniprot_IDs(sequence_length)
+    
+
+def bootstrap():
+    """
+    Take the link lengths and bootstrap the data to get means and 33% confidence intervals.
+
+    Saves the mean and lower and upper bound C.L. of the bootstrapped data and the PDB ids belonging to
+    that length range as a csv file.
+
+    Parameters
+    ----------
+    None
+
+    Return
+    ------
+    None
+    """
+    input_file = input('Please enter the input file including the full path to the file: ')
+  #  link_length_file = is_file(link_length_file)
+
+    # Get the link length file and concatenate arrays in the file into one
[…]

### 2022-01-14 src/test_methods.py

@@ -4,45 +4,63 @@ import pandas as pd
 import PCN
 import os
 import MDAnalysis as mda
-PDB_IDs = ['5EUB']
+import functions
+import urllib
 
-for PDB_ID in PDB_IDs:
-	PDB_ID_lower = str(PDB_ID).lower()
-	PDB_file = f'/Volumes/Seagate_Extension_Plus/PDBs/pdb{PDB_ID_lower}.ent'
-	if os.path.isfile(PDB_file):
-		u = mda.Universe(PDB_file)
+sequence_length = '100'
+uniprot_IDs = functions.get_uniprot_IDs(sequence_length)
 
+download = f'https://alphafold.ebi.ac.uk/files/AF-{uniprot_IDs[0]}-F1-model_v2.pdb'
 
-		# for seg_index in range(number_of_segments):
+try:
+    file_name = f'../data/alphafold_data/temp/AF-{uniprot_IDs[0]}-F1-model_v2.pdb'
+    #bio.retrieve_pdb_file(pdb,pdir='../data/tmep/', file_format='pbd')
+    urllib.request.urlretrieve(download, file_name)
+except Exception:
+    print("failed at %s" % pdb, flush=True)
+    traceback.print_exc(file=log)
+ 
+u = mda.Universe(file_name)
+print(u)
 
-		segments = u.universe.residues.segments
-			
-		number_of_segments = len(segments)
+# PDB_IDs = ['5EUB']
 
-		for index in range(number_of_segments):
-			segment_id = u.universe.residues.segments.segids[index]
-
-			try:
-				C_alphas = u.universe.select_atoms(f'name CA and segid {segment_id}')
-				
-				print(len(C_alphas))
-				break
-			
-			except:
-				print('Error in first segment id. Trying the next one.')
-				continue
-
+# for PDB_ID in PDB_IDs:
+#     PDB_ID_lower = str(PDB_ID).lower()
+#     PDB_file = f'/Volumes/Seagate_Extension_Plus/PDBs/pdb{PDB_ID_lower}.ent'
+#     if os.path.isfile(PDB_file):
+#         u = mda.Universe(PDB_file)
 
 
+#         # for seg_index in range(number_of_segments):
 
-		# if segment_id:
-		# 	print(segment_id)
-		# segments = self.universe.residues.segments
-		# number_of_segments = len(segments)
-		# for seg_index in range(number_of_segments):
-		# 	str_seg_ids = str(segments[seg_index])
-			
[…]

### 2022-03-17 code/scripts/adjacency.py
put adjacency matrix plotting code in separate file

@@ -1,0 +1,77 @@
+""" Functions for plotting an adjacency matrix from 3D simulations or PDBs"""
+import matplotlib
+import numpy as np
+from matplotlib import pyplot as plt
+import seaborn as sns
+from sequence_distance_distribution.code.scripts import protein_contact_map, theory_functions
+from sequence_distance_distribution.code.scripts.plot_functions import _COLOUR_PALETTE
+
+
+def get_simulation_matrix(matrix_file: str) -> np.ndarray:
+    """
+    Open a 3D simulation file and return the adjacency matrix
+    @param matrix_file: matrix_<length_range>_<file_no>.txt
+    @return: adjacency matrix as a Numpy array
+    """
+    adjacency_matrix = np.loadtxt(matrix_file)
+    adjacency_matrix[adjacency_matrix > 1] = 0
+    return adjacency_matrix
+
+
+def get_pdb_matrix(pdb_file: str) -> np.ndarray:
+    """
+    Open a PDB file, create a graph and return its adjacency matrix
+    @param pdb_file: PDB file from either RCSB or AlphaFold (default)
+    @return: adjacency matrix as a Numpy array
+    """
+    pcm = protein_contact_map.ProteinContactMap(pdb_file)
+    alpha_carbons = pcm.get_alpha_carbons
+    protein_graph = pcm.get_protein_graph(alpha_carbons)
+    return theory_functions.get_adjacency_matrix(protein_graph)
+
+
+def set_adjacency_matrix_ticks(plot: matplotlib.axes.Axes) -> None:
+    """
+    Format tick labels for adjacency matrix plots
+    @param plot: seaborn heatmap
+    @return:
+    """
+    for index, label in enumerate(plot.get_xticklabels()):
+        if index % 5 == 0:
+            label.set_visible(True)
+            label.set_rotation(360)
+            label.set_font("Helvetica")
+            label.set_fontsize(14)
+        else:
+            label.set_visible(False)
+    for index, label in enumerate(plot.get_yticklabels()):
+        if index % 5 == 0:
+            label.set_visible(True)
+            label.set_font("Helvetica")
+            label.set_fontsize(14)
+        else:
+            label.set_visible(False)
+
+
+def plot_adjacency_matrix(file: str, data_type: str) -> None:
+    """
[…]

### 2022-03-17 code/scripts/protein_contact_map.py

@@ -1,5 +1,5 @@
 """
-Base class for PCNs. 
+Base class for PCMs.
 
 """
 import MDAnalysis

### 2022-03-17 code/scripts/bars.py
put bar plotting code in separate file

@@ -1,0 +1,76 @@
+import matplotlib.pyplot as plt
+import pandas as pd
+import seaborn as sns
+from sequence_distance_distribution.code.scripts.plot_functions import _COLOUR_PALETTE
+import numpy as np
+
+
+def get_data_for_bars(path_to_data: str) -> tuple:
+    """
+    Read csv file from path and return numpy arrays as tuple
+    @param path_to_data:
+    @return:
+    """
+    dataframe = pd.read_csv(path_to_data)
+    bins = dataframe["Bins"].to_numpy()
+    frequencies = dataframe["Number"].to_numpy()
+    return bins, frequencies
+
+
+def calculate_adjusted_frequency(frequency_to_adjust: np.ndarray, bottom_frequency: np.ndarray) -> np.ndarray:
+    """
+
+    @param frequency_to_adjust: "top" bars i.e. either all SwissProt or all RCSB frequencies
+    @param bottom_frequency: "bottom" bars i.e. the used PDBs' frequencies
+    @return: difference between top and bottom bars
+    """
+    return frequency_to_adjust - bottom_frequency
+
+
+def create_bar_plots() -> None:
+    """
+    Bring all stats together and plot subplots of PDB statistics
+    @return: None
+    """
+    path = "../data/pdb_statistics/"
+    alphafold_file = "AlphaFold_used.csv"
+    pdb_file = "ll_used.csv"
+    swissprot_file = "UniProt_Swiss_Prot.csv"
+    rcsb_file = "RCSB_by_length.csv"
+
+    alpha_bins, alpha_frequencies = get_data_for_bars(path + alphafold_file)
+    swiss_bins, swiss_frequencies = get_data_for_bars(path + swissprot_file)
+    pdb_bins, pdb_frequencies = get_data_for_bars(path + pdb_file)
+    bins, rcsb_frequencies = get_data_for_bars(path + rcsb_file)
+
+    adjusted_swiss = calculate_adjusted_frequency(swiss_frequencies, alpha_frequencies)
+    adjusted_rcsb = calculate_adjusted_frequency(rcsb_frequencies, pdb_frequencies)
+
+    fig = plt.figure(figsize=(10, 8))
+    ax = fig.subplots(2, 1, sharex=True)
+
+    ax[0].bar(bins, adjusted_swiss,
+              color=_COLOUR_PALETTE["DATABANK"],
+              label="Swiss-Prot Frequencies",
[…]

### 2022-03-17 code/scripts/plot_bars.py

@@ -1,55 +1,0 @@
-import matplotlib.pyplot as plt
-import pandas as pd
-import seaborn as sns
-
-alphafold_data = pd.read_csv("/Users/jasminguven/Documents/GitHub/sequence_distance_distribution/data/pdb_statistics"
-                             "/AlphaFold_used.csv")
-alpha_bins = alphafold_data["Bins"].to_numpy()
-alpha_frequency = alphafold_data["Number"].to_numpy()
-
-swissprot_data = pd.read_csv("/Users/jasminguven/Documents/GitHub/sequence_distance_distribution/data/pdb_statistics"
-                             "/UniProt_Swiss_Prot.csv")
-swiss_bins = swissprot_data["Bins"].to_numpy()
-swiss_frequency = swissprot_data["Number"].to_numpy()
-swiss_adjusted_frequency = swiss_frequency - alpha_frequency
-
-# trembl_data = pd.read_csv("/Users/jasminguven/Documents/GitHub/sequence_distance_distribution/data/pdb_statistics"
-#                           "/UniProt_TrEMBL_sorted.csv")
-# trembl_bins = trembl_data["Bins"].to_numpy()
-# trembl_frequency = trembl_data["Number"].to_numpy()
-# trembl_adjusted_frequency = trembl_frequency - swiss_frequency - alpha_frequency
-
-ll_data = pd.read_csv("/Users/jasminguven/Documents/GitHub/sequence_distance_distribution/data/pdb_statistics"
-                      "/ll_used.csv")
-ll_bins = ll_data["Bins"].to_numpy()
-ll_frequency = ll_data["Number"].to_numpy()
-
-rcsb_data = pd.read_csv("/Users/jasminguven/Documents/GitHub/sequence_distance_distribution/data/pdb_statistics"
-                        "/RCSB_by_length.csv")
-
-rcsb_bins = rcsb_data["Bins"].to_numpy()
-rcsb_frequency = rcsb_data["Number"].to_numpy() - ll_frequency
-
-sns.set_theme(context="notebook", palette="colorblind", style="ticks", font_scale=1.5, font="Helvetica")
-fig = plt.figure(figsize=(10, 8))
-ax = fig.subplots(2, 1, sharex=True)
[…]

### 2022-03-17 code/scripts/bars.py
put comparison plotting code in separate file

@@ -1,3 +1,4 @@
+"""Functions for plotting bar plots"""
 import matplotlib.pyplot as plt
 import pandas as pd
 import seaborn as sns

### 2022-03-17 code/scripts/comparison.py

@@ -1,0 +1,85 @@
+"""Functions for plotting comparison plot (Fig.4.)"""
+import argparse
+import numpy as np
+import sequence_distance_distribution.code.scripts.theory_functions as theory_functions
+from sequence_distance_distribution.code.scripts.plot_functions import _COLOUR_PALETTE
+import matplotlib.pyplot as plt
+import seaborn as sns
+import pandas as pd
+
+
+def get_comparison_data(path_to_file: str, is_simulation: bool) -> tuple:
+    """
+    Get csv data from path
+    @param is_simulation: boolean to determine if file contains simulation data
+    @param path_to_file: RCSB, AlphaFold or Simulation folder
+    @return: tuple containing data required for comparison plot
+    """
+    dataframe = pd.read_csv(path_to_file)
+    if is_simulation:
+        distances = dataframe["index"].to_numpy()
+    else:
+        distances = dataframe["variable"].to_numpy()
+    means = dataframe["mean"].to_numpy()
+    normalised_means = means / np.sum(means)
+    lower_bound = dataframe["lower_bound"].to_numpy()
+    normalised_lower_bound = lower_bound / np.sum(means)
+    upper_bound = dataframe["upper_bound"].to_numpy()
+    normalised_upper_bound = upper_bound / np.sum(means)
+    return distances, normalised_means, normalised_lower_bound, normalised_upper_bound
+
+
+def create_comparison_plot(arguments: argparse.Namespace) -> None:
+    """
+    Create amino acid distance distribution plot (Fig.4)
+    @param arguments: command line arguments from user
+    @return: None
+    """
+    rcsb_tuple = get_comparison_data(f"../data/rcsb/bootstrap_{arguments.length_range}_stats.csv",
+                                     is_simulation=False)
+    alphafold_tuple = get_comparison_data(f"../data/alphafold/chunk_{arguments.length_range}_stats.csv",
+                                          is_simulation=False)
+    sim_tuple = get_comparison_data(f"../data/simulations/3d/lls_{arguments.length_range}.csv",
+                                    is_simulation=True)
+    rcsb_means = rcsb_tuple[1]
+    alphafold_means = alphafold_tuple[1]
+    rcsb_n_points = len(rcsb_means)
+    alphafold_n_points = len(alphafold_means)
+
+    rcsb_half_n_harmonic = theory_functions.harmonic_number(rcsb_n_points // 2)
[…]

### 2022-03-17 code/scripts/2d_simulation.py
put 2dsim plotting into separate file

@@ -1,0 +1,149 @@
+import glob
+import numpy as np
+import pandas as pd
+import sequence_distance_distribution.code.scripts.theory_functions as theory_functions
+import matplotlib.pyplot as plt
+import seaborn as sns
+from sequence_distance_distribution.code.scripts.plot_functions import _COLOUR_PALETTE
+
+
+def get_2d_sim_files():
+    """
+    Get all the files for 2D simulations
+    @return: list containing all simulation files
+    """
+    return glob.glob(f"../data/simulations/2d/2d_sim_no_circle_*")
+
+
+def get_2d_distance_frequencies(files: list) -> np.ndarray:
+    """
+    Get the frequencies from 2D simulation files
+    @param files: 2D simulation files
+    @return: array containing frequencies for amino acid distances
+    """
+    distances_list = []
+    for file in files:
+        distances = np.loadtxt(file)
+        distances_list.append(distances)
+    return np.asarray(distances_list)
+
+
+def get_2d_simulation_stats() -> pd.DataFrame:
+    """
+    Compute the mean and confidence interval for 2D simulations
+    @return: None
+    """
+    files = get_2d_sim_files()
+    frequencies = get_2d_distance_frequencies(files)
+    simulation_dataframe = pd.DataFrame(frequencies)
+    melted_dataframe = simulation_dataframe.melt()
+    simulation_stats_dataframe = melted_dataframe.groupby("variable",
+                                                          as_index=False).agg(mean=("value", np.mean),
+                                                                              lower_bound=("value", lambda val:
+                                                                              np.quantile(val, q=0.05)),
+                                                                              upper_bound=("value", lambda val:
+                                                                              np.quantile(val, q=0.95)))
+    simulation_stats_dataframe.to_csv(f"../data/simulations/2d/simulation_stats.csv")
+    return simulation_stats_dataframe
+
+
+def get_2d_plotting_data(simulation_dataframe: pd.DataFrame) -> tuple:
+    """
+    Get 2D simulation data from dataframe
+    @param simulation_dataframe: dataframe containing 2D simulation stats
+    @return: tuple containing the mean frequencies of amino acid distances, distances and the CI bounds
+    """
+    mean_frequencies = simulation_dataframe["mean"].to_numpy()[2:]
[…]

### 2022-03-17 code/scripts/2d_simulation.py
simulation grid plots

@@ -1,149 +1,0 @@
-import glob
-import numpy as np
-import pandas as pd
-import sequence_distance_distribution.code.scripts.theory_functions as theory_functions
-import matplotlib.pyplot as plt
-import seaborn as sns
-from sequence_distance_distribution.code.scripts.plot_functions import _COLOUR_PALETTE
-
-
-def get_2d_sim_files():
-    """
-    Get all the files for 2D simulations
-    @return: list containing all simulation files
-    """
-    return glob.glob(f"../data/simulations/2d/2d_sim_no_circle_*")
-
-
-def get_2d_distance_frequencies(files: list) -> np.ndarray:
-    """
-    Get the frequencies from 2D simulation files
-    @param files: 2D simulation files
-    @return: array containing frequencies for amino acid distances
-    """
-    distances_list = []
-    for file in files:
-        distances = np.loadtxt(file)
-        distances_list.append(distances)
-    return np.asarray(distances_list)
-
-
-def get_2d_simulation_stats() -> pd.DataFrame:
-    """
-    Compute the mean and confidence interval for 2D simulations
-    @return: None
-    """
-    files = get_2d_sim_files()
-    frequencies = get_2d_distance_frequencies(files)
-    simulation_dataframe = pd.DataFrame(frequencies)
-    melted_dataframe = simulation_dataframe.melt()
-    simulation_stats_dataframe = melted_dataframe.groupby("variable",
-                                                          as_index=False).agg(mean=("value", np.mean),
-                                                                              lower_bound=("value", lambda val:
-                                                                              np.quantile(val, q=0.05)),
-                                                                              upper_bound=("value", lambda val:
-                                                                              np.quantile(val, q=0.95)))
-    simulation_stats_dataframe.to_csv(f"../data/simulations/2d/simulation_stats.csv")
-    return simulation_stats_dataframe
-
-
-def get_2d_plotting_data(simulation_dataframe: pd.DataFrame) -> tuple:
-    """
-    Get 2D simulation data from dataframe
-    @param simulation_dataframe: dataframe containing 2D simulation stats
-    @return: tuple containing the mean frequencies of amino acid distances, distances and the CI bounds
-    """
-    mean_frequencies = simulation_dataframe["mean"].to_numpy()[2:]
[…]

### 2022-03-17 code/scripts/adjacency.py
split plot_functions to separate files

@@ -3,8 +3,9 @@ import matplotlib
 import numpy as np
 from matplotlib import pyplot as plt
 import seaborn as sns
-from sequence_distance_distribution.code.scripts import protein_contact_map, theory_functions
-from sequence_distance_distribution.code.scripts.plot_functions import _COLOUR_PALETTE
+import protein_contact_map
+import theory_functions
+from plot_functions import _COLOUR_PALETTE
 
 
 def get_simulation_matrix(matrix_file: str) -> np.ndarray:

### 2022-03-17 code/scripts/bars.py

@@ -2,7 +2,7 @@
 import matplotlib.pyplot as plt
 import pandas as pd
 import seaborn as sns
-from sequence_distance_distribution.code.scripts.plot_functions import _COLOUR_PALETTE
+from plot_functions import _COLOUR_PALETTE
 import numpy as np
 
 

### 2022-03-17 code/scripts/colour_palette.py

@@ -1,0 +1,3 @@
+class ColourPalette:
+    def __init__(self):
+        pass

### 2022-03-17 code/scripts/comparison.py

@@ -1,8 +1,8 @@
 """Functions for plotting comparison plot (Fig.4.)"""
 import argparse
 import numpy as np
-import sequence_distance_distribution.code.scripts.theory_functions as theory_functions
-from sequence_distance_distribution.code.scripts.plot_functions import _COLOUR_PALETTE
+import theory_functions
+from plot_functions import _COLOUR_PALETTE
 import matplotlib.pyplot as plt
 import seaborn as sns
 import pandas as pd

### 2022-03-17 code/scripts/functions.py

@@ -6,7 +6,7 @@ import traceback
 
 import networkx as nx
 
-import pcm
+import protein_contact_map
 import numpy as np
 import os
 import pandas as pd

### 2022-03-17 code/scripts/plot.py

@@ -3,6 +3,11 @@ Handle all plotting functions and actually make the plots
 """
 import pandas as pd
 import plot_functions
+import adjacency
+import bars
+import comparison
+import simulation
+import distances
 
 
 def plot() -> None:
@@ -14,26 +19,32 @@ def plot() -> None:
     length_range = arguments.length_range
     algorithm = arguments.algorithm
 
-    if algorithm == "BOTH":
-        plot_functions.create_comparison_plot(arguments)
-    elif algorithm == "B":
-        plot_functions.create_bar_plots()
-    elif algorithm == "BS" or algorithm == "C":
+    if algorithm == "comp":
+        comparison.create_comparison_plot(arguments)
+    elif algorithm == "bar":
+        bars.create_bar_plots()
+    elif algorithm == "rcsb" or algorithm == "alpha":
         dimensionality_start = arguments.start_dimensionality
         dimensionality_end = arguments.end_dimensionality
         exponent_start = arguments.start_exponent
         exponent_end = arguments.end_exponent
-        pdb_dataframe = plot_functions.get_dataframe(arguments)
-        sim_dataframe = pd.read_csv(f"../data/simulations/3d/lls_{length_range}.csv")
-
+        pdb_dataframe = distances.get_dataframe(arguments)
+        sim_dataframe = pd.read_csv(f"../data/simulations/3d/simulation_{length_range}_stats.csv")
         if dimensionality_start == dimensionality_end and exponent_start == exponent_end:
-            pass
+            distances.create_plots(arguments, exponent_start, dimensionality_start, pdb_dataframe, sim_dataframe)
         elif dimensionality_start != dimensionality_end and exponent_start != exponent_end:
-            plot_functions.create_grid_plots(arguments, pdb_dataframe, sim_dataframe)
-    elif algorithm == "A":
-        plot_functions.plot_adjacency_matrix(arguments.file, arguments.data_type)
-    elif algorithm == "2D-SIM":
-        pass
+            distances.create_grid_plots(arguments, pdb_dataframe, sim_dataframe)
+    elif algorithm == "adj":
+        adjacency.plot_adjacency_matrix(arguments.file, arguments.data_type)
+    elif algorithm == "2d-sim":
+        dimensionality_start = arguments.start_dimensionality
+        dimensionality_end = arguments.end_dimensionality
+        exponent_start = arguments.start_exponent
+        exponent_end = arguments.end_exponent
[…]

### 2022-03-17 code/scripts/plot_functions.py

@@ -8,14 +8,16 @@ import matplotlib.axes
 import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
-import pcm
+import protein_contact_map
 import theory_functions
 import seaborn as sns
 
 _COLOUR_PALETTE = {"PDB_SCATTER": "#006374",
-                   "SIM_SCATTER": "#fbafe4",
+                   "2D_SIM_SCATTER": "#006374",
+                   "3D_SIM_SCATTER": "#fbafe4",
                    "ALPHA_SCATTER": "#1ac938",
                    "THEORY": "#006374",
+                   "CL": "#fbafe4",
                    "RESIDUALS": "#fbafe4",
                    "DATABANK": "#006374",
                    "USED": "#fbafe4",
@@ -23,109 +25,7 @@ _COLOUR_PALETTE = {"PDB_SCATTER": "#006374",
                    "NO_CONTACT": "#006374"}
 
 
-def create_plot_label(length_range: str, algorithm: str):
-    """
-    Create custom labels for Alphafold PDB data and RCSB PDB data with the given link length
-    @param algorithm: data source from RCSB (bootstrapped) or AlphaFold (chunked)
-    @param length_range: chain length range
-    @return:
-    """
-    plot_label = " "
-    if algorithm == "BS" or algorithm == "BOTH":
-        plot_label = f"PDB {length_range}s"
-    elif algorithm == "C":
-        plot_label = f"AF PDB {length_range}"
-    return plot_label
-
-
-def grid_plot_distances(dimensionality_range: np.ndarray, exponent_range: np.ndarray, n_points: int,
-                        half_n_harmonic: float, plotting_sumrange: np.ndarray, normalised_means: np.ndarray,
-                        distances: np.ndarray, normalised_sim_means: np.ndarray, sim_distances: np.ndarray,
-                        lower_cl: np.ndarray, upper_cl: np.ndarray, length_range: str, algorithm: str) -> None:
-    """
-    Plot amino acid distances and theory plot with different values of exponent and dimensionality scaling constant
-    @param algorithm: data source from RCSB (bootstrapped) or AlphaFold (chunked)
-    @param length_range: chain length range
-    @param sim_distances: amino acid distances for simulated data (x-axis)
-    @param upper_cl: upper confidence level
-    @param lower_cl: lower confidence level
-    @param normalised_sim_means: normalised simulated means of amino acid distance frequencies
[…]

### 2022-03-21 paper/main.tex
shameful commit dump: WIP alphafold stuff
updating with latest info from overleaf

--- 2021-12-02 paper/main.tex
@@
 \usepackage[colorinlistoftodos]{todonotes}
+\newcommand{\Pin}{P_{\mathrm{in}}}
+\newcommand{\fhigh}{f_{\mathrm{high}}}
+\newcommand{\flow}{f_{\mathrm{low}}}
 
@@
 \begin{abstract}
-The mechanisms by which the folded structures of proteins are determined have long been one of the key mysteries of Biophysics.
-Here we approach this problem from a network ensemble perspective, analyzing the relative frequencies with which residues at different distances along the chain come into contact.
-We define and analyze the residue distance distribution for measured protein structures and structures resulting from a geometrically constrained random interaction model and find that the distributions are compatible.
-We furthermore derive an analytic approximation for the two dimensional analog of the model and find this approximation to also describe the same distribution.
-We therefore propose the conclusion that the approximate power-law behaviour of the residue distance distribution is largely caused by geometric constraints of chains rather than specifics of protein sequences.
+The mechanisms by which we can determine a protein's 3D structure based on the protein's amino acid sequence have long been one of the key mysteries of biophysics.
+With the advent of AlphaFold2, making predictions for protein structures based on amino acid sequences at an accuracy similar to experiments is now possible, yet requires a lot of available structural data for training the model.
+Predicting contact maps from sequence \textit{de novo} is an appealing alternative.
+Here, we are interested in understanding the emergent behavior of protein contact maps derived from a network ensemble perspective of a geometrically constrained random interaction model.
+We show that the frequency distributions of amino acid distances obtained from simulations of the geometrically constrained interaction model can be described by an analytical approximation of the two-dimensional analogue of this model.
+To this end, we look at the scaling behavior of the simulation with the analytical approximation for different chain lengths typically found in proteins with good agreement.
+Furthermore, we compare our constrained model with frequency distributions of amino acid distances from contact maps of real proteins---from the protein data bank and AlphaFold2 database---showing that these follow a similar trend as the geometrically constrained model.
+ We therefore suggest the conclusion that the approximate power-law behavior of the amino acid distance distribution can be attributed to geometric constraints of protein chains and amino acid sequences only play a secondary role.
 %The distribution of chain separations of interacting amino acids in proteins roughly follows a power law.
-Here we show, analytically and in simulations, that a geometrical stochastic model of folding chains can explain this behaviour.
+Here we show, analytically and in simulations, that a geometrical stochastic model of folding chains can explain this behavior.
 \end{abstract}
@@
 
-\section*{Introduction}
+\section{Introduction}
 Proteins, the molecular machines of every living organism perform vital tasks required for life to persist, ranging from transport (e.g.
-hemoglobin), signal transduction (e.g.
-rhodopsin), immune responses (e.g.
+hemoglobin)~\cite{ahmed2020hemoglobin}, signal transduction (e.g.
+rhodopsin)~\cite{nagata2021rhodopsins}, immune responses (e.g.
 antibodies), and hormonal regulation (e.g.
-insulin)~\cite{dill2008protein,Dill1042} \red{(Actually in the last paper we focused more on modelling of proteins rather than why they are so important so i didn't find much)}.
-All natural proteins are made up of 20 different amino acids which dictate the three dimensional conformations the proteins adopt in order to function~\cite{scheraga2007protein}.
-One way of looking at the functional forms of proteins and classifying their structural properties is by looking at protein residue or contact networks~\cite{Vendruscolo2002,DiPaola2013,Estrada2011}.
-Protein Contact Networks (PCN) have shown promise in understanding protein folding patterns as well as revealing allosteric communication pathways~\cite{yao2019establishing, menichetti2016network,dokholyan2002topological} \red{(i found some papers that use PCN's but not necessarily for allosteric communication)}.
-From a physicist's perspective it is interesting to understand the emergent behaviour in protein structure from PCNs, by investigating common network measures and their behaviour for PCNs.
-For example Bartoli et al.~\cite{bartoli2008effect} build a model in which they assume that proteins have a residue distance distribution that follows $\frac{1}{s}$, where $s$ is the separation of two connected amino acids along the backbone chain (in the following called \emph{residue distance}) as the simplest implementation of a connection probability, thus decreasing with the distance along the sequence.
+insulin)~\cite{dill2008protein,Dill1042}.
+All natural proteins are made of 20 different amino acids which dictate the three-dimensional conformations the proteins adopt in order to function~\cite{scheraga2007protein}.
+One of the great challenges has been to understand how the primary structure, the amino acid sequence, can lead to the fully folded functional protein.
+The last 50 years has seen many different routes to computationally predict biologically active protein structures without having to solve a crystal structure of the protein.
+[input information from Rohan's presentation].
+AlphaFold2, a machine learning model for structure prediction, has shown predictiveness at a level of experimental accuracy in the most recent protein structure prediction challenge~\cite{jumper2021highly, kryshtafovych2021critical}.
+This has lead to an explosion of available structural protein models beyond the protein database, containing structures from crystallographic, NMR, and cryo-EM experiments~\cite{berman2000protein}.
+Beyond the prediction of structure understanding emergent patterns in proteins on a more fundamental level is of interest, as it opens up a deeper understanding of proteins structure and function relationship.
+One way of looking at the functional forms of proteins and classifying their structural properties is by looking at protein contact maps (PCM) or protein contact networks (PCN)~\cite{Vendruscolo2002,DiPaola2013,Estrada2011}.
+A PCM---we will use PCMs in favor of PCN in this paper---is the resulting adjacency matrix of spatially close amino acids, given a certain cut-off as determined typically from a protein structure.
+PCMs have shown promise in understanding protein folding patterns as well as revealing allosteric communication pathways~\cite{yao2019establishing, menichetti2016network,dokholyan2002topological}.
+Their use is also prominently featured in machine learning approaches.
+One challenge is to predict a PCM \textit{de novo} from sequence alone~\cite{bassot2019using, rives2021biological, rao2021msa}, without using structural data as part of the training set, as was done for AlphaFold2.
+These PCMs can then be used in e.g.
+graph convolutional machine learning approaches to train a model able to predict binding affinities for certain types of proteins~\cite{jiang2020drug},
+From a physicist's perspective it is interesting to understand the emergent behavior in protein structure by looking at PCMs.
+Since a PCM is an adjacency matrix and therefore a graph, it is possible to investigate PCMs using common network measures.
+For example, Bartoli \textit{et al.}~\cite{bartoli2008effecta} build a model in which they assume that proteins have an amino acid distance distribution that follows $\frac{1}{s}$, where $s$ is the separation of two connected amino acids along the backbone chain (in the following called \emph{amino acid distance}) as the simplest implementation of a connection probability, thus decreasing with the distance along the sequence.
 However, they do not provide a quantitative explanation for this assumption.
 
-Here, we show how a simple geometrical model can reproduce this heuristically found behaviour well, and even provide a simple analytical approximation.
-The three-dimensional geometrical model and the analytical approximation of its two-dimensional analogue both build on the idea that inherently geometrical objects, such as amino acids, imply that any resulting PCN network ensemble modelling them has to be spatially embedded~\cite{molkenthin2016scaling, molkenthin2020self}.
+Here, we show how a simple geometrical model can reproduce this heuristically found behavior well, and even provide a simple analytical approximation.
+The three-dimensional geometrical model and the analytical approximation of its two-dimensional analogue both build on the idea that inherently geometrical objects, such as amino acids, imply that any resulting PCM network ensemble modelling them has to be spatially embedded~\cite{molkenthin2016scaling, molkenthin2020self}.
 Different approaches have been used in the past for geometrical models describing protein folding, some of which derived characteristics of the secondary structure from constraints on bond and torsion angles~\cite{bhattacharjee2013flory,Danielsson2010,Molkenthin2011} and others focused on the formation mechanisms of the tertiary structure~\cite{molkenthin2016scaling, molkenthin2020self}.
 
-Based on such considerations, the network formation process repeatedly connects random pairs of units, without causing any intersections along the chain ~\cite{molkenthin2016scaling}.
-Such a model can now be used to investigate the emergent behaviour of residue distance distributions as used heuristically by Bartoli et al.~\cite{bartoli2008effect}.
-To this end, we define and analyze the residue distance distributions $P(s)$ that emerge from the geometric model, where the \emph{residue distance} $s$ of a link is defined as the distance of the two end units along the original chain.
-This distribution is then compared to the residue distance distribution extracted from protein contact networks generated from protein data bank data.
+Based on such considerations, the network formation process repeatedly connects random pairs of units, without causing any intersections along the chain~\cite{molkenthin2016scaling}.
+Such a model can now be used to investigate the emergent behavior of amino acid distance distributions as used heuristically by Bartoli \textit{et al.}~\cite{bartoli2008effecta}.
+To this end, we define and analyze the amino acid distance distributions $P(s)$ that emerge from the geometric model, where the \emph{amino acid distance} $s$ of a link is defined as the distance of the two end units along the original chain.
 
-We find that, despite the highly simplified model, the generated residue distance distribution is in agreement with the one observed in networks generated from PDB data and roughly consistent with the heuristic $1/s$ finding.
+This distribution is then compared to the amino acid distance distribution extracted from protein contact networks generated from protein data bank data.
+
+We find that, despite the highly simplified model, the generated amino acid distance distribution is in agreement with the one observed in networks generated from PDB data and roughly consistent with the heuristic $1/s$ finding.
 The analytical approximation provides a correction term on top of the $1/s$ heuristic.
 
-\section*{Protein Residue Networks and their synthetic counterparts}
-The complex interaction pattern between the amino acids in a protein can be very naturally expressed as a network, in which each amino acid is represented by a node and spatial proximity is encoded as a link.
-Whenever two central $C_\alpha$ atoms are closer together than a threshold $d_c$, they are connected.
-These connections are encoded in the adjacency matrix of the folded protein, which is the binary matrix
+%%%%%%%%%%%%%%%%%
+% Section       %
+%%%%%%%%%%%%%%%%%
+\section{Protein contact maps}
+The complex interaction pattern between the amino acids in a protein can be very naturally expressed as a network or graph, in which each amino acid is represented by a node and spatial proximity is encoded as a link.
+Whenever two central $C_\alpha$ atoms are closer together than a threshold $d_c$, they are connected, linked, or in `close contact'.
+These connections can be determined from the 3D functional or folded structure of the protein and are encoded in a so called contact map.
+This contact map is effectively an adjacency matrix $\textbf{A}^{\mathrm{struc}}$ or binary matrix
 \begin{equation}
-  A^{\textsf{PDB}}_{ij}=
+  A^{\mathrm{struc}}_{ij}=
   \begin{cases}
@@
 \end{equation}
-Adjacency matrices are commonly used to describe networks and in this case represent the PCN.
+Adjacency matrices are commonly used to describe networks and make up a \textit{protein contact map}. 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.8\columnwidth]{figures/Fig1/Fig1.pdf}
+    \caption{a) The amino acid distance $s$ is defined as the separation along the chain (teal) of two connected units (pink).
+b)~The solid green link prohibits some connections from node~$A$ (dotted pink), while others are still permitted (dotted green).
+Counting all newly excluded links, we find that a link of length $s$ prohibits $2(l-1)$ links of length $l$ up to a maximum of $s$ and $2(s-1)$ links of length $s$ or more.
+c)~Example of a protein contact map taken from a 2D simulation.
+Pink squares indicate a contact is present between two amino acids.}
+    \label{fig:schematic}
+\end{figure}
 
-The underlying formation mechanisms of individual protein folds are incredibly complex and depend on the surrounding solvent as well as the specific amino acid sequence and their quantum-mechanical interactions.
-Here we want to study the ensemble of all folded proteins, however, and will thus "average" over such specifics.
+%%%%%%%%%%%%
+% subsection
+%%%%%%%%%%%%
+\subsection{Contact maps in a geometrically constrained protein model}
 
-Based on the basic observation that amino acids are objects in space which can not overlap indefinitely, we introduced a model in \cite{molkenthin2020self} starting from a chain of identical spheres, each in contact only with the neighbours it is connected to.
+Based on the basic observation that amino acids are objects in space which cannot overlap indefinitely, we introduced a model in~\cite{molkenthin2020self} starting from a chain of identical spheres, each in contact only with the neighbours it is connected to.
 Then additional connections form by randomly selecting two spheres and moving them towards each other until they touch, the new links formed that way can not be broken in later steps and function as constraints on subsequent links.
 If contact of the two selected spheres is geometrically impossible without breaking previously made connections or leading to overlaps of any spheres, the link is not made and taken out of the pool of possible connections.
-This process is repeated until no more links can be formed without violating the geometric constraints.
+This process is repeated until no more links can be formed without violating the geometric constraints and the final structure can be expressed as the adjacency matrix
+\begin{equation}
+  \mathbf{A}^{\mathrm{sim}}_{ij}=
+  \begin{cases}
+   1, & \text{ if }|i-j|=1 \text{ or i and j  randomly connected} \\
+      0, & \text{ otherwise }  .
+      \end{cases}
+    \label{eq:aij}
+\end{equation}
 
-In \cite{molkenthin2016scaling}, we have introduced a simplified, two-dimensional version of this model, which starts from a closed chain (i.e.
+In~\cite{molkenthin2016scaling}, we have introduced a simplified, two-dimensional version of this model, which starts from a closed chain (i.e.
 a ring) of $N$ unit discs and subsequently adds links, such that connected discs touch, yet no discs intersect.
 The advantage of this model is that it can be approximated by a purely topological simplification, which can be treated analytically.
+Fig.~\ref{fig:schematic} illustrates the construction of the 2D geometric model.
 
@@
 
-\section*{Analytic approximation}
-Here we want to analytically approximate the residue distance distribution for the simplified 2D model, that is the separation along the backbone chain of two nodes connected by a link, as illustrated in Fig~\ref{fig:schematic}~a).
+%%%%Missing%%%%%
+% TO DO: @Nora
+% Describe how the protein contact maps are generated in the geometric constrained model %
+% Let's call the adjacency matrix in this case \mathbf{A}^{\mathrm{sim}}_{ij} where sim stands for simulation. 
+%%%%%%%%%%%%%%%
 
-\begin{figure}[h]
+\subsection{Protein contact maps from structural protein data}
+
+The underlying formation mechanisms of individual protein folds are incredibly complex and depend on the surrounding solvent, as well as the specific amino acid sequence and their quantum-mechanical interactions.
+Here we want to assess if the geometrically constrained protein model's contact map distributions capture real protein behavior as suggested in~\cite{bartoli2008effecta}.
+To understand this better, we want to study the ensemble of PCMs from real protein structures.
+To this end we will look at `averaged' contact maps over many different proteins that have different amino acid sequences, but their overall sequence or chain length is the same.
+We will use the RCSB protein data bank (PDB), a database of structurally resolved protein amino acid sequences through X-ray crystallography, NMR or cryo-EM experiments and the synthetically generated structural data base from the machine learned structures of AlphaFold2~\cite{jumper2021highly}.
+The PDB contains just under 200,000 protein structures to date~\cite{PDB}.
+The distribution of amino acid chain lengths in the PDB is illustrated in Fig.~\ref{fig:pdb_stats} b.
+Protein chain lengths in the PDB vary from small fragments of less than 10 amino acids to large agglomerates of over 2000 amino acids.
+The most frequently occurring chain lengths, however, are between 100 and 300 amino acids as shown in Fig.~\ref{fig:pdb_stats} b.
+The pink subset of the barplot shown in Fig.~\ref{fig:pdb_stats} b, are the set of structures used that fulfill the criterion on amino acid chain lengths between 85 and 115 amino acids, 185 and 215 amino acids and 285 and 315 amino acids and represent all structures for the analysis.
+The second database consists of synthetically generated structures using AlphaFold2 with sequences taken from SwissProt~\cite{alpha fold 2 database}.
+SwissProt is a manually curated database of protein sequences containing over 500,000 protein sequences~\cite{SwissProt}, whose amino acid chain length distribution is shown in Fig.~\ref{fig:pdb_stats} a in blue-green.
+The pink bars again indicate the subset of AlphaFold2 structures used for the contact map analysis.
+
+ \begin{figure}[t]
         \centering
-    \includegraphics[width=0.7\columnwidth]{figures/Fig1/Fig1.pdf}
-    \caption{a) The residue distance $s$ is defined as the separation along the chain (blue) of two connected units (red).
-b) The solid blue link prohibits some connections from node $A$ (dotted red), while others are still permitted (dotted blue).
-Counting all newly excluded links, we find that a link of length $s$ prohibits $2(l-1)$ links of length $l$ up to a maximum of $s$ and $2(s-1)$ links of length $s$ or more.}
-    \label{fig:schematic}
+	\includegraphics[width=0.8\columnwidth]{paper/figures/Fig2/Fig2.pdf}
+	    \caption{Histogram of protein sequence lengths for protein structures from two structure databases.
+a) AlphaFold2 predicted protein structures ($\alpha$F) taken from all protein sequences in the reviewed Uniprot Knowledge Base (Swiss-Prot) database based on their Uniprot ids.
+Pink bars shows fraction of structures used for the average contact map analysis.
+b) RSCB PDB chain length frequencies based on sequence length.
+Pink bars in plot show proteins structures whose sequence lengths matches the resolution of structure accounting for the whole sequence.
+Only pink PDB samples were used for the average contact map analysis.
+c) Example of a protein contact map taken from PDB ID: 1AKI, with a cut-off of 8~Å.
+Pink squares indicate a contact present between amino acids.}
+        \label{fig:pdb_stats}
 \end{figure}
+All structure based contact maps are generated in the same way for AlphaFold2 structures and PDB structures.
+Structures used were downloaded in pdb format, read in with MDAnalysis version 2.0.0~\cite{gowers2016mdanalysis} and only C$_{\alpha}$ were selected and used for the analysis.
+Proteins were categorised by length and placed into three groups.
+Amino acid chain lengths between 85 and 115 amino acids, 185-215 amino acids and 285-315.
+For each of the used sequences, all pairwise C$_{\alpha}$ distances were assessed and two C$_{\alpha}$ atoms were in contact if their distance was less than 8~Å.
+All pairwise distances were evaluated and in this way, building up the binary contact map or adjacency $A_{ij}^{\mathrm{struc}}$.
+All code for the construction of the contact maps both for the geometrically constrained system and structural protein data and their analysis can be found on github at: \url{https://github.com/meyresearch/sequence_distance_distribution}.
 
-Let us start by introducing the auxiliary variable $F_k(s)$, defined to be the number of possible links with a residue distance of $s$ if $k$ links have been added before.
-Before any links are added all residue distances are equally likely, as there are $F_0(s)=N$ possibilities of making a link of each separation $s$ with $2\leq s < N/2$.
+%%%%%%%%%%%%
+% Section. %
+%%%%%%%%%%%%
+\section{Analytic approximation for the geometrically constrained model}
 
-As we add more links, not only are links taken out of this pool, because they have already been realized, each existing link can also geometrically prohibit other connections.
+To assess how we can best describe the distribution of the probability of amino acid distances as obtained by a PCM for the geometrically constrained model, we propose the following analytical approximation.
+To be able to do this we will look at the amino acid distance distribution for the simplified 2D model.
+Fig~\ref{fig:schematic} summarizes the idea behind this two-dimensional model.
+
+Let us start by introducing the auxiliary variable $F_k(s)$, defined to be the number of possible links with an amino acid distance of $s$ if $k$ links have been added before.
+Before any links are added, all amino acid distances are equally likely, as there are $F_0(s)=N$ possibilities of making a link of each distance $s$ with $2\leq s < N/2$.
+
+As we add more links, not only are links taken out of this pool, because they have already been realized, \red{but} each existing link can also geometrically prohibit other connections.
 Adding a link of length $s_1$ prohibits $2(s-1)$ links of length $s<s_1$ and $2(s_1-1)$ links of length $s>s_1$ (see Fig.~\ref{fig:schematic}~b).
-To find the expected number of links still available in subsequent stops, we average over all possible residue distances $s_1$ of the first link added.
+To find the expected number of links still available in subsequent steps, we average over all possible amino acid distances $s_1$ of the first link added.
 
-The expected distribution of possible links after one step is given thus by the average over available link pools taken over all possible lengths of the initial link
+The expected distribution of possible links after one step is \red{thus given} by the average over available link pools taken over all possible values of $s_1$
 \begin{align}
@@
 
-In subsequent steps we take the weighted average over new links being added at step $k-1$, where the probability that a link of length $s$ can still be made is
+In subsequent steps, we take the weighted average over new links being added at step $k-1$, where the probability that a link amino acid distance $s$ can still be made is
 \begin{equation}
-    P_{in}^k(s)=F_k(s)/C_k,
+    \Pin^k(s)=F_k(s)/C_k,
 \end{equation}
-where $C_k=\sum_{s=2}^{N/2}F_k(s)$ and get the same reduction (compare Fig.~\ref{fig:schematic}) but starting from the pool $F_{k-1}(s)$ of the step before rather than the full $N$ links for each length.
-To account for possible overlaps between newly removed links and links that were removed previously, we multiply by
+where $C_k=\sum_{s=2}^{N/2}F_k(s)$ and get the same reduction (compare Fig.~\ref{fig:schematic}) but starting from the pool $F_{k-1}(s)$ of the step before rather than the full $N$ links for each amino acid distance.
+To account for possible overlaps between newly removed links and links that were removed previously, we multiply the number of blocked links by
 \begin{equation}
-    \frac{F_k(s)}{F_0(s)} \approx F_k(s)\frac{1}{N}.
+    \frac{F_k(s)}{F_0(s)} = F_k(s)\frac{1}{N}.
     \label{eq.Pk}
@@
 %with $C_k=\sum_{s=2}^{N/2}F_k(s)$, 
-This ignores any bias in which links are removed, however, in small, manageable examples this assumption seemed to hold well. 
+This ignores any bias in which links are removed, however, examples suggest this assumption seemed to hold well. 
 This leads to the recursion
 \begin{align}
-    F_k(s)&= \sum_{s_k=2}^{N/2} P_{in}^k(s_k)\nonumber \\
-    & (F_{k-1}(s)-  F_k(s)\frac{1}{N}
+    F_k(s)&= \sum_{s_k=2}^{N/2} \Pin^k(s_k)\nonumber \\
+    & \left(F_{k-1}(s)-  F_{k-1}(s)\frac{1}{N}
     {\begin{cases}
@@
      2(s_k -1))\text{ , for } s\geq s_k
-    \end{cases}} )
+    \end{cases}} \right)
 \end{align}
@@
 \begin{align}
-   F_k(s)&=F_{k-1}(s)(1-\frac{1}{N} \sum_{s_k=2}^{s} (2 (s_k-1) P_{in}^k(s_k))\nonumber \\
-   &-\frac{1}{N} \sum_{s_k=s+1}^{N/2} (2 (s-1) P_{in}^k(s_k)))\nonumber \\
+   F_k(s)&=F_{k-1}(s)(1-\frac{1}{N} \sum_{s_k=2}^{s} (2 (s_k-1) \Pin^k(s_k))\nonumber \\
+   &-\frac{1}{N} \sum_{s_k=s+1}^{N/2} (2 (s-1) \Pin^k(s_k)))\nonumber \\
    &=F_{k-1}(s)(1-\frac{1}{N}
-   \sum_{s_k=2}^{s} (2 (s_k-1) P_{in}^k(s_k))\nonumber \\
-   &-\frac{2 (s-1)}{N} (\sum_{s_k=2}^{N/2} P_{in}^k(s_k) - \sum_{s_k=2}^{s} P_{in}^k(s_k)))\nonumber \\
+   \sum_{s_k=2}^{s} (2 (s_k-1) \Pin^k(s_k))\nonumber \\
+   &-\frac{2 (s-1)}{N} (\sum_{s_k=2}^{N/2} \Pin^k(s_k) - \sum_{s_k=2}^{s} \Pin^k(s_k)))\nonumber \\
    &=F_{k-1}(s)(1-\frac{2}{N} (s-1) 
    +\frac{2}{N} s
-   \sum_{s_k=2}^{s}P_{in}^k(s_k)\nonumber \\
+   \sum_{s_k=2}^{s}\Pin^k(s_k)\nonumber \\
    &-\frac{2}{N}
-   \sum_{s_k=2}^{s} s_k P_{in}^k(s_k))
+   \sum_{s_k=2}^{s} s_k \Pin^k(s_k))
    \label{eq.Fk_rec}
@@
 
-To approximately evaluate this we make a guess for $P_{in}^k(s)$ at low $k$ and at high $k$, namely
+To approximately evaluate this we make a guess for $\Pin^k(s)$ at low $k$ and at high $k$, namely
 \begin{equation}
-    P_{in}^{k\ll N/2}(s)\approx P_{in}^{k=0}(s)=2/N.
+    \Pin^{k\ll N/2}(s)\approx \Pin^{k=0}(s)=\frac{2}{N}.
     \label{eq.P_small}
 \end{equation}
-For later steps $k$, the probability of a link to still be in the pool decreases with $s$, we thus make the ansatz, that it drops off as $1/s$, leading to
+For later steps $k$, the probability of a link to still be in the pool decreases with $s$, 
+we thus approximate $\Pin(s)$ with an ansatz for the $k-$average, namely, that it drops off as $\frac{1}{s}$, leading to
 \begin{equation}
-    P_{in}^{k\gg 1}(s)\approx \frac{s^{-1}}{ (H_{\frac{N}{2}}-1)},
+    \Pin^{k\gg 1}(s)\approx \sum_{k=a}^{N-4}\Pin^{k}(s)
+    \approx \frac{s^{-1}}{ (H_{\frac{N}{2}}-1)},
     \label{eq.P_large}
@@
 where $H_{\frac{N}{2}}$ is the harmonic number. 
-We suspect that the approximation for low $k$ only holds for a couple of steps while Eq.~\ref{eq.P_large} holds, at least on average for larger $k$.
-We have no approximation for the intermediate steps but assume that the errors roughly cancel each other out.
-Inserting Eq.\ref{eq.P_small} and Eq.\ref{eq.P_large} into Eq.\ref{eq.Fk_rec} results in the following two recursions for the early and late evolution of the sequence distance distribution respectively.
+As we will see later this approximation holds most precisely for intermediate values of $k$, with the $k-$average of $\Pin^{k\gg 1}(s)\approx P(s)$, making it self-consistent.
+Errors for smaller and larger values of $k$ are thought to approximately cancel out.
+Substituting Eq.~\ref{eq.P_small} and Eq.~\ref{eq.P_large} into Eq.~\ref{eq.Fk_rec} results in the following two recursions for the early and late evolution of the amino acid distance distribution respectively.
 \begin{align}
@@
    &-\frac{2}{N^2}(s^2+s-2)\nonumber \\
-   &=F_{k-1}(s)(1-f_{low}(s)),
+   &=F_{k-1}(s)(1-\flow(s)),
    \label{eq.Fk_rec_low}
@@
 where 
-\[f_{low}(s)=-\frac{2}{N^2}s^2+(\frac{2}{N}+\frac{6}{N^2})s-(\frac{2}{N}+\frac{4}{N^2}),\]
+\[\flow(s)=-\frac{2}{N^2}s^2+\left(\frac{2}{N}+\frac{6}{N^2}\right)s-\left(\frac{2}{N}+\frac{4}{N^2}\right),\]
 and
+
 \begin{align}
@@
    F_{k-1}(s)(1-\frac{2}{N} (s-1) 
-   +\frac{2}{N} s \sum_{s_k=2}^{s}\frac{1}{s_k(H_{\frac{N}{2}}-1)}\nonumber \\
+   +\frac{2}{N} s \sum_{s_k=2}^{s}\frac{1}{s_k(H_{\frac{N}{2}}-1)}\nonumber\\
    &-\frac{2}{N}
@@
    &+ \frac{2}{N(H_{\frac{N}{2}}-1)} + \frac{2}{N})\nonumber \\
-   &=F_{k-1}(s)(1-f_{high}(s)).
+   &=F_{k-1}(s)(1-\fhigh(s)).
    \label{eq.Fk_rec_high}
@@
 where 
-\[f_{high}(s)=\frac{2}{N}(\frac{H_{\frac{N}{2}}-H_{s}+1}{H_{\frac{N}{2}}-1}s-\frac{H_{\frac{N}{2}}}{H_{\frac{N}{2}}-1}).\]
+\[\fhigh(s)=\frac{2}{N}(\frac{H_{\frac{N}{2}}-H_{s}+1}{H_{\frac{N}{2}}-1}s-\frac{H_{\frac{N}{2}}}{H_{\frac{N}{2}}-1}).\]
 
@@
 \begin{align}
-    F_k(s)&=N\prod_{i_1}^a\left(1-f_{low}(s)\right)\prod_{i=a+1}^k\left(1-f_{high}(s)\right) \nonumber \\
-    &= N\left(\frac{1-f_{low}(s)}{1-f_{high}(s)}\right)^a\left(1-f_{high}(s)\right)^{k}
+    F_k(s)&=N\prod_{i_1}^a\left(1-\flow(s)\right)\prod_{i=a+1}^k\left(1-\fhigh(s)\right) \nonumber \\
+    &= N\left(\frac{1-\flow(s)}{1-\fhigh(s)}\right)^a\left(1-\fhigh(s)\right)^{k}
 \end{align}
@@
 In 2D this is repeated for $N-3$ steps until no more links can be added \cite{molkenthin2016scaling}.
-The probability distribution of the realized residue distances of all added links is then given by the average over the available pools at each link addition step, leading to
+The probability distribution of the realized amino acid distances of all added links is then given by the average over the available pools at each link addition step, leading to
 \begin{align}
     P(s)&=\frac{1}{N-3}\sum_{k=0}^{N-4} \frac{F_k(s)}{C_k} \nonumber \\
-    &= \frac{N}{N-3}\left(\frac{1-f_{low}(s)}{1-f_{high}(s)}\right)^a
-\sum_{k=0}^{N-4} \frac{1}{C_k} \left(1-f_{high}(s)\right)^{k} \nonumber \\
-    &\approx \left(\frac{1-f_{low}(s)}{1-f_{high}(s)}\right)^a\frac{A}{f_{high}(s)},
+    &\approx \frac{N}{N-3}\left(\frac{1-\flow(s)}{1-\fhigh(s)}\right)^a
+\sum_{k=0}^{N-4} \frac{1}{\evdel{C_k}_k} \left(1-\fhigh(s)\right)^{k} \nonumber \\
+    &\approx \left(\frac{1-\flow(s)}{1-\fhigh(s)}\right)^a\frac{A}{\fhigh(s)},
     \label{eq.Ps}
 \end{align}
-
-\red{The assumption that the sum is basically a harmonic series seems like a bit of a stretch I shall check it again but the results look good}
-
-for $1<s<\frac{N}{2}$.
-The approximation sets all $C_k=C$ to allow the use of the geometric series for solving the equation.
-We can do this because $C_k$ monotonically decreases with $k$ and $N>C_k>C_{N-2}$ meaning we can obtain an upper and lower bound by replacing all $C_k$ in the expression with either $N$ (underestimates all probabilities) or $C_{N-2}$ (overestimates all probabilities), however, as neither of those depend on $s$, these bounds are merely in the constant factor $A$ and do not affect the scaling as such.
+where $\evdel{C_k}_k$ is the average of $C_k$ over $k$.
+This approximation allows the use of the geometric series for solving the equation. 
+The resulting expression in Eq.~\ref{eq.Ps} is close to a power-law in $s$ with an exponent of \red{something close to -1.2 depending on a}, thus justifying the earlier approximation of $\Pin^k(s)$.
 %which makes use of Eq.~\ref{eq.Pk}. 
@@
 % where $A$ is an unknown constant, containing an approximation of the various normalization constants $C_i$, as well as the direct effects of the chain length.
-Since for long chains and $s \ll N/2$, $f(s)$ is dominated by the linear term, the distribution is well approximated by a power law with an exponent of $-1$.
 
@@
 
-%\begin{equation}
- %   P_{corrected}(s)\approx A f(s)^{-1+\alpha},
-%\end{equation}
-%where $0<\alpha \ll 1$.
-\blue{The parameters $A$ and $\alpha$ in this approximation are treated as free parameters and used for fitting the theoretical result to data and simulations.
-While they could be estimated by evaluating the middle line of Eq.~\ref{eq.Ps}, this estimate is for the 2D model, in which the exclusion of connections is faster, leading to a steeper decrease of $C_k$, as well as a different factor $A$.}
-
-\red{We could add a figure with simulated 2D results (which i should have...
-somewhere) and the exact evaluation with the sums and all that.
-I'm not super motivated to do that but i guess it might be a nice addition and the only place for the exact formula really.}
-
-\section*{Data and simulations}
-For large values of the chain length $N$ we thus find an approximate power law probability distribution for the residue distance with an exponent of $-1+\alpha$.
-Here we use simulations, as introduced in~\cite{molkenthin2020self}, and networks extracted from measured protein data from the PDB \cite{PDB} to compare this behaviour with the more realistic 3D settings and show that the basic principle persists, indicating that much of the residue distance distribution can be explained by generic geometric constraints.
- \begin{figure}[h]
+\section{Results}
+\subsection{Analytic approximation and 2D simulations}
+Fig.~\ref{fig:2d_sim} compares Eq.~\ref{eq.Ps} with respect to the two-dimensional simulation data introduced in~\cite{molkenthin2020self}.
+The simulation data is generated from 10 repeats with chain lengths of up to 498 amino acids long.
+The simulation data is plotted showing the mean frequencies of each amino acid distance up to $N/2\sim250$.
+The shaded area represent the 95\% confidence interval calcualted from the 10 repeats.
+The theoretical approximation from Eq.~\ref{eq.Ps} was fitted to the simulation data.
+The two fitting parameters were chosen to be $a=6$ and $A=0.39$, to best capture the trend of the simulation data.
+The fitted approximation captures the behavior of the simulation data well, both describe two dimensional systems.
+ \begin{figure}[tb]
         \centering
-	\includegraphics[width=0.45\textwidth]{paper/figures/Fig2/pdb_statistics.pdf}
-        \caption{Statistics on all PDB entries based on the number of amino acid residues, with most entries found in the interval of 100-400 residues as of May 26th 2020.}
-        \label{fig:pdb_stats}
+	\includegraphics[width=0.45\textwidth]{paper/figures/Fig3/Fig3.pdf}
+	    \caption{Amino acid distance distribution for two-dimensional simulated data plotted against the analytic approximation. Green area shows the 95\% confidence interval.}
+        \label{fig:2d_sim}
 \end{figure}
 
-Protein chain lengths in the PDB vary from small fragments of less than 10 amino acids to large agglomerates of over 2000 residues.
-The most frequently occurring chain lengths, however, are between 100 and 400 residues as shown in Fig.~\ref{fig:pdb_stats}.
-\red{For our analysis, we have thus chosen ............
-structures of around 200, ........
-structures of around 300 and ..........
-structures of around 400 residues from the PDB.} Those have been converted into PCN's based on the measured locations of their residues to analyze the distribution of residue distances between connected amino acids.
-The resulting distribution is compared to those generated in simulations in~\cite{molkenthin2020self} as well as the above derived theoretical expression scaled with a factor to account for the different dimensionality, as well as approximations made in the derivation.
+\subsection{3D simulations and PCMs from structural data }
+Next, we investigate if and how the analytical approximation can also capture trends in 3-D simulation data, as well as experimentally determined protein structures.
+This seeks to answer the question if indeed a power law can be used to describe the probability distribution of amino acid distances.
+Fig.~\ref{fig:sdd} summarises the results for protein chain lengths of $\approx$100 (a), $\approx$200, and $\approx$ 300 amino acids, protein chain lengths, most commonly found in living organism (see.
+Fig.~\ref{fig:pdb_stats} a).
+We are comparing data from repeats of 30 simulation runs of the three-dimensional simulations, as introduced in~\cite{molkenthin2020self}, and protein contact maps from RCSB~\cite{}, and AlphaFold2~\cite{}.
 
-The results are shown logarithmically scaled in Fig.\ref{fig:sdd}.
-\red{For all three lengths, the distributions show an approximate power law decay with slopes of $...\pm ...$, which is consistent with both, the simulated results ans the theoretical approximation.}
+We have chosen a subset of structures from the PDB and curated alphaFold2 database for the analysis.
+7,376 structures with chain lengths that fall between 85-100, 6,822 structures that fall between 185-215, and 4,160 structures that fall between 285-315 amino acids were selected from the RCSB database~\cite{PDB}.
+Additionally, 33,229 structures with chain lengths in the range of 85-115, 37,430 in the range of 185-215 and 32,137 in the range of 285-315 amino acids were chosen from the AlphaFold2 database~\cite{}.
+All structures were converted into PCMs as described in section~\ref{sec:methods}.
+The resulting distributions for each chain length interval are plotted logarithmically scaled in Fig.\ref{fig:sdd} for intervals 85-100 in a, 185-215 in b, and 285-315 in c. 
+For all three lengths, the distributions show an approximate power law decay, for both the 3-D simulations according to~\cite{molkenthin2020self}, as well as the structural data.
+This allows a fit  which is consistent with both the simulated results and the theoretical approximation.
 
-\begin{figure}[t]
+is compared to those generated in simulations in~\cite{molkenthin2020self} as well as the above derived theoretical expression scaled with a factor to account for the different dimensionality, i.e.
+constant $A$, as well as approximations made in the derivation.
+
+\begin{figure}
         \centering
-	\includegraphics[width=\columnwidth]{paper/figures/Fig3/Fig3.pdf}
-        \caption{residue distance distributions for real and simulated folds for a) Chain lengths around 200 b) chain lengths around 300 c) chain lengths around 400.
-\red{Depending on what this looks like when we have all the data, perhaps adding error bars could be nice.
-Also, since i'm mentioning the overrepresentation of separations 2 to 4, maybe those should be shown.
-If it fits better we can take them out for the fits and stuff.}
+	\includegraphics[scale=0.65]{paper/figures/Fig4/Fig4.pdf}
+        \caption{Amino acid distance distributions for contact maps derived from the RCSB~\cite{PDB} and AlphaFold2~\cite{} databases as well as 3D simulated folds for protein chains in the range of a) 85-115,  b) 185-215 and c) 285-315 amino acids long.
         }
@@
 \end{figure}
+For large values of the chain length $N$, we thus find an approximate power law probability distribution for the amino acid distance with an exponent of $-1+\alpha$.
 
-We observe that, while the general shape seems to match, in real proteins residue distances of 2 to 4 amino acids are slightly over-represented (see Fig.\ref{fig:sdd}).
-The over-representation of very short residue distances is an artefact of the different network construction for simulated (and theoretically approximated) networks and PCNs.
-While in the simulation all links are equally long and links are made until no longer possible, the parameter, determining the connectivity in PCNs is the cutoff distance $d_c$\blue{(mention used value?)}, which is chosen such that the number of links in the PCN is comparable to those in a simulated network of the same $N$.
+We observe that, while the general shape seems to match, in real proteins amino acid distances of two to four amino acids are slightly over-represented (see Fig.\ref{fig:sdd}).
+The over-representation of very short amino acid distances is an artefact of the different network construction for simulated (and theoretically approximated) networks and PCMs.
+While in the simulation all links are equally long and links are made until no longer possible, the parameter, determining the connectivity in PCMs is the cutoff distance $d_c$\blue{(mention used value?)}, which is chosen such that the number of links in the PCM is comparable to those in a simulated network of the same $N$.
 This leads to a larger range of bond angles randomly resulting in connections for small $s$ and thus an over-representation of those links.
 
-\section*{Conclusion}
+Here we use 30 runs of three-dimensional simulations, as introduced in~\cite{molkenthin2020self}, and networks extracted from measured protein data from both the RCSB~\cite{PDB} and AlphaFold2~\cite{} databases to compare this behavior with the more realistic 3D settings and show that the basic principle persists, indicating that much of the amino acid distance distribution can be attributed to generic geometric constraints.
+\section{Conclusion}
 The tertiary structures of folded proteins have long been one of the most important mysteries of Biophysics.
@@
 
-Here we have introduced and analyzed the residue distance $s$ and its probability distribution $P(s)$ in measured and simulated protein structures.
-For the equivalent model in two dimensions we have used a mean-field approach to derive an analytic approximation for the residue distance distribution, which appears to agree with the simulated and measured distributions.
-We have thus demonstrated in analytical calculations as well as direct simulations how a random linking model with geometric constraints, such as the ones introduced in \cite{molkenthin2016scaling, molkenthin2020self}, generates a residue distance distribution that resembles a power-law, similar to the one found in folded proteins.
-This feature of protein network structures has previously been used in a heuristic manner to model and generate protein-like structures in \cite{bartoli2008effect}.
+Here we have introduced and analyzed the amino acid distance $s$ and its probability distribution $P(s)$ in measured and simulated protein structures.
+For the equivalent model in two dimensions we have used a mean-field approach to derive an analytic approximation for the amino acid distance distribution, which appears to agree with the simulated and measured distributions.
+We have thus demonstrated in analytical calculations as well as direct simulations how a random linking model with geometric constraints, such as the ones introduced in \cite{molkenthin2016scaling, molkenthin2020self}, generates a amino acid distance distribution that resembles a power-law, similar to the one found in folded proteins.
+This feature of protein network structures has previously been used in a heuristic manner to model and generate protein-like structures in \cite{bartoli2008effecta}.
 
@@
 Together with an understanding of secondary structure principles, such as those in
-\cite{Danielsson2010, Molkenthin2011}, this can help to narrow down the complex energy landscapes and find paths through them more effectively in the future.
+~\cite{Danielsson2010, Molkenthin2011}, this can help to narrow down the complex energy landscapes and find paths through them more effectively in the future.
 

### 2022-03-21 paper/supplementary_material.tex
updating with latest info from overleaf

%\documentclass{article}
%\usepackage[a4paper, total={6in, 8in}]{geometry}
%\documentclass[reprint,amsmath,amssymb,rmp,onecolumn,notitlepage,11pt]{revtex4-1}
\documentclass[notitlepage,
reprint,%twocolumn,
%preprint
onecolumn,
amsmath,amssymb,superscriptaddress,aps,
pre]{revtex4-1}
\usepackage[utf8]{inputenc}
%\usepackage{authblk}
%\usepackage{natbib}
\usepackage[normalem]{ulem}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\newcommand{\red}[1]{\textcolor{red!80!black}{#1}}
\newcommand{\blue}[1]{\textcolor{blue!80!black}{#1}}
\newcommand{\green}[1]{\textcolor{green!70!black}{#1}}
\newcommand{\gray}[1]{\textcolor{gray!80!black}{#1}}
\usepackage{mathtools}
\DeclarePairedDelimiter{\evdel}{\langle}{\rangle}
\usepackage[colorinlistoftodos]{todonotes}
\newcommand{\Pin}{P_{\mathrm{in}}}
\newcommand{\fhigh}{f_{\mathrm{high}}}
\newcommand{\flow}{f_{\mathrm{low}}}
\setlength {\marginparwidth }{2cm}

\begin{document}
\title{Supplementary material accompanying "Theory of chain separation distributions gives insights in real protein contact networks"}
\author{Nora Molkenthin}
\affiliation{Potsdam Institut für Klimafolgenforschung}
\affiliation{Center for Advancing Electronics Dresden (cfaed), Technical University of Dresden}
\author{J. J.  Güven}
\affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
\author{Steffen Mühle}
\affiliation{Uni Göttingen, 37077 Göttingen, Germany}
\author{Antonia S J S Mey}
\email[Electronic Address: ]{antonia.mey@ed.ac.uk}
\affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
\maketitle

\section*{Optimizing theory parameters}

 Fig.~\ref{fig:2d} shows a grid of the two-dimensional simulations for amino acid distance distributions for chains of around 498 amino acids long. The solid line shows the theoretical distribution given in Eq.~(12) of the main article. Each column corresponds to a different value of the scaling parameter $A$ and each row to a different value of the exponent parameter $a$ for this theoretical distribution. Plotted also is the 95\% confidence level for the 2D amino acid distance distribution.

In addition to plotting this amino acid distance distribution grid, we also plotted the residual plots for each combination of parameters. These are shown in Fig.~\ref{fig:2d_r}. The residual sum of squares $\Sigma \sigma$, $R^2$ statistic and mean of residuals, $\bar{\sigma}$, were calculated for each combination of parameters $a$ and $A$.

The choice of parameters was first narrowed down by looking at how well the theoretical distribution appeared to fit the amino acid distance distribution of the simulations in Fig.~\ref{fig:2d}, as well as the spread of residuals around the theoretical line in Fig.~\ref{fig:2d_r}. From Fig.~\ref{fig:2d}, we chose a smaller group of plots where the theoretical distribution fitted through a majority of simulation data points. Similarly, from Fig.~\ref{fig:2d_r}, we chose a group of plots where the majority of residuals were randomly scattered around zero. The non-linear parts of the residuals both at the low and high ends of the amino acid distances were expected from Fig.~\ref{fig:2d}. In the next step, we found the optimal parameters by finding the best combination of statistics, i.e. a minimized residual sum of squares, $\Sigma \sigma^2$, maximized constant of determination, $R^2$ and the mean of residuals, $\bar{\sigma} \approx 0$. This was done to avoid relying on a single statistic to determine the optimal parameters. For this 2D simulation data, we chose $A=0.39$ and $a=6$, which produced the following statistics: $\Sigma \sigma^2 = 17.4\times10^3$, $R^2 = 90.0$~\% and $\bar{\sigma} = -0.265$.

The above steps were repeated for both the AlphaFold 2 and RCSB PDBs, for each chain length range. Tables~\ref{table:af}~and~\ref{table:rcsb} show the optimized parameters and their statistics for PDBs obtained from AlphaFold 2 and RCSB, respectively. Figs.~\ref{fig:AF100},~\ref{fig:AF200}~and~\ref{fig:AF300} show the parameter grid plots of amino acid distance distributions for chains of 100, 200 and 300 amino acids long, respectively. Similarly, Figs.~\ref{fig:AF100_r},~\ref{fig:AF200_r}~and~\ref{fig:AF300_r} show the residual plots. While the 95\% confidence intervals are plotted for the PDBs, these were not plotted in the main article for clarity, as the confidence level appears smaller than the scatter markers.

\begin{table}[tb]
\centering
\begin{tabular}{|c c c c c c|}
\hline
Chain length  & $A$ ($10^{-3}$) & $a$ & $\Sigma \sigma^{2}$ & $R^2$ (\%) & $\bar{\sigma}$ \\
\hline
100 & 2.1 & 3 & 0.0002 & 57 & 0.0001 \\
200 & 1.1 & 3 & 0.0135 & 62 & -0.0005 \\
300 &1.0 & 5 & 0.1478 & 76  & -0.0012 \\
\hline
\end{tabular}
 \caption{Detailed parameters and statistics for PDBs obtained from AlphaFold 2. Shown are the optimized parameters $A$ and $a$ used to fit the theoretical amino acid distance distribution given in Eq.(12) of the main article, for chains of 100, 200 and 300 amino acids long.  Shown also are the residual sum of squares ($\Sigma \sigma^{2}$), constant of determination ($R^2$) and the mean of residuals ($\bar{\sigma}$) for each fit.}
\label{table:af}
\end{table}

\begin{table}[h]
\centering
\begin{tabular}{|c c c c c c|}
\hline
Chain length & $A$ ($10^{-3}$) & $a$ & $\Sigma \sigma^{2}$ & $R^2$ (\%) & $\bar{\sigma}$ \\
\hline
100 & 1.9 & 2 & 0.0002 & 59 & 0.0001 \\
200 & 1.2 & 3 & 0.0474 & 70 & -0.0010\\
300 & 0.9 & 4 & 0.1102 & 76 & -0.0011\\
\hline
\end{tabular}
 \caption{Detailed parameters and statistics for PDBs obtained from RCSB. Shown are the optimized parameters $A$ and $a$ used to fit the theoretical amino acid distance distribution given in Eq.(12) of the main article, for chains of 100, 200 and 300 amino acids long. Shown also are the residual sum of squares ($\Sigma \sigma^{2}$), constant of determination ($R^2$) and the mean of residuals ($\bar{\sigma}$) for each fit.}
\label{table:rcsb}
\end{table}

\begin{figure*}[htb]
    \centering{
    \includegraphics[scale=0.65]{paper/figures/SM_figures/Fig1/Fig1.pdf}
    \caption{a) A grid of the amino acid distance distribution in protein chains of 498 residues long in two-dimensional simulations. The solid line shows the theoretical distribution for different values of parameters $a$ and $A$. b) The residual plots of the amino acid distance distribution around the theoretical distribution. The residual sum of squares, $\Sigma \sigma$, $R^2$ statistic and the mean of residuals, $\bar{\sigma}$, were calculated for each combination of parameters.}
    \label{fig:2d}}
\end{figure*}

\begin{figure*}[htb]
    \centering{
    \includegraphics[width=0.8\textwidth]{paper/figures/SM_figures/Fig2/Fig2.pdf}
    \caption{a) A grid of the amino acid distance distributions in chains of around 100 residues long in PDBs obtained from AlphaFold 2. The solid line shows the theoretical distribution for different values of parameters $a$ and $A$. b) The residual plots of the amino acid distance distribution around the theoretical distribution. The residual sum of squares, $\Sigma \sigma$, $R^2$ statistic and the mean of residuals, $\bar{\sigma}$, were calculated for each combination of parameters.}
    \label{fig:AF100}}

\end{figure*}

\begin{figure*}[htb]
    \centering{
    \includegraphics[width=0.8\textwidth]{paper/figures/SM_figures/Fig3/Fig3.pdf}
    \caption{a) A grid of the amino acid distance distributions in chains of around 200 residues long in PDBs obtained from AlphaFold 2. The solid line shows the theoretical distribution for different values of parameters $a$ and $A$. b) The residual plots of the amino acid distance distributions around the theoretical distribution. The residual sum of squares, $\Sigma \sigma$, $R^2$ statistic and the mean of residuals, $\bar{\sigma}$, were calculated for each combination of parameters.}
    \label{fig:AF200}}
\end{figure*}

\begin{figure*}[htb]
    \centering{
    \includegraphics[width=0.8\textwidth]{paper/figures/SM_figures/Fig4/Fig4.pdf}
    \caption{a) A grid of the distribution of link lengths in chains of around 300 residues long in PDBs obtained from AlphaFold 2. The solid line shows the theoretical distribution for different values of parameters $a$ and $A$. b) The residual plots of the amino acid distance distributions around the theoretical distribution. The residual sum of squares, $\Sigma \sigma$, $R^2$ statistic and the mean of residuals, $\bar{\sigma}$, were calculated for each combination of parameters.}
    \label{fig:AF300}}

\end{figure*}

\begin{figure*}[htb]
    \centering{
    \includegraphics[width=0.8\textwidth]{paper/figures/SM_figures/Fig5/Fig5.pdf}
    \caption{a) A grid of the distribution of link lengths in chains of around 100 residues long in PDBs obtained from RCSB. The solid line shows the theoretical distribution for different values of parameters $a$ and $A$. b) The residual plots of the amino acid distance distributions around the theoretical distribution. The residual sum of squares, $\Sigma \sigma$, $R^2$ statistic and the mean of residuals, $\bar{\sigma}$, were calculated for each combination of parameters.}
    \label{fig:PDB100}}
\end{figure*}

\begin{figure*}[htb]
    \centering{
    \includegraphics[width=0.8\textwidth]{paper/figures/SM_figures/Fig6/Fig6.pdf}
    \caption{A grid of the distribution of link lengths in chains of around 200 residues long in PDBs obtained from RCSB. The solid line shows the theoretical distribution for different values of parameters $a$ and $A$. b) The residual plots of the amino acid distance distributions around the theoretical distribution. The residual sum of squares, $\Sigma \sigma$, $R^2$ statistic and the mean of residuals, $\bar{\sigma}$, were calculated for each combination of parameters.}
    \label{fig:PDB200}}

\end{figure*}

\begin{figure*}[htb]
    \centering{
    \includegraphics[width=0.8\textwidth]{paper/figures/SM_figures/Fig7/Fig7.pdf}
    \caption{A grid of the distribution of link lengths in chains of around 300 residues long in PDBs obtained from RCSB. The solid line shows the theoretical distribution for different values of parameters $a$ and $A$. b) The residual plots of the amino acid distance distributions around the theoretical distribution. The residual sum of squares, $\Sigma \sigma$, $R^2$ statistic and the mean of residuals, $\bar{\sigma}$, were calculated for each combination of parameters.}
    \label{fig:PDB300}}
\end{figure*}

\bibliographystyle{unsrt}
% \bibliography{proteins}

\end{document}

### 2022-03-21 paper/main.tex
testing if this works.

@@
 For the equivalent model in two dimensions we have used a mean-field approach to derive an analytic approximation for the amino acid distance distribution, which appears to agree with the simulated and measured distributions.
-We have thus demonstrated in analytical calculations as well as direct simulations how a random linking model with geometric constraints, such as the ones introduced in \cite{molkenthin2016scaling, molkenthin2020self}, generates a amino acid distance distribution that resembles a power-law, similar to the one found in folded proteins.
+We have thus demonstrated in analytical calculations as well as direct simulations how a random linking model with geometric constraints, such as the ones introduced in~\cite{molkenthin2016scaling, molkenthin2020self}, generates a amino acid distance distribution that resembles a power-law, similar to the one found in folded proteins.
 This feature of protein network structures has previously been used in a heuristic manner to model and generate protein-like structures in \cite{bartoli2008effecta}.

### 2022-04-05 paper/main.tex
Updates from Overleaf

@@
 \author{Steffen Mühle}
-\affiliation{Uni Göttingen, 37077 Göttingen, Germany}
+\affiliation{Max-Planck Institute for Dynamics and Self-Organization, 37077 Göttingen, Germany}
 \author{Antonia S J S Mey}
@@
 The expected distribution of possible links after one step is \red{thus given} by the average over available link pools taken over all possible values of $s_1$
-\begin{align}
+\begin{equation}
+\begin{aligned}
     F_1(s)&=\frac{1}{N/2-1} \sum_{s_1=2}^{N/2}{ \begin{cases}
@@
     \end{cases}}
-\end{align}
+\end{aligned}
+    \label{eq:reduction_first_step}
+\end{equation}
 
-In subsequent steps, we take the weighted average over new links being added at step $k-1$, where the probability that a link amino acid distance $s$ can still be made is
+After $k-1$ links were made, the probability that the $k$-th link, being randomly chosen from the pool of available links, has length $s$ obeys
 \begin{equation}
@@
 \end{equation}
-where $C_k=\sum_{s=2}^{N/2}F_k(s)$ and get the same reduction (compare Fig.~\ref{fig:schematic}) but starting from the pool $F_{k-1}(s)$ of the step before rather than the full $N$ links for each amino acid distance.
-To account for possible overlaps between newly removed links and links that were removed previously, we multiply the number of blocked links by
+where $C_k=\sum_{s=2}^{N/2}F_k(s)$.
+In subsequent steps, we use this probability to perform a weighted average over new links being added, and get the same reduction (compare Fig.~\ref{fig:schematic}) but starting from the pool $F_{k-1}(s)$ of the step before rather than the initially available $N$ links.
+However, we must subtract fewer links to account for some of them having left the pool in earlier steps.
+To this end, we multiply the number of blocked links by
 \begin{equation}
-    \frac{F_k(s)}{F_0(s)} = F_k(s)\frac{1}{N}.
+    \frac{F_{k-1}(s)}{F_0(s)} = F_{k-1}(s)\frac{1}{N}.
     \label{eq.Pk}
@@
 %with $C_k=\sum_{s=2}^{N/2}F_k(s)$, 
-This ignores any bias in which links are removed, however, examples suggest this assumption seemed to hold well. 
+This ignores any length-dependent bias in which links have been removed, however, examples suggest this assumption seemed to hold well. 
 This leads to the recursion
@@
 \end{align}
-which simplifies to
+which generalizes equation~\ref{eq:reduction_first_step} and simplifies to
 \begin{align}
@@
 
-To approximately evaluate this we make a guess for $\Pin^k(s)$ at low $k$ and at high $k$, namely
+Together with the initial condition $F_0(s)=N$, this expression constitutes an iteration rule from which $F_k(s)$ can be found for all $k$ and $s$.
+Aiming to decouple those dynamics for different $s$, we make a heuristic guess for $\Pin^k(s_k)$.
+It consists of two separate approximation which are used in steps $k<a$ and steps $k\geq a$ respectively, where the threshold $1\leq a\leq N/2$ is a free parameter.
+In the early stage we use a uniform probability
 \begin{equation}
@@
 \end{equation}
-For later steps $k$, the probability of a link to still be in the pool decreases with $s$, 
-we thus approximate $\Pin(s)$ with an ansatz for the $k-$average, namely, that it drops off as $\frac{1}{s}$, leading to
+For later steps, the probability of a link to still be in the pool decreases with $s$.
+We thus approximate $\Pin(s)$ with an ansatz for the $k-$average, namely, that it drops off as $\frac{1}{s}$, leading to
 \begin{equation}
@@
 \end{equation}
-where $H_{\frac{N}{2}}$ is the harmonic number. 
+where $H_{\frac{N}{2}}$ is the harmonic number serving as a normalization factor. 
 As we will see later this approximation holds most precisely for intermediate values of $k$, with the $k-$average of $\Pin^{k\gg 1}(s)\approx P(s)$, making it self-consistent.
@@
 
-We can now use the recursive expressions to write down a closed expression for $F_k(s)$
+We can now use the recursive expressions to write down a closed expression for $F_k(s)$, using $\flow$ for the first $a$ steps and $\fhigh$ for the rest:
 

### 2022-04-14 scripts/adjacency.py
changing size of plots

@@ -53,8 +53,8 @@ def plot_adjacency_matrix(file: str, data_type: str) -> None:
         adjacency_matrix = functions.get_3d_simulation_adjacency_matrix(file)
     elif data_type == "pdb":
         adjacency_matrix = get_pdb_matrix(file)
-    plt.figure(figsize=(8, 8))
-    sns.set(context="notebook", palette="colorblind", style="ticks", font_scale=2.88, font="Helvetica")
+    plt.figure(figsize=(6, 6))
+    sns.set(context="notebook", palette="colorblind", style="ticks", font_scale=1.8, font="Helvetica")
     colormap = [_COLOUR_PALETTE["NO_CONTACT"], _COLOUR_PALETTE["CONTACT"]]
     heatmap = sns.heatmap(adjacency_matrix, cmap=colormap, cbar=False)
     # heatmap.set_xticks()

### 2022-04-14 scripts/comparison.py

@@ -62,8 +62,8 @@ def create_comparison_plot(arguments: argparse.Namespace, rcsb_histogram: np.nda
     ks_condition = theory_functions.ks_critical_condition(len(rcsb_measure), len(alpha_measure), c_alpha_multiplier)
     is_accepted = theory_functions.accept_null_hypothesis(ks_statistics.statistic, ks_condition)
     fig = plt.figure()
-    fig.set_size_inches((8, 8))
-    sns.set(context="notebook", palette="colorblind", style="ticks", font_scale=2.88, font="Helvetica")
+    fig.set_size_inches((6, 6))
+    sns.set(context="notebook", palette="colorblind", style="ticks", font_scale=1.88, font="Helvetica")
 
     plt.scatter(rcsb_distances, rcsb_measure, label=f"RCSB {arguments.length_range}",
                 color=_COLOUR_PALETTE["PDB_SCATTER"], marker="o")
@@ -94,10 +94,10 @@ def create_comparison_plot(arguments: argparse.Namespace, rcsb_histogram: np.nda
     plt.yscale("log")
     plt.xscale("log")
     plt.xlim(4, arguments.end_point)
-    plt.ylim(0.001, 0.1)
+    plt.ylim(0.0005, 0.1)
     plt.xlabel("s")
     plt.ylabel("P(s)")
-    plt.legend(fontsize=18)
+    plt.legend(loc="upper right", fontsize=16)
     sns.despine()
     plt.tight_layout()
     plt.show()

### 2022-04-14 scripts/contour.py

@@ -25,7 +25,7 @@ def contour_cloud(alphas, betas, colour, plot_alpha) -> matplotlib.contour.Conto
     values = np.vstack([alphas, betas])
     kernel = scipy.stats.gaussian_kde(values)
     z = np.reshape(kernel(positions).T, xx.shape)
-    contour = plt.contour(xx, yy, z, colors=colour, alpha=plot_alpha)
+    contour = plt.contour(xx, yy, z, colors=colour, alpha=plot_alpha, linewidths=2)
     return contour
 
 
@@ -38,7 +38,7 @@ def create_contour_plots() -> None:
     indices = [50, 50, 70]
     colours = ["#682860", "#286830", "#fbafe4", '#006374']
     for i in range(len(lengths)):
-        sns.set(context="notebook", style="ticks", font_scale=2.88, font="Helvetica")
+        sns.set(context="notebook", style="ticks", font_scale=1.88, font="Helvetica")
         sns.set_palette(sns.color_palette(colours))
         alphafold_df = pd.read_csv(f"../data/alphafold/unique_secondary_structures_{lengths[i]}.csv")
         rcsb_df = pd.read_csv(f"../data/rcsb/secondary_structures_{lengths[i]}.csv")
@@ -54,17 +54,13 @@ def create_contour_plots() -> None:
                     color=colours[1], alpha=0.3)
         fig = plt.gcf()
         ax = plt.gca()
-        fig.set_size_inches((8, 8))
-        rcsb_contour = contour_cloud(alphas=rcsb_df["alpha"], betas=rcsb_df["beta"],
-                                     colour=colours[0], plot_alpha=0.6)
-        alphafold_contour = contour_cloud(alphas=alphafold_df["alpha"], betas=alphafold_df["beta"],
-                                          colour=colours[1], plot_alpha=0.7)
-        # rcsb_contour.collections[0].set_label(f"RCSB {lengths[i]}")
-        # alphafold_contour.collections[0].set_label(f"AlphaFold {lengths[i]}")
-        ax.legend([lines.Line2D([0], [0], marker="", c=colours[0]),
-                   lines.Line2D([0], [0], marker="", c=colours[1])],
+        fig.set_size_inches((6, 6))
+        contour_cloud(alphas=rcsb_df["alpha"], betas=rcsb_df["beta"], colour=colours[0], plot_alpha=0.6)
[…]

### 2022-04-14 scripts/distances.py

@@ -37,10 +37,10 @@ def create_plots(arguments: argparse.Namespace) -> None:
     if arguments.algorithm == "rcsb":
         colours = ["#e2b6dc", "#b847a9", "#682860"]
     elif arguments.algorithm == "alpha":
-        colours = ["#a3dbab", "#47b856", "#286830"]
+        colours = ["#a3dbab", "#47b856", "#15371a"]
     markers = ["o", "s", "^"]
-    plt.figure(figsize=(8, 8))
-    sns.set(context="notebook", palette="colorblind", style="ticks", font_scale=2.88, font="Helvetica")
+    plt.figure(figsize=(6,6))
+    sns.set(context="notebook", palette="colorblind", style="ticks", font_scale=1.8, font="Helvetica")
     for i in range(len(length_ranges)):
         histogram = plot_functions.get_histogram(length_ranges[i], arguments.algorithm)
         plotting_tuple = plot_functions.get_data_for_plotting(histogram, arguments, length_ranges[i])
@@ -54,7 +54,7 @@ def create_plots(arguments: argparse.Namespace) -> None:
         plt.ylabel("P(s)")
         plt.xlim(2, 315)
         plt.ylim(0.0001, 0.1)
-        plt.legend(loc="upper right", fontsize=24, frameon=False)
+        plt.legend(loc="upper right", fontsize=20, frameon=False)
         sns.despine()
         plt.tight_layout()
     plt.show()

### 2022-04-14 scripts/simulation.py

@@ -217,11 +217,11 @@ def plot_3d_distances(distances: np.ndarray, measure: np.ndarray,
     sns.set(context="notebook", palette="colorblind", style="ticks", font_scale=2.88, font="Helvetica")
     plt.scatter(distances, measure, label=f"3D Simulation {arguments.length_range}",
                 color=_COLOUR_PALETTE["3D_SIM_SCATTER"], marker="o")
-    plt.fill_between(distances, upper_confidence, lower_confidence,
-                     color=_COLOUR_PALETTE["CL"], alpha=0.4,
-                     label=r"3D Simulation %i$\sigma$ C.L." % arguments.quantile, zorder=-99)
     plt.plot(distances[start_point:end_point], theory, label="Theory", color=_COLOUR_PALETTE["3D_SIM_SCATTER"],
              lw=1.5, ls="--")
+    plt.fill_between(distances, upper_confidence, lower_confidence,
+                     color=_COLOUR_PALETTE["CL"], alpha=0.4,
+                     label="95% C.L.", zorder=-99)
     print("\n")
     print("----------------Parameters----------------")
     print(f"Chain length: {parameters[0]} +/- {sigma[0]}")

### 2022-04-14 paper/main.tex
Updates from Overleaf

@@
 \author{Antonia S J S Mey}
-\email[Electronic Address: ]{antonia.mey@ed.ac.uk}
+\email[Author to whom correspondence should be addressed:]{antonia.mey@ed.ac.uk}
 \affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
@@
 \begin{abstract}
-The mechanisms by which we can determine a protein's 3D structure based on the protein's amino acid sequence have long been one of the key mysteries of biophysics.
-With the advent of AlphaFold2, making predictions for protein structures based on amino acid sequences at an accuracy similar to experiments is now possible, yet requires a lot of available structural data for training the model.
+The mechanisms by which we can determine a protein's 3D structure based on its amino acid sequence have long been one of the key mysteries of biophysics.
+With the advent of AlphaFold~2, making predictions for protein structures based on amino acid sequences at an accuracy similar to experiments is now possible, yet requires a lot of available structural data for training the model.
 Predicting contact maps from sequence \textit{de novo} is an appealing alternative.
@@
 To this end, we look at the scaling behavior of the simulation with the analytical approximation for different chain lengths typically found in proteins with good agreement.
-Furthermore, we compare our constrained model with frequency distributions of amino acid distances from contact maps of real proteins---from the protein data bank and AlphaFold2 database---showing that these follow a similar trend as the geometrically constrained model.
+Furthermore, we compare our constrained model with frequency distributions of amino acid distances from contact maps of real proteins---from the protein data bank and AlphaFold~2 database---showing that these follow a similar trend as the geometrically constrained model.
  We therefore suggest the conclusion that the approximate power-law behavior of the amino acid distance distribution can be attributed to geometric constraints of protein chains and amino acid sequences only play a secondary role.
@@
 insulin)~\cite{dill2008protein,Dill1042}.
-All natural proteins are made of 20 different amino acids which dictate the three-dimensional conformations the proteins adopt in order to function~\cite{scheraga2007protein}.
+All natural proteins are made of 20 different amino acids which dictate the three-dimensional conformations the proteins adopt in order to function~\cite{scheraga2007proteinfolding}.
 One of the great challenges has been to understand how the primary structure, the amino acid sequence, can lead to the fully folded functional protein.
@@
 [input information from Rohan's presentation].
-AlphaFold2, a machine learning model for structure prediction, has shown predictiveness at a level of experimental accuracy in the most recent protein structure prediction challenge~\cite{jumper2021highly, kryshtafovych2021critical}.
-This has lead to an explosion of available structural protein models beyond the protein database, containing structures from crystallographic, NMR, and cryo-EM experiments~\cite{berman2000protein}.
+AlphaFold~2, a machine learning model for structure prediction, has shown predictiveness at a level of experimental accuracy in the most recent protein structure prediction challenge~\cite{jumper2021highly, kryshtafovych2021critical}.
+This has led to an explosion of available structural protein models beyond the protein database, containing structures from crystallographic, NMR, and cryo-EM experiments~\cite{berman2000protein}.
 Beyond the prediction of structure understanding emergent patterns in proteins on a more fundamental level is of interest, as it opens up a deeper understanding of proteins structure and function relationship.
-One way of looking at the functional forms of proteins and classifying their structural properties is by looking at protein contact maps (PCM) or protein contact networks (PCN)~\cite{Vendruscolo2002,DiPaola2013,Estrada2011}.
+One way of looking at the functional forms of proteins and classifying their structural properties is by looking at protein contact maps (PCM) or protein contact networks (PCN)~\cite{Vendruscolo2002,dipaola2013protein,Estrada2011}.
 A PCM---we will use PCMs in favor of PCN in this paper---is the resulting adjacency matrix of spatially close amino acids, given a certain cut-off as determined typically from a protein structure.
@@
 Their use is also prominently featured in machine learning approaches.
-One challenge is to predict a PCM \textit{de novo} from sequence alone~\cite{bassot2019using, rives2021biological, rao2021msa}, without using structural data as part of the training set, as was done for AlphaFold2.
+One challenge is to predict a PCM \textit{de novo} from sequence alone~\cite{bassot2019using, rives2021biological, rao2021msa}, without using structural data as part of the training set, as was done for AlphaFold~2.
 These PCMs can then be used in e.g.
@@
 Since a PCM is an adjacency matrix and therefore a graph, it is possible to investigate PCMs using common network measures.
-For example, Bartoli \textit{et al.}~\cite{bartoli2008effecta} build a model in which they assume that proteins have an amino acid distance distribution that follows $\frac{1}{s}$, where $s$ is the separation of two connected amino acids along the backbone chain (in the following called \emph{amino acid distance}) as the simplest implementation of a connection probability, thus decreasing with the distance along the sequence.
+For example, Bartoli \textit{et al.}~\cite{bartoli2008effecta} build a model in which they assume that proteins have an amino acid distance distribution that follows $s^{-1}$, where $s$ is the separation of two connected amino acids along the backbone chain (in the following called \emph{amino acid distance}) as the simplest implementation of a connection probability, thus decreasing with the distance along the sequence.
 However, they do not provide a quantitative explanation for this assumption.
@@
 To this end, we define and analyze the amino acid distance distributions $P(s)$ that emerge from the geometric model, where the \emph{amino acid distance} $s$ of a link is defined as the distance of the two end units along the original chain.
+This distribution is then compared to the amino acid distance distribution extracted from protein contact networks generated from protein data bank and AlphaFold~2 data.
 
-This distribution is then compared to the amino acid distance distribution extracted from protein contact networks generated from protein data bank data.
-
-We find that, despite the highly simplified model, the generated amino acid distance distribution is in agreement with the one observed in networks generated from PDB data and roughly consistent with the heuristic $1/s$ finding.
-The analytical approximation provides a correction term on top of the $1/s$ heuristic.
+We find that, despite the highly simplified model, the generated amino acid distance distribution is in agreement with the one observed in networks generated from PDB data and roughly consistent with the heuristic $s^{-1}$ finding.
+The analytical approximation provides a correction term on top of the $s^{-1}$ heuristic.
 
@@
 \section{Protein contact maps}
+\label{sec:methods}
 The complex interaction pattern between the amino acids in a protein can be very naturally expressed as a network or graph, in which each amino acid is represented by a node and spatial proximity is encoded as a link.
 Whenever two central $C_\alpha$ atoms are closer together than a threshold $d_c$, they are connected, linked, or in `close contact'.
-These connections can be determined from the 3D functional or folded structure of the protein and are encoded in a so called contact map.
+These connections can be determined from the 3D functional or folded structure of the protein and are encoded in a so-called contact map.
 This contact map is effectively an adjacency matrix $\textbf{A}^{\mathrm{struc}}$ or binary matrix
@@
    0, & \text{ if } d_{i,j}>d_c \text{ or } i=j\\
-      1, & \text{ if } d_{i,j}\leq d_c .
+      1, & \text{ if } d_{i,j}\leq d_c.
       \end{cases}
@@
 b)~The solid green link prohibits some connections from node~$A$ (dotted pink), while others are still permitted (dotted green).
-Counting all newly excluded links, we find that a link of length $s$ prohibits $2(l-1)$ links of length $l$ up to a maximum of $s$ and $2(s-1)$ links of length $s$ or more.
-c)~Example of a protein contact map taken from a 2D simulation.
-Pink squares indicate a contact is present between two amino acids.}
+Counting all newly excluded links, we find that a link of length $s$ prohibits $2(l-1)$ links of length $l$ up to a maximum of $s$ and $2(s-1)$ links of length $s$ or more.}
     \label{fig:schematic}
@@
 To this end we will look at `averaged' contact maps over many different proteins that have different amino acid sequences, but their overall sequence or chain length is the same.
-We will use the RCSB protein data bank (PDB), a database of structurally resolved protein amino acid sequences through X-ray crystallography, NMR or cryo-EM experiments and the synthetically generated structural data base from the machine learned structures of AlphaFold2~\cite{jumper2021highly}.
+We will use the RCSB protein data bank (PDB), a database of structurally resolved protein amino acid sequences through X-ray crystallography, NMR or cryo-EM experiments and the synthetically generated structural database from the machine learned structures of AlphaFold~2~\cite{jumper2021highly}.
 The PDB contains just under 200,000 protein structures to date~\cite{PDB}.
@@
 The pink subset of the barplot shown in Fig.~\ref{fig:pdb_stats} b, are the set of structures used that fulfill the criterion on amino acid chain lengths between 85 and 115 amino acids, 185 and 215 amino acids and 285 and 315 amino acids and represent all structures for the analysis.
-The second database consists of synthetically generated structures using AlphaFold2 with sequences taken from SwissProt~\cite{alpha fold 2 database}.
-SwissProt is a manually curated database of protein sequences containing over 500,000 protein sequences~\cite{SwissProt}, whose amino acid chain length distribution is shown in Fig.~\ref{fig:pdb_stats} a in blue-green.
-The pink bars again indicate the subset of AlphaFold2 structures used for the contact map analysis.
+The second database consists of synthetically generated structures using AlphaFold~2 with sequences taken from SwissProt~\cite{alpha fold 2 database}.
+SwissProt is a manually curated database of protein sequences containing over 500,000 protein sequences~\cite{SwissProt}, whose amino acid chain length distribution is shown in Fig.~\ref{fig:pdb_stats} a in teal.
+The pink bars again indicate the subset of AlphaFold~2 structures used for the contact map analysis.
+In this case, the used structures were filtered first with the condition that the AlphaFold~2 pre-residue confidence score~\cite{jumper2021highly} was `Very high', i.e.
+above 90 out of 100, and then by removing any structures describing the same protein to remove duplicates.
 
@@
 	    \caption{Histogram of protein sequence lengths for protein structures from two structure databases.
-a) AlphaFold2 predicted protein structures ($\alpha$F) taken from all protein sequences in the reviewed Uniprot Knowledge Base (Swiss-Prot) database based on their Uniprot ids.
-Pink bars shows fraction of structures used for the average contact map analysis.
-b) RSCB PDB chain length frequencies based on sequence length.
-Pink bars in plot show proteins structures whose sequence lengths matches the resolution of structure accounting for the whole sequence.
+a) AlphaFold~2~\cite{jumper2021highly} predicted protein structures taken from protein sequences in the reviewed Uniprot Knowledge Base (Swiss-Prot) database based on their Uniprot IDs.
+Pink bars show fraction of structures used for the average contact map analysis.
+b) RSCB~\cite{PDB} PDB chain length frequencies based on sequence length.
+Pink bars in plot show protein structures whose sequence lengths matches the resolution of structure accounting for the whole sequence.
 Only pink PDB samples were used for the average contact map analysis.
-c) Example of a protein contact map taken from PDB ID: 1AKI, with a cut-off of 8~Å.
+c) Example of a protein contact map taken from PDB ID: 3UFG, with a cut-off of 8~Å.
 Pink squares indicate a contact present between amino acids.}
@@
 \end{figure}
-All structure based contact maps are generated in the same way for AlphaFold2 structures and PDB structures.
+All structure-based contact maps are generated in the same way for AlphaFold~2 structures and PDB structures.
 Structures used were downloaded in pdb format, read in with MDAnalysis version 2.0.0~\cite{gowers2016mdanalysis} and only C$_{\alpha}$ were selected and used for the analysis.
@@
 All pairwise distances were evaluated and in this way, building up the binary contact map or adjacency $A_{ij}^{\mathrm{struc}}$.
-All code for the construction of the contact maps both for the geometrically constrained system and structural protein data and their analysis can be found on github at: \url{https://github.com/meyresearch/sequence_distance_distribution}.
+All code for the construction of the contact maps both for the geometrically constrained system and structural protein data and their analysis can be found on GitHub at: \url{https://github.com/meyresearch/sequence_distance_distribution}.
 
@@
 Let us start by introducing the auxiliary variable $F_k(s)$, defined to be the number of possible links with an amino acid distance of $s$ if $k$ links have been added before.
-Before any links are added, all amino acid distances are equally likely, as there are $F_0(s)=N$ possibilities of making a link of each distance $s$ with $2\leq s < N/2$.
+Before any links are added, all amino acid distances are equally likely, as there are $F_0(s)=N$ possibilities of making a link of each distance $s$ with $2\leq s < \frac{N}{2}$.
 
@@
 Aiming to decouple those dynamics for different $s$, we make a heuristic guess for $\Pin^k(s_k)$.
-It consists of two separate approximation which are used in steps $k<a$ and steps $k\geq a$ respectively, where the threshold $1\leq a\leq N/2$ is a free parameter.
+It consists of two separate approximations which are used in steps $k<a$ and steps $k\geq a$ respectively, where the threshold $1\leq a\leq N/2$ is a free parameter.
 In the early stage we use a uniform probability
@@
 This approximation allows the use of the geometric series for solving the equation. 
-The resulting expression in Eq.~\ref{eq.Ps} is close to a power-law in $s$ with an exponent of \red{something close to -1.2 depending on a}, thus justifying the earlier approximation of $\Pin^k(s)$.
+The resulting expression in Eq.~\ref{eq.Ps} is by visual inspection of Fig.\ref{fig:2d_sim} resembles a power law with an exponent of -1, thus justifying the earlier approximation of $\Pin^k(s)$.
 %which makes use of Eq.~\ref{eq.Pk}. 
@@
 \subsection{Analytic approximation and 2D simulations}
-Fig.~\ref{fig:2d_sim} compares Eq.~\ref{eq.Ps} with respect to the two-dimensional simulation data introduced in~\cite{molkenthin2020self}.
-The simulation data is generated from 10 repeats with chain lengths of up to 498 amino acids long.
-The simulation data is plotted showing the mean frequencies of each amino acid distance up to $N/2\sim250$.
-The shaded area represent the 95\% confidence interval calcualted from the 10 repeats.
-The theoretical approximation from Eq.~\ref{eq.Ps} was fitted to the simulation data.
-The two fitting parameters were chosen to be $a=6$ and $A=0.39$, to best capture the trend of the simulation data.
-The fitted approximation captures the behavior of the simulation data well, both describe two dimensional systems.
- \begin{figure}[tb]
+Fig.~\ref{fig:2d_sim} compares Eq.~\ref{eq.Ps} with respect to the two-dimensional (a) and three-dimensional (b) simulation data introduced in~\cite{molkenthin2020self}.
+The 2D simulation data, generated from 10 repeats with chain lengths of up to 498 amino acids, consists of frequencies of amino acid distances in each simulation run, while the 3D simulation data consists of adjacency matrices computed from 30 simulation runs, and was analyzed the same way as the PDB and AlphaFold~2 protein data.
+These are plotted showing the mean frequencies of each amino acid distance up to $\frac{N}{2}\sim250$ and $\frac{N}{2}\sim150$ for 2D and 3D simulations, respectively.
+The shaded areas represent the 95\% confidence intervals.
+The theoretical approximation from Eq.~\ref{eq.Ps} was fitted to both the 2D and 3D simulation data separately.
+The two parameters, the exponent $a$, and the dimensionality constant $A$, were given by the fit as $a_{\mathrm{2D}}=6 \pm 1$, $A_{\mathrm{2D}}=(3.0 \pm 0.5)\cdot 10^{-4}$ and $a_{\mathrm{3D}}=1 \pm 1$, $A_{\mathrm{3D}}=(7 \pm 2)\cdot 10^{-3}$ for 2D and 3D simulations, respectively.
+The fitted approximation captures the behavior of the simulation data well, as in the case for 2D simulations, both describe two-dimensional systems.
+
+ \begin{figure}[htb]
         \centering
-	\includegraphics[width=0.45\textwidth]{paper/figures/Fig3/Fig3.pdf}
-	    \caption{Amino acid distance distribution for two-dimensional simulated data plotted against the analytic approximation. Green area shows the 95\% confidence interval.}
+	\includegraphics[width=\columnwidth]{paper/figures/Fig3/Fig3.pdf}
+	    \caption{Amino acid distance distributions for a) two-dimensional and b) three-dimensional simulated data plotted against the analytic approximation.
+Pink area shows the 95\% confidence interval.
+c) Example of a protein contact map taken from one of the three-dimensional simulation runs.
+The pink squares indicate a contact between two nodes.}
         \label{fig:2d_sim}
@@
 \subsection{3D simulations and PCMs from structural data }
-Next, we investigate if and how the analytical approximation can also capture trends in 3-D simulation data, as well as experimentally determined protein structures.
+%Next, we investigate if and how the analytical approximation can also capture trends in 3D simulation data, as well as experimentally determined protein structures. 
+Next, we investigate if and how the analytical approximation can also capture trends in the experimentally determined protein structures.
 This seeks to answer the question if indeed a power law can be used to describe the probability distribution of amino acid distances.
-Fig.~\ref{fig:sdd} summarises the results for protein chain lengths of $\approx$100 (a), $\approx$200, and $\approx$ 300 amino acids, protein chain lengths, most commonly found in living organism (see.
-Fig.~\ref{fig:pdb_stats} a).
-We are comparing data from repeats of 30 simulation runs of the three-dimensional simulations, as introduced in~\cite{molkenthin2020self}, and protein contact maps from RCSB~\cite{}, and AlphaFold2~\cite{}.
+We have chosen a subset of structures from the RCSB~\cite{PDB} and AlphaFold~2~\cite{jumper2021highly} databases for the analysis.
+From the RCSB database, 12,206 structures with chain lengths that fall between 85-115, 11,031 structures that fall between 185-215, and 10,091 structures that fall between 285-315 amino acids were selected.
+Additionally, 17,370 structures with chain lengths in the range of 85-115, 24,665 in the range of 185-215 and 21,804 in the range of 285-315 amino acids were chosen from the AlphaFold~2 database~\cite{jumper2021highly}.
+All structures were converted into PCMs as described in Section~\ref{sec:methods}.
 
-We have chosen a subset of structures from the PDB and curated alphaFold2 database for the analysis.
-7,376 structures with chain lengths that fall between 85-100, 6,822 structures that fall between 185-215, and 4,160 structures that fall between 285-315 amino acids were selected from the RCSB database~\cite{PDB}.
-Additionally, 33,229 structures with chain lengths in the range of 85-115, 37,430 in the range of 185-215 and 32,137 in the range of 285-315 amino acids were chosen from the AlphaFold2 database~\cite{}.
-All structures were converted into PCMs as described in section~\ref{sec:methods}.
-The resulting distributions for each chain length interval are plotted logarithmically scaled in Fig.\ref{fig:sdd} for intervals 85-100 in a, 185-215 in b, and 285-315 in c. 
+Fig.~\ref{fig:sdd} a summarises the results for protein chain lengths of $\approx100$, $\approx200$ and $\approx300$ amino acids, most commonly found in living organisms (see Fig.~\ref{fig:pdb_stats} a).
+This also shows how the amino acid distribution is scaled with respect to increasing chain lengths.
+
+In Fig.~\ref{fig:sdd} b, we compare data from  protein contact maps from RCSB~\cite{PDB}, and AlphaFold~2~\cite{jumper2021highly} for protein chain lengths of $\approx300$ amino acids.
+The resulting distributions for each chain length interval are plotted logarithmically scaled in Fig.\ref{fig:sdd} a for intervals 85-115 (100), 185-215 (200), and 285-315 (300).
 For all three lengths, the distributions show an approximate power law decay, for both the 3-D simulations according to~\cite{molkenthin2020self}, as well as the structural data.
 This allows a fit  which is consistent with both the simulated results and the theoretical approximation.
+We also fit the analytic approximation as well as a general power law to the RCSB structures in this range.
+The shaded area shows the 95\% confidence level (C.L.).
+The parameters given by the theoretical approximation are $a=4 \pm 1$ and $A=(7.4 \pm 0.8)\cdot 10^{-4}$.
+As for the power law, the exponent is $\gamma=1.44 \pm 0.02$, meaning that this is a scale-free network ($1\leq\gamma\leq2$).
+We also analyze the amino acid distributions from the chosen RCSB and AlphaFold~2 structures by calculating the Kolmogorov-Smirnov (KS) statistic.
+The KS statistic for the protein chain lengths of $\approx$300 amino acids was 0.12 and thus the null hypothesis was accepted at the $\alpha=0.01$ level.
+%Fig.~\ref{fig:sdd} summarises the results for protein chain lengths of $\approx$100 (a), $\approx$200, and $\approx$ 300 amino acids, protein chain lengths, most commonly found in living organism (see.
+Fig.~\ref{fig:pdb_stats} a).
+%We are comparing data from repeats of 30 simulation runs of the three-dimensional simulations, as introduced in~\cite{molkenthin2020self}, and protein contact maps from RCSB~\cite{}, and AlphaFold~2~\cite{}.
 
-is compared to those generated in simulations in~\cite{molkenthin2020self} as well as the above derived theoretical expression scaled with a factor to account for the different dimensionality, i.e.
+\begin{figure}[h!]
+        \centering
+	\includegraphics[width=\columnwidth]{paper/figures/Fig4/Fig4.pdf}
+        \caption{a) Amino acid distance distributions for contact maps derived from the RCSB~\cite{PDB} database for protein chains in the ranges of 85-115 (100), 185-215 (200) and 285-315 (300) amino acids.
+b) Comparison between the distributions of RCSB~\cite{PDB} and AlphaFold~2~\cite{jumper2021highly} databases for protein chains in the $\approx300$ range.
+The analytic approximation (solid line) and power law (dashed line) are fitted to the tail of the RCSB data.
+Pink shaded area shows the 95\% confidence level.
+c) Secondary structure content distribution in the RCSB~\cite{PDB} and AlphaFold~2~\cite{jumper2021highly} structures in the $\approx300$ chain length range.}
+        \label{fig:sdd}
+\end{figure}
+
+Finally, Fig.~\ref{fig:sdd} c shows the secondary structure content distributions for proteins in the $\approx300$ chain length range from RCSB and AlphaFold~2.
+This shows us that the proteins in this range seem to mostly contain secondary structures in the combined $\alpha$-helix---$\beta$-sheet region.
+Similar plots for the protein chain length regions of $\approx100$ and $\approx200$ amino acids can be found in the Supplementary Materials.
+
+%is compared to those generated in simulations in~\cite{molkenthin2020self} as well as the above derived theoretical expression scaled with a factor to account for the different dimensionality, i.e.
 constant $A$, as well as approximations made in the derivation.
 
-\begin{figure}
-        \centering
-	\includegraphics[scale=0.65]{paper/figures/Fig4/Fig4.pdf}
-        \caption{Amino acid distance distributions for contact maps derived from the RCSB~\cite{PDB} and AlphaFold2~\cite{} databases as well as 3D simulated folds for protein chains in the range of a) 85-115,  b) 185-215 and c) 285-315 amino acids long.
-        }
-        \label{fig:sdd}
-\end{figure}
-For large values of the chain length $N$, we thus find an approximate power law probability distribution for the amino acid distance with an exponent of $-1+\alpha$.
+%For large values of the chain length $N$, we thus find an approximate power law probability distribution for the amino acid distance with an exponent of $-1+\alpha$.
 
-We observe that, while the general shape seems to match, in real proteins amino acid distances of two to four amino acids are slightly over-represented (see Fig.\ref{fig:sdd}).
+We observe that, while the general shape seems to match, in real proteins amino acid distances of two to four amino acids are slightly over-represented (see Fig.\ref{fig:sdd} a and b).
 The over-representation of very short amino acid distances is an artefact of the different network construction for simulated (and theoretically approximated) networks and PCMs.
@@
 
-Here we use 30 runs of three-dimensional simulations, as introduced in~\cite{molkenthin2020self}, and networks extracted from measured protein data from both the RCSB~\cite{PDB} and AlphaFold2~\cite{} databases to compare this behavior with the more realistic 3D settings and show that the basic principle persists, indicating that much of the amino acid distance distribution can be attributed to generic geometric constraints.
+%Here we use 30 runs of three-dimensional simulations, as introduced in~\cite{molkenthin2020self}, and networks extracted from measured protein data from both the RCSB~\cite{PDB} and AlphaFold~2~\cite{} databases to compare this behavior with the more realistic 3D settings and
+
+\red{...show that the basic principle persists, indicating that much of the amino acid distance distribution can be attributed to generic geometric constraints.}
+
 \section{Conclusion}
@@
 \subsection{Availability of data and materials}
-All data generated or analysed during this study are included in this published article [and its supplementary information files].
+All data generated or analysed during this study are included in this published article.
+All raw data and their analysis can be found on GitHub at: \url{https://github.com/meyresearch/sequence_distance_distribution}, including information required for their reproduction.
 \subsection{Competing interests}
@@
 \subsection{Authors' contributions}
-
+N.M. and J.J.G. contributed equally to this work.
 \subsection{Acknowledgements}
-We thank Marc Timme for fruitful discussions.
+We thank Marc Timme, and Matteo Degiacomi for fruitful discussions.
 
@@
 
-\appendix
-\section{Supplemental Material Collection}
-\red{Do we have anything that needs to go here or can we just leave this?}
-\section{Random ideas}
-\begin{itemize}
-    \item Why is the beginning different for simulations?
-    \item Does the dip disappear for IDPs
-    \item Can we use all NMR structures for IDPs
-    \item Would it help using more simulated structures than just the Crystal structure. 
-\end{itemize}
-
 \end{document}

### 2022-04-14 paper/supplementary_material.tex
Updates from Overleaf

@@
 amsmath,amssymb,superscriptaddress,aps,
-pre]{revtex4-1}
+pre,floatfix]{revtex4-1}
 \usepackage[utf8]{inputenc}
@@
 \author{Antonia S J S Mey}
-\email[Electronic Address: ]{antonia.mey@ed.ac.uk}
+\email[Author to whom correspondence should be addressed: ]{antonia.mey@ed.ac.uk}
 \affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
@@
 
-\section*{Optimizing theory parameters}
+\section*{Amino acid distance distributions of other chain lengths}
+\subsection{AlphaFold 2}
+Fig.~\ref{fig:alphafold} shows the amino acid distance distributions for protein contact maps calculated from AlphaFold 2~\cite{jumper2021highly} sequences in the chain length ranges of 85-115 (100), 185-215 (200) and 285-315 (300) amino acids.
+The plot shows the approximate power law behaviour of the distribution as well as the scaling with increasing chain lengths.
 
- Fig.~\ref{fig:2d} shows a grid of the two-dimensional simulations for amino acid distance distributions for chains of around 498 amino acids long.
-The solid line shows the theoretical distribution given in Eq.~(12) of the main article.
-Each column corresponds to a different value of the scaling parameter $A$ and each row to a different value of the exponent parameter $a$ for this theoretical distribution.
-Plotted also is the 95\% confidence level for the 2D amino acid distance distribution.
+\begin{figure*}[htb]
+    \centering{
+    \includegraphics[scale=0.5]{paper/figures/SM_figures/alpha_chain_comparison.pdf}
+    \caption{Amino acid distance distributions for contact maps derived from the AlphaFold 2~\cite{jumper2021highly} database for protein chains in the ranges of 85-115 (100), 185-215 (200) and 285-315 (300) amino acids.}
+    \label{fig:alphafold}}
+\end{figure*}
 
-In addition to plotting this amino acid distance distribution grid, we also plotted the residual plots for each combination of parameters.
-These are shown in Fig.~\ref{fig:2d_r}.
-The residual sum of squares $\Sigma \sigma$, $R^2$ statistic and mean of residuals, $\bar{\sigma}$, were calculated for each combination of parameters $a$ and $A$.
+\subsection{Protein Data Bank and three-dimensional simulations}
 
-The choice of parameters was first narrowed down by looking at how well the theoretical distribution appeared to fit the amino acid distance distribution of the simulations in Fig.~\ref{fig:2d}, as well as the spread of residuals around the theoretical line in Fig.~\ref{fig:2d_r}.
-From Fig.~\ref{fig:2d}, we chose a smaller group of plots where the theoretical distribution fitted through a majority of simulation data points.
-Similarly, from Fig.~\ref{fig:2d_r}, we chose a group of plots where the majority of residuals were randomly scattered around zero.
-The non-linear parts of the residuals both at the low and high ends of the amino acid distances were expected from Fig.~\ref{fig:2d}.
-In the next step, we found the optimal parameters by finding the best combination of statistics, i.e.
-a minimized residual sum of squares, $\Sigma \sigma^2$, maximized constant of determination, $R^2$ and the mean of residuals, $\bar{\sigma} \approx 0$.
-This was done to avoid relying on a single statistic to determine the optimal parameters.
-For this 2D simulation data, we chose $A=0.39$ and $a=6$, which produced the following statistics: $\Sigma \sigma^2 = 17.4\times10^3$, $R^2 = 90.0$~\% and $\bar{\sigma} = -0.265$.
+Fig.~\ref{fig:rcsb_100} a shows the amino acid distance distributions for protein contact maps calculated from structures in the chain length range $\approx100$ for RCSB~\cite{PDB} (dark purple circles) and AlphaFold 2~\cite{jumper2021highly} (dark green squares).
+The shaded area shows the 95\% confidence level.
+Also shown are the analytic approximation and a general power law fitted to the RCSB data and their fit parameters, exponent $a$ and the dimensionality constant $A$ for the analytic approximation and exponent $\gamma$ for the power law,  are shown in~Table~\ref{table:stats} a.
+Similarly, Fig.~\ref{fig:rcsb_200} shows the amino acid distance distributions for protein contact maps calculated from structures in the chain length range $\approx200$ for RCSB~\cite{PDB}.
+We also computed the Kolmogorov-Smirnov (KS) statistic between the RCSB and AlphaFold 2 distributions for each length range.
+The KS statistics were 0.07 and 0.10 for the $\approx100$ and $\approx200$ ranges, respectively.
+For both of these, the null hypothesis was accepted at the $\alpha=0.01$ level.
 
-The above steps were repeated for both the AlphaFold 2 and RCSB PDBs, for each chain length range.
-Tables~\ref{table:af}~and~\ref{table:rcsb} show the optimized parameters and their statistics for PDBs obtained from AlphaFold 2 and RCSB, respectively.
-Figs.~\ref{fig:AF100},~\ref{fig:AF200}~and~\ref{fig:AF300} show the parameter grid plots of amino acid distance distributions for chains of 100, 200 and 300 amino acids long, respectively.
-Similarly, Figs.~\ref{fig:AF100_r},~\ref{fig:AF200_r}~and~\ref{fig:AF300_r} show the residual plots.
-While the 95\% confidence intervals are plotted for the PDBs, these were not plotted in the main article for clarity, as the confidence level appears smaller than the scatter markers.
+Fig.~\ref{fig:sim_3d} shows the amino acid distance distributions for protein contact maps calculated from three-dimensional simulations obtained from~\cite{molkenthin2020self}.
+As before, the analytic approximation was fitted to the data in each range and the shaded area shows the 95\% confidence interval.
+The fit parameters $a$ and $A$ are given in Table~\ref{table:stats} c and d for ranges $\approx100$ and $\approx200$, respectively.
 
@@
 \centering
-\begin{tabular}{|c c c c c c|}
+\begin{tabular}{|c c c c c |}
 \hline
-Chain length  & $A$ ($10^{-3}$) & $a$ & $\Sigma \sigma^{2}$ & $R^2$ (\%) & $\bar{\sigma}$ \\
+Index & Data  & $A$ ($10^{-3}$) & $a$ & $\gamma$ \\
 \hline
-100 & 2.1 & 3 & 0.0002 & 57 & 0.0001 \\
-200 & 1.1 & 3 & 0.0135 & 62 & -0.0005 \\
-300 &1.0 & 5 & 0.1478 & 76  & -0.0012 \\
+a & RCSB 100 & 1.5 $\pm$ 0.3  & 2 $\pm$ 1 &  0.83 $\pm$ 0.02\\
+b & RCSB 200 & 0.8 $\pm$ 0.2 & 2 $\pm$ 1 & 0.95 $\pm$ 0.02 \\
+c & 3D SIM 100 & 3.1 $\pm$ 0.6 & 2 $\pm$ 1 & --  \\
+d & 3D SIM 200 & 1.2 $\pm$ 0.4 & 0.9 $\pm$ 0.9  & --  \\
 \hline
 \end{tabular}
- \caption{Detailed parameters and statistics for PDBs obtained from AlphaFold 2.
-Shown are the optimized parameters $A$ and $a$ used to fit the theoretical amino acid distance distribution given in Eq.(12) of the main article, for chains of 100, 200 and 300 amino acids long.
-Shown also are the residual sum of squares ($\Sigma \sigma^{2}$), constant of determination ($R^2$) and the mean of residuals ($\bar{\sigma}$) for each fit.}
-\label{table:af}
-\end{table}
-
-\begin{table}[h]
-\centering
-\begin{tabular}{|c c c c c c|}
-\hline
-Chain length & $A$ ($10^{-3}$) & $a$ & $\Sigma \sigma^{2}$ & $R^2$ (\%) & $\bar{\sigma}$ \\
-\hline
-100 & 1.9 & 2 & 0.0002 & 59 & 0.0001 \\
-200 & 1.2 & 3 & 0.0474 & 70 & -0.0010\\
-300 & 0.9 & 4 & 0.1102 & 76 & -0.0011\\
-\hline
-\end{tabular}
- \caption{Detailed parameters and statistics for PDBs obtained from RCSB.
-Shown are the optimized parameters $A$ and $a$ used to fit the theoretical amino acid distance distribution given in Eq.(12) of the main article, for chains of 100, 200 and 300 amino acids long.
-Shown also are the residual sum of squares ($\Sigma \sigma^{2}$), constant of determination ($R^2$) and the mean of residuals ($\bar{\sigma}$) for each fit.}
-\label{table:rcsb}
+ \caption{Parameters (exponent $a$ and dimensionality constant $A$) for the analytic approximation, exponent $\gamma$ for the power law fit, and their uncertainties from fitting to amino acid distributions obtained from RCSB proteins in the chain length range $\approx100$ a and $\approx200$ b amino acids.
+Shown also are the fit parameters $a$ and $A$ from fitting the analytic approximation to the 3D simulations~\cite{molkenthin2020self}.
+of chains in the $\approx100$ c and $\approx200$ d ranges.}
+\label{table:stats}
 \end{table}
@@
     \centering{
-    \includegraphics[scale=0.65]{paper/figures/SM_figures/Fig1/Fig1.pdf}
-    \caption{a) A grid of the amino acid distance distribution in protein chains of 498 residues long in two-dimensional simulations.
-The solid line shows the theoretical distribution for different values of parameters $a$ and $A$.
-b) The residual plots of the amino acid distance distribution around the theoretical distribution.
-The residual sum of squares, $\Sigma \sigma$, $R^2$ statistic and the mean of residuals, $\bar{\sigma}$, were calculated for each combination of parameters.}
-    \label{fig:2d}}
+    \includegraphics[scale=0.35]{paper/figures/SM_figures/distributions_combined.pdf}
+    \caption{a) Comparison between the distributions of RCSB~\cite{PDB} and AlphaFold 2~\cite{jumper2021highly} databases for protein chains in the $\approx100$ range.
+The analytic approximation (solid line) and power law (dashed line) are fitted to the the RCSB data.
+Pink shaded area shows the 95\% confidence level.
+b) Secondary structure content distribution in the RCSB~\cite{PDB} and AlphaFold 2~\cite{jumper2021highly} structures in the $\approx100$ chain length range.}
+    \label{fig:rcsb_100}}
 \end{figure*}
+
+% \begin{figure*}[htb]
+%     \centering{
+%     \includegraphics[scale=0.35]{paper/figures/SM_figures/distributions_200.pdf}
+%     \caption{Comparison between the distributions of RCSB~\cite{PDB} and AlphaFold 2~\cite{jumper2021highly} databases for protein chains in the $\approx200$ range.
+The analytic approximation (solid line) and power law (dashed line) are fitted to the tail of the RCSB data.
+Pink shaded area shows the 95\% confidence level.}
+%     \label{fig:rcsb_200}}
+% \end{figure*}
 
@@
     \centering{
-    \includegraphics[width=0.8\textwidth]{paper/figures/SM_figures/Fig2/Fig2.pdf}
-    \caption{a) A grid of the amino acid distance distributions in chains of around 100 residues long in PDBs obtained from AlphaFold 2.
-The solid line shows the theoretical distribution for different values of parameters $a$ and $A$.
-b) The residual plots of the amino acid distance distribution around the theoretical distribution.
-The residual sum of squares, $\Sigma \sigma$, $R^2$ statistic and the mean of residuals, $\bar{\sigma}$, were calculated for each combination of parameters.}
-    \label{fig:AF100}}
-
+    \includegraphics[scale=0.35]{paper/figures/SM_figures/simulations_3d.pdf}
+    \caption{Amino acid distance distributions obtained from three-dimensional simulations~\cite{molkenthin2020self} for chains in the ranges $\approx100$ (a) and $\approx200$ (b) amino acids.
+The dotted line shows the analytic approximation fitted to the data and the shaded area shows the 95\% confidence level.}
+    \label{fig:sim_3d}}
 \end{figure*}
 
-\begin{figure*}[htb]
-    \centering{
-    \includegraphics[width=0.8\textwidth]{paper/figures/SM_figures/Fig3/Fig3.pdf}
-    \caption{a) A grid of the amino acid distance distributions in chains of around 200 residues long in PDBs obtained from AlphaFold 2.
-The solid line shows the theoretical distribution for different values of parameters $a$ and $A$.
-b) The residual plots of the amino acid distance distributions around the theoretical distribution.
-The residual sum of squares, $\Sigma \sigma$, $R^2$ statistic and the mean of residuals, $\bar{\sigma}$, were calculated for each combination of parameters.}
-    \label{fig:AF200}}
-\end{figure*}
-
-\begin{figure*}[htb]
-    \centering{
-    \includegraphics[width=0.8\textwidth]{paper/figures/SM_figures/Fig4/Fig4.pdf}
-    \caption{a) A grid of the distribution of link lengths in chains of around 300 residues long in PDBs obtained from AlphaFold 2.
-The solid line shows the theoretical distribution for different values of parameters $a$ and $A$.
-b) The residual plots of the amino acid distance distributions around the theoretical distribution.
-The residual sum of squares, $\Sigma \sigma$, $R^2$ statistic and the mean of residuals, $\bar{\sigma}$, were calculated for each combination of parameters.}
-    \label{fig:AF300}}
-
-\end{figure*}
-
-\begin{figure*}[htb]
-    \centering{
-    \includegraphics[width=0.8\textwidth]{paper/figures/SM_figures/Fig5/Fig5.pdf}
-    \caption{a) A grid of the distribution of link lengths in chains of around 100 residues long in PDBs obtained from RCSB.
-The solid line shows the theoretical distribution for different values of parameters $a$ and $A$.
-b) The residual plots of the amino acid distance distributions around the theoretical distribution.
-The residual sum of squares, $\Sigma \sigma$, $R^2$ statistic and the mean of residuals, $\bar{\sigma}$, were calculated for each combination of parameters.}
-    \label{fig:PDB100}}
-\end{figure*}
-
-\begin{figure*}[htb]
-    \centering{
-    \includegraphics[width=0.8\textwidth]{paper/figures/SM_figures/Fig6/Fig6.pdf}
-    \caption{A grid of the distribution of link lengths in chains of around 200 residues long in PDBs obtained from RCSB.
-The solid line shows the theoretical distribution for different values of parameters $a$ and $A$.
-b) The residual plots of the amino acid distance distributions around the theoretical distribution.
-The residual sum of squares, $\Sigma \sigma$, $R^2$ statistic and the mean of residuals, $\bar{\sigma}$, were calculated for each combination of parameters.}
-    \label{fig:PDB200}}
-
-\end{figure*}
-
-\begin{figure*}[htb]
-    \centering{
-    \includegraphics[width=0.8\textwidth]{paper/figures/SM_figures/Fig7/Fig7.pdf}
-    \caption{A grid of the distribution of link lengths in chains of around 300 residues long in PDBs obtained from RCSB.
-The solid line shows the theoretical distribution for different values of parameters $a$ and $A$.
-b) The residual plots of the amino acid distance distributions around the theoretical distribution.
-The residual sum of squares, $\Sigma \sigma$, $R^2$ statistic and the mean of residuals, $\bar{\sigma}$, were calculated for each combination of parameters.}
-    \label{fig:PDB300}}
-\end{figure*}
-
+\newpage
 \bibliographystyle{unsrt}
-% \bibliography{proteins}
+\bibliography{proteins}
 

### 2022-04-14 scripts/distances.py
some additional plot inspections

@@ -36,6 +36,7 @@ def create_plots(arguments: argparse.Namespace) -> None:
     colours = ["#a3dbab", "#47b856", "#286830"]
     if arguments.algorithm == "rcsb":
         colours = ["#e2b6dc", "#b847a9", "#682860"]
+
     elif arguments.algorithm == "alpha":
         colours = ["#a3dbab", "#47b856", "#15371a"]
     markers = ["o", "s", "^"]
@@ -44,10 +45,12 @@ def create_plots(arguments: argparse.Namespace) -> None:
     for i in range(len(length_ranges)):
         histogram = plot_functions.get_histogram(length_ranges[i], arguments.algorithm)
         plotting_tuple = plot_functions.get_data_for_plotting(histogram, arguments, length_ranges[i])
+        chain_length = plotting_tuple[0]
         distance_bins = plotting_tuple[1]
         measure = plotting_tuple[2]
         plt.scatter(distance_bins, measure, label=create_plot_label(arguments, length_ranges[i]),
                     color=colours[i], marker=markers[i])
+
         plt.yscale("log")
         plt.xscale("log")
         plt.xlabel("s")

### 2022-05-23 paper/main.tex

--- 2022-04-05 paper/main.tex
@@
 \usepackage[colorinlistoftodos]{todonotes}
-\newcommand{\Pin}{P_{\mathrm{in}}}
+\newcommand{\Pin}{P_{\mathrm{next}}}
 \newcommand{\fhigh}{f_{\mathrm{high}}}
@@
 \title{Theoretical approximation of chain separation distribution in protein contact networks}
-\title{Theory of chain separation distributions gives insights in real protein contact networks}
+\title{What geometrically constrained protein models can tell us about real-world protein contact maps}
 %\title{}
 \author{Nora Molkenthin}
-\affiliation{Potsdam Institut für Klimafolgenforschung}
-\affiliation{Center for Advancing Electronics Dresden (cfaed), Technical University of Dresden}
+\email[Author to whom correspondence should be addressed:]{nora.molkenthin@pik-potsdam.de}
+\affiliation{Potsdam Institute for Climate Impact Research, Germany}
 \author{J. J. Güven}
@@
 \author{Steffen Mühle}
-\affiliation{Max-Planck Institute for Dynamics and Self-Organization, 37077 Göttingen, Germany}
+\affiliation{Max-Planck Institute for Dynamics and Self-Organization (MPIDS), Am Faßberg 17, 37077 Göttingen, Germany}
 \author{Antonia S J S Mey}
-\email[Electronic Address: ]{antonia.mey@ed.ac.uk}
+\email[Author to whom correspondence should be addressed:]{antonia.mey@ed.ac.uk}
 \affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
@@
 \begin{abstract}
-The mechanisms by which we can determine a protein's 3D structure based on the protein's amino acid sequence have long been one of the key mysteries of biophysics.
-With the advent of AlphaFold2, making predictions for protein structures based on amino acid sequences at an accuracy similar to experiments is now possible, yet requires a lot of available structural data for training the model.
-Predicting contact maps from sequence \textit{de novo} is an appealing alternative.
-Here, we are interested in understanding the emergent behavior of protein contact maps derived from a network ensemble perspective of a geometrically constrained random interaction model.
-We show that the frequency distributions of amino acid distances obtained from simulations of the geometrically constrained interaction model can be described by an analytical approximation of the two-dimensional analogue of this model.
-To this end, we look at the scaling behavior of the simulation with the analytical approximation for different chain lengths typically found in proteins with good agreement.
-Furthermore, we compare our constrained model with frequency distributions of amino acid distances from contact maps of real proteins---from the protein data bank and AlphaFold2 database---showing that these follow a similar trend as the geometrically constrained model.
- We therefore suggest the conclusion that the approximate power-law behavior of the amino acid distance distribution can be attributed to geometric constraints of protein chains and amino acid sequences only play a secondary role.
-%The distribution of chain separations of interacting amino acids in proteins roughly follows a power law.
-Here we show, analytically and in simulations, that a geometrical stochastic model of folding chains can explain this behavior.
+The mechanisms by which a protein's 3D structure can be determined based on its amino acid sequence have long been one of the key mysteries of biophysics.
+Often simplistic models, e.g.
+derived from geometric constraints, capture bulk real-world 3D protein--protein properties well.
+One approach is using protein contact maps to better understand protein's properties.
+Here, we investigate the emergent behavior of protein contact maps derived from a geometrically constrained random interaction model in comparison to real-world proteins.
+Deriving an analytical approximation for the distribution of amino acid distances $s$ of a 2D geometric model, by means of a mean-field approach, we present a theoretical basis for the distribution of amino acid distances in proteins.
+To validate this approximation we compare it to simulations of a 2D and 3D geometrically constrained model, as well as contact maps of real proteins.
+Using data from the RCSB Protein Data Bank (PDB) and AlphaFold~2 database, the analytical approximation can be accurately fit to both datasets.
+This holds for protein chain lengths of $L\approx100$, $L\approx200$, and $L\approx300$.
+While, a universal scaling behavior for the different chain lengths could not be deduced, it seems that the amino acid distance distributions can be attributed to geometric constraints of protein chains in bulk and amino acid sequences only play a secondary role.
+
 \end{abstract}
@@
 \section{Introduction}
-Proteins, the molecular machines of every living organism perform vital tasks required for life to persist, ranging from transport (e.g.
+Proteins, the molecular machines of every living organism, perform vital tasks required for life to persist.
+These range from transport (e.g.
 hemoglobin)~\cite{ahmed2020hemoglobin}, signal transduction (e.g.
@@
 insulin)~\cite{dill2008protein,Dill1042}.
-All natural proteins are made of 20 different amino acids which dictate the three-dimensional conformations the proteins adopt in order to function~\cite{scheraga2007protein}.
-One of the great challenges has been to understand how the primary structure, the amino acid sequence, can lead to the fully folded functional protein.
-The last 50 years has seen many different routes to computationally predict biologically active protein structures without having to solve a crystal structure of the protein.
-[input information from Rohan's presentation].
-AlphaFold2, a machine learning model for structure prediction, has shown predictiveness at a level of experimental accuracy in the most recent protein structure prediction challenge~\cite{jumper2021highly, kryshtafovych2021critical}.
-This has lead to an explosion of available structural protein models beyond the protein database, containing structures from crystallographic, NMR, and cryo-EM experiments~\cite{berman2000protein}.
-Beyond the prediction of structure understanding emergent patterns in proteins on a more fundamental level is of interest, as it opens up a deeper understanding of proteins structure and function relationship.
-One way of looking at the functional forms of proteins and classifying their structural properties is by looking at protein contact maps (PCM) or protein contact networks (PCN)~\cite{Vendruscolo2002,DiPaola2013,Estrada2011}.
-A PCM---we will use PCMs in favor of PCN in this paper---is the resulting adjacency matrix of spatially close amino acids, given a certain cut-off as determined typically from a protein structure.
-PCMs have shown promise in understanding protein folding patterns as well as revealing allosteric communication pathways~\cite{yao2019establishing, menichetti2016network,dokholyan2002topological}.
+All natural proteins are made of 20 different amino acids which dictate the 3D conformations the proteins adopt in order to function~\cite{scheraga2007proteinfolding}.
+One of the great challenges has been to understand how the primary structure, i.e.
+the amino acid sequence, can lead to the fully folded functional protein, known as the protein folding problem~\cite{nassar2021protein}.
+The last 50 years have seen many different routes to computationally predict biologically active protein structures without having to solve a crystal structure of the protein~\cite{marks2012protein}.
+AlphaFold~2, a machine learning model for structure prediction, has shown predictiveness at a level of experimental accuracy in the most recent protein structure prediction challenge~\cite{jumper2021highly, kryshtafovych2021critical}.
+It was trained on structural data from the Research Collaboratory for Structural Bioinformatics Protein Data Bank~\cite{berman2000protein} (RCSB PDB).
+The RCSB PDB contains structures from crystallographic, NMR, and cryo-EM experiments.
+AlphaFold~2 certainly was a game changer in protein structure prediction~\cite{callaway2022what}, but it has not solved the protein folding problem in the sense that the mechanisms of the folding or structure function relationship remain unexplained.
+However, it does provide a rich resource of structures to understand emergent patterns in proteins on a more fundamental level.
+
+One way of mapping the functional forms of proteins and classifying their structural properties is by looking at protein contact maps (PCM)~\cite{Vendruscolo2002,dipaola2013protein,Estrada2011}.
+A PCM is a matrix representation of spatially close amino acids, given a certain cut-off distance as determined typically from a protein structure.
+PCMs have shown promise in understanding protein folding patterns, as well as revealing allosteric communication pathways~\cite{yao2019establishing, menichetti2016network,dokholyan2002topological}.
 Their use is also prominently featured in machine learning approaches.
-One challenge is to predict a PCM \textit{de novo} from sequence alone~\cite{bassot2019using, rives2021biological, rao2021msa}, without using structural data as part of the training set, as was done for AlphaFold2.
-These PCMs can then be used in e.g.
-graph convolutional machine learning approaches to train a model able to predict binding affinities for certain types of proteins~\cite{jiang2020drug},
-From a physicist's perspective it is interesting to understand the emergent behavior in protein structure by looking at PCMs.
-Since a PCM is an adjacency matrix and therefore a graph, it is possible to investigate PCMs using common network measures.
-For example, Bartoli \textit{et al.}~\cite{bartoli2008effecta} build a model in which they assume that proteins have an amino acid distance distribution that follows $\frac{1}{s}$, where $s$ is the separation of two connected amino acids along the backbone chain (in the following called \emph{amino acid distance}) as the simplest implementation of a connection probability, thus decreasing with the distance along the sequence.
-However, they do not provide a quantitative explanation for this assumption.
+For example, similarly to AlphaFold~2, one challenge is to predict a PCM \textit{de novo} from sequence alone~\cite{bassot2019using, rives2021biological, rao2021msa}, without using structural data as part of the training set.
+Another is the use PCMs in machine learning approaches to e.g.
+train a model able to predict binding affinities between a protein and a drug-like molecule~\cite{jiang2020drug}.
 
-Here, we show how a simple geometrical model can reproduce this heuristically found behavior well, and even provide a simple analytical approximation.
-The three-dimensional geometrical model and the analytical approximation of its two-dimensional analogue both build on the idea that inherently geometrical objects, such as amino acids, imply that any resulting PCM network ensemble modelling them has to be spatially embedded~\cite{molkenthin2016scaling, molkenthin2020self}.
-Different approaches have been used in the past for geometrical models describing protein folding, some of which derived characteristics of the secondary structure from constraints on bond and torsion angles~\cite{bhattacharjee2013flory,Danielsson2010,Molkenthin2011} and others focused on the formation mechanisms of the tertiary structure~\cite{molkenthin2016scaling, molkenthin2020self}.
+From a physicist's perspective, it is interesting to understand the emergent behavior, in terms of folding or function, in many protein structures.
+Therefore studying properties of underlying features of a PCM can reveal information on the scale of the proteome and not just individual proteins.
+Thus, having a reliable model to generate a PCM not from real data is invaluable.
+To this end, different heuristic models have been indtroduced in the past by Atilgan \textit{et al.}~\cite{atilgan2004smallworld} and Bartoli \textit{et al.}~\cite{bartoli2008effecta}.
+Atilgan \textit{et al.} propose a random shuffle model to generate artificial PCMs with similar properties to real protein models.
+Bartoli \textit{et al.} built a model for PCMs in which they assume ~\textit{ad hoc} that proteins have a distribution of amino acid distances $P(s)$ that follows $P(s) \approx s^{-1}$.
+The \textit{amino acid distance} $s$ is the separation of two connected amino acids along the backbone chain, and gives rise to the simplest implementation of a connection probability in a protein.
+Bartoli \textit{et al.} justify their approximation only heuristically in the sense that resulting PCMs have similar properties to real world PCMs.
 
-Based on such considerations, the network formation process repeatedly connects random pairs of units, without causing any intersections along the chain~\cite{molkenthin2016scaling}.
-Such a model can now be used to investigate the emergent behavior of amino acid distance distributions as used heuristically by Bartoli \textit{et al.}~\cite{bartoli2008effecta}.
-To this end, we define and analyze the amino acid distance distributions $P(s)$ that emerge from the geometric model, where the \emph{amino acid distance} $s$ of a link is defined as the distance of the two end units along the original chain.
+In this paper, we provide an explanation why the $s^{-1}$ heuristic is a good initial assumption.
+To do so, we derived an analytical approximation from a 2D geometrically constrained protein model to get an expression for $P(s)$.
+The approximation provides additional correction terms.
+We then show that the approximation can be used to fit simulations of the 2D contstrained model~\cite{molkenthin2016scaling}, a 3D version of the constrained model~\cite{molkenthin2020self} and even to real world proteins.
+This means that the analytical approximation is a good starting point for artificially modelling PCMs in the future.
 
-This distribution is then compared to the amino acid distance distribution extracted from protein contact networks generated from protein data bank data.
+%We find that, despite the highly simplified 2D approximation and 2D and 3D models, the generated amino acid distance distribution $P(s)$ is in agreement with the one observed in PCMs generated from structural data and roughly consistent with the heuristic of $s^{-1}$.
+From the derived analytical approximation a correction term for the $s^{-1}$ heuristic, can be derived.
 
-We find that, despite the highly simplified model, the generated amino acid distance distribution is in agreement with the one observed in networks generated from PDB data and roughly consistent with the heuristic $1/s$ finding.
-The analytical approximation provides a correction term on top of the $1/s$ heuristic.
-
+The paper is structured in the following way: In Sec.~\ref{sec:models}, we give a brief overview of the geometrically constrained models used, In Sec.~\ref{sec:methods}, we describe the construction of PCMs from the 2- and 3D simulations of the geometrically constrained model, as well as from structural data in the RCSB PDB and AlphaFold2 databases.
+In Sec.~\ref{sec:theory} we then derive an analytical approximation for the 2D analogue of the geometrically constrained model.
+Sec.~\ref{sec:results} highlights the results for the 2D, and 3D simulations, RCSB PDB, and AlphaFold~2 structural data by making use of the 2D analytical approximation.
 %%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%
-\section{Protein contact maps}
-The complex interaction pattern between the amino acids in a protein can be very naturally expressed as a network or graph, in which each amino acid is represented by a node and spatial proximity is encoded as a link.
-Whenever two central $C_\alpha$ atoms are closer together than a threshold $d_c$, they are connected, linked, or in `close contact'.
-These connections can be determined from the 3D functional or folded structure of the protein and are encoded in a so called contact map.
-This contact map is effectively an adjacency matrix $\textbf{A}^{\mathrm{struc}}$ or binary matrix
-\begin{equation}
-  A^{\mathrm{struc}}_{ij}=
-  \begin{cases}
-   0, & \text{ if } d_{i,j}>d_c \text{ or } i=j\\
-      1, & \text{ if } d_{i,j}\leq d_c .
-      \end{cases}
-    \label{eq:aij}
-\end{equation}
-Adjacency matrices are commonly used to describe networks and make up a \textit{protein contact map}. 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.8\columnwidth]{figures/Fig1/Fig1.pdf}
-    \caption{a) The amino acid distance $s$ is defined as the separation along the chain (teal) of two connected units (pink).
-b)~The solid green link prohibits some connections from node~$A$ (dotted pink), while others are still permitted (dotted green).
-Counting all newly excluded links, we find that a link of length $s$ prohibits $2(l-1)$ links of length $l$ up to a maximum of $s$ and $2(s-1)$ links of length $s$ or more.
-c)~Example of a protein contact map taken from a 2D simulation.
-Pink squares indicate a contact is present between two amino acids.}
-    \label{fig:schematic}
-\end{figure}
+\section{2D and 3D geometrically constrained models}
+\label{sec:models}
+Different approaches have been used in the past for geometrical models describing protein folding, some of which derived characteristics of the secondary structure from constraints on bond and torsion angles~\cite{bhattacharjee2013flory,Danielsson2010,Molkenthin2011} and others focused on the formation mechanisms of the tertiary structure~\cite{molkenthin2016scaling, molkenthin2020self} and even others are modelled by self avoiding random walks~\cite{mey2014rareevent, hills2009insights}.
+Here we will focus on models representing maximally folded structures in 2D, and 3D as introduced in~\cite{molkenthin2016scaling} and~\cite{molkenthin2020self} respectively.
+The 3D geometrical model and the analytical approximation of its 2D analogue both build on the idea that inherently geometrical objects, such as amino acids, imply that any resulting PCM network ensemble modeling them has to be spatially embedded.
 
-%%%%%%%%%%%%
-% subsection
-%%%%%%%%%%%%
-\subsection{Contact maps in a geometrically constrained protein model}
-
-Based on the basic observation that amino acids are objects in space which cannot overlap indefinitely, we introduced a model in~\cite{molkenthin2020self} starting from a chain of identical spheres, each in contact only with the neighbours it is connected to.
-Then additional connections form by randomly selecting two spheres and moving them towards each other until they touch, the new links formed that way can not be broken in later steps and function as constraints on subsequent links.
-If contact of the two selected spheres is geometrically impossible without breaking previously made connections or leading to overlaps of any spheres, the link is not made and taken out of the pool of possible connections.
-This process is repeated until no more links can be formed without violating the geometric constraints and the final structure can be expressed as the adjacency matrix
-\begin{equation}
-  \mathbf{A}^{\mathrm{sim}}_{ij}=
-  \begin{cases}
-   1, & \text{ if }|i-j|=1 \text{ or i and j  randomly connected} \\
-      0, & \text{ otherwise }  .
-      \end{cases}
-    \label{eq:aij}
-\end{equation}
-
-In~\cite{molkenthin2016scaling}, we have introduced a simplified, two-dimensional version of this model, which starts from a closed chain (i.e.
-a ring) of $N$ unit discs and subsequently adds links, such that connected discs touch, yet no discs intersect.
+In~\cite{molkenthin2016scaling}, we have introduced a simplified, 2D version of this geometrically constrained model.
+It starts from a closed chain (i.e.
+a ring) of $L$ unit discs and subsequently adds links, such that connected discs touch, yet no discs intersect.
 The advantage of this model is that it can be approximated by a purely topological simplification, which can be treated analytically.
@@
 (d) Once connected by a link, pairs of units do not disconnect.
+In~\cite{molkenthin2020self} this geometrically constrained model was extended to 3D spheres using simulations to generate compact artificial protein structures.
+%%%%%%%%%%%%%%%%%
+% Section  %
+%%%%%%%%%%%%%%%%%
+\section{Protein contact maps}
+\label{sec:methods}
+The complex interaction pattern between the amino acids in a protein can be naturally expressed as a network or graph, in which each amino acid is represented by a node and spatial proximity is encoded as a link.
+Whenever two central $\mathrm{C}_\alpha$ atoms are closer together than a threshold $d_c$, they are connected, linked, or in `close contact'.
+These connections can be determined from the 3D functional or folded structure of the protein and are encoded in a so-called (protein) contact map.
+This contact map is effectively an adjacency matrix $\textbf{A}^{\mathrm{struc}}$
+\begin{equation}
+  A^{\mathrm{struc}}_{ij}=
+  \begin{cases}
+   0, & \text{ if } d_{i,j}>d_c \text{ or } i=j\\
+      1, & \text{ if } d_{i,j}\leq d_c.
+      \end{cases}
+    \label{eq:aij}
+\end{equation}
 
-%%%%Missing%%%%%
-% TO DO: @Nora
-% Describe how the protein contact maps are generated in the geometric constrained model %
-% Let's call the adjacency matrix in this case \mathbf{A}^{\mathrm{sim}}_{ij} where sim stands for simulation. 
-%%%%%%%%%%%%%%%
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{figures/Fig1/Fig1.pdf}
+    \caption{a) The amino acid distance $s$ (green) is defined as the number of amino acids (AS) (here represented by a circle) along the protein backbone chain (black) between two connected AS (pink).
+b) Shows an example AS distance of length $s=3$ (solid green line).
+When adding a new connection from node A, the previously created connection prohibits new ones (dashed pink lines).
+The green dotted link from A is still permitted and can be made.
+c) Shows an example of an existing AS distance $s=7$ (solid green), where there are now more permitted (dotted green) links and fewer prohibited (dotted pink) ones.}
+    \label{fig:schematic}
+\end{figure}
+
+%%%%%%%%%%%%
+% subsection
+%%%%%%%%%%%%
+\subsection{Contact maps in a geometrically constrained protein model}
+
+Based on the simple observation that amino acids are objects in space which cannot overlap indefinitely, we introduced the 3D version of the 2D geometrically constrained model in~\cite{molkenthin2020self}.
+Starting from a chain of identical spheres, each in contact only with the neighbors it is connected to, additional connections are made by randomly selecting two spheres and moving them towards each other until they touch.
+The new links formed that way cannot be broken in later steps and function as constraints on subsequent link formation steps.
+If the contact of the two selected spheres is geometrically impossible without breaking previously made connections or leading to overlaps of any spheres, the link is not made and is taken out of the pool of possible connections.
+This process is repeated until no more links can be formed without violating the geometric constraints and the final structure can be expressed as the adjacency matrix
+\begin{equation}
+  A^{\mathrm{sim}}_{ij}=
+  \begin{cases}
+   1, & \text{ if }|i-j|=1 \text{ or i and j connected} \\
+      0, & \text{ otherwise}.
+      \end{cases}
+    \label{eq:sim_aij}
+\end{equation}
+
+%In~\cite{molkenthin2016scaling}, we have introduced a simplified, two-dimensional version of this model, which starts from a closed chain (i.e.
+a ring) of $L$ unit discs and subsequently adds links, such that connected discs touch, yet no discs intersect.
+The advantage of this model is that it can be approximated by a purely topological simplification, which can be treated analytically.
+Fig.~\ref{fig:schematic} illustrates the construction of the 2D geometric model.
+
+%The topological constraints in the resulting network model are:
+%(a) New links always form between two units that are part of the same face of the graph (region enclosed by a cycle in the network).
+This prevents the overlapping of discs.
+(b) No links form across the outer face.
+This prevents the enclosure of a unit by less than six other units (which is geometrically impossible) such that (c) the maximum degree of each unit is six, as six is the maximum number of unit discs one central unit disc can touch.
+(d) Once connected by a link, pairs of units do not disconnect.
 
@@
 The underlying formation mechanisms of individual protein folds are incredibly complex and depend on the surrounding solvent, as well as the specific amino acid sequence and their quantum-mechanical interactions.
-Here we want to assess if the geometrically constrained protein model's contact map distributions capture real protein behavior as suggested in~\cite{bartoli2008effecta}.
-To understand this better, we want to study the ensemble of PCMs from real protein structures.
-To this end we will look at `averaged' contact maps over many different proteins that have different amino acid sequences, but their overall sequence or chain length is the same.
-We will use the RCSB protein data bank (PDB), a database of structurally resolved protein amino acid sequences through X-ray crystallography, NMR or cryo-EM experiments and the synthetically generated structural data base from the machine learned structures of AlphaFold2~\cite{jumper2021highly}.
-The PDB contains just under 200,000 protein structures to date~\cite{PDB}.
-The distribution of amino acid chain lengths in the PDB is illustrated in Fig.~\ref{fig:pdb_stats} b.
-Protein chain lengths in the PDB vary from small fragments of less than 10 amino acids to large agglomerates of over 2000 amino acids.
+Here we want to assess if the geometrically constrained protein model's contact map distributions capture real protein behavior and make it a viable model for PCMs and provide a better explanation for the heuristic used in Bartoli \textit{et al.}~\cite{bartoli2008effecta}.
+To understand this better, we study the ensemble of PCMs from real protein structures and will look at `averaged' contact maps over many different proteins that have different amino acid sequences, but their overall sequence or chain length is the same.
+We will use the RCSB Protein Data Bank (RCSB PDB), a database of structurally resolved protein amino acid sequences through X-ray crystallography, NMR or cryo-EM experiments and the synthetically generated structural database from the machine learned structures from the AlphaFold~2 database~\cite{varadi2022alphafold}.
+The RCSB PDB contains just under 200,000 protein structures to date~\cite{berman2000protein}.
+The distribution of protein chain lengths in the RCSB PDB is illustrated in Fig.~\ref{fig:pdb_stats} b.
+These chain lengths vary from small fragments of less than 10 amino acids to large agglomerates of over 2000 amino acids.
 The most frequently occurring chain lengths, however, are between 100 and 300 amino acids as shown in Fig.~\ref{fig:pdb_stats} b.
-The pink subset of the barplot shown in Fig.~\ref{fig:pdb_stats} b, are the set of structures used that fulfill the criterion on amino acid chain lengths between 85 and 115 amino acids, 185 and 215 amino acids and 285 and 315 amino acids and represent all structures for the analysis.
-The second database consists of synthetically generated structures using AlphaFold2 with sequences taken from SwissProt~\cite{alpha fold 2 database}.
-SwissProt is a manually curated database of protein sequences containing over 500,000 protein sequences~\cite{SwissProt}, whose amino acid chain length distribution is shown in Fig.~\ref{fig:pdb_stats} a in blue-green.
-The pink bars again indicate the subset of AlphaFold2 structures used for the contact map analysis.
+The pink subset of the bar plot shown in Fig.~\ref{fig:pdb_stats} b, is the set of structures used that fulfill the criterion on protein chain lengths of $L\approx100$ ($85\le L\le 115$) amino acids, $L\approx 200$ ($185 \le L \le 215$) amino acids and $L\approx300$ ($285 \le L\le 315$) amino acids and represent all RCSB PDB structures used in our analysis.
+The second database consists of synthetically generated structures using AlphaFold~2 with sequences taken from SwissProt~\cite{bairoch2000swissprot}.
+SwissProt is a manually curated database of protein sequences containing over 500,000 protein sequences, whose protein chain length distribution is shown in Fig.~\ref{fig:pdb_stats} a in teal.
+The pink bars indicate the subset of AlphaFold~2 structures used for the contact map analysis.
+In this case, the initial structures were selected based on the condition of chain lengths in the same three intervals as used for the RCSB PDB.
+Second, the AlphaFold~2 per-residue confidence scores~\cite{jumper2021highly} were used to select proteins that had an average score of 90 or higher.
+Lastly, we removed structures that described the same protein but originated from different organisms, by selecting unique protein names from the pandas data table.
 
- \begin{figure}[t]
+ \begin{figure}[htb]
         \centering
-	\includegraphics[width=0.8\columnwidth]{paper/figures/Fig2/Fig2.pdf}
-	    \caption{Histogram of protein sequence lengths for protein structures from two structure databases.
-a) AlphaFold2 predicted protein structures ($\alpha$F) taken from all protein sequences in the reviewed Uniprot Knowledge Base (Swiss-Prot) database based on their Uniprot ids.
-Pink bars shows fraction of structures used for the average contact map analysis.
-b) RSCB PDB chain length frequencies based on sequence length.
-Pink bars in plot show proteins structures whose sequence lengths matches the resolution of structure accounting for the whole sequence.
-Only pink PDB samples were used for the average contact map analysis.
-c) Example of a protein contact map taken from PDB ID: 1AKI, with a cut-off of 8~Å.
+	\includegraphics[width=0.9\columnwidth]{paper/figures/Fig2/Fig2.pdf}
+	    \caption{Bar plot of protein sequence lengths for protein structures from two structure databases.
+a) AlphaFold~2 predicted protein structures taken from protein sequences in the manually curated SwissProt database of protein sequences.
+Pink bars show fraction of structures used for the average contact map analysis.
+b) RSCB PDB chain length frequencies.
+Pink bars show protein structures whose chain lengths matches the resolution of structure accounting for the whole sequence.
+Only pink RCSB PDB samples were used for the average contact map analysis.
+c) Example of a protein contact map taken from RCSB PDB ID: 3UFG, with a cut-off of 8~Å.
 Pink squares indicate a contact present between amino acids.}
@@
 \end{figure}
-All structure based contact maps are generated in the same way for AlphaFold2 structures and PDB structures.
+All structure-based PCMs were generated in the same way for AlphaFold~2 structures and RCSB PDB structures.
 Structures used were downloaded in pdb format, read in with MDAnalysis version 2.0.0~\cite{gowers2016mdanalysis} and only C$_{\alpha}$ were selected and used for the analysis.
-Proteins were categorised by length and placed into three groups.
-Amino acid chain lengths between 85 and 115 amino acids, 185-215 amino acids and 285-315.
-For each of the used sequences, all pairwise C$_{\alpha}$ distances were assessed and two C$_{\alpha}$ atoms were in contact if their distance was less than 8~Å.
-All pairwise distances were evaluated and in this way, building up the binary contact map or adjacency $A_{ij}^{\mathrm{struc}}$.
-All code for the construction of the contact maps both for the geometrically constrained system and structural protein data and their analysis can be found on github at: \url{https://github.com/meyresearch/sequence_distance_distribution}.
+Proteins were categorised by length and placed into the same three groups as for the RCSB PDB data.
+For each of the sequences, all pairwise C$_\alpha$ distances were calculated, and two atoms were said to be in contact if their distance was less than 8~Å.
+The contacts (1 or 0) were then recorded in the adjacency matrix $\mathbf{A}^{\mathrm{struc}}$.
+An example of a protein contact map, or adjacency matrix, is shown in Fig.~\ref{fig:pdb_stats}~c.
+In order to study the amino acid distance distributions of real and simulated protein structures, the distances were calculated from the adjacency matrices.
+These distances were then histogrammed to calculate the mean amino acid distance distributions shown in Sec.~\ref{sec:results}.
+The pseudoalgorithms describing the above two processes are shown in the Supplemental Material~(SI)~\cite{SI}.
+All code for the construction of the contact maps and their analysis is available on GitHub at~\cite{2022sequence}.
+The derived analytical approximation in Sec.~\ref{sec:theory}, was used for data fitting.
+All fits were carried out using Scipy 1.7.0 and with scripts for the fits available on GitHub at~\cite{2022sequence}.
 
@@
 %%%%%%%%%%%%
-\section{Analytic approximation for the geometrically constrained model}
+\section{Analytical approximation for the geometrically constrained model}\label{sec:theory}
 
-To assess how we can best describe the distribution of the probability of amino acid distances as obtained by a PCM for the geometrically constrained model, we propose the following analytical approximation.
-To be able to do this we will look at the amino acid distance distribution for the simplified 2D model.
+To assess how we can best describe the distribution of the probability of amino acid distances $P(s)$ obtained from a PCM for the geometrically constrained model, we consider the amino acid distance distribution for the simplified 2D model as proposed in~\cite{molkenthin2016scaling}.
+We are thus able to propose the following analytical approximation.
 Fig~\ref{fig:schematic} summarizes the idea behind this two-dimensional model.
@@
 Let us start by introducing the auxiliary variable $F_k(s)$, defined to be the number of possible links with an amino acid distance of $s$ if $k$ links have been added before.
-Before any links are added, all amino acid distances are equally likely, as there are $F_0(s)=N$ possibilities of making a link of each distance $s$ with $2\leq s < N/2$.
+Before any links are added, all amino acid distances are equally likely, as there are $F_0(s)=N$ possibilities of making a link of distance $s$ each, with $2\leq s < \frac{N}{2}$.
 
-As we add more links, not only are links taken out of this pool, because they have already been realized, \red{but} each existing link can also geometrically prohibit other connections.
-Adding a link of length $s_1$ prohibits $2(s-1)$ links of length $s<s_1$ and $2(s_1-1)$ links of length $s>s_1$ (see Fig.~\ref{fig:schematic}~b).
+As we add more links, not only are links taken out of this pool, because they have already been realized, but each existing link can also geometrically prohibit other connections.
+Adding a link of length $s_1$ prohibits $2(s-1)$ links of length $s<s_1$ and $2(s_1-1)$ links of length $s\geq s_1$ (see Fig.~\ref{fig:schematic}~b).
 To find the expected number of links still available in subsequent steps, we average over all possible amino acid distances $s_1$ of the first link added.
 
-The expected distribution of possible links after one step is \red{thus given} by the average over available link pools taken over all possible values of $s_1$
-\begin{equation}
-\begin{aligned}
-    F_1(s)&=\frac{1}{N/2-1} \sum_{s_1=2}^{N/2}{ \begin{cases}
+The expected distribution of possible links after one step is thus given by the average over available link pools taken over all possible values of $s_1$:
+
+\begin{align}
+    F_1(s)&=\frac{1}{\frac{N}{2}-1} \sum_{s_1=2}^{N/2}{ \begin{cases}
     F_0(s)-2(s-1) \text{ , for } s<s_1\\
     F_0(s)-2(s_1 -1)\text{ , for } s\geq s_1
-    \end{cases}}
-    \nonumber \\
+    \end{cases}}\nonumber\\
     &\approx\frac{2}{N} \sum_{s_1=2}^{N/2} { \begin{cases}
     N-2(s-1) \text{ , for } s<s_1\\
-    N-2(s_1 -1)\text{ , for } s\geq s_1
+    N-2(s_1 -1)\text{ , for } s\geq s_1.
     \end{cases}}
-\end{aligned}
 \label{eq:reduction_first_step}
-\end{equation}
+\end{align}
 
-After $k-1$ links were made, the probability that the $k$-th link, being randomly chosen from the pool of available links, has length $s$ obeys
+After $k-1$ links are made, the probability that the $k^{\mathrm{th}}$ link, being randomly chosen from the pool of available links, has length $s$ obeys
 \begin{equation}
-    \Pin^k(s)=F_k(s)/C_k,
+    \Pin^k(s)=\frac{F_k(s)}{C_k},
 \end{equation}
@@
 \end{equation}
-% \begin{equation}
-%     P_k(s)=\frac{F_k(s)}{C_k},
-%     \label{eq.Pk}
-% \end{equation}
-%with $C_k=\sum_{s=2}^{N/2}F_k(s)$, 
-This ignores any length-dependent bias in which links have been removed, however, examples suggest this assumption seemed to hold well. 
+
+This ignores any length-dependent bias in which links have been removed, however, examples suggest this assumption holds well. 
 This leads to the recursion
+
 \begin{align}
-    F_k(s)&= \sum_{s_k=2}^{N/2} \Pin^k(s_k)\nonumber \\
-    & \left(F_{k-1}(s)-  F_{k-1}(s)\frac{1}{N}
+    F_k(s)&= \sum_{s_k=2}^{N/2} \Pin^k(s_k)\nonumber \bigg(F_{k-1}(s) \\
+    & - F_{k-1}(s)\frac{1}{N}
     {\begin{cases}
@@
      2(s_k -1))\text{ , for } s\geq s_k
-    \end{cases}} \right)
+    \end{cases}}\bigg),
 \end{align}
-which generalizes equation~\ref{eq:reduction_first_step} and simplifies to
+
+which generalizes Eq.~\ref{eq:reduction_first_step} and simplifies to
 \begin{align}
-   F_k(s)&=F_{k-1}(s)(1-\frac{1}{N} \sum_{s_k=2}^{s} (2 (s_k-1) \Pin^k(s_k))\nonumber \\
-   &-\frac{1}{N} \sum_{s_k=s+1}^{N/2} (2 (s-1) \Pin^k(s_k)))\nonumber \\
-   &=F_{k-1}(s)(1-\frac{1}{N}
-   \sum_{s_k=2}^{s} (2 (s_k-1) \Pin^k(s_k))\nonumber \\
-   &-\frac{2 (s-1)}{N} (\sum_{s_k=2}^{N/2} \Pin^k(s_k) - \sum_{s_k=2}^{s} \Pin^k(s_k)))\nonumber \\
-   &=F_{k-1}(s)(1-\frac{2}{N} (s-1) 
+   F_k(s)&=F_{k-1}(s)\Big(1-\frac{1}{N} \sum_{s_k=2}^{s} 2 (s_k-1) \Pin^k(s_k)\nonumber \\
+   &-\frac{1}{N} \sum_{s_k=s+1}^{N/2} 2 (s-1) \Pin^k(s_k)\Big)\nonumber \\
+   &=F_{k-1}(s)\bigg(1-\frac{1}{N}
+   \sum_{s_k=2}^{s} 2 (s_k-1) \Pin^k(s_k)\nonumber \\
+   &-\frac{2 (s-1)}{N} \Big[\sum_{s_k=2}^{N/2} \Pin^k(s_k) - \sum_{s_k=2}^{s} \Pin^k(s_k)\Big]\bigg)\nonumber \\
+   &=F_{k-1}(s)\Big(1-\frac{2}{N} (s-1) 
    +\frac{2}{N} s
@@
    &-\frac{2}{N}
-   \sum_{s_k=2}^{s} s_k \Pin^k(s_k))
+   \sum_{s_k=2}^{s} s_k \Pin^k(s_k)\Big).
    \label{eq.Fk_rec}
@@
 Aiming to decouple those dynamics for different $s$, we make a heuristic guess for $\Pin^k(s_k)$.
-It consists of two separate approximation which are used in steps $k<a$ and steps $k\geq a$ respectively, where the threshold $1\leq a\leq N/2$ is a free parameter.
+It consists of two separate approximations which are used in steps $k<a$ and steps $k\geq a$ respectively, where the threshold $1\leq a\leq \frac{N}{2}$ is a free parameter.
 In the early stage we use a uniform probability
@@
 For later steps, the probability of a link to still be in the pool decreases with $s$.
-We thus approximate $\Pin(s)$ with an ansatz for the $k-$average, namely, that it drops off as $\frac{1}{s}$, leading to
+We thus approximate $\Pin(s)$ with an Ansatz for the $k-$average, namely, that it drops off as $s^{-1}$, leading to
 \begin{equation}
     \Pin^{k\gg 1}(s)\approx \sum_{k=a}^{N-4}\Pin^{k}(s)
-    \approx \frac{s^{-1}}{ (H_{\frac{N}{2}}-1)},
+    \approx \frac{s^{-1}}{ H_{\frac{N}{2}}-1},
     \label{eq.P_large}
@@
 where $H_{\frac{N}{2}}$ is the harmonic number serving as a normalization factor. 
-As we will see later this approximation holds most precisely for intermediate values of $k$, with the $k-$average of $\Pin^{k\gg 1}(s)\approx P(s)$, making it self-consistent.
+As we will see later, this approximation holds most precisely for intermediate values of $k$, with the $k-$average of $\Pin^{k\gg 1}(s)\approx P(s)$, making it self-consistent.
 Errors for smaller and larger values of $k$ are thought to approximately cancel out.
@@
 \begin{align}
-   F_{k\ll N/2}(s)&\approx 
-   F_{k-1}(s)(1-\frac{2}{N} (s-1) 
+   F_{k\ll \frac{N}{2}}(s)&\approx 
+   F_{k-1}(s)\Big(1-\frac{2}{N} (s-1) 
    +\frac{4}{N^2} s (s-1)\nonumber \\
-   &-\frac{2}{N^2}(s^2+s-2)\nonumber \\
+   &-\frac{2}{N^2}(s^2+s-2)\Big)\nonumber \\
    &=F_{k-1}(s)(1-\flow(s)),
@@
 \[\flow(s)=-\frac{2}{N^2}s^2+\left(\frac{2}{N}+\frac{6}{N^2}\right)s-\left(\frac{2}{N}+\frac{4}{N^2}\right),\]
+
 and
@@
 \begin{align}
-   F_{k\gg1}(s)&\approx 
-   F_{k-1}(s)(1-\frac{2}{N} (s-1) 
-   +\frac{2}{N} s \sum_{s_k=2}^{s}\frac{1}{s_k(H_{\frac{N}{2}}-1)}\nonumber\\
-   &-\frac{2}{N}
-   \sum_{s_k=2}^{s} s_k \frac{1}{s_k(H_{\frac{N}{2}}-1)})\nonumber \\
-   &=
-   F_{k-1}(s)(1 -(\frac{2}{N}-\frac{2}{N}\frac{H_{s}-1}{H_{\frac{N}{2}}-1}+\frac{2}{N(H_{\frac{N}{2}}-1)})s \nonumber \\ 
-   &+ \frac{2}{N(H_{\frac{N}{2}}-1)} + \frac{2}{N})\nonumber \\
+   F_{k\gg1}(s)&\approx F_{k-1}(s)\Big(1-\frac{2}{N} (s-1) \nonumber\\
+   &+\frac{2}{N} s \sum_{s_k=2}^{s}\frac{1}{s_k(H_{\frac{N}{2}}-1)}\nonumber\\
+   &-\frac{2}{N} \sum_{s_k=2}^{s} s_k \frac{1}{s_k(H_{\frac{N}{2}}-1)}\Big)\nonumber \\
+   &=F_{k-1}(s)\bigg[1 -\Big(\frac{2}{N}-\frac{2}{N}\frac{H_{s}-1}{H_{\frac{N}{2}}-1} \nonumber\\
+   &+\frac{2}{N(H_{\frac{N}{2}}-1)}\Big)s \nonumber + \frac{2}{N(H_{\frac{N}{2}}-1)} + \frac{2}{N}\bigg]\nonumber \\
    &=F_{k-1}(s)(1-\fhigh(s)).
@@
 where 
-\[\fhigh(s)=\frac{2}{N}(\frac{H_{\frac{N}{2}}-H_{s}+1}{H_{\frac{N}{2}}-1}s-\frac{H_{\frac{N}{2}}}{H_{\frac{N}{2}}-1}).\]
+\[\fhigh(s)=\frac{2}{N}\left(\frac{H_{\frac{N}{2}}-H_{s}+1}{H_{\frac{N}{2}}-1}s-\frac{H_{\frac{N}{2}}}{H_{\frac{N}{2}}-1}\right).\]
 
@@
     F_k(s)&=N\prod_{i_1}^a\left(1-\flow(s)\right)\prod_{i=a+1}^k\left(1-\fhigh(s)\right) \nonumber \\
-    &= N\left(\frac{1-\flow(s)}{1-\fhigh(s)}\right)^a\left(1-\fhigh(s)\right)^{k}
+    &= N\left(\frac{1-\flow(s)}{1-\fhigh(s)}\right)^a\left(1-\fhigh(s)\right)^{k}.
 \end{align}
 
-In 2D this is repeated for $N-3$ steps until no more links can be added \cite{molkenthin2016scaling}.
+In 2D, this is repeated for $N-3$ steps until no more links can be added \cite{molkenthin2016scaling}.
 The probability distribution of the realized amino acid distances of all added links is then given by the average over the available pools at each link addition step, leading to
 \begin{align}
-    P(s)&=\frac{1}{N-3}\sum_{k=0}^{N-4} \frac{F_k(s)}{C_k} \nonumber \\
-    &\approx \frac{N}{N-3}\left(\frac{1-\flow(s)}{1-\fhigh(s)}\right)^a
-\sum_{k=0}^{N-4} \frac{1}{\evdel{C_k}_k} \left(1-\fhigh(s)\right)^{k} \nonumber \\
-    &\approx \left(\frac{1-\flow(s)}{1-\fhigh(s)}\right)^a\frac{A}{\fhigh(s)},
+    P(s)&=\frac{1}{N-3}\sum_{k=0}^{N-4} \Pin^k(s) \nonumber \\
+    &\approx \frac{N}{N-3}\left(\frac{1-\flow(s)}{1-\fhigh(s)}\right)^a \nonumber\\
+&\times \sum_{k=0}^{N-4} \frac{1}{\evdel{C_k}_k} \left(1-\fhigh(s)\right)^{k} \nonumber \\
+    &\approx \left(\frac{1-\flow(s)}{1-\fhigh(s)}\right)^a\frac{\Gamma}{\fhigh(s)},
     \label{eq.Ps}
@@
 This approximation allows the use of the geometric series for solving the equation. 
-The resulting expression in Eq.~\ref{eq.Ps} is close to a power-law in $s$ with an exponent of \red{something close to -1.2 depending on a}, thus justifying the earlier approximation of $\Pin^k(s)$.
-%which makes use of Eq.~\ref{eq.Pk}. 
-%The only unknown in this expression is $C_k$, \red{which we can not compute exactly...?}.
-However, since $C_k$ monotonically decreases with $k$, $N>C_k>C_{N-2}$ always holds and we can obtain an upper and lower bound by replacing all $C_k$ in the expression with either $N$ (underestimates all probabilities) or $C_{N-2}$ (overestimates all probabilities).
-Each time we get an expression of the form
-% \begin{align}
-%      P(s)&\approx\frac{N}{N-3}\sum_{k=0}^{N-3} \frac{1}{C_k}\prod_{i}^k\left(1-\frac{f(s)}{(\frac{N}{2}-2)C} \right)\nonumber \\
-%      &=\frac{N}{N-3} \sum_{k=0}^{N-3}\frac{1}{C_k}\left(1-\frac{f(s)}{(\frac{N}{2}-2)C} \right)^k \nonumber \\
-%      &\approx \frac{A}{f(s)}
-% \end{align}
-% for $1<s<\frac{N}{2}$,
-% where $A$ is an unknown constant, containing an approximation of the various normalization constants $C_i$, as well as the direct effects of the chain length.
+The resulting expression in Eq.~\ref{eq.Ps} by visual inspection of Fig.\ref{fig:2d_sim} resembles a power law with an exponent of $-1$, thus justifying the earlier approximation of $\Pin^k(s)$.
 
-%In the last step of this approximation, we took the $C_k$ to all be approximately equal in order to apply the geometric series.
-However, in reality we know $1/C_k$ to increase with $k$, giving large $k$ summands relatively more weight than in a geometric series.
-This can be expressed as a small correction $\alpha$ in the exponent.
+ \begin{figure*}[htb]
+        \centering
+	\includegraphics[width=\textwidth]{paper/figures/Fig3/Fig3.pdf}
+	    \caption{The analytical approximation fitted to amino acid distance distributions for a) 2D and b) 3D simulated data.
+Pink area shows the 95\% confidence interval.
+c) Example of a PCM taken from one of the 3D simulation runs.
+The pink squares indicate a contact between two nodes.}
+        \label{fig:2d_sim}
+\end{figure*}
 
-\section{Results}
-\subsection{Analytic approximation and 2D simulations}
-Fig.~\ref{fig:2d_sim} compares Eq.~\ref{eq.Ps} with respect to the two-dimensional simulation data introduced in~\cite{molkenthin2020self}.
-The simulation data is generated from 10 repeats with chain lengths of up to 498 amino acids long.
-The simulation data is plotted showing the mean frequencies of each amino acid distance up to $N/2\sim250$.
-The shaded area represent the 95\% confidence interval calcualted from the 10 repeats.
-The theoretical approximation from Eq.~\ref{eq.Ps} was fitted to the simulation data.
-The two fitting parameters were chosen to be $a=6$ and $A=0.39$, to best capture the trend of the simulation data.
-The fitted approximation captures the behavior of the simulation data well, both describe two dimensional systems.
- \begin{figure}[tb]
+\section{Results and Discussion}\label{sec:results}
+The results analysis centers around understanding how the analytical approximation can be used to understand the simulations in 2D and 3D from~\cite{molkenthin2016scaling,molkenthin2020self}, and structural data from the RCSB PDB and AlphaFold~2 databases.
+To this end, Eq.~\ref{eq.Ps} is used as a fitting function in order to determine the two parameters $\Gamma$ and $a$.
+
+\subsection{The analytical approximation fits 2D and 3D simulation data}
+Fig.~\ref{fig:2d_sim} compares Eq.~\ref{eq.Ps} with respect to the two-dimensional (a) and three-dimensional (b) simulation data introduced in~\cite{molkenthin2016scaling} and ~\cite{molkenthin2020self}, respectively.
+Fig.~\ref{fig:2d_sim} c, shows an example of an adjacency matrix $\mathbf{A}^{\mathrm{sim}}$ as generated from a 3D simulation run.
+The 2D simulation data was generated from 10 repeats with chain lengths of 498 amino acids taken from~\cite{molkenthin2016scaling}.
+The 3D simulation data consists of adjacency matrices computed from 30 simulation runs, and was analyzed as was explained in Sec.~\ref{sec:methods}.
+The amino acid distance distributions are plotted showing the mean frequencies of each amino acid distance up to $\frac{L}{2}\sim250$ and $\frac{L}{2}\sim150$ for 2D and 3D simulations, respectively.
+The shaded areas represent the 95\% confidence intervals.
+The theoretical approximation from Eq.~\ref{eq.Ps} was fitted to both 2D and 3D simulation data separately.
+The two parameters, $a$ and $\Gamma$ in Eq.(\ref{eq.Ps}), are given in Table~\ref{table:sim-results}.
+The fitted approximation captures the behavior of the simulation data well for both 2D (Fig.~\ref{fig:2d_sim}a) and 3D (Fig.~\ref{fig:2d_sim}b) simulations.
+
+\begin{table}[htb]
 \centering
-	\includegraphics[width=0.45\textwidth]{paper/figures/Fig3/Fig3.pdf}
-	    \caption{Amino acid distance distribution for two-dimensional simulated data plotted against the analytic approximation. Green area shows the 95\% confidence interval.}
-        \label{fig:2d_sim}
-\end{figure}
+\setlength{\tabcolsep}{25pt}
+% \resizebox{\columnwidth}{!}{%
+\begin{tabular}{ccc}
+\hline
+Simulation & a &$\Gamma\; (10^{-3})$ \\
+\hline
+2D& $6 \pm 1$ & $0.3 \pm 0.1$ \\
+3D 300& $1 \pm 1$ & $7.0 \pm 0.2$ \\
+\hline
+\end{tabular}%
+% }
+\caption{Fit parameters from fitting the analytical approximation to the 2D simulated data as well as the 3D simulated data in the $L\approx 300$ chain length range.}
+\label{table:sim-results}
+\end{table}
 
-\subsection{3D simulations and PCMs from structural data }
-Next, we investigate if and how the analytical approximation can also capture trends in 3-D simulation data, as well as experimentally determined protein structures.
-This seeks to answer the question if indeed a power law can be used to describe the probability distribution of amino acid distances.
-Fig.~\ref{fig:sdd} summarises the results for protein chain lengths of $\approx$100 (a), $\approx$200, and $\approx$ 300 amino acids, protein chain lengths, most commonly found in living organism (see.
-Fig.~\ref{fig:pdb_stats} a).
-We are comparing data from repeats of 30 simulation runs of the three-dimensional simulations, as introduced in~\cite{molkenthin2020self}, and protein contact maps from RCSB~\cite{}, and AlphaFold2~\cite{}.
+\subsection{$P(s)$ from RCSB PDB and AlphaFold~2 agree and cover similar secondary structure content}
+As described in Sec.~\ref{sec:methods}, we have chosen a subset of structures from the RCSB PDB and AlphaFold~2 database for the analysis.
+The number of structures in each chain length range for both databases are shown in Table~\ref{table:subsets}.
+All structures were converted into PCMs, as described in Section~\ref{sec:methods}.
+From the PCMs we computed the amino acid distance distribution in the same way as was done for the 3D simulations.
+Since AlphaFold~2 generates structures from an AI model, we wanted to understand if the distributions generated from it and RCSB PDB at lengths $L\approx 100$, $L \approx 200$ and $L \approx 300$ are statistically the same.
+To this end, we use the non-parametric test of the Kolmogorov-Smirnov (KS) statistic to compare the probability distributions of AlphaFold~2 and RCSB PDB for each of the lengths~\cite{2008kolmogorov}.
+The KS statistic for the lengths of $L \approx 100$ $L \approx 200$, and $L \approx 300$ amino acids was 0.07, 0.10, and 0.12 respectively.
+Therefore, the null hypotheses for the distributions being the same was accepted at the $\alpha=0.01$ level.
+For more details see the SI~\cite{SI}.
 
-We have chosen a subset of structures from the PDB and curated alphaFold2 database for the analysis.
-7,376 structures with chain lengths that fall between 85-100, 6,822 structures that fall between 185-215, and 4,160 structures that fall between 285-315 amino acids were selected from the RCSB database~\cite{PDB}.
-Additionally, 33,229 structures with chain lengths in the range of 85-115, 37,430 in the range of 185-215 and 32,137 in the range of 285-315 amino acids were chosen from the AlphaFold2 database~\cite{}.
-All structures were converted into PCMs as described in section~\ref{sec:methods}.
-The resulting distributions for each chain length interval are plotted logarithmically scaled in Fig.\ref{fig:sdd} for intervals 85-100 in a, 185-215 in b, and 285-315 in c. 
-For all three lengths, the distributions show an approximate power law decay, for both the 3-D simulations according to~\cite{molkenthin2020self}, as well as the structural data.
-This allows a fit  which is consistent with both the simulated results and the theoretical approximation.
+\begin{table}[htb]
+\centering
+\setlength{\tabcolsep}{5pt}
+\resizebox{\columnwidth}{!}{
+\begin{tabular}{ccc}
+\hline
+Protein chain length $L$ & RCSB PDB & AlphaFold 2 \\
+\hline
+100 & 12,206 & 4,396 \\
+200 & 11,031 & 7,053 \\
+300 & 10,091 & 6,231 \\
+\hline
+\end{tabular}
+}
+ \caption{The number of structures used for the analysis from the RCSB PDB and AlphaFold 2 database, after filtering.}
+\label{table:subsets}
+\end{table}
 
-is compared to those generated in simulations in~\cite{molkenthin2020self} as well as the above derived theoretical expression scaled with a factor to account for the different dimensionality, i.e.
-constant $A$, as well as approximations made in the derivation.
+\subsection{The analytical approximation can describe the amino acid distance distributions of the structural data}
+Next, we investigated if the amino acid distance distribution from the structural data (RCSB PDB and AlphaFold~2) can be modeled by the analytical approximation.
+In order to answer this, we look at the scaling of the the distribution of RCSB PDB data in the first instance.
+We will present the results for the AlphaFold~2 data in the SI, since the KS statistics suggests that the RCSB PDB distributions and AlphaFold~2 of $P(s)$ agree.
+As a result, we observe similar trends.
 
-\begin{figure}
+\begin{figure*}[htb]
         \centering
-	\includegraphics[scale=0.65]{paper/figures/Fig4/Fig4.pdf}
-        \caption{Amino acid distance distributions for contact maps derived from the RCSB~\cite{PDB} and AlphaFold2~\cite{} databases as well as 3D simulated folds for protein chains in the range of a) 85-115,  b) 185-215 and c) 285-315 amino acids long.
-        }
+	\includegraphics[width=\textwidth]{paper/figures/Fig4/Fig4.pdf}
+        \caption{a) Amino acid distance distributions for contact maps derived from the RCSB PDB database for protein chains in the ranges of $L\approx100$, $L\approx200$ and $L\approx300$ amino acids.
+b) Comparison between the distributions of RCSB PDB and AlphaFold~2 database for protein chains in the $L\approx300$ range.
+The analytical approximation (solid line) and power law (dashed line) are fitted to the tail of the RCSB data.
+Pink shaded area shows the 95\% confidence level.
+c) Secondary structure content distribution in the RCSB PDB and AlphaFold~2 structures in the $L\approx300$ chain length range.}
         \label{fig:sdd}
-\end{figure}
-For large values of the chain length $N$, we thus find an approximate power law probability distribution for the amino acid distance with an exponent of $-1+\alpha$.
+\end{figure*}
 
-We observe that, while the general shape seems to match, in real proteins amino acid distances of two to four amino acids are slightly over-represented (see Fig.\ref{fig:sdd}).
-The over-representation of very short amino acid distances is an artefact of the different network construction for simulated (and theoretically approximated) networks and PCMs.
-While in the simulation all links are equally long and links are made until no longer possible, the parameter, determining the connectivity in PCMs is the cutoff distance $d_c$\blue{(mention used value?)}, which is chosen such that the number of links in the PCM is comparable to those in a simulated network of the same $N$.
-This leads to a larger range of bond angles randomly resulting in connections for small $s$ and thus an over-representation of those links.
+Fig.~\ref{fig:sdd} a shows the distributions of the amino acid distances for RCSB PDB logarithmically scaled for the different protein chain lengths of $L\approx100$ (light purple circles), $L\approx200$ (medium purple squares) and $L\approx300$ (palatinate triangles).
+For all three lengths, the distributions show an approximate power law decay in their tail, which is consistent with the simulated results.
 
-Here we use 30 runs of three-dimensional simulations, as introduced in~\cite{molkenthin2020self}, and networks extracted from measured protein data from both the RCSB~\cite{PDB} and AlphaFold2~\cite{} databases to compare this behavior with the more realistic 3D settings and show that the basic principle persists, indicating that much of the amino acid distance distribution can be attributed to generic geometric constraints.
+In addition to a general power law, we fit the analytical approximation to the RCSB PDB amino acid distance distribution in Fig.~\ref{fig:sdd} b in the $L\approx300$ range.
+We also show the AlphaFold~2 amino acid distance distribution for comparison.
+The theoretical and power law fits are shown in black solid and dotted lines, respectively.
+The parameters given by the analytical approximation are $a=4 \pm 1$ and $\Gamma=(7.4 \pm 0.8)\cdot 10^{-4}$.
+As for the power law, the exponent is $\gamma=1.45 \pm 0.02$.
+The shaded area shows the 95\% confidence level (C.L.).
+Similar fits for $L\approx100$ and $L\approx200$ are presented in the SI.
+
+We observe that, for intermediate values of $s$, i.e.
+$30\lesssim s\lesssim \frac{L}{2}$, the distributions are described well by the analytical approximation.
+However, the curves deviate for amino acid distances in the $5\lesssim s \lesssim 30$ range (see Fig.\ref{fig:sdd} b).
+Comparing this to the simulation distributions and their theoretical fits (Fig~\ref{fig:2d_sim}), we see an under-representation of amino acid distances calculated from PCMs from RCSB PDB data.
+
+One possible explanation of the under-representation can stem from secondary structure properties, as it appears in a specific distance range, where contacts arise from the proteins' secondary structure elements.
+In order to try to explore this feature, Fig.~\ref{fig:sdd} c shows the secondary structure content distributions for proteins in the $L\approx300$ chain length range from RCSB PDB and AlphaFold~2.
+Proteins mostly contain secondary structures in the combined $\alpha$-helix---$\beta$-sheet region.
+For protein chain length regions of $L\approx100$ and $L\approx200$ proteins with predominently $\alpha$ and $\beta$-sheet regions exist, as shown in the SI~\cite{SI}.
+However, in all data, a high percentage of mixed secondary structure is observed.
+This is also in agreement with~\cite{michie1996analysis}, suggesting that any secondary structure features should be evened out.
+
+Next, we looked at changing the threshold value $d_c$ from the original $8 \; \mathrm{\AA}$ to 10, 15 and 21 $\mathrm{\AA}$, see Fig.
+4 in the SI~\cite{SI}.
+For 10 $\mathrm{\AA}$, the shape of the distribution was similar to that of in Fig.~\ref{fig:sdd}~b, but for 15 $\mathrm{\AA}$ the under-representation starts to disappear slightly whereas for 21 $\mathrm{\AA}$ it disappears almost completely.
+Due to the high number of these mixed $\alpha$-helix---$\beta$-sheet secondary structures, it is likely that with a threshold of 8 $\mathrm{\AA}$ some inter-secondary structure contacts are being excluded, e.g.
+for large $\beta$-sheet content.
+This is a likely cause for the under-representation of amino acid distances in the $5 \lesssim s \lesssim 30$ region.
+Comparing to simulations where this dip is not observed, links are made until no longer possible, whereas in the PCMs, the connectivity is determined by the threshold value.
+Therefore it is not surprising this feature cannot be captured in the simulations.
+
 \section{Conclusion}
-The tertiary structures of folded proteins have long been one of the most important mysteries of Biophysics.
-While for individual structures, detailed molecular dynamics or statistical models are essential in approaching these questions, general statements for the ensemble of folded proteins can be made with simple models.
+The tertiary structures of folded proteins have long been one of the most important mysteries of biophysics.
+While for individual structures, detailed molecular dynamics or statistical models are essential in approaching these questions, general statements for the ensemble of folded proteins can be made with much simpler models.
 
 Here we have introduced and analyzed the amino acid distance $s$ and its probability distribution $P(s)$ in measured and simulated protein structures.
-For the equivalent model in two dimensions we have used a mean-field approach to derive an analytic approximation for the amino acid distance distribution, which appears to agree with the simulated and measured distributions.
-We have thus demonstrated in analytical calculations as well as direct simulations how a random linking model with geometric constraints, such as the ones introduced in~\cite{molkenthin2016scaling, molkenthin2020self}, generates a amino acid distance distribution that resembles a power-law, similar to the one found in folded proteins.
-This feature of protein network structures has previously been used in a heuristic manner to model and generate protein-like structures in \cite{bartoli2008effecta}.
+For the equivalent model in 2D, we have used a mean-field approach to derive an analytical approximation for the amino acid distance distribution, and we show its agreement with the simulated and measured distributions from real proteins.
+Therefore we managed to demonstrate that the geometrically constrained model's sequence distance distribution $P(s)$ match real world data well.
+In addition, the derived analytical approximation could serve as a good basis to model and generate protein-like adjacency matrices as was done in ~\cite{bartoli2008effecta}.
+It also highlights that the proposed heuristic of $P(s)\approx s^{-1}$, is a good starting point, but with an  approximation from Eq.~\ref{eq.Ps}.
 
@@
 \subsection{Availability of data and materials}
-All data generated or analysed during this study are included in this published article [and its supplementary information files].
+All data generated or analyzed during this study are included in this published article.
+All raw data and their analysis can be found on GitHub at~\cite{2022sequence}, including information required for their reproduction.
 \subsection{Competing interests}
@@
 \subsection{Funding}
-This research was supported by ...
+This research was supported by a Chancellor's Fellowship from the University of Edinburgh. 
 \subsection{Authors' contributions}
-
+N.M.
+and J.J.G.
+contributed equally to this work.
+A.M.
+and N.M.
+designed the study.
+N.M.
+and S.M.
+did the analytical approximation.
+J.J.G.
+gathered data and analyzed PCMs from the AlphaFold 2 database and RCSB PDB.
+Everyone wrote the manuscript.
 \subsection{Acknowledgements}
-We thank Marc Timme for fruitful discussions.
+We thank Marc Timme, and Matteo Degiacomi for fruitful discussions.
 

### 2022-05-23 paper/supplemental_material.tex

%\documentclass{article}
%\usepackage[a4paper, total={6in, 8in}]{geometry}
%\documentclass[reprint,amsmath,amssymb,rmp,onecolumn,notitlepage,11pt]{revtex4-1}
\documentclass[notitlepage,
reprint,%twocolumn,
%preprint
onecolumn,
amsmath,amssymb,superscriptaddress,aps,
pre,floatfix]{revtex4-1}
\usepackage[utf8]{inputenc}
%\usepackage{authblk}
%\usepackage{natbib}
\usepackage[normalem]{ulem}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{xcolor}
\newcommand{\red}[1]{\textcolor{red!80!black}{#1}}
\newcommand{\blue}[1]{\textcolor{blue!80!black}{#1}}
\newcommand{\green}[1]{\textcolor{green!70!black}{#1}}
\newcommand{\gray}[1]{\textcolor{gray!80!black}{#1}}
\usepackage{mathtools}
\DeclarePairedDelimiter{\evdel}{\langle}{\rangle}
\usepackage[colorinlistoftodos]{todonotes}
\newcommand{\Pin}{P_{\mathrm{in}}}
\newcommand{\fhigh}{f_{\mathrm{high}}}
\newcommand{\flow}{f_{\mathrm{low}}}
\setlength {\marginparwidth }{2cm}

\begin{document}
\title{Supplemental material accompanying "What geometrically constrained protein models can tell us about real-world protein contact maps"}
\author{Nora Molkenthin}
\email[Author to whom correspondence should be addressed:]{nora.molkenthin@pik-potsdam.de}
\affiliation{Potsdam Institute for Climate Impact Research, Germany}
\author{J. J.  Güven}
\affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
\author{Steffen Mühle}
\affiliation{Max-Planck Institute for Dynamics and Self-Organization (MPIDS), Am Faßberg 17, 37077 Göttingen, Germany}
\author{Antonia S J S Mey}
\email[Author to whom correspondence should be addressed:]{antonia.mey@ed.ac.uk}
\affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
\maketitle

\section*{Pseudoalgorithms for creating protein contact maps from protein data}

Algorithm~\ref{alg:pcm} describes how we construct PCMs from structural pdb files. We loop through all files and find all C$_\alpha$ atoms from each structure to calculate the pairwise C$_\alpha$ distances. If the distance is less than 8~Å, a contact ($A_{ij}=1$) is recorded in the adjacency matrix. Similarly, Algorithm~\ref{alg:dist} describes how we compute the amino acid distances from amino acids who are in close contact.

\begin{algorithm}[htb]
\caption{Creating PCMs from pdb structure files}\label{alg:pcm}
\begin{algorithmic}[1]
\For{\texttt{structure} in \texttt{structure}}\Comment{Read each pdb file}
\For{C$_\alpha^{(i)}$ and C$_\alpha^{(j)}$ in \texttt{pdb}}\Comment{Get all C$_\alpha$}
\State Compute $d_{ij}=|d_i-d_j|$ \Comment{Distance in Å}
\If{$d_{ij}\leq 8\;$Å}
    \State \texttt{$A_{ij}=1$} \Comment{Contact}
\Else
    \State \texttt{$A_{ij}=0$} \Comment{No contact}
\EndIf
\EndFor
\EndFor
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[htb]
\caption{Amino acid distances from PCMs}\label{alg:dist}
\begin{algorithmic}[1]
\For{i in $\mathbf{A}^{\mathrm{struc}}$}\Comment{Loop over rows}
\For{j in $\mathbf{A}^{\mathrm{struc}}$}\Comment{Loop over columns}
\If{$A_{ij}=1$} \Comment{If contact, get distance}
    \State s = $|i - j|$
\EndIf
\EndFor
\EndFor
\end{algorithmic}
\end{algorithm}

\section*{Amino acid distance distributions of other chain lengths}
\subsection{AlphaFold 2}
Fig.~\ref{fig:alphafold} shows the amino acid distance distributions for protein contact maps calculated from AlphaFold~2 data in the chain length ranges of $L \approx 100$ amino acids ($85 \leq L \leq 115$, light green circles), $L \approx 200$ amino acids ($185 \leq L \leq 215$, medium green squares) and $L \approx 300$ amino acids ($285 \leq L \leq 315$, dark green triangles). The plot shows the approximate power law behaviour of the distribution.

\begin{figure*}[htb]
    \centering{
    \includegraphics[width=\columnwidth]{paper/figures/SM_figures/alpha_chain_comparison.pdf}
    \caption{Amino acid distance distributions for contact maps derived from the AlphaFold 2 database for protein chains in the ranges of $L \approx 100$ amino acids ($85 \leq L \leq 115$, light green circles), $L \approx 200$ amino acids ($185 \leq L \leq 215$, medium green squares) and $L \approx 300$ amino acids ($285 \leq L \leq 315$, dark green triangles).}
    \label{fig:alphafold}}
\end{figure*}

\subsection{RCSB PDB and three-dimensional simulations}

Fig.~\ref{fig:rcsb_combined} a shows the amino acid distance distributions for protein contact maps calculated from structures in the chain length range $L\approx100$ amino acids for RCSB PDB (dark purple circles) and AlphaFold~2 (dark green squares). The shaded area shows the 95\% confidence level. Also shown are the analytic approximation and a general power law fitted to the RCSB data and their fit parameters, $a$ and $\Gamma$ for the analytical approximation and exponent $\gamma$ for the power law,  are shown in~Table~\ref{table:stats}. Similarly, Fig.~\ref{fig:rcsb_combined} shows the amino acid distance distributions for protein contact maps calculated from structures in the chain length range $L\approx200$ amino acids for RCSB PDB. We also computed the Kolmogorov-Smirnov (KS) statistic between the RCSB PDB and AlphaFold 2 distributions for each length range. The KS statistics were 0.07 and 0.10 for the ranges $L\approx100$ amino acids and $L\approx200$ amino acids, respectively. For both of these, the null hypothesis was accepted at the $\alpha=0.01$ level.

Fig.~\ref{fig:sim_3d} shows the amino acid distance distributions for protein contact maps calculated from the three-dimensional simulations. As before, the analytic approximation was fitted to the data in each range and the shaded area shows the 95\% confidence interval. The fit parameters $a$ and $\Gamma$ are given in Table~\ref{table:stats} for ranges $L\approx100$ amino acids,  $L\approx200$ amino acids and $L\approx300$ amino acids.

Finally, Fig.~\ref{fig:dc} shows the effect of changing the threshold value for determining contacts between amino acids for contact maps calculated from RCSB PDB, for a)  $L\approx100$ amino acids,  b) $L\approx200$ amino acids and c) $L\approx300$ amino acids.

% \begin{table}[tb]
% \centering
% \setlength{\tabcolsep}{10pt}
% \begin{tabular}{ c c c c}
% \hline
%  Data  & $\Gamma$ ($10^{-3}$) & $a$ & $\gamma$ \\
% \hline
%  RCSB 100 & 1.57 $\pm$ 0.51  & 2 $\pm$ 1 &  0.88 $\pm$ 0.02\\
%  RCSB 200 & 0.81 $\pm$ 0.19 & 2 $\pm$ 1 & 0.95 $\pm$ 0.02 \\
%  RCSB 300 & 0.74 $\pm$ 0.08 & 4 $\pm$ 1 & 1.45 $\pm$ 0.02 \\
%  3D SIM 100 & 3.1 $\pm$ 0.6 & 2 $\pm$ 1 & --  \\
%  3D SIM 200 & 1.2 $\pm$ 0.4 & 1 $\pm$ 1  & --  \\
%  3D SIM 300 & 7.0 $\pm$ 0.2 & 1 $\pm$ 1 & -- \\
%  2D SIM & 0.3 $\pm$ 0.1 & 6 $\pm$ 1 & -- \\
% \hline
% \end{tabular}
%  \caption{Parameters $a$ and $\Gamma$ for the analytical approximation, exponent $\gamma$ for the power law fit, and their uncertainties from fitting to amino acid distributions obtained from RCSB proteins in the chain length range $L\approx100$  and $L\approx200$ amino acids. Shown also are the fit parameters $a$ and $\Gamma$ from fitting the analytic approximation to the 3D simulations of chains in the $L\approx100$ and $L\approx200$ ranges.}
% \label{table:stats}
% \end{table}

\begin{table}[tb] 
\centering
\setlength{\tabcolsep}{10pt}
{\renewcommand{\arraystretch}{1.5}
\begin{tabular}{c c c c c c c c}
\hline
\multirow{2}{*}{Parameter}&\multicolumn{3}{c}{RCSB}& \multicolumn{3}{c}{3D Simulation}&\multirow{2}{*}{2D Simulation}
\\&100&200&300&100&200&300&\\
\hline
$\Gamma\; (10^{-3})$ & $1.57 \pm 0.51$ & $0.81 \pm 0.19 $ & $0.74 \pm 0.08$ & $3.1 \pm 0.6$ & $1.2 \pm 0.4$ & $7.0 \pm 0.2$ & $0.3 \pm 0.1$ \\
$a$ & $2 \pm 1$ & $2 \pm 1$ & $4 \pm 1$ & $2 \pm 1$ & $1 \pm 1$ & $1 \pm 1$ & $6 \pm 1$ \\
$\gamma$ & $0.88 \pm 0.02$ & $0.95 \pm 0.02$ & $1.45 \pm 0.02$ & - & - & - & - \\
 \hline
\end{tabular}
}
 \caption{Parameters $a$ and $\Gamma$ for the analytical approximation, exponent $\gamma$ for the power law fit, and their uncertainties from fitting to amino acid distributions obtained from RCSB PDB structures in ranges $L\approx100$ amino acids, $L\approx200$ amino acids and $L\approx300$ amino acids. Shown also are the fit parameters $a$ and $\Gamma$ from fitting the analytic approximation to the 2D simulation and 3D simulations of chains in the $L\approx100$, $L\approx200$ and $L\approx300$ ranges.}
\label{table:stats}
\end{table}

\begin{figure*}[htb]
    \centering{
    \includegraphics[width=\columnwidth]{paper/figures/SM_figures/distributions_combined.pdf}
    \caption{Comparison between the distributions of RCSB PDB and AlphaFold 2 database. The analytic approximation (solid line) and power law (dashed line) are fitted to the the RCSB PDB data. Pink shaded area shows the 95\% confidence level. The comparison is shown for two protein chain length ranges: a) $L\approx 100$ and b) $L\approx200$ range. Secondary structure content distribution in the RCSB and AlphaFold 2 structures in the  c) $L\approx100$ and the d) $L\approx200$ chain length range.}
    \label{fig:rcsb_combined}}
\end{figure*}

% \begin{figure*}[htb]
%     \centering{
%     \includegraphics[scale=0.35]{paper/figures/SM_figures/distributions_200.pdf}
%     \caption{Comparison between the distributions of RCSB~\cite{PDB} and AlphaFold 2~\cite{jumper2021highly} databases for protein chains in the $L\approx200$ range. The analytic approximation (solid line) and power law (dashed line) are fitted to the tail of the RCSB data. Pink shaded area shows the 95\% confidence level.}
%     \label{fig:rcsb_200}}
% \end{figure*}

\begin{figure*}[htb]
    \centering{
    \includegraphics[width=\columnwidth]{paper/figures/SM_figures/simulations_3d.pdf}
    \caption{Amino acid distance distributions obtained from three-dimensional simulations for chains in the ranges a) $L\approx100$  and b) $\approx200$ amino acids. The dotted line shows the analytical approximation fitted to the data and the shaded area shows the 95\% confidence level.}
    \label{fig:sim_3d}}
\end{figure*}

\begin{figure*}[htb]
    \centering{
    \includegraphics[width=\columnwidth]{paper/figures/SM_figures/threshold_combined.pdf}
    \caption{Amino acid distance distributions from RCSB PDB and 3D simulations for a) $L\approx100$, b) $L\approx200$ and c) $L\approx300$. Figure shows the effect of changing the threshold value $d_c$ from the default 8~Å (dark purple squares) to 10 (light purple upwards-triangles), 15 (medium purple downwards-triangles) and 21~Å (green crosses).}
    \label{fig:dc}}
\end{figure*}

\end{document}

### 2022-12-29 paper/main.tex
Updates from Overleaf

@@
 %\documentclass[reprint,amsmath,amssymb,rmp,onecolumn,notitlepage,11pt]{revtex4-1}
-\documentclass[
-reprint,
-twocolumn,
-%preprint
-%onecolumn,
-amsmath,amssymb,superscriptaddress,aps,
-pre]{revtex4-1}
+%\documentclass[reprint,twocolumn,amsmath,amssymb,superscriptaddress,aps,pre]{revtex4-1}
+\documentclass[10pt]{iopart}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{xcolor}
+\usepackage{iopams}
 \newcommand{\red}[1]{\textcolor{red!80!black}{#1}}
@@
 \begin{document}
-\title{Chain separation distribution in protein contact networks}
-\title{Theoretical approximation of chain separation distribution in protein contact networks}
-\title{What geometrically constrained protein models can tell us about real-world protein contact maps}
-%\title{}
-\author{Nora Molkenthin}
-\email[Author to whom correspondence should be addressed:]{nora.molkenthin@pik-potsdam.de}
-\affiliation{Potsdam Institute for Climate Impact Research, Germany}
-\author{J. J. Güven}
-\affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
-\author{Steffen Mühle}
-\affiliation{Max-Planck Institute for Dynamics and Self-Organization (MPIDS), Am Faßberg 17, 37077 Göttingen, Germany}
-\author{Antonia S J S Mey}
-\email[Author to whom correspondence should be addressed:]{antonia.mey@ed.ac.uk}
-\affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
+
+\title[Geometrically constrained folding models]{What geometrically constrained folding models can tell us about real-world protein contact maps}
+\author{Nora Molkenthin$^{1*\dagger}$, J. Jasmin Güven$^{2*}$, Steffen Mühle$^{3}$, Antonia S. J. S. Mey$^{2\dagger}$}
+\address{$^1$Potsdam Institute for Climate Impact Research, Germany}
+\address{$^2$EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
+\address{$^3$Max-Planck Institute for Dynamics and Self-Organization (MPIDS), Am Faßberg 17, 37077 Göttingen, Germany}
+\address{$^*$ These authors contributed equally}
+\address{$^\dagger$ To whom correspondence should be addressed}
+\ead{nora.molkenthin@pik-potsdam.de,antonia.mey@ed.ac.uk}
 
@@
 The mechanisms by which a protein's 3D structure can be determined based on its amino acid sequence have long been one of the key mysteries of biophysics.
-Often simplistic models, e.g.
-derived from geometric constraints, capture bulk real-world 3D protein--protein properties well.
-One approach is using protein contact maps to better understand protein's properties.
-Here, we investigate the emergent behavior of protein contact maps derived from a geometrically constrained random interaction model in comparison to real-world proteins.
-Deriving an analytical approximation for the distribution of amino acid distances $s$ of a 2D geometric model, by means of a mean-field approach, we present a theoretical basis for the distribution of amino acid distances in proteins.
-To validate this approximation we compare it to simulations of a 2D and 3D geometrically constrained model, as well as contact maps of real proteins.
-Using data from the RCSB Protein Data Bank (PDB) and AlphaFold~2 database, the analytical approximation can be accurately fit to both datasets.
-This holds for protein chain lengths of $L\approx100$, $L\approx200$, and $L\approx300$.
-While, a universal scaling behavior for the different chain lengths could not be deduced, it seems that the amino acid distance distributions can be attributed to geometric constraints of protein chains in bulk and amino acid sequences only play a secondary role.
+Often simplistic models, such as those derived from geometric constraints, capture bulk real-world 3D protein-protein properties well.
+One approach is using protein contact maps to better understand proteins' properties.
+Here, we investigate the emergent behaviour of contact maps for different geometrically constrained models and real-world protein systems.
+We derive an analytical approximation for the distribution of model amino acid distances, $s$, by means of a mean-field approach.
+This approximation is then validated for simulations using a 2D and 3D random interaction model, as well as from contact maps of real-world protein data.
+Using data from the RCSB Protein Data Bank (PDB) and AlphaFold~2 database, the analytical approximation is fitted to protein chain lengths of $L\approx100$, $L\approx200$, and $L\approx300$.
+While a universal scaling behaviour for protein chains of different lengths could not be deduced, we present evidence that the amino acid distance distributions can be attributed to geometric constraints of protein chains in bulk and amino acid sequences only play a secondary role.
 
 \end{abstract}
+\noindent{\it Biophysics, Geometric constraints, Protein folding\/}
 \maketitle
-
 \section{Introduction}
@@
 antibodies), and hormonal regulation (e.g.
-insulin)~\cite{dill2008protein,Dill1042}.
+insulin)~\cite{dill2008protein, Dill1042}.
 All natural proteins are made of 20 different amino acids which dictate the 3D conformations the proteins adopt in order to function~\cite{scheraga2007proteinfolding}.
@@
 The RCSB PDB contains structures from crystallographic, NMR, and cryo-EM experiments.
-AlphaFold~2 certainly was a game changer in protein structure prediction~\cite{callaway2022what}, but it has not solved the protein folding problem in the sense that the mechanisms of the folding or structure function relationship remain unexplained.
-However, it does provide a rich resource of structures to understand emergent patterns in proteins on a more fundamental level.
+AlphaFold~2 enabled many new avenues of research that require good protein structure prediction~\cite{callaway2022what}, but it has not solved the mechanisms by which proteins fold yet.
+A clear step change is that it does provide a rich resource of structures to understand emergent patterns in proteins on a more fundamental level.
 
-One way of mapping the functional forms of proteins and classifying their structural properties is by looking at protein contact maps (PCM)~\cite{Vendruscolo2002,dipaola2013protein,Estrada2011}.
+One way of mapping the functional forms of proteins and classifying their structural properties is by studying protein contact maps (PCMs)~\cite{Vendruscolo2002,dipaola2013protein,Estrada2011}.
 A PCM is a matrix representation of spatially close amino acids, given a certain cut-off distance as determined typically from a protein structure.
@@
 For example, similarly to AlphaFold~2, one challenge is to predict a PCM \textit{de novo} from sequence alone~\cite{bassot2019using, rives2021biological, rao2021msa}, without using structural data as part of the training set.
-Another is the use PCMs in machine learning approaches to e.g.
+Another is the use of PCMs in machine learning approaches to e.g.
 train a model able to predict binding affinities between a protein and a drug-like molecule~\cite{jiang2020drug}.
 
-From a physicist's perspective, it is interesting to understand the emergent behavior, in terms of folding or function, in many protein structures.
-Therefore studying properties of underlying features of a PCM can reveal information on the scale of the proteome and not just individual proteins.
-Thus, having a reliable model to generate a PCM not from real data is invaluable.
-To this end, different heuristic models have been indtroduced in the past by Atilgan \textit{et al.}~\cite{atilgan2004smallworld} and Bartoli \textit{et al.}~\cite{bartoli2008effecta}.
+From a physicist's perspective, it is interesting to understand the emergent behaviour, in terms of folding or function, in the ensemble of all protein structures.
+The underlying features of a PCM can reveal information on the scale of the proteome rather than the individual protein and therefore, having a reliable model to generate the spatial structure hidden beneath a PCM is invaluable.
+As such, we consider the set of all proteins of a given length and try to find general principles to allow the characterization of this set.
+Effectively, this corresponds to taking an average over all proteins of a given length.
+This approach was already taken in~\cite{molkenthin2020self}, where it was shown that many of the simpler features of PCMs can in fact be reproduced by only considering geometric constraints of folding chains, disregarding all additional complexities arising from the specific features of a sequence or even secondary structures.
+
+Different heuristic models for such simplified PCMs have been introduced in the past by Atilgan \textit{et al.}~\cite{atilgan2004smallworld} and Bartoli \textit{et al.}~\cite{bartoli2008effecta}.
 Atilgan \textit{et al.} propose a random shuffle model to generate artificial PCMs with similar properties to real protein models.
 Bartoli \textit{et al.} built a model for PCMs in which they assume ~\textit{ad hoc} that proteins have a distribution of amino acid distances $P(s)$ that follows $P(s) \approx s^{-1}$.
-The \textit{amino acid distance} $s$ is the separation of two connected amino acids along the backbone chain, and gives rise to the simplest implementation of a connection probability in a protein.
-Bartoli \textit{et al.} justify their approximation only heuristically in the sense that resulting PCMs have similar properties to real world PCMs.
+The \textit{amino acid distance} $s$ is the separation of two connected amino acids along the backbone chain and gives rise to the simplest implementation of a connection probability in a protein.
+Bartoli \textit{et al.} justify their approximation only heuristically in the sense that resulting PCMs have similar properties to real-world PCMs.
 
-In this paper, we provide an explanation why the $s^{-1}$ heuristic is a good initial assumption.
-To do so, we derived an analytical approximation from a 2D geometrically constrained protein model to get an expression for $P(s)$.
+In this paper, we provide an explanation for why the $s^{-1}$ heuristic is a good initial assumption.
+To do so, we derived an analytical approximation from a 2D geometrically constrained model to get an expression for $P(s)$.
 The approximation provides additional correction terms.
-We then show that the approximation can be used to fit simulations of the 2D contstrained model~\cite{molkenthin2016scaling}, a 3D version of the constrained model~\cite{molkenthin2020self} and even to real world proteins.
+We then show that the approximation can be used to fit simulations of the 2D constrained model~\cite{molkenthin2016scaling}, a 3D version of the constrained model~\cite{molkenthin2020self}, and even to real-world proteins.
 This means that the analytical approximation is a good starting point for artificially modelling PCMs in the future.
@@
 
-The paper is structured in the following way: In Sec.~\ref{sec:models}, we give a brief overview of the geometrically constrained models used, In Sec.~\ref{sec:methods}, we describe the construction of PCMs from the 2- and 3D simulations of the geometrically constrained model, as well as from structural data in the RCSB PDB and AlphaFold2 databases.
-In Sec.~\ref{sec:theory} we then derive an analytical approximation for the 2D analogue of the geometrically constrained model.
-Sec.~\ref{sec:results} highlights the results for the 2D, and 3D simulations, RCSB PDB, and AlphaFold~2 structural data by making use of the 2D analytical approximation.
+The paper is structured in the following way: In section~\ref{sec:models}, we give a brief overview of the geometrically constrained models used.
+In section~\ref{sec:methods}, we describe the construction of PCMs from the 2D and 3D simulations of the geometrically constrained model, as well as from structural data in the RCSB PDB and AlphaFold2 databases.
+In section~\ref{sec:theory} we then derive an analytical approximation for the 2D analog of the geometrically constrained model.
+Section~\ref{sec:results} highlights the results for the 2D, and 3D simulations, RCSB PDB, and AlphaFold~2 structural data by making use of the 2D analytical approximation.
 %%%%%%%%%%%%%%%%%
@@
 \label{sec:models}
-Different approaches have been used in the past for geometrical models describing protein folding, some of which derived characteristics of the secondary structure from constraints on bond and torsion angles~\cite{bhattacharjee2013flory,Danielsson2010,Molkenthin2011} and others focused on the formation mechanisms of the tertiary structure~\cite{molkenthin2016scaling, molkenthin2020self} and even others are modelled by self avoiding random walks~\cite{mey2014rareevent, hills2009insights}.
-Here we will focus on models representing maximally folded structures in 2D, and 3D as introduced in~\cite{molkenthin2016scaling} and~\cite{molkenthin2020self} respectively.
-The 3D geometrical model and the analytical approximation of its 2D analogue both build on the idea that inherently geometrical objects, such as amino acids, imply that any resulting PCM network ensemble modeling them has to be spatially embedded.
+Different approaches have been used in the past for geometrical models describing protein folding, some of which derived characteristics of the secondary structure from constraints on bond and torsion angles~\cite{bhattacharjee2013flory, Danielsson2010, Molkenthin2011} or on the formation mechanisms of the tertiary structure~\cite{molkenthin2016scaling, molkenthin2020self}, and others are modeled by self-avoiding random walks~\cite{mey2014rareevent, hills2009insights}.
+Here we will focus on models representing maximally folded structures in 2D, and 3D as introduced in~\cite{molkenthin2016scaling} and~\cite{molkenthin2020self}, respectively.
+The 3D geometrical model and the analytical approximation of its 2D analogue both build on the idea that inherently geometrical objects, such as amino acids, imply that any resulting PCM network ensemble modelling them has to be spatially embedded.
 
-In~\cite{molkenthin2016scaling}, we have introduced a simplified, 2D version of this geometrically constrained model.
+In~\cite{molkenthin2016scaling}, we introduced a simplified, 2D version of this geometrically constrained model.
 It starts from a closed chain (i.e.
@@
 The advantage of this model is that it can be approximated by a purely topological simplification, which can be treated analytically.
-Fig.~\ref{fig:schematic} illustrates the construction of the 2D geometric model.
+Figure~\ref{fig:schematic} illustrates the construction of the 2D geometric model.
+The 3D version gives rise to the geometric constraints directly, rather than through an approximation by means of a topological constraint.
 
@@
 (d) Once connected by a link, pairs of units do not disconnect.
-In~\cite{molkenthin2020self} this geometrically constrained model was extended to 3D spheres using simulations to generate compact artificial protein structures.
+In~\cite{molkenthin2020self} this geometrically constrained model was extended to 3D spheres using simulations to generate compact artificial protein-like polymer structures.
 %%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%
-\section{Protein contact maps}
+\section{Protein contact maps (PCM)}
 \label{sec:methods}
@@
     \centering
-    \includegraphics[width=0.9\columnwidth]{figures/Fig1/Fig1.pdf}
-    \caption{a) The amino acid distance $s$ (green) is defined as the number of amino acids (AS) (here represented by a circle) along the protein backbone chain (black) between two connected AS (pink).
-b) Shows an example AS distance of length $s=3$ (solid green line).
+    \includegraphics[width=0.5\columnwidth]{figures/Fig1/Fig1.pdf}
+    \caption{a) The amino acid distance $s$ (green) is defined as the number of amino acids (AAs) (here represented by a circle) along the protein backbone chain (black) between two connected AAs (pink).
+b) Shows an example AA distance of length $s=3$ (solid green line).
 When adding a new connection from node A, the previously created connection prohibits new ones (dashed pink lines).
@@
 
-Based on the simple observation that amino acids are objects in space which cannot overlap indefinitely, we introduced the 3D version of the 2D geometrically constrained model in~\cite{molkenthin2020self}.
-Starting from a chain of identical spheres, each in contact only with the neighbors it is connected to, additional connections are made by randomly selecting two spheres and moving them towards each other until they touch.
+Based on the simple observation that amino acids are objects in space that cannot overlap indefinitely, in~\cite{molkenthin2020self} we introduced the 3D version of the 2D geometrically constrained model.
+Starting from a chain of identical spheres, each in contact only with the neighbours it is connected to, additional connections are made by randomly selecting two spheres and moving them towards each other until they touch.
 The new links formed that way cannot be broken in later steps and function as constraints on subsequent link formation steps.
@@
 
-The underlying formation mechanisms of individual protein folds are incredibly complex and depend on the surrounding solvent, as well as the specific amino acid sequence and their quantum-mechanical interactions.
-Here we want to assess if the geometrically constrained protein model's contact map distributions capture real protein behavior and make it a viable model for PCMs and provide a better explanation for the heuristic used in Bartoli \textit{et al.}~\cite{bartoli2008effecta}.
-To understand this better, we study the ensemble of PCMs from real protein structures and will look at `averaged' contact maps over many different proteins that have different amino acid sequences, but their overall sequence or chain length is the same.
-We will use the RCSB Protein Data Bank (RCSB PDB), a database of structurally resolved protein amino acid sequences through X-ray crystallography, NMR or cryo-EM experiments and the synthetically generated structural database from the machine learned structures from the AlphaFold~2 database~\cite{varadi2022alphafold}.
+The underlying formation mechanisms of individual protein folds are incredibly complex and depend on the surrounding solvent, as well as the specific amino acid sequence and their interactions.
+Here, we assess if the geometrically constrained protein model's contact map distributions capture real protein behaviour and make it a viable model for PCMs and provide a better explanation for the heuristic used by Bartoli \textit{et al.}~\cite{bartoli2008effecta}.
+We study the ensemble of PCMs from real protein structures and will look at `averaged' contact maps over many different proteins that have different amino acid sequences, but their overall sequence or chain length is the same.
+We use the RCSB Protein Data Bank (RCSB PDB), a database of structurally resolved protein amino acid sequences through X-ray crystallography, NMR or cryo-EM experiments and the synthetically generated structural database from the machine-learned structures from the AlphaFold~2 database~\cite{varadi2022alphafold}.
 The RCSB PDB contains just under 200,000 protein structures to date~\cite{berman2000protein}.
-The distribution of protein chain lengths in the RCSB PDB is illustrated in Fig.~\ref{fig:pdb_stats} b.
+The distribution of protein chain lengths in the RCSB PDB is illustrated in Figure.~\ref{fig:pdb_stats} b.
 These chain lengths vary from small fragments of less than 10 amino acids to large agglomerates of over 2000 amino acids.
-The most frequently occurring chain lengths, however, are between 100 and 300 amino acids as shown in Fig.~\ref{fig:pdb_stats} b.
-The pink subset of the bar plot shown in Fig.~\ref{fig:pdb_stats} b, is the set of structures used that fulfill the criterion on protein chain lengths of $L\approx100$ ($85\le L\le 115$) amino acids, $L\approx 200$ ($185 \le L \le 215$) amino acids and $L\approx300$ ($285 \le L\le 315$) amino acids and represent all RCSB PDB structures used in our analysis.
+The most frequently occurring chain lengths, however, are between 100 and 300 amino acids as shown in figure~\ref{fig:pdb_stats} b.
+The pink subset of the bar plot shown in figure~\ref{fig:pdb_stats} b is the set of structures used that fulfil the criterion on protein chain lengths of $L\approx100$ ($85\le L\le 115$) amino acids, $L\approx 200$ ($185 \le L \le 215$) amino acids and $L\approx300$ ($285 \le L\le 315$) amino acids and represent all RCSB PDB structures used in our analysis.
 The second database consists of synthetically generated structures using AlphaFold~2 with sequences taken from SwissProt~\cite{bairoch2000swissprot}.
-SwissProt is a manually curated database of protein sequences containing over 500,000 protein sequences, whose protein chain length distribution is shown in Fig.~\ref{fig:pdb_stats} a in teal.
+SwissProt is a manually curated database of protein sequences containing over 500,000 protein sequences, whose protein chain length distribution is shown in figure~\ref{fig:pdb_stats} in teal.
 The pink bars indicate the subset of AlphaFold~2 structures used for the contact map analysis.
@@
 Second, the AlphaFold~2 per-residue confidence scores~\cite{jumper2021highly} were used to select proteins that had an average score of 90 or higher.
-Lastly, we removed structures that described the same protein but originated from different organisms, by selecting unique protein names from the pandas data table.
+Lastly, we removed structures that described the same protein but originated from different organisms, by selecting only unique protein names.
 
@@
         \centering
-	\includegraphics[width=0.9\columnwidth]{paper/figures/Fig2/Fig2.pdf}
-	    \caption{Bar plot of protein sequence lengths for protein structures from two structure databases.
+	\includegraphics[width=0.5\columnwidth]{paper/figures/Fig2/Fig2.pdf}
+	    \caption{Bar plot of protein sequence lengths for protein structures from two protein structure databases.
 a) AlphaFold~2 predicted protein structures taken from protein sequences in the manually curated SwissProt database of protein sequences.
-Pink bars show fraction of structures used for the average contact map analysis.
+Pink bars show the fraction of structures used for the average contact map analysis.
 b) RSCB PDB chain length frequencies.
-Pink bars show protein structures whose chain lengths matches the resolution of structure accounting for the whole sequence.
+Pink bars show protein structures whose chain lengths match the resolution of the structure accounting for the whole sequence.
 Only pink RCSB PDB samples were used for the average contact map analysis.
@@
 \end{figure}
-All structure-based PCMs were generated in the same way for AlphaFold~2 structures and RCSB PDB structures.
+All structure-based PCMs were generated in the same way for both AlphaFold~2 structures and RCSB PDB structures.
 Structures used were downloaded in pdb format, read in with MDAnalysis version 2.0.0~\cite{gowers2016mdanalysis} and only C$_{\alpha}$ were selected and used for the analysis.
@@
 The contacts (1 or 0) were then recorded in the adjacency matrix $\mathbf{A}^{\mathrm{struc}}$.
-An example of a protein contact map, or adjacency matrix, is shown in Fig.~\ref{fig:pdb_stats}~c.
-In order to study the amino acid distance distributions of real and simulated protein structures, the distances were calculated from the adjacency matrices.
-These distances were then histogrammed to calculate the mean amino acid distance distributions shown in Sec.~\ref{sec:results}.
-The pseudoalgorithms describing the above two processes are shown in the Supplemental Material~(SI)~\cite{SI}.
+An example of a protein contact map is shown in figure~\ref{fig:pdb_stats}~c.
+We study the distribution of distances between connected amino acids in real and simulated proteins calculated from adjacency matrices.
+These distances were then histogrammed to calculate the mean amino acid distance distributions shown in section~\ref{sec:results}.
+The pseudo-algorithms describing the above two processes are shown in the Supplemental Material~(SI)~\cite{SI}.
 All code for the construction of the contact maps and their analysis is available on GitHub at~\cite{2022sequence}.
-The derived analytical approximation in Sec.~\ref{sec:theory}, was used for data fitting.
-All fits were carried out using Scipy 1.7.0 and with scripts for the fits available on GitHub at~\cite{2022sequence}.
+The derived analytical approximation in section~\ref{sec:theory}, was used for data fitting.
+All fits were carried out using SciPy 1.7.0 and with scripts for the fits available on GitHub at~\cite{2022sequence}.
+The secondary structure protein analysis was done using a local DSSP server~\cite{1983kabsch}, and the scripts for the processing of the DSSP output are available in the same GitHub repository~\cite{2022sequence}.
 
@@
 
-To assess how we can best describe the distribution of the probability of amino acid distances $P(s)$ obtained from a PCM for the geometrically constrained model, we consider the amino acid distance distribution for the simplified 2D model as proposed in~\cite{molkenthin2016scaling}.
-We are thus able to propose the following analytical approximation.
-Fig~\ref{fig:schematic} summarizes the idea behind this two-dimensional model.
+To assess how we can best describe the distribution of the probability of amino acid distances $P(s)$ obtained from an average PCM for the geometrically constrained model, we consider the amino acid distance distribution for the simplified 2D model as proposed in~\cite{molkenthin2016scaling}.
+We make the assumption that each distance is equally likely, from which we propose an analytical approximation for $P(s)$, derived from the geometric models, and schematically summarised in figure~\ref{fig:schematic}.
 
-Let us start by introducing the auxiliary variable $F_k(s)$, defined to be the number of possible links with an amino acid distance of $s$ if $k$ links have been added before.
-Before any links are added, all amino acid distances are equally likely, as there are $F_0(s)=N$ possibilities of making a link of distance $s$ each, with $2\leq s < \frac{N}{2}$.
+We introduce the auxiliary variable $F_k(s)$, defined to be the number of possible links with an amino acid distance of $s$ if $k$ links have been added before.
+Before any links are added, because we are looking for an average PCM over the whole proteome restricted to a particular chain length, we make the assumption that all amino acid distances are equally likely, so there are $F_0(s)=N$ possibilities of making a link of distance $s$ each, with $2\leq s < \frac{N}{2}$.
 
 As we add more links, not only are links taken out of this pool, because they have already been realized, but each existing link can also geometrically prohibit other connections.
-Adding a link of length $s_1$ prohibits $2(s-1)$ links of length $s<s_1$ and $2(s_1-1)$ links of length $s\geq s_1$ (see Fig.~\ref{fig:schematic}~b).
-To find the expected number of links still available in subsequent steps, we average over all possible amino acid distances $s_1$ of the first link added.
+Adding a link of length $s_1$ prohibits $2(s-1)$ links of length $s<s_1$ and $2(s_1-1)$ links of length $s\geq s_1$ (see figure~\ref{fig:schematic}~b).
+To find the expected number of links still available in the second step, we average over all possible amino acid distances $s_1$ of the first link added.
 
@@
     \Pin^k(s)=\frac{F_k(s)}{C_k},
+    \label{eq.Pnext}
 \end{equation}
 where $C_k=\sum_{s=2}^{N/2}F_k(s)$.
-In subsequent steps, we use this probability to perform a weighted average over new links being added, and get the same reduction (compare Fig.~\ref{fig:schematic}) but starting from the pool $F_{k-1}(s)$ of the step before rather than the initially available $N$ links.
-However, we must subtract fewer links to account for some of them having left the pool in earlier steps.
-To this end, we multiply the number of blocked links by
+In subsequent steps, we use this probability to perform a weighted average over new links being added and get the same reduction (compare figure~\ref{fig:schematic}) but starting from the pool $F_{k-1}(s)$ of the step before rather than the initially available $N$ links.
+The probability of a link of length $s$ is thus the average overall added link probabilities from the first to the last added link.
+In 2D, this is repeated for $N-3$ steps until no more links can be added \cite{molkenthin2016scaling}, giving
+\begin{equation}
+    P(s)=\frac{1}{N-3}\sum_{k=0}^{N-4} \Pin^k(s).
+    \label{eq.P_next_sum}
+\end{equation}
+However, we must subtract fewer links to account for some of them having left the pool in earlier steps. To this end, we multiply the number of blocked links by
 \begin{equation}
     \frac{F_{k-1}(s)}{F_0(s)} = F_{k-1}(s)\frac{1}{N}.
-    \label{eq.Pk}
+    %\label{eq.Pk}
 \end{equation}
 
-This ignores any length-dependent bias in which links have been removed, however, examples suggest this assumption holds well. 
-This leads to the recursion
+This ignores any length-dependent bias in which links have been removed. Having considered some examples this assumption seems to hold well. 
+This leads to the following recursion:
 
 \begin{align}
-    F_k(s)&= \sum_{s_k=2}^{N/2} \Pin^k(s_k)\nonumber \bigg(F_{k-1}(s) \\
-    & - F_{k-1}(s)\frac{1}{N}
+    F_k(s)&= \sum_{s_k=2}^{N/2} \Pin^k(s_k)\nonumber \bigg(F_{k-1}(s) - F_{k-1}(s)\frac{1}{N}
     {\begin{cases}
@@
 
-which generalizes Eq.~\ref{eq:reduction_first_step} and simplifies to
+which generalizes~\ref{eq:reduction_first_step} and simplifies to
+
 \begin{align}
-   F_k(s)&=F_{k-1}(s)\Big(1-\frac{1}{N} \sum_{s_k=2}^{s} 2 (s_k-1) \Pin^k(s_k)\nonumber \\
-   &-\frac{1}{N} \sum_{s_k=s+1}^{N/2} 2 (s-1) \Pin^k(s_k)\Big)\nonumber \\
+   F_k(s)&=F_{k-1}(s)\Big(1-\frac{1}{N} \sum_{s_k=2}^{s} 2 (s_k-1) \Pin^k(s_k) -\frac{1}{N} \sum_{s_k=s+1}^{N/2} 2 (s-1) \Pin^k(s_k)\Big) \nonumber \\
    &=F_{k-1}(s)\bigg(1-\frac{1}{N}
-   \sum_{s_k=2}^{s} 2 (s_k-1) \Pin^k(s_k)\nonumber \\
-   &-\frac{2 (s-1)}{N} \Big[\sum_{s_k=2}^{N/2} \Pin^k(s_k) - \sum_{s_k=2}^{s} \Pin^k(s_k)\Big]\bigg)\nonumber \\
-   &=F_{k-1}(s)\Big(1-\frac{2}{N} (s-1) 
+   \sum_{s_k=2}^{s} 2 (s_k-1) \Pin^k(s_k)-\frac{2 (s-1)}{N} \Big[\sum_{s_k=2}^{N/2} \Pin^k(s_k) - \sum_{s_k=2}^{s} \Pin^k(s_k)\Big]\bigg) \nonumber \\
+   \nonumber
+\end{align}
+and thus:
+\begin{equation}
+   F_k(s) =  F_{k-1}(s)\Big(1-\frac{2}{N} (s-1) 
    +\frac{2}{N} s
-   \sum_{s_k=2}^{s}\Pin^k(s_k)\nonumber \\
-   &-\frac{2}{N}
-   \sum_{s_k=2}^{s} s_k \Pin^k(s_k)\Big).
-   \label{eq.Fk_rec}
-\end{align}
+   \sum_{s_k=2}^{s}\Pin^k(s_k) -\frac{2}{N}
+   \sum_{s_k=2}^{s} s_k \Pin^k(s_k)\Big).\label{eq.Fk_rec}
+\end{equation}
 
 Together with the initial condition $F_0(s)=N$, this expression constitutes an iteration rule from which $F_k(s)$ can be found for all $k$ and $s$.
-Aiming to decouple those dynamics for different $s$, we make a heuristic guess for $\Pin^k(s_k)$.
+Aiming to decouple those dynamics for different $s$, we make a heuristic guess $\tilde{P}_{\text{next}}^k(s_k)$ for $\Pin^k(s_k)$.
 It consists of two separate approximations which are used in steps $k<a$ and steps $k\geq a$ respectively, where the threshold $1\leq a\leq \frac{N}{2}$ is a free parameter.
-In the early stage we use a uniform probability
+In the early stage, we use a uniform probability
 \begin{equation}
-    \Pin^{k\ll N/2}(s)\approx \Pin^{k=0}(s)=\frac{2}{N}.
+    \tilde{P}_{\text{next}}^{k\ll N/2}(s)\approx \Pin^{k=0}(s)=\frac{2}{N}.
     \label{eq.P_small}
 \end{equation}
-For later steps, the probability of a link to still be in the pool decreases with $s$.
-We thus approximate $\Pin(s)$ with an Ansatz for the $k-$average, namely, that it drops off as $s^{-1}$, leading to
+For later steps, the probability of a link to still being in the pool decreases with $s$.
+We thus approximate $\Pin(s)$ with an ansatz for the $k-$average, namely, that it drops off as $s^{-1}$, leading to
 \begin{equation}
-    \Pin^{k\gg 1}(s)\approx \sum_{k=a}^{N-4}\Pin^{k}(s)
+    \tilde{P}_{\text{next}}^{k\gg 1}(s)\approx \sum_{k=a}^{N-4}\Pin^{k}(s)
     \approx \frac{s^{-1}}{ H_{\frac{N}{2}}-1},
@@
 where $H_{\frac{N}{2}}$ is the harmonic number serving as a normalization factor. 
-As we will see later, this approximation holds most precisely for intermediate values of $k$, with the $k-$average of $\Pin^{k\gg 1}(s)\approx P(s)$, making it self-consistent.
+As we will see later, this approximation holds most precisely for intermediate values of $k$, with the $k-$average of $\tilde{P}_{\text{next}}^{k\gg 1}(s)\approx P(s)$, making it self-consistent.
 Errors for smaller and larger values of $k$ are thought to approximately cancel out.
-Substituting Eq.~\ref{eq.P_small} and Eq.~\ref{eq.P_large} into Eq.~\ref{eq.Fk_rec} results in the following two recursions for the early and late evolution of the amino acid distance distribution respectively.
+Substituting~\ref{eq.P_small} and~\ref{eq.P_large} into~\ref{eq.Fk_rec} results in the following two recursions for the early and late evolution of the amino acid distance distribution respectively.
 \begin{align}
@@
    F_{k-1}(s)\Big(1-\frac{2}{N} (s-1) 
-   +\frac{4}{N^2} s (s-1)\nonumber \\
-   &-\frac{2}{N^2}(s^2+s-2)\Big)\nonumber \\
+   +\frac{4}{N^2} s (s-1)\nonumber -\frac{2}{N^2}(s^2+s-2)\Big)\nonumber \\
    &=F_{k-1}(s)(1-\flow(s)),
@@
 \begin{align}
-   F_{k\gg1}(s)&\approx F_{k-1}(s)\Big(1-\frac{2}{N} (s-1) \nonumber\\
-   &+\frac{2}{N} s \sum_{s_k=2}^{s}\frac{1}{s_k(H_{\frac{N}{2}}-1)}\nonumber\\
-   &-\frac{2}{N} \sum_{s_k=2}^{s} s_k \frac{1}{s_k(H_{\frac{N}{2}}-1)}\Big)\nonumber \\
-   &=F_{k-1}(s)\bigg[1 -\Big(\frac{2}{N}-\frac{2}{N}\frac{H_{s}-1}{H_{\frac{N}{2}}-1} \nonumber\\
-   &+\frac{2}{N(H_{\frac{N}{2}}-1)}\Big)s \nonumber + \frac{2}{N(H_{\frac{N}{2}}-1)} + \frac{2}{N}\bigg]\nonumber \\
-   &=F_{k-1}(s)(1-\fhigh(s)).
+   F_{k\gg1}(s)&\approx F_{k-1}(s)\Big(1-\frac{2}{N} (s-1) \nonumber+\frac{2}{N} s \sum_{s_k=2}^{s}\frac{1}{s_k(H_{\frac{N}{2}}-1)}\nonumber-\frac{2}{N} \sum_{s_k=2}^{s} s_k \frac{1}{s_k(H_{\frac{N}{2}}-1)}\Big)\nonumber \\
+   &=F_{k-1}(s)\bigg[1 -\Big(\frac{2}{N}-\frac{2}{N}\frac{H_{s}-1}{H_{\frac{N}{2}}-1} \nonumber +\frac{2}{N(H_{\frac{N}{2}}-1)}\Big)s \nonumber + \frac{2}{N(H_{\frac{N}{2}}-1)} + \frac{2}{N}\bigg]\nonumber \\
+   &=F_{k-1}(s)(1-\fhigh(s)),
    \label{eq.Fk_rec_high}
@@
 
-We can now use the recursive expressions to write down a closed expression for $F_k(s)$, using $\flow$ for the first $a$ steps and $\fhigh$ for the rest:
+We can now use the recursive expressions~\ref{eq.Fk_rec_low} and~\ref{eq.Fk_rec_high} to write down a closed expression for $F_k(s)$, using $\flow$ for the first $a$ steps and $\fhigh$ for the rest:
 
@@
     &= N\left(\frac{1-\flow(s)}{1-\fhigh(s)}\right)^a\left(1-\fhigh(s)\right)^{k}.
+    \label{eq.F_final}
 \end{align}
 
-In 2D, this is repeated for $N-3$ steps until no more links can be added \cite{molkenthin2016scaling}.
-The probability distribution of the realized amino acid distances of all added links is then given by the average over the available pools at each link addition step, leading to
+This can now be used to find an approximation for the amino acid distance distribution $P(s)$.
+By inserting~\ref{eq.F_final} into~\ref{eq.P_next_sum}, using~\ref{eq.Pnext}, we can state that the probability distribution of the realized amino acid distances of all added links is then given by the average over the available pools at each link addition step.
+
 \begin{align}
-    P(s)&=\frac{1}{N-3}\sum_{k=0}^{N-4} \Pin^k(s) \nonumber \\
-    &\approx \frac{N}{N-3}\left(\frac{1-\flow(s)}{1-\fhigh(s)}\right)^a \nonumber\\
-&\times \sum_{k=0}^{N-4} \frac{1}{\evdel{C_k}_k} \left(1-\fhigh(s)\right)^{k} \nonumber \\
-    &\approx \left(\frac{1-\flow(s)}{1-\fhigh(s)}\right)^a\frac{\Gamma}{\fhigh(s)},
+    P(s)&=\frac{1}{N-3}\sum_{k=0}^{N-4} \Pin^k(s)  \\
+    &\approx \frac{N}{N-3}\left(\frac{1-\flow(s)}{1-\fhigh(s)}\right)^a %\nonumber\\
+\times \sum_{k=0}^{N-4} \frac{1}{\evdel{C_k}_k}\nonumber \left(1-\fhigh(s)\right)^{k} 
+\end{align}
+where $\evdel{C_k}_k$ approximates the individual $C_k$'s as the average of $C_k$ over $k$.
+This approximation allows the use of the geometric series for solving the equation:
+\begin{align}
+    P(s)&\approx \left(\frac{1-\flow(s)}{1-\fhigh(s)}\right)^a\frac{\Gamma}{\fhigh(s)},
     \label{eq.Ps}
 \end{align}
-where $\evdel{C_k}_k$ is the average of $C_k$ over $k$.
-This approximation allows the use of the geometric series for solving the equation. 
-The resulting expression in Eq.~\ref{eq.Ps} by visual inspection of Fig.\ref{fig:2d_sim} resembles a power law with an exponent of $-1$, thus justifying the earlier approximation of $\Pin^k(s)$.
+where $\Gamma=\frac{N}{(N-1)}\evdel{C_k}_k$ collects all constant factors and is used as a free fitting parameter, as the details of the evolution of $C_k$, as well as the role of dimensionality (3D vs.
+2D) are unknown.
+The resulting expression in~\ref{eq.Ps} by visual inspection of figure~\ref{fig:2d_sim} resembles a power law, thus justifying the earlier approximation of $\Pin^k(s)\approx\tilde{P}_{\text{next}}^k(s)$ introduced in~\ref{eq.P_large}.
 
@@
 	    \caption{The analytical approximation fitted to amino acid distance distributions for a) 2D and b) 3D simulated data.
-Pink area shows the 95\% confidence interval.
+The pink area shows the 95\% confidence interval.
 c) Example of a PCM taken from one of the 3D simulation runs.
@@
 \section{Results and Discussion}\label{sec:results}
-The results analysis centers around understanding how the analytical approximation can be used to understand the simulations in 2D and 3D from~\cite{molkenthin2016scaling,molkenthin2020self}, and structural data from the RCSB PDB and AlphaFold~2 databases.
-To this end, Eq.~\ref{eq.Ps} is used as a fitting function in order to determine the two parameters $\Gamma$ and $a$.
+Our analysis centres around understanding how the analytical approximation can be used to understand the simulations in 2D and 3D from~\cite{molkenthin2016scaling,molkenthin2020self}, and structural data from the RCSB PDB and AlphaFold~2 databases.
+To this end,~\ref{eq.Ps} is used as a fitting function in order to determine the two parameters $\Gamma$ and $a$.
 
 \subsection{The analytical approximation fits 2D and 3D simulation data}
-Fig.~\ref{fig:2d_sim} compares Eq.~\ref{eq.Ps} with respect to the two-dimensional (a) and three-dimensional (b) simulation data introduced in~\cite{molkenthin2016scaling} and ~\cite{molkenthin2020self}, respectively.
-Fig.~\ref{fig:2d_sim} c, shows an example of an adjacency matrix $\mathbf{A}^{\mathrm{sim}}$ as generated from a 3D simulation run.
+Figure~\ref{fig:2d_sim} compares~\ref{eq.Ps} with respect to (a) 2D and (b) 3D simulation data introduced in~\cite{molkenthin2016scaling} and ~\cite{molkenthin2020self}, respectively.
+Figure~\ref{fig:2d_sim}~c shows an example of an adjacency matrix $\mathbf{A}^{\mathrm{sim}}$ as generated from a 3D simulation run.
 The 2D simulation data was generated from 10 repeats with chain lengths of 498 amino acids taken from~\cite{molkenthin2016scaling}.
-The 3D simulation data consists of adjacency matrices computed from 30 simulation runs, and was analyzed as was explained in Sec.~\ref{sec:methods}.
-The amino acid distance distributions are plotted showing the mean frequencies of each amino acid distance up to $\frac{L}{2}\sim250$ and $\frac{L}{2}\sim150$ for 2D and 3D simulations, respectively.
+The 3D simulation data consists of adjacency matrices computed from 30 simulation runs and was analyzed as was explained in section~\ref{sec:methods}.
+The amino acid distance distributions are plotted to show the mean frequencies of each amino acid distance up to $\frac{L}{2}\sim250$ and $\frac{L}{2}\sim150$ for 2D and 3D simulations, respectively.
 The shaded areas represent the 95\% confidence intervals.
-The theoretical approximation from Eq.~\ref{eq.Ps} was fitted to both 2D and 3D simulation data separately.
-The two parameters, $a$ and $\Gamma$ in Eq.(\ref{eq.Ps}), are given in Table~\ref{table:sim-results}.
-The fitted approximation captures the behavior of the simulation data well for both 2D (Fig.~\ref{fig:2d_sim}a) and 3D (Fig.~\ref{fig:2d_sim}b) simulations.
+The theoretical approximation from~\ref{eq.Ps} was fitted to both 2D and 3D simulation data separately.
+The two parameters, $a$~and~$\Gamma$ in~\ref{eq.Ps}, are given in Table~\ref{table:sim-results}.
+See the SI~\cite{SI} for the parameters for 3D simulation data in the $L\approx 100$ and $L \approx 200$ ranges.
+The fitted approximation captures the behaviour of the simulation data well for both~2D and~3D simulations.
 
@@
 \subsection{$P(s)$ from RCSB PDB and AlphaFold~2 agree and cover similar secondary structure content}
-As described in Sec.~\ref{sec:methods}, we have chosen a subset of structures from the RCSB PDB and AlphaFold~2 database for the analysis.
-The number of structures in each chain length range for both databases are shown in Table~\ref{table:subsets}.
+Since AlphaFold~2 generates structures from an AI model, the database contains a variety of structures whose 3D folded structures are more or less confident (as measured by the confidence score measure provided by the database).
+Therefore, before analyzing the amino acid distance distributions from RCSB PDB and AlphaFold~2, we wanted to understand if the distributions generated from the two databases at lengths $L\approx 100$, $L \approx 200$, and $L \approx 300$ are statistically the same, given the selection of a mean confidence score of 90 or above.
+To this end, we used the non-parametric test of the Kolmogorov-Smirnov (KS) statistic to compare the probability distributions of AlphaFold~2 and RCSB PDB for each of the lengths~\cite{2008kolmogorov}.
+The KS statistic for the lengths of $L \approx 100$, $L \approx 200$, and $L \approx 300$ amino acids was 0.07, 0.10, and 0.12 respectively.
+Therefore, the null hypotheses for the distributions being the same were accepted at the $\alpha=0.01$ level.
+
+In addition to calculating the KS statistic between the filtered AlphaFold~2 structures (confidence score above 90) and RCSB PDB structures, we also used a subset of AlphaFold~2 structures in the $L \approx 300$ range with per-residue confidence scores below 70.
+In this case, the KS statistic was 0.50 and the null hypothesis was rejected at the $\alpha=0.01$ level.
+For more details see the SI~\cite{SI}.
+
+The number of structures used for the analysis in each chain length range for both databases is shown in Table~\ref{table:subsets}.
 All structures were converted into PCMs, as described in Section~\ref{sec:methods}.
 From the PCMs we computed the amino acid distance distribution in the same way as was done for the 3D simulations.
-Since AlphaFold~2 generates structures from an AI model, we wanted to understand if the distributions generated from it and RCSB PDB at lengths $L\approx 100$, $L \approx 200$ and $L \approx 300$ are statistically the same.
-To this end, we use the non-parametric test of the Kolmogorov-Smirnov (KS) statistic to compare the probability distributions of AlphaFold~2 and RCSB PDB for each of the lengths~\cite{2008kolmogorov}.
-The KS statistic for the lengths of $L \approx 100$ $L \approx 200$, and $L \approx 300$ amino acids was 0.07, 0.10, and 0.12 respectively.
-Therefore, the null hypotheses for the distributions being the same was accepted at the $\alpha=0.01$ level.
-For more details see the SI~\cite{SI}.
 
@@
 \setlength{\tabcolsep}{5pt}
-\resizebox{\columnwidth}{!}{
+\resizebox{0.5\columnwidth}{!}{
 \begin{tabular}{ccc}
@@
 Next, we investigated if the amino acid distance distribution from the structural data (RCSB PDB and AlphaFold~2) can be modeled by the analytical approximation.
-In order to answer this, we look at the scaling of the the distribution of RCSB PDB data in the first instance.
-We will present the results for the AlphaFold~2 data in the SI, since the KS statistics suggests that the RCSB PDB distributions and AlphaFold~2 of $P(s)$ agree.
-As a result, we observe similar trends.
+In order to answer this, we look at the scaling of the distribution of RCSB PDB data in the first instance.
 
@@
 	\includegraphics[width=\textwidth]{paper/figures/Fig4/Fig4.pdf}
-        \caption{a) Amino acid distance distributions for contact maps derived from the RCSB PDB database for protein chains in the ranges of $L\approx100$, $L\approx200$ and $L\approx300$ amino acids.
-b) Comparison between the distributions of RCSB PDB and AlphaFold~2 database for protein chains in the $L\approx300$ range.
-The analytical approximation (solid line) and power law (dashed line) are fitted to the tail of the RCSB data.
-Pink shaded area shows the 95\% confidence level.
+        \caption{a) Amino acid distance distributions for contact maps derived from the RCSB PDB database for protein chains in the ranges of $L \approx 100$, $L\approx200$, and $L\approx300$ amino acids.
+b) Comparison between the distributions of RCSB PDB and AlphaFold~2 database for protein chains in the $L \approx 300$ range.
+The analytical approximation (solid line) is fitted to the tail of the RCSB data.
+The pink shaded area shows a 95\% confidence level.
 c) Secondary structure content distribution in the RCSB PDB and AlphaFold~2 structures in the $L\approx300$ chain length range.}
@@
 
-Fig.~\ref{fig:sdd} a shows the distributions of the amino acid distances for RCSB PDB logarithmically scaled for the different protein chain lengths of $L\approx100$ (light purple circles), $L\approx200$ (medium purple squares) and $L\approx300$ (palatinate triangles).
+Figure~\ref{fig:sdd} a shows the distributions of the amino acid distances for RCSB PDB logarithmically scaled for the different protein chain lengths of $L\approx100$ (light purple circles), $L\approx200$ (medium purple squares), and $L\approx300$ (palatinate triangles).
 For all three lengths, the distributions show an approximate power law decay in their tail, which is consistent with the simulated results.
 
-In addition to a general power law, we fit the analytical approximation to the RCSB PDB amino acid distance distribution in Fig.~\ref{fig:sdd} b in the $L\approx300$ range.
-We also show the AlphaFold~2 amino acid distance distribution for comparison.
-The theoretical and power law fits are shown in black solid and dotted lines, respectively.
+We fit the analytical approximation to the RCSB PDB amino acid distance distribution in figure~\ref{fig:sdd} b in the $L\approx300$ range.
+In addition, we show the AlphaFold~2 amino acid distance distribution for comparison.
+The theoretical fit is represented by a solid black line.
 The parameters given by the analytical approximation are $a=4 \pm 1$ and $\Gamma=(7.4 \pm 0.8)\cdot 10^{-4}$.
-As for the power law, the exponent is $\gamma=1.45 \pm 0.02$.
-The shaded area shows the 95\% confidence level (C.L.).
-Similar fits for $L\approx100$ and $L\approx200$ are presented in the SI.
+The shaded area shows a 95\% confidence level (C.L.).
+Similar fits for $L\approx100$ and $L\approx200$ as well as the AlfaFold~2 results are presented in the SI.
 
-We observe that, for intermediate values of $s$, i.e.
-$30\lesssim s\lesssim \frac{L}{2}$, the distributions are described well by the analytical approximation.
-However, the curves deviate for amino acid distances in the $5\lesssim s \lesssim 30$ range (see Fig.\ref{fig:sdd} b).
-Comparing this to the simulation distributions and their theoretical fits (Fig~\ref{fig:2d_sim}), we see an under-representation of amino acid distances calculated from PCMs from RCSB PDB data.
+We observe that for intermediate values of $s$, i.e.
+$30\lesssim s \lesssim \frac{L}{2}$, the distributions are described well by the analytical approximation.
+However, the curves deviate for amino acid distances in the $5\lesssim s \lesssim 30$ range (see figure~\ref{fig:sdd} b).
+Comparing this to the simulation distributions and their theoretical fits (Figure~\ref{fig:2d_sim}), we see an under-representation of amino acid distances calculated from PCMs from RCSB PDB data.
 
-One possible explanation of the under-representation can stem from secondary structure properties, as it appears in a specific distance range, where contacts arise from the proteins' secondary structure elements.
-In order to try to explore this feature, Fig.~\ref{fig:sdd} c shows the secondary structure content distributions for proteins in the $L\approx300$ chain length range from RCSB PDB and AlphaFold~2.
-Proteins mostly contain secondary structures in the combined $\alpha$-helix---$\beta$-sheet region.
-For protein chain length regions of $L\approx100$ and $L\approx200$ proteins with predominently $\alpha$ and $\beta$-sheet regions exist, as shown in the SI~\cite{SI}.
+To understand this under-representation better, we looked at the secondary structure content of the PCMs we studied.
+The reasoning behind this is that the range in which the under-representation is observed is at a specific distance range, where contacts arise from the proteins' secondary structure elements.
+In figure~\ref{fig:sdd}~c we show the secondary structure content distributions for proteins in the $L\approx300$ chain length range from RCSB PDB and AlphaFold~2.
+We see that proteins mostly contain secondary structures in the combined $\alpha$-helix---$\beta$-sheet region.
+For protein chain length regions of $L\approx100$ and $L\approx200$ proteins with predominantly $\alpha$ and $\beta$-sheet regions exist, are shown in the SI~\cite{SI}.
 However, in all data, a high percentage of mixed secondary structure is observed.
-This is also in agreement with~\cite{michie1996analysis}, suggesting that any secondary structure features should be evened out.
+This is also in agreement with~\cite{michie1996analysis}.
+As all three chain length ranges contain high percentages of both mixed and predominantly $\alpha$ or predominantly $\beta$ secondary structures.
+This rules out that a certain type of secondary structure, or absence thereof, gives rise to this under-representation the analysis does not give insights if the secondary structure, in the form of an $\alpha$-helix---$\beta$-sheet as a structure per se leads to this under-representation.
 
-Next, we looked at changing the threshold value $d_c$ from the original $8 \; \mathrm{\AA}$ to 10, 15 and 21 $\mathrm{\AA}$, see Fig.
-4 in the SI~\cite{SI}.
-For 10 $\mathrm{\AA}$, the shape of the distribution was similar to that of in Fig.~\ref{fig:sdd}~b, but for 15 $\mathrm{\AA}$ the under-representation starts to disappear slightly whereas for 21 $\mathrm{\AA}$ it disappears almost completely.
-Due to the high number of these mixed $\alpha$-helix---$\beta$-sheet secondary structures, it is likely that with a threshold of 8 $\mathrm{\AA}$ some inter-secondary structure contacts are being excluded, e.g.
+Threfore we next looked at different threshold used that define a contact between two amino acids.
+We changed the threshold value $d_c$ from the original $8 \; \mathrm{\AA}$ to 10, 15 and 21 $\mathrm{\AA}$, see figure 4 in the SI~\cite{SI}.
+For 10 $\mathrm{\AA}$, the shape of the distribution was similar to that of in figure~\ref{fig:sdd}~b, but for 15 $\mathrm{\AA}$ the under-representation starts to disappear, whereas for 21 $\mathrm{\AA}$ it disappears almost completely.
+It is thus likely that with a threshold of 8 $\mathrm{\AA}$ some inter-secondary structure contacts are being excluded, e.g.
 for large $\beta$-sheet content.
-This is a likely cause for the under-representation of amino acid distances in the $5 \lesssim s \lesssim 30$ region.
-Comparing to simulations where this dip is not observed, links are made until no longer possible, whereas in the PCMs, the connectivity is determined by the threshold value.
-Therefore it is not surprising this feature cannot be captured in the simulations.
+Tuning this threshold value allows for the modulation of the under-representation of amino acid distances in the $5 \lesssim s \lesssim 30$ region.
+Compared to simulations where this is not observed due to the neglect of the existence of secondary structure.
 
@@
 
-Here we have introduced and analyzed the amino acid distance $s$ and its probability distribution $P(s)$ in measured and simulated protein structures.
-For the equivalent model in 2D, we have used a mean-field approach to derive an analytical approximation for the amino acid distance distribution, and we show its agreement with the simulated and measured distributions from real proteins.
-Therefore we managed to demonstrate that the geometrically constrained model's sequence distance distribution $P(s)$ match real world data well.
-In addition, the derived analytical approximation could serve as a good basis to model and generate protein-like adjacency matrices as was done in ~\cite{bartoli2008effecta}.
-It also highlights that the proposed heuristic of $P(s)\approx s^{-1}$, is a good starting point, but with an  approximation from Eq.~\ref{eq.Ps}.
+Here, we introduced and analyzed the amino acid distance $s$ and its probability distribution $P(s)$ for both real-world and gemoetrically constrained model simulations.
+For the equivalent model in 2D, we have used a mean-field approach to derive an analytical approximation for the amino acid distance distribution $P(s)$, and we show its agreement with the simulated geometrically constrained folding model and measured distributions from real-world proteins.
+Therefore we demonstrate that the geometrically constrained model's amino acid distance distribution $P(s)$ can match real-world data well.
+In addition, the derived analytical approximation can serve as a good basis to model and generate protein-like adjacency matrices as was done in~\cite{bartoli2008effecta}.
+It also highlights that the proposed heuristic of $P(s)\approx s^{-1}$ is a good starting point to describe amino acid distance distributions.
+However, here we managed to derive a more broad approximation for which a power law can be seen as a special case~\ref{eq.Ps}.
 
 Gaining a better understanding of the ensemble of folded protein structures can help guide the way to a better understanding of the constraints within which structures may occur.
-Together with an understanding of secondary structure principles, such as those in
-~\cite{Danielsson2010, Molkenthin2011}, this can help to narrow down the complex energy landscapes and find paths through them more effectively in the future.
+Together with an understanding of secondary structure principles, such as those in~\cite{Danielsson2010, Molkenthin2011}, this can help to narrow down the complex energy landscapes and find paths through them more effectively in the future.
 
-\section*{Declarations}
-\subsection{Availability of data and materials}
+\section*{Data availability statement}
 All data generated or analyzed during this study are included in this published article.
 All raw data and their analysis can be found on GitHub at~\cite{2022sequence}, including information required for their reproduction.
-\subsection{Competing interests}
+
+\section*{Conflict of interests}
 The authors declare no competing interests.
-\subsection{Funding}
-This research was supported by a Chancellor's Fellowship from the University of Edinburgh. 
-\subsection{Authors' contributions}
-N.M.
+\section*{Acknowledgements}
+We thank Marc Timme and Matteo Degiacomi for fruitful discussions.
+%\subsection{Authors' contributions}
+%N.M.
 and J.J.G.
@@
 Everyone wrote the manuscript.
-\subsection{Acknowledgements}
-We thank Marc Timme, and Matteo Degiacomi for fruitful discussions.
-
+\section*{References}
 \bibliographystyle{unsrt}

### 2022-12-29 paper/supplemental_material.tex
Updates from Overleaf

@@
 %\documentclass[reprint,amsmath,amssymb,rmp,onecolumn,notitlepage,11pt]{revtex4-1}
-\documentclass[notitlepage,
-reprint,%twocolumn,
-%preprint
-onecolumn,
-amsmath,amssymb,superscriptaddress,aps,
-pre,floatfix]{revtex4-1}
+\documentclass[10pt]{iopart}
 \usepackage[utf8]{inputenc}
-%\usepackage{authblk}
-%\usepackage{natbib}
 \usepackage[normalem]{ulem}
@@
 \begin{document}
-\title{Supplemental material accompanying "What geometrically constrained protein models can tell us about real-world protein contact maps"}
-\author{Nora Molkenthin}
-\email[Author to whom correspondence should be addressed:]{nora.molkenthin@pik-potsdam.de}
-\affiliation{Potsdam Institute for Climate Impact Research, Germany}
-\author{J. J.  Güven}
-\affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
-\author{Steffen Mühle}
-\affiliation{Max-Planck Institute for Dynamics and Self-Organization (MPIDS), Am Faßberg 17, 37077 Göttingen, Germany}
-\author{Antonia S J S Mey}
-\email[Author to whom correspondence should be addressed:]{antonia.mey@ed.ac.uk}
-\affiliation{EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
-\maketitle
+\title[Supplementary material]{Supplemental material accompanying "What geometrically constrained folding models can tell us about real-world protein contact maps"}
+\author{Nora Molkenthin$^{1*\dagger}$, J. Jasmin Güven$^{2*}$, Steffen Mühle$^{3}$, Antonia S. J. S. Mey$^{2\dagger}$}
+\address{$^1$Potsdam Institute for Climate Impact Research, Germany}
+\address{$^2$EaStCHEM School of Chemistry, University of Edinburgh, Edinburgh, United Kingdom}
+\address{$^3$Max-Planck Institute for Dynamics and Self-Organization (MPIDS), Am Faßberg 17, 37077 Göttingen, Germany}
+\address{$^*$ These authors contributed equally}
+\address{$^\dagger$ To whom correspondence should be addressed}
+\ead{nora.molkenthin@pik-potsdam.de,antonia.mey@ed.ac.uk}
 
-\section*{Pseudoalgorithms for creating protein contact maps from protein data}
+%\maketitle
+
+\section{Pseudo-algorithms for creating protein contact maps from protein data}
 
@@
 If the distance is less than 8~Å, a contact ($A_{ij}=1$) is recorded in the adjacency matrix.
-Similarly, Algorithm~\ref{alg:dist} describes how we compute the amino acid distances from amino acids who are in close contact.
+Similarly, Algorithm~\ref{alg:dist} describes how we compute the amino acid distances from amino acids that are in close contact.
 
@@
 
-\section*{Amino acid distance distributions of other chain lengths}
-\subsection{AlphaFold 2}
-Fig.~\ref{fig:alphafold} shows the amino acid distance distributions for protein contact maps calculated from AlphaFold~2 data in the chain length ranges of $L \approx 100$ amino acids ($85 \leq L \leq 115$, light green circles), $L \approx 200$ amino acids ($185 \leq L \leq 215$, medium green squares) and $L \approx 300$ amino acids ($285 \leq L \leq 315$, dark green triangles).
-The plot shows the approximate power law behaviour of the distribution.
+\section*{Combined protein contact map from RCSB PDB}
+
+Figure~\ref{fig:average} shows a contact map obtained from 100 RCSB PDB protein structures in the range $L \approx 300$.
 
@@
     \centering{
-    \includegraphics[width=\columnwidth]{paper/figures/SM_figures/alpha_chain_comparison.pdf}
+    \includegraphics[width=0.5\columnwidth]{paper/figures/SM_figures/average_matrix_300.jpeg}
+    \caption{Combined contact map of a 100 RCSB PDB protein structures in the range $L \approx 300$. }
+    \label{fig:average}}
+\end{figure*}
+
+\section{Amino acid distance distributions of other chain lengths}
+\subsection{AlphaFold 2}
+Figure~\ref{fig:alphafold} shows the amino acid distance distributions for protein contact maps calculated from AlphaFold~2 data in the chain length ranges of $L \approx 100$ amino acids ($85 \leq L \leq 115$, light green circles), $L \approx 200$ amino acids ($185 \leq L \leq 215$, medium green squares) and $L \approx 300$ amino acids ($285 \leq L \leq 315$, dark green triangles).
+The plot shows the approximate power law behavior of the distribution.
+
+\begin{figure*}[htb]
+    \centering{
+    \includegraphics[width=0.5\columnwidth]{paper/figures/SM_figures/alpha_chain_comparison.pdf}
     \caption{Amino acid distance distributions for contact maps derived from the AlphaFold 2 database for protein chains in the ranges of $L \approx 100$ amino acids ($85 \leq L \leq 115$, light green circles), $L \approx 200$ amino acids ($185 \leq L \leq 215$, medium green squares) and $L \approx 300$ amino acids ($285 \leq L \leq 315$, dark green triangles).}
@@
 
+\subsection{Low-confidence AlphaFold 2 structures}
+
+In addition to studying AlphaFold~2 structures with per-residue confidence scores of above 90, we investigate the low-confidence AlphaFold~2 structures.
+A model structure is said to be low confidence if its per-residue confidence score is less than 70 but greater than 50, and very low, if the score is less than 50.
+In this work, the low-confidence structures are made up of both the above categories.
+Figure~\ref{fig:low-conf} shows the amino acid distance distribution for protein contact maps derived from these low-confidence AlphaFold~2 structures.
+
+\begin{figure*}[htb]
+    \centering{
+    \includegraphics[width=\columnwidth]{paper/figures/SM_figures/low_confidence_combined.pdf}
+    \caption{a) Comparison between the amino acid distance distributions of RCSB PDB and low-confidence AlphaFold~2 structures.
+Pink shaded area shows the 95\% confidence level.
+The comparison is shown for the $L\approx300$ range.
+b) Example of a low-confidence structure (UniProt ID: O04609)from the AlphaFold~2 database.
+Colouring indicates the confidence score of each part of the structure: dark blue - very high, light blue - confident, yellow - low, and orange - very low.
+}
+    \label{fig:low-conf}}
+\end{figure*}
+
 \subsection{RCSB PDB and three-dimensional simulations}
 
-Fig.~\ref{fig:rcsb_combined} a shows the amino acid distance distributions for protein contact maps calculated from structures in the chain length range $L\approx100$ amino acids for RCSB PDB (dark purple circles) and AlphaFold~2 (dark green squares).
+Figure~\ref{fig:rcsb_combined} a shows the amino acid distance distributions for protein contact maps calculated from structures in the chain length range $L\approx100$ amino acids for RCSB PDB (dark purple circles) and AlphaFold~2 (dark green squares).
 The shaded area shows the 95\% confidence level.
 Also shown are the analytic approximation and a general power law fitted to the RCSB data and their fit parameters, $a$ and $\Gamma$ for the analytical approximation and exponent $\gamma$ for the power law,  are shown in~Table~\ref{table:stats}.
-Similarly, Fig.~\ref{fig:rcsb_combined} shows the amino acid distance distributions for protein contact maps calculated from structures in the chain length range $L\approx200$ amino acids for RCSB PDB.
+Similarly, figure~\ref{fig:rcsb_combined} shows the amino acid distance distributions for protein contact maps calculated from structures in the chain length range $L\approx200$ amino acids for RCSB PDB.
 We also computed the Kolmogorov-Smirnov (KS) statistic between the RCSB PDB and AlphaFold 2 distributions for each length range.
@@
 
-Fig.~\ref{fig:sim_3d} shows the amino acid distance distributions for protein contact maps calculated from the three-dimensional simulations.
+Figure~\ref{fig:sim_3d} shows the amino acid distance distributions for protein contact maps calculated from the three-dimensional simulations.
 As before, the analytic approximation was fitted to the data in each range and the shaded area shows the 95\% confidence interval.
@@
 
-Finally, Fig.~\ref{fig:dc} shows the effect of changing the threshold value for determining contacts between amino acids for contact maps calculated from RCSB PDB, for a)  $L\approx100$ amino acids,  b) $L\approx200$ amino acids and c) $L\approx300$ amino acids.
+Finally, figure~\ref{fig:dc} shows the effect of changing the threshold value for determining contacts between amino acids for contact maps calculated from RCSB PDB, for a)  $L\approx100$ amino acids,  b) $L\approx200$ amino acids and c) $L\approx300$ amino acids.
 
@@
 
-\begin{table}[tb] 
+\begin{table}[h!] 
 \centering
-\setlength{\tabcolsep}{10pt}
-{\renewcommand{\arraystretch}{1.5}
+\resizebox{\textwidth}{!}{
+%\setlength{\tabcolsep}{9pt}
+%{\renewcommand{\arraystretch}{1.0}
 \begin{tabular}{c c c c c c c c}
@@
  \hline
-\end{tabular}
-}
- \caption{Parameters $a$ and $\Gamma$ for the analytical approximation, exponent $\gamma$ for the power law fit, and their uncertainties from fitting to amino acid distributions obtained from RCSB PDB structures in ranges $L\approx100$ amino acids, $L\approx200$ amino acids and $L\approx300$ amino acids.
+\end{tabular}}
+%}
+ \caption{Parameters $a$ and $\Gamma$ for the analytical approximation, exponent $\gamma$ for the power law fit, and their uncertainties from fitting to amino acid distributions obtained from RCSB PDB structures in ranges $L\approx100$ amino acids, $L\approx200$ amino acids, and $L\approx300$ amino acids.
 Shown also are the fit parameters $a$ and $\Gamma$ from fitting the analytic approximation to the 2D simulation and 3D simulations of chains in the $L\approx100$, $L\approx200$ and $L\approx300$ ranges.}
@@
     \centering{
-    \includegraphics[width=\columnwidth]{paper/figures/SM_figures/distributions_combined.pdf}
+    \includegraphics[width=0.8\columnwidth]{paper/figures/SM_figures/distributions_combined.pdf}
     \caption{Comparison between the distributions of RCSB PDB and AlphaFold 2 database.
 The analytic approximation (solid line) and power law (dashed line) are fitted to the the RCSB PDB data.
-Pink shaded area shows the 95\% confidence level.
+The pink shaded area shows the 95\% confidence level.
 The comparison is shown for two protein chain length ranges: a) $L\approx 100$ and b) $L\approx200$ range.
@@
 
-\begin{figure*}[htb]
+\begin{figure*}[!h]
     \centering{
-    \includegraphics[width=\columnwidth]{paper/figures/SM_figures/simulations_3d.pdf}
+    \includegraphics[width=0.8\columnwidth]{paper/figures/SM_figures/simulations_3d.pdf}
     \caption{Amino acid distance distributions obtained from three-dimensional simulations for chains in the ranges a) $L\approx100$  and b) $\approx200$ amino acids.
@@
 
-\begin{figure*}[htb]
+\begin{figure*}[!h]
     \centering{
-    \includegraphics[width=\columnwidth]{paper/figures/SM_figures/threshold_combined.pdf}
+    \includegraphics[width=0.8\columnwidth]{paper/figures/SM_figures/threshold_combined.pdf}
     \caption{Amino acid distance distributions from RCSB PDB and 3D simulations for a) $L\approx100$, b) $L\approx200$ and c) $L\approx300$.

### 2022-12-29 code/scripts/bars.py

@@ -1,0 +1,77 @@
+"""Functions for plotting bar plots"""
+import matplotlib.pyplot as plt
+import pandas as pd
+import seaborn as sns
+from plot_functions import _COLOUR_PALETTE
+import numpy as np
+
+
+def get_data_for_bars(path_to_data: str) -> tuple:
+    """
+    Read csv file from path and return numpy arrays as tuple
+    @param path_to_data:
+    @return:
+    """
+    dataframe = pd.read_csv(path_to_data)
+    bins = dataframe["Bins"].to_numpy()
+    frequencies = dataframe["Number"].to_numpy()
+    return bins, frequencies
+
+
+def calculate_adjusted_frequency(frequency_to_adjust: np.ndarray, bottom_frequency: np.ndarray) -> np.ndarray:
+    """
+
+    @param frequency_to_adjust: "top" bars i.e. either all SwissProt or all RCSB frequencies
+    @param bottom_frequency: "bottom" bars i.e. the used PDBs' frequencies
+    @return: difference between top and bottom bars
+    """
+    return frequency_to_adjust - bottom_frequency
+
+
+def create_bar_plots() -> None:
+    """
+    Bring all stats together and plot subplots of PDB statistics
+    @return: None
+    """
+    path = "../data/pdb_statistics/"
+    alphafold_file = "AlphaFold_used.csv"
+    pdb_file = "ll_used.csv"
+    swissprot_file = "UniProt_Swiss_Prot.csv"
+    rcsb_file = "RCSB_by_length.csv"
+
+    alpha_bins, alpha_frequencies = get_data_for_bars(path + alphafold_file)
+    swiss_bins, swiss_frequencies = get_data_for_bars(path + swissprot_file)
+    pdb_bins, pdb_frequencies = get_data_for_bars(path + pdb_file)
+    bins, rcsb_frequencies = get_data_for_bars(path + rcsb_file)
+
+    adjusted_swiss = calculate_adjusted_frequency(swiss_frequencies, alpha_frequencies)
+    adjusted_rcsb = calculate_adjusted_frequency(rcsb_frequencies, pdb_frequencies)
+
+    fig = plt.figure(figsize=(10, 8))
+    ax = fig.subplots(2, 1, sharex=True)
+
+    ax[0].bar(bins, adjusted_swiss,
+              color=_COLOUR_PALETTE["DATABANK"],
+              label="Swiss-Prot Frequencies",
[…]

### 2022-12-29 code/scripts/protein_contact_map.py

@@ -1,0 +1,169 @@
+"""
+Base class for PCMs.
+
+"""
+import MDAnalysis
+import networkx as nx
+import MDAnalysis as mdA
+import numpy as np
+import warnings
+
+warnings.filterwarnings("ignore")
+
+
+def create_connected_component_subgraphs(graph: nx.Graph) -> None:
+    """
+    Create connected component subgraphs within protein graph.
+    @param graph: protein graph
+    @return:
+    """
+    for components in nx.connected_components(graph):
+        yield graph.subgraph(components)
+
+
+def get_chain_length(alpha_carbons: MDAnalysis.AtomGroup) -> int:
+    """
+    Calculate the chain length of protein from alpha carbons.
+    @param alpha_carbons
+    @return: length of chain
+    """
+    chain_length = 0
+    if alpha_carbons:
+        chain_length = len(alpha_carbons)
+    return chain_length
+
+
+class ProteinContactMap:
+    """
+    Base class for ProteinContactNetworks.
+    """
+
+    def __init__(self, pdb_file, default_threshold=8.0):
+        """
+        Construct the ProteinContactMap and initialise default threshold.
+        @param pdb_file: str
+        @param default_threshold: float
+        """
+        self.universe = mdA.Universe(pdb_file)
+        self.threshold = default_threshold
+
+    @property
+    def threshold(self):
+        """
+        Get threshold
+        @return:
+        """
+        return self._threshold
+
+    @threshold.setter
+    def threshold(self, threshold_value: float) -> None:
+        """
+        Set threshold value
+        @param threshold_value:
+        @return: None
+        """
+        if threshold_value != float(threshold_value):
+            raise TypeError('The threshold value must be a float.')
+
+        if threshold_value >= 0:
+            self._threshold = threshold_value
+
+        else:
+            raise ValueError('The threshold value must be larger than 0.')
+
+    @threshold.deleter
+    def threshold(self) -> None:
+        """
+        Deleter for threshold
+        @return: None
+        """
+        raise AttributeError("The threshold value cannot be deleted. You can set it to 0.")
+
+    @property
+    def get_alpha_carbons(self) -> None:
+        """
+        Take a universe, get segment A and return the C-alphas for the PDB.
+        @return:
+        """
+        segments = self.universe.residues.segments
+        n_segments = len(segments)
+        alpha_carbons = None
+        for i in range(n_segments):
+            segment_id = self.universe.residues.segments.segids[i]
+            try:
[…]

### 2022-12-29 code/scripts/retrieve_rcsb_pdbs.py

@@ -1,0 +1,18 @@
+"""
+Opens each chunked ID file and retrieves PDB files from RCSB to ../data/rcsb/pdb_files/.
+"""
+
+from Bio.PDB.PDBList import PDBList
+import functions
+
+
+pdblist = PDBList()
+pdb_ids = functions.concatenate_rcsb_id_files()
+counter = 1
+for pdb in pdb_ids:
+    print(f"At entry {counter}")
+    pdblist.retrieve_pdb_file(pdb_code=pdb,
+                              file_format="pdb",
+                              pdir="../data/rcsb/pdb_files/")
+    counter += 1
+

### 2022-12-29 code/scripts/trivial_parallelisation/retrieve_pdbs_parallel.py

@@ -1,0 +1,36 @@
+"""
+Exclusively to be ran alongside the embarrasing_parallel.sh script! 
+Opens each chunked ID file and retrieves PDB files from RCSB to ../../pdb_files/.
+"""
+import numpy as np
+from Bio.PDB.PDBList import PDBList
+import os
+import pandas as pd
+
+
+def concatenate_id_files() -> np.array:
+    """
+    Opens id_file_N.txt as a DF and concatenates them all into one DF
+    @return: pdb_array
+    """
+    frames = []
+    for filename in os.listdir("."):
+        if filename.endswith(".txt"):
+            temp_df = pd.read_csv(filename, header=None)
+            frames.append(temp_df)
+    dataframe = pd.concat(frames).reset_index()
+    dataframe = dataframe.drop(columns='index')
+    pdb_array = dataframe[0].to_numpy()
+    return pdb_array
+
+
+pdblist = PDBList()
+pdb_ids = concatenate_id_files()
+counter = 1
+for pdb in pdb_ids:
+    print(f"At entry {counter}")
+    pdblist.retrieve_pdb_file(pdb_code=pdb,
+                              file_format="pdb",
+                              pdir="../../pdb_files/")
+    counter += 1
+

### 2022-12-29 scripts/adjacency.py

@@ -1,78 +1,0 @@
-""" Functions for plotting an adjacency matrix from 3D simulations or PDBs"""
-import matplotlib
-import numpy as np
-from matplotlib import pyplot as plt
-import seaborn as sns
-import protein_contact_map
-import functions
-from colour_palette import _COLOUR_PALETTE
-
-
-def get_pdb_matrix(pdb_file: str) -> np.ndarray:
-    """
-    Open a PDB file and return its adjacency matrix
-    @param pdb_file: PDB file from either RCSB or AlphaFold (default)
-    @return: adjacency matrix as a Numpy array
-    """
-    return functions.pdb_to_adjacency(pdb_file)[1]
-
-
-def set_adjacency_matrix_ticks(plot: matplotlib.axes.Axes) -> None:
-    """
-    Format tick labels for adjacency matrix plots
-    @param plot: seaborn heatmap
-    @return:
-    """
-    for index, label in enumerate(plot.get_xticklabels()):
-        if index % 4 == 0:
-            label.set_visible(True)
-            label.set_rotation(360)
-            label.set_font("Helvetica")
-        else:
-            label.set_visible(False)
-    for index, label in enumerate(plot.get_yticklabels()):
-        if index % 4 == 0:
-            label.set_visible(True)
-            label.set_font("Helvetica")
-        else:
-            label.set_visible(False)
-
-
-def remove_diagonal_elements(matrix: np.ndarray):
-    """
-    Remove diagonal elements and add contacts to first off-diagonals
-    @param matrix: adjacency matrix from simulations
-    @return: simulation matrix with diagonals set to 1
-    """
-    np.fill_diagonal(matrix, 0)
-    off_diagonal = np.ones(len(np.diag(matrix, 1)))
-    np.fill_diagonal(matrix[1:], off_diagonal)
-    np.fill_diagonal(matrix[:, 1:], off_diagonal)
-    return matrix
-
-
-def plot_adjacency_matrix(file: str, data_type: str) -> None:
-    """
-    Plot adjacency matrix for given protein file (or simulation matrix)
-    @param file: PDB or matrix file
-    @param data_type: PDB or SIM
-    @return: None
-    """
-    adjacency_matrix = []
-    if data_type == "sim":
-        simulation_matrix = functions.get_3d_simulation_adjacency_matrix(file)
[…]

### 2022-12-29 scripts/adjacency_3d_simulations.sh

@@ -1,7 +1,0 @@
-#!/bin/bash
-# Bash script to plot different 3D adjacency matrix plots
-
-for i in `seq 0 1 27`
-do
-	python plot.py adj -t sim -f ../data/simulations/3d/matrices/matrix_300_$i.txt &
-done

### 2022-12-29 scripts/bars.py

@@ -1,79 +1,0 @@
-"""Functions for plotting bar plots"""
-import matplotlib.pyplot as plt
-import pandas as pd
-import seaborn as sns
-import numpy as np
-from colour_palette import _COLOUR_PALETTE
-
-
-def get_data_for_bars(path_to_data: str) -> tuple:
-    """
-    Read csv file from path and return numpy arrays as tuple
-    @param path_to_data:
-    @return:
-    """
-    dataframe = pd.read_csv(path_to_data)
-    bins = dataframe["Bins"].to_numpy()
-    frequencies = dataframe["Number"].to_numpy()
-    return bins, frequencies
-
-
-def calculate_adjusted_frequency(frequency_to_adjust: np.ndarray, bottom_frequency: np.ndarray) -> np.ndarray:
-    """
-    Calculate the adjusted frequency of the "top" bar to create a stack bar plot
-    @param frequency_to_adjust: "top" bars i.e. either all SwissProt or all RCSB frequencies
-    @param bottom_frequency: "bottom" bars i.e. the used PDBs' frequencies
-    @return: difference between top and bottom bars
-    """
-    return frequency_to_adjust - bottom_frequency
-
-
-def create_bar_plots() -> None:
-    """
-    Bring all stats together and plot subplots of PDB statistics
-    @return: None
-    """
-    path = "../data/pdb_statistics/"
-    alphafold_file = "alphafold_unique.csv"
-    pdb_file = "RCSB_used.csv"
-    swissprot_file = "UniProt_Swiss_Prot.csv"
-    rcsb_file = "RCSB_by_length.csv"
-
-    alpha_bins, alpha_frequencies = get_data_for_bars(path + alphafold_file)
-    swiss_bins, swiss_frequencies = get_data_for_bars(path + swissprot_file)
-    pdb_bins, pdb_frequencies = get_data_for_bars(path + pdb_file)
-    bins, rcsb_frequencies = get_data_for_bars(path + rcsb_file)
-
-    adjusted_swiss = calculate_adjusted_frequency(swiss_frequencies, alpha_frequencies)
-    adjusted_rcsb = calculate_adjusted_frequency(rcsb_frequencies, pdb_frequencies)
-
-    fig = plt.figure(figsize=(8, 8))
-    ax = fig.subplots(2, 1, sharex=True)
-
-    ax[0].bar(bins, adjusted_swiss,
[…]

### 2022-12-29 scripts/chunk_id_files.py

@@ -1,15 +1,0 @@
-"""
-Read in combined_ids_tr.csv and save files containing chunks of 1000 IDs.
-"""
-
-import pandas as pd
-import numpy as np
-
-infile = "../data/rcsb/combined_ids_tr.csv"
-chunk = 1000
-filename_counter = 0
-
-for chunk in pd.read_csv(infile, chunksize=chunk, lineterminator="\n", header=None):
-    chunk_array = chunk.to_numpy()
-    np.savetxt(f"../data/rcsb/ids/id_file_{filename_counter}.txt", chunk_array, fmt="%s")
-    filename_counter += 1

### 2022-12-29 scripts/colour_palette.py

@@ -1,12 +1,0 @@
-_COLOUR_PALETTE = {"PDB_SCATTER": "#682860",
-                   "2D_SIM_SCATTER": "#006374",
-                   "3D_SIM_SCATTER": "#006374",
-                   "ALPHA_SCATTER": "#286830",
-                   "POWER": "#1ac938",
-                   "THEORY": "#006374",
-                   "CL": "#fbafe4",
-                   "RESIDUALS": "#fbafe4",
-                   "DATABANK": "#006374",
-                   "USED": "#fbafe4",
-                   "CONTACT": "#fbafe4",
-                   "NO_CONTACT": "#006374"}

### 2022-12-29 scripts/comparison.py

@@ -1,102 +1,0 @@
-"""Functions for plotting comparison plot (Fig.4.b.)"""
-import argparse
-import numpy as np
-import scipy.optimize
-import matplotlib.pyplot as plt
-import seaborn as sns
-import pandas as pd
-import plot_functions
-import theory_functions
-from colour_palette import _COLOUR_PALETTE
-
-
-def create_comparison_plot(arguments: argparse.Namespace, rcsb_histogram: np.ndarray,
-                           alpha_histogram: np.ndarray) -> None:
-    """
-    Create amino acid distance distribution plot (Fig.4.b) and compare to theory and power law
-    @param arguments: command line arguments from user
-    @param rcsb_histogram: numpy array containing RCSB PDB data
-    @param alpha_histogram: numpy array containing AlphaFold 2 PDB data
-    @return: None
-    """
-    rcsb_plotting_tuple = plot_functions.get_data_for_plotting(rcsb_histogram, arguments, arguments.length_range)
-    rcsb_chain_length = rcsb_plotting_tuple[0]
-    rcsb_distances = rcsb_plotting_tuple[1]
-    rcsb_measure = rcsb_plotting_tuple[2]
-    rcsb_lower_bound, rcsb_upper_bound = rcsb_plotting_tuple[3], rcsb_plotting_tuple[4]
-    half_n_harmonic = theory_functions.harmonic_number(n_numbers=(rcsb_chain_length // 2))
-    alpha_plotting_tuple = plot_functions.get_data_for_plotting(alpha_histogram, arguments, arguments.length_range)
-    alpha_distances = alpha_plotting_tuple[1]
-    alpha_measure = alpha_plotting_tuple[2]
-
-    dimensionality_range = np.arange(arguments.rcsb_startd,
-                                     arguments.rcsb_endd,
-                                     arguments.step_dimensionality)
-    exponent_range = np.arange(arguments.rcsb_starte,
-                               arguments.rcsb_ende,
-                               arguments.step_exponent)
-    start = arguments.start_point
-    end = arguments.end_point
-    weights = rcsb_plotting_tuple[4] - rcsb_plotting_tuple[3]
-    rcsb_parameters, rcsb_cov = scipy.optimize.curve_fit(f=theory_functions.amino_acid_distance_distribution,
[…]

### 2022-12-29 scripts/confidences.py

@@ -1,49 +1,0 @@
-import glob
-import pandas as pd
-import numpy as np
-
-
-def get_confidences(pdb_files: list) -> list:
-    """
-    Read AlphaFold PDB file contents to find confidence info
-    @param: pdbs glob list of AlphaFold PDB files
-    @return: confidence_info list containing the dictionary of confidence info
-    """
-    counter = 1
-    confidence_info = []
-    for pdb_file in pdb_files:
-        print(f"At {counter}/{len(pdb_files)}")
-        file = open(pdb_file)
-        lines = file.readlines()
-        file.close()
-        confidence_scores = []
-        confidence_dict = {}
-        for line in lines:
-            if line.startswith("ATOM"):
-                current_line = line.split()
-                confidence_score = float(current_line[-2])
-                confidence_scores.append(confidence_score)
-        confidence_dict["mean"] = np.mean(confidence_scores)
-        confidence_dict["std"] = np.std(confidence_scores)
-        confidence_dict["filename"] = pdb_file
-        confidence_info.append(confidence_dict)
-        counter += 1
-    return confidence_info
-                                          
-                                      
-def filter_confidences(pdb_files: list) -> pd.DataFrame:
-    """
-    Filter out AlphaFold structures with less than 90% confidence
-    @param pdb_files: glob list of AlphaFold PDBs
-    @return good_confidence_df: dataframe containing PDBs with 90% confidence or above
-    """
-    confidence_list = get_confidences(pdb_files)
-    confidence_dataframe = pd.DataFrame.from_dict(confidence_list)
-    confidence_filter = confidence_dataframe["mean"] > 90.0
-    good_confidence_df = confidence_dataframe.where(confidence_filter).dropna()
-    return good_confidence_df
-
-
-pdbs = glob.glob("../data/alphafold/pdb_files/*.pdb")
-confidence_df = dssp_functions.filter_confidences(pdbs)
-confidence_df.to_csv("../data/alphafold/confidences.csv")

### 2022-12-29 scripts/contour.py

@@ -1,71 +1,0 @@
-"""Functions for plotting contour plots of alpha and beta protein content"""
-import numpy as np
-import matplotlib.pyplot as plt
-import seaborn as sns
-import matplotlib
-import pandas as pd
-import scipy
-from matplotlib import lines
-
-
-def contour_cloud(alphas, betas, colour, plot_alpha) -> matplotlib.contour.ContourSet:
-    """
-    Take alphas and betas and return a contour cloud in an alpha-beta plot
-    @param alphas: alpha-helix content
-    @param betas: beta sheet content
-    @param colour: colour to use for the contour plot
-    @param plot_alpha: transparency of plot
-    @return:
-    """
-    x_min, x_max = 0, 1
-    y_min, y_max = 0, 1
-
-    xx, yy = np.mgrid[x_min:x_max:100j, y_min:y_max:100j]
-    positions = np.vstack([xx.ravel(), yy.ravel()])
-    values = np.vstack([alphas, betas])
-    kernel = scipy.stats.gaussian_kde(values)
-    z = np.reshape(kernel(positions).T, xx.shape)
-    contour = plt.contour(xx, yy, z, colors=colour, alpha=plot_alpha, linewidths=2)
-    return contour
-
-
-def create_contour_plots() -> None:
-    """
-    Create contour distribution plots of the secondary structure content in PDB files
-    @return: None
-    """
-    lengths = ["100", "200", "300"]
-    indices = [50, 50, 70]
-    colours = ["#682860", "#286830", "#fbafe4", '#006374']
-    for i in range(len(lengths)):
-        sns.set(context="notebook", style="ticks", font_scale=1.88, font="Helvetica")
-        sns.set_palette(sns.color_palette(colours))
-        alphafold_df = pd.read_csv(f"../data/alphafold/unique_secondary_structures_{lengths[i]}.csv")
-        rcsb_df = pd.read_csv(f"../data/rcsb/secondary_structures_{lengths[i]}.csv")
-        rcsb_df = rcsb_df.loc[(rcsb_df[['alpha', 'beta']] != 0).all(axis=1)]
-        alphafold_df = alphafold_df.loc[(alphafold_df[["alpha", "beta"]] != 0).all(axis=1)]
-        rcsb_x = rcsb_df["alpha"]
-        rcsb_y = rcsb_df["beta"]
-        alphafold_x = alphafold_df["alpha"]
[…]

### 2022-12-29 scripts/distances.py

@@ -1,55 +1,0 @@
-"""Functions for plotting RCSB and AlphaFold amino acid distance distributions for each length in same plot"""
-import matplotlib.pyplot as plt
-import seaborn as sns
-import argparse
-import plot_functions
-
-
-def create_plot_label(arguments: argparse.Namespace, length_range: str) -> str:
-    """
-    Create legend label for the plots
-    @param arguments: command line arguments
-    @param length_range: 100, 200 or 300
-    @return:
-    """
-    label = ""
-    if arguments.algorithm == "rcsb":
-        label = f"RCSB {length_range}"
-    elif arguments.algorithm == "alpha":
-        label = f"AlphaFold 2 {length_range}"
-    return label
-
-
-def create_plots(arguments: argparse.Namespace) -> None:
-    """
-    Plot amino acid distances in each length range on the same plot
-    @param arguments: command line arguments
-    @return: None
-    """
-    length_ranges = ["100", "200", "300"]
-
-    colours = ["#a3dbab", "#47b856", "#286830"]
-    if arguments.algorithm == "rcsb":
-        colours = ["#e2b6dc", "#b847a9", "#682860"]
-    elif arguments.algorithm == "alpha":
-        colours = ["#a3dbab", "#47b856", "#15371a"]
-    markers = ["o", "s", "^"]
-    plt.figure(figsize=(6, 6))
-    sns.set(context="notebook", palette="colorblind", style="ticks", font_scale=1.8, font="Helvetica")
-    for i in range(len(length_ranges)):
-        histogram = plot_functions.get_histogram(length_ranges[i], arguments.algorithm)
-        plotting_tuple = plot_functions.get_data_for_plotting(histogram, arguments, length_ranges[i])
-        distance_bins = plotting_tuple[1]
-        measure = plotting_tuple[2]
-        plt.scatter(distance_bins, measure, label=create_plot_label(arguments, length_ranges[i]),
-                    color=colours[i], marker=markers[i])
-        plt.yscale("log")
-        plt.xscale("log")
-        plt.xlabel("s")
-        plt.ylabel("P(s)")
-        plt.xlim(2, 315)
-        plt.ylim(0.0001, 0.1)
-        plt.legend(loc="upper right", fontsize=20, frameon=False)
-        sns.despine()
-        plt.tight_layout()
-    plt.show()

### 2022-12-29 scripts/dssp.py

@@ -1,29 +1,0 @@
-# Load data from AlphaFold for chain-lengths 100, 200, 300
-
-import dssp_functions
-import glob
-import numpy as np
-import pandas as pd
-import argparse
-
-
-def parse_arguments() -> argparse.Namespace:
-    """
-    Parse command line arguments
-    """
-    parser = argparse.ArgumentParser(description="Get secondary structure information from AlphaFold 2 or RCSB PDBs.")
-    parser.add_argument("algorithm", type=str, choices=["alpha", "rcsb"])
-    return parser.parse_args()
-
-
-algorithm = parse_arguments().algorithm
-
-if algorithm == "alpha":
-    dssp_functions.save_alpha_secondary_info()
-
-elif algorithm == "rcsb":
-    dssp_functions.save_rcsb_secondary_info()
-
-
-                
-            

### 2022-12-29 scripts/dssp_functions.py

@@ -1,119 +1,0 @@
-# Functions for DSSP code
-import subprocess
-import numpy as np
-import os
-import protein_contact_map
-import pandas as pd
-import collections
-import traceback
-
-
-def get_secondary_structure(filename: str, dssp_path: str, log_file: str) -> np.ndarray:
-    """
-    Get the following secondary structure information for each PDB file
-    H = α-helix
-    B = residue in isolated β-bridge
-    E = extended strand, participates in β ladder
-    G = 3-helix (310 helix)
-    I = 5 helix (π-helix)
-    T = hydrogen bonded turn
-    S = bend
-
-    @param filename: pdb file + path
-    @param dssp_path: full path to dssp executable
-    @param log_file: text file to record any exceptions
-    @return: numpy array of secondary structures in pdb file
-    """
-    with open(log_file, "w") as log_file:
-        secondary_structures = []
-        try:
-            subprocess.check_call(f"{dssp_path} {filename} -o result.dssp", shell=True)
-            infile = open("result.dssp", "r")
-            readit = False
-            for line in infile:
-                if readit:
-                    try:
-                        if line[13:15] == '!*' or line[13] == '!':
-                            continue
-                        else:
-                            secondary_structure = line[16]
-                            if line[16] == " ":
-                                secondary_structure = "-"
-                            secondary_structures.append(secondary_structure)
-                    except:
-                        continue
-                if "#" in line:
-                    readit = True
-            infile.close()
-            os.remove("result.dssp")
-        except subprocess.CalledProcessError:
-            traceback.print_exc(file=log_file)
-        except Exception:
-            traceback.print_exc(file=log_file)
-    return np.array(secondary_structures)
-
-
-def save_alpha_secondary_info():
-    """
-    Get secondary structure and confidence information for AlphaFold 2 from DSSP and save into csv
-    @return:
-    """
-    length_ranges = ["100", "200", "300"]
-    chain_counter = 1
-    for length in length_ranges:
-        secondary_structures = []
-        print(f"Chains {chain_counter}/{len(length_ranges)}")
-        pdb_counter = 1
-        confidence_data = pd.read_csv(f"../data/alphafold/confidences_{length}.csv")
-        pdbs = confidence_data["filename"].to_numpy()
[…]

### 2022-12-29 scripts/functions.py

@@ -1,330 +1,0 @@
-"""
-Functions for getting link lengths and bootstrapping.
-"""
-import argparse
-import glob
-import os
-import traceback
-import networkx as nx
-import numpy as np
-import pandas as pd
-import protein_contact_map
-np.warnings.filterwarnings('ignore', category=np.VisibleDeprecationWarning)
-
-
-def concatenate_rcsb_id_files() -> np.array:
-    """
-    Opens id_file_N.txt as a DF and concatenates them all into one DF for RCSB PDBs.
-    @return: pdb_array
-    """
-    frames = []
-    for filename in os.listdir("../data/rcsb/ids/"):
-        if filename.endswith(".txt"):
-            temp_df = pd.read_csv(f"../data/rcsb/ids/{filename}", header=None)
-            frames.append(temp_df)
-    dataframe = pd.concat(frames).reset_index()
-    dataframe = dataframe.drop(columns='index')
-    pdb_array = dataframe[0].to_numpy()
-    return pdb_array
-
-
-def get_uniprot_ids(sequence_length: str) -> np.ndarray:
-    """
-    Return Uniprot IDs in the given sequence length range.
-    @param sequence_length
-    @return: np.ndarray of Uniprot IDs
-    """
-    id_dataframe = pd.read_excel(f"../data/alphafold/uniprot_ids/uniprot_{sequence_length}s.xlsx")
-    return id_dataframe["Entry"].to_numpy()
-
-
-def commandline_arguments() -> argparse.Namespace:
-    """
-    Parser for commandline arguments
-    @return: commandline arguments from user
-    """
-
-    parser = argparse.ArgumentParser(description="Get amino acid distance distributions.")
-    parser.add_argument("algorithm", type=str, choices=["rcsb", "alpha", "3d-sim"],
-                        help="Get amino acid distances from rcsb, [alpha]fold or 3D simulations")
-    parser.add_argument("-r", "--range", dest="length_range", type=str, choices=[None, "100", "200", "300"],
-                        help="Chain length range")
-    parser.add_argument("-p", "--path", dest="path_to_pdbs", type=str,
-                        help="Full path to file containing PDB ID csv file")
-    cl_arguments = parser.parse_args()
-    check_arguments(cl_arguments, parser)
-    return cl_arguments
-
-
-def check_arguments(arguments: argparse.Namespace, argument_parser: argparse.ArgumentParser) -> None:
-    """
-    Check the requirements for command line arguments
[…]

### 2022-12-29 scripts/get_alpha_distances.sh

@@ -1,5 +1,0 @@
-#!/bin/bash
-for i in `seq 100 100 300`
-do
-	nohup python run.py alpha -r=$i -p=../data/alphafold/unique_secondary_structures_$i.csv > nohup_run_not_normed_$i.out &
-done

### 2022-12-29 scripts/get_alphafold_pdbs.sh

@@ -1,4 +1,0 @@
-for i in `seq 100 100 300`
-do
-nohup python retrieve_alphafold_pdbs.py $i &
-done

### 2022-12-29 scripts/get_rcsb_chains.sh

@@ -1,4 +1,0 @@
-for i in `seq 100 100 300`
-do
-nohup python run.py rcsb -r=$i -p=../data/rcsb/chains_$i.csv &
-done

### 2022-12-29 scripts/get_rcsb_distances.sh

@@ -1,5 +1,0 @@
-#!/bin/bash
-for i in `seq 100 100 300`
-do
-	nohup python run.py rcsb -r=$i -p=../data/rcsb/secondary_structures_$i.csv > ../data/rcsb/nohup_run_not_normed_$i.out &
-done

### 2022-12-29 scripts/get_threshold_histograms.py

@@ -1,81 +1,0 @@
-import pandas as pd
-import functions
-import numpy as np
-from tqdm import tqdm
-import protein_contact_map
-
-
-def pdb_to_adjacency(pdb_file: str, threshold: float) -> tuple:
-    """
-    Convert given PDB file to an adjacency matrix
-    @param pdb_file: PDB file from RCSB or AlphaFold in each length range
-    @param threshold: threshold value to use for counting contacts
-    @return: tuple of the adjacency and distance matrices as numpy arrays
-    """
-    pcm = protein_contact_map.ProteinContactMap(pdb_file, default_threshold=threshold)
-    alpha_carbons = pcm.get_alpha_carbons
-    distance_array = protein_contact_map.get_distance_array(alpha_carbons)
-    distance_matrix = protein_contact_map.get_distance_matrix(alpha_carbons, distance_array)
-    adjacency_matrix = pcm.get_adjacency_matrix(alpha_carbons, distance_array)
-    return distance_matrix, adjacency_matrix
-
-
-def return_distance_histogram(log_file: str, given_algorithm: str, length_range: str,
-                              path_to_csvs: str, threshold: float) -> np.ndarray:
-    """
-    Compute the amino acid distance distribution for PDB files in given range from adjacency matrix
-    and save in a numpy file.
-    @param log_file: file to save exceptions in
-    @param given_algorithm: alphafold or rcsb
-    @param length_range: 100, 200 or 300
-    @param path_to_csvs: full path to csv files
-    @param threshold: threshold value to use for counting contacts
-    @return: None
-    """
-    dataframe = pd.read_csv(path_to_csvs)
-    pdb_files = dataframe["filename"].to_numpy()
-    histogram_list = []
-    counter = 1
-    with open(log_file, "w") as log_file:
-
-        for pdb_file in tqdm(pdb_files):
-            # print(f"Progress: {counter}/{len(pdb_files)}")
-            if given_algorithm == "alphafold":
-                clean_pdb_filename = pdb_file.replace("/home/jguven/Projects/sequence_distance_distribution", "..")
-            else:
-                clean_pdb_filename = pdb_file
-            try:
-                adjacency_matrix = pdb_to_adjacency(clean_pdb_filename, threshold)[1]
-                distances = functions.get_distances(adjacency_matrix)
-                bins = np.linspace(start=1, stop=350, num=350)
-                histogram = np.histogram(distances, bins=bins, density=False)[0]
[…]

### 2022-12-29 scripts/plot.py

@@ -1,43 +1,0 @@
-"""
-Handle all plotting functions and actually make the plots
-"""
-import adjacency
-import bars
-import comparison
-import contour
-import distances
-import plot_functions
-import simulation
-
-
-def plot() -> None:
-    """
-    Handle command line arguments and call required plotting function
-    @return: None
-    """
-    arguments = plot_functions.parse_command_line_arguments()
-    algorithm = arguments.algorithm
-    if algorithm == "comp":
-        rcsb_histogram = plot_functions.get_histogram(arguments.length_range, "rcsb")
-        alpha_histogram = plot_functions.get_histogram(arguments.length_range, "alpha")
-        comparison.create_comparison_plot(arguments, rcsb_histogram, alpha_histogram)
-    elif algorithm == "bar":
-        bars.create_bar_plots()
-    elif algorithm == "rcsb" or algorithm == "alpha":
-        distances.create_plots(arguments)
-    elif algorithm == "adj":
-        adjacency.plot_adjacency_matrix(arguments.file, arguments.data_type)
-    elif algorithm == "2d-sim":
-        simulation.create_2d_plots(arguments)
-    elif algorithm == "3d-sim":
-        simulation.create_3d_simulation_plot(arguments)
-    elif algorithm == "cont":
-        contour.create_contour_plots()
-
-
-def main():
-    plot()
-
-
-if __name__ == "__main__":
-    main()

### 2022-12-29 scripts/plot.sh

@@ -1,36 +1,0 @@
-#!/bin/bash
-
-# RCSB ranges 
-python plot.py rcsb
-
-# AlphaFold 2 ranges
-python plot.py alpha
-
-# RCSB 100 comparison
-python plot.py comp -r=100 --rcsb-startd 0.0015 --rcsb-endd 0.0016 --rcsb-starte 1 --rcsb-ende 3 -q 2 -m mean -startp 10 -endp 50
-
-# RCSB 200 comparison
-python plot.py comp -r=200 --rcsb-startd 0.0007 --rcsb-endd 0.0009 --rcsb-starte 2 --rcsb-ende 4 -q 2 -m mean -startp 20 -endp 100
-
-# RCSB 300 comparison
-python plot.py comp -r=300 --rcsb-startd 0.0007 --rcsb-endd 0.0009 --rcsb-starte 4 --rcsb-ende 6 -q 2 -m mean -startp 20 -endp 150
-
-# Contour plots
-python plot.py cont
-
-# 3D simulation 100
-python plot.py 3d-sim -r=100 -m mean -q 2 -startd 0.003 -endd 0.0032 -starte 2 -ende 4 -startp 4 -endp 50
-
-# 3D simulation 200
-python plot.py 3d-sim -r=200 -m mean -q 2 -startd 0.001 -endd 0.0013 -starte 0 -ende 2 -startp 4 -endp 100
-
-# 3D simulation 300
-python plot.py 3d-sim -r=300 -m mean -q 2 -startd 0.0006 -endd 0.0008 -starte 0 -ende 2 -startp 4 -endp 150
-
-# 2D simulation 
-python plot.py 2d-sim -m mean -q 2 -startd 0.0002 -endd 0.0004 -starte 6 -ende 8 -startp 4 -endp 250
-
-
-
-
-

### 2022-12-29 scripts/plot_comparison.py

@@ -1,208 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import scipy.optimize
-import seaborn as sns
-from sequence_distance_distribution.scripts import theory_functions
-from sequence_distance_distribution.scripts.colour_palette import _COLOUR_PALETTE
-import argparse
-
-
-def get_histogram(arguments: argparse.Namespace, algorithm: str) -> np.ndarray:
-    """
-    Get distribution of amino acid distances from histogram.
-    @param arguments: command line arguments
-    @param algorithm: rcsb or alpha from command line arguments
-    @return: histogram of RCSB/AlphaFold 2 data
-    """
-    histogram = None
-    if algorithm == "rcsb":
-        histogram = np.load(f"../data/rcsb/histogram_{arguments.length_range}_not_normed.npy", allow_pickle=True)
-    elif algorithm == "alpha":
-        histogram = np.load(f"../data/alphafold/histogram_{arguments.length_range}_not_normed.npy", allow_pickle=True)
-    return histogram
-
-
-def get_data_for_plotting(histogram: np.ndarray, arguments: argparse.Namespace) -> tuple:
-    """
-    Get protein data for plotting from dataframe
-    @param arguments: command line arguments
-    @param histogram: numpy array that contains amino acid distances
-    @return: tuple of number of datapoints, normalised measure of central tendency and confidence level bounds
-    """
-    distances = np.linspace(start=1, stop=300, num=300)[:-1]
-    lower_bound, upper_bound = theory_functions.get_confidence_interval(histogram, arguments.quantile)
-    measure = theory_functions.get_measure_of_central_tendency(histogram, arguments.measure)
-    normalised_measure = measure / np.sum(measure)
-    normalised_lower_bound = lower_bound / np.sum(measure)
-    normalised_upper_bound = upper_bound / np.sum(measure)
-    chain_length = int(arguments.length_range)
-    return chain_length, distances, normalised_measure, normalised_lower_bound, normalised_upper_bound
-
-
-def create_comparison_plot(arguments: argparse.Namespace, rcsb_histogram: np.ndarray,
-                           alpha_histogram: np.ndarray) -> None:
-    """
-    Create amino acid distance distribution plot (Fig.4)
-    @param arguments: command line arguments from user
-    @param rcsb_histogram: numpy array containing RCSB PDB data
[…]

### 2022-12-29 scripts/plot_comparison.sh

@@ -1,5 +1,0 @@
-#!/bin/bash
-
-python plot.py -r=100 comp -rcsbd 0.0019 -rcsbe 2 -alphad 0.0021 -alphae 3 -m=mean -q=1
-python plot.py -r=200 comp -rcsbd 0.0012 -rcsbe 3 -alphad 0.0011 -alphae 3 -m=mean -q=1
-python plot.py -r=300 comp -rcsbd 0.0009 -rcsbe 4 -alphad 0.0010 -alphae 5 -m=mean -q=1

### 2022-12-29 scripts/plot_functions.py

@@ -1,183 +1,0 @@
-"""
-Common functions used by all plotting codes, such as for handling command line arguments
-"""
-import argparse
-import numpy as np
-import theory_functions
-
-
-def get_histogram(length_range: str, algorithm: str) -> np.ndarray:
-    """
-    Get distribution of amino acid distances from histogram.
-    @param length_range: chain length range
-    @param algorithm: rcsb or alpha from command line arguments
-    @return: histogram of RCSB/AlphaFold 2 data
-    """
-    histogram = None
-    if algorithm == "rcsb":
-        histogram = np.load(f"../data/rcsb/histogram_{length_range}_not_normed.npy", allow_pickle=True)
-    elif algorithm == "alpha":
-        histogram = np.load(f"../data/alphafold/histogram_{length_range}_not_normed.npy", allow_pickle=True)
-    return histogram
-
-
-def get_data_for_plotting(histogram: np.ndarray, arguments: argparse.Namespace, length_range: str) -> tuple:
-    """
-    Get protein data for plotting from dataframe
-    @param histogram: numpy array that contains amino acid distance_bins
-    @param arguments: command line arguments
-    @param length_range: chain length range
-    @return: tuple of chain length, distances, normalised measure of central tendency and confidence level bounds
-    """
-    distance_bins = np.linspace(start=1, stop=350, num=350)[:-1]
-    lower_bound, upper_bound = theory_functions.get_confidence_interval(histogram, arguments.quantile)
-    measure = theory_functions.get_measure_of_central_tendency(histogram, arguments.measure)
-    normalised_measure = measure / np.sum(measure)
-    normalised_lower_bound = lower_bound / np.sum(measure)
-    normalised_upper_bound = upper_bound / np.sum(measure)
-    chain_length = int(length_range)
-    return chain_length, distance_bins, normalised_measure, normalised_lower_bound, normalised_upper_bound
-
-
-def parse_command_line_arguments() -> argparse.Namespace:
-    """
-    Parse CL arguments for plotting.
-    @return: Namespace containing CL arguments
-    """
-    parser = argparse.ArgumentParser(description="Plot amino acid distances and residuals.")
-    sim_group = parser.add_argument_group("2d-sim, 3d-sim")
-    compare_group = parser.add_argument_group("comp")
-    adjacency_group = parser.add_argument_group("adj")
-    parser.add_argument("-r", "--range", dest="length_range", type=str, choices=[None, "100", "200", "300"],
[…]

### 2022-12-29 scripts/protein_contact_map.py

@@ -1,240 +1,0 @@
-"""
-Base class for PCMs.
-
-"""
-import networkx as nx
-import MDAnalysis as mdA
-import numpy as np
-import warnings
-from MDAnalysis.analysis import distances, contacts
-warnings.filterwarnings("ignore")
-
-
-def create_connected_component_subgraphs(graph: nx.Graph) -> None:
-    """
-    Create connected component subgraphs within protein graph.
-    @param graph: protein graph
-    @return:
-    """
-    for components in nx.connected_components(graph):
-        yield graph.subgraph(components)
-
-
-def get_chain_length(alpha_carbons: mdA.AtomGroup) -> int:
-    """
-    Calculate the chain length of protein from alpha carbons.
-    @param alpha_carbons
-    @return: length of chain
-    """
-    chain_length = 0
-    if alpha_carbons:
-        chain_length = len(alpha_carbons)
-    return chain_length
-
-
-def get_distance_matrix(alpha_carbons: mdA.AtomGroup, distance_array: np.ndarray) -> np.ndarray:
-    """
-    Get distance matrix for PDBs from distance array and C-alphas
-    @param: alpha_carbons C-alphas from PDB
-    @param: distance_array distance array from C-alphas
-    @return: distance_matrix as a numpy array
-    """
-    n_alpha_carbons = get_chain_length(alpha_carbons)
-    distance_matrix = np.zeros((n_alpha_carbons, n_alpha_carbons))
-    n_rows = distance_matrix.shape[0]
-    diagonal = 1
-    upper_triangle = np.triu_indices(n_rows, k=diagonal)
-    distance_matrix[upper_triangle] = distance_array
-    np.fill_diagonal(distance_matrix, 0)
-    for i in range(len(distance_matrix)):
-        for j in range(i):
-            distance_matrix[i, j] = distance_matrix[j, i]
-    return distance_matrix
-
-
-def get_distance_array(alpha_carbons: mdA.AtomGroup) -> np.ndarray:
-    """
-    Use the C-alphas to get the distance array
-    @return: distance array
-    """
-    return distances.self_distance_array(alpha_carbons.positions)
-
-
-def get_discreet_distances(adjacency_matrix: np.ndarray) -> np.ndarray:
-    """
-    Get the amino acid distances from the adjacency matrix
-    @param: adjacency_matrix from PDB
-    @return list of discrete distances from adjacency matrix
-    """
-    distance_list = []
-    rows = len(adjacency_matrix)
[…]

### 2022-12-29 scripts/remove_duplicates.py

@@ -1,50 +1,0 @@
-"""
-Remove duplicates from AlphaFold 2 PDB files
-"""
-import glob
-import pandas as pd
-import numpy as np
-
-
-def get_protein_names(filenames: np.ndarray) -> np.ndarray:
-    """
-    Loop through the filenames in the secondary structure dataframe and get the protein names
-    @param filenames: array of full paths to pdb files in secondary structure dataframes
-    @return: array of protein names
-    """
-    protein_names = []
-    counter = 1
-    for file in filenames:
-        try:
-            with open(file) as pdb:
-                print(f"Progress: {counter}/{len(filenames)}") 
-                header = [next(pdb) for i in range(2)]
-                protein_name = header[-1].strip().split('FOR')[-1]
-                protein_names.append(protein_name)
-                counter += 1
-        except TypeError as e:
-            print(f"NaN in filename, excluding from list: {e}")
-    return np.asarray(protein_names)
-            
-
-def filter_unique_secondary_structures() -> None:
-    """
-    Open secondary structures csv and find unique protein structures and save as a csv.
-    @return: None
-    """
-    chain_lengths = ["100", "200", "300"]
-    for length in chain_lengths:
-        secondary_structures_df = pd.read_csv(f"../data/alphafold/secondary_structures_{length}.csv")
-        filenames = secondary_structures_df["filename"]
-        secondary_structures_df["protein_name"] = get_protein_names(filenames)
-        unique_structures_df = secondary_structures_df.drop_duplicates(subset="protein_name")
-        unique_structures_df.to_csv(f"../data/alphafold/unique_secondary_structures_{length}.csv")
-
-        
-def main():
-    filter_unique_secondary_structures()
-
-    
-if __name__ == "__main__":
-    main()
-    

### 2022-12-29 scripts/retrieve_alphafold_pdbs.py

@@ -1,38 +1,0 @@
-"""
-Get UniProt ids and loop through them to save AlphaFold PDBs in ../data/alphafold/pdbs/
-"""
-import urllib.request
-import functions
-import argparse
-
-
-def commandline_arguments() -> argparse.Namespace:
-    """
-    Parser for commandline arguments
-    @return: commandline arguments from user
-    """
-
-    parser = argparse.ArgumentParser(description="Get AlphaFold 2 PDBs.")
-    parser.add_argument("range", type=str, choices=["100", "200", "300"], help="Chain length range")
-    cl_arguments = parser.parse_args()
-    return cl_arguments
-
-
-sequence_length = commandline_arguments().range
-
-counter = 0
-
-uniprot_IDs = functions.get_uniprot_ids(sequence_length)
-for uniprot_ID in uniprot_IDs:
-    print(f"At entry {counter}/{len(uniprot_IDs)}")
-    print(f"ID: {uniprot_ID}")
-    download = f"https://alphafold.ebi.ac.uk/files/AF-{uniprot_ID}-F1-model_v2.pdb"
-        
-    try:
-        file_name = f"../data/alphafold/pdb_files/AF-{uniprot_ID}-F1-model_v2.pdb"
-        urllib.request.urlretrieve(download, file_name)
-    except urllib.error.HTTPError:
-        print("No such file.")
-
-        counter += 1
-         

### 2022-12-29 scripts/retrieve_rcsb_pdbs.py

@@ -1,17 +1,0 @@
-"""
-Opens each chunked ID file and retrieves PDB files from RCSB to ../data/rcsb/pdbs/.
-"""
-
-from Bio.PDB.PDBList import PDBList
-import functions
-
-pdblist = PDBList()
-pdb_ids = functions.concatenate_rcsb_id_files()
-counter = 1
-for pdb in pdb_ids:
-    print(f"At entry {counter}")
-    pdblist.retrieve_pdb_file(pdb_code=pdb,
-                              file_format="pdb",
-                              pdir="../data/rcsb/pdb_files/")
-    counter += 1
-

### 2022-12-29 scripts/run.py

@@ -1,37 +1,0 @@
-"""
-Compute amino acid distances or bootstrap.
-"""
-import argparse
-import functions
-
-
-def handle_commandline_options(cl_arguments: argparse.Namespace) -> None:
-    """
-    Read in CL arguments and make a choice depending on given option
-    @param cl_arguments from user
-    @return: None
-    """
-    given_algorithm = cl_arguments.algorithm
-    given_linklength = cl_arguments.length_range
-    given_path = cl_arguments.path_to_pdbs
-    if given_algorithm == "rcsb":
-        functions.return_distance_histogram(log_file="log.txt",
-                                            given_algorithm=given_algorithm,
-                                            length_range=given_linklength,
-                                            path_to_csvs=given_path)
-    elif given_algorithm == "alpha":
-        functions.return_distance_histogram(log_file="log.txt",
-                                            given_algorithm=given_algorithm,
-                                            length_range=given_linklength,
-                                            path_to_csvs=given_path)
-    elif given_algorithm == "3d-sim":
-        functions.return_3d_simulation_distance_histogram(length_range=given_linklength)
-
-
-def main():
-    arguments = functions.commandline_arguments()
-    handle_commandline_options(arguments)
-
-
-if __name__ == "__main__":
-    main()

### 2022-12-29 scripts/save_chain_lengths.py

@@ -1,108 +1,0 @@
-# save chain lengths
-import subprocess
-import numpy as np
-import os
-import protein_contact_map
-import glob 
-import pandas as pd
-import argparse
-
-
-def save_alpha_chain_lengths(confidence_dataframe: pd.DataFrame) -> None:
-    """
-    Open the confidence dataframe and sort files into length ranges
-    @param confidence_dataframe: dataframe containing confidence info
-    return: None
-    """
-    counter = 1
-    pdb_files = confidence_dataframe["filename"].to_numpy()
-    mean_confidences = confidence_dataframe["mean"].to_numpy()
-    std_confidences = confidence_dataframe["std"].to_numpy()
-    proteins_100 = {"filename": [], "mean_conf": [], "std_conf": []}
-    proteins_200 = {"filename": [], "mean_conf": [], "std_conf": []}
-    proteins_300 = {"filename": [], "mean_conf": [], "std_conf": []}
-    for i in range(len(pdb_files)):
-        print(f"At {counter}/{len(pdb_files)}")
-        pcm = protein_contact_map.ProteinContactMap(pdb_files[i])
-        alpha_carbons = pcm.get_alpha_carbons
-        chain_length = protein_contact_map.get_chain_length(alpha_carbons)
-        if chain_length in range(85, 116):
-            print("In 100s")
-            proteins_100["filename"].append(pdb_files[i])
-            proteins_100["mean_conf"].append(mean_confidences[i])
-            proteins_100["std_conf"].append(std_confidences[i])
-        elif chain_length in range(185, 216):
-            print("In 200s")
-            proteins_200["filename"].append(pdb_files[i])
-            proteins_200["mean_conf"].append(mean_confidences[i])
-            proteins_200["std_conf"].append(std_confidences[i])
-        elif chain_length in range(285, 316):
-            print("In 300s")
-            proteins_300["filename"].append(pdb_files[i])
-            proteins_300["mean_conf"].append(mean_confidences[i])
-            proteins_300["std_conf"].append(std_confidences[i])
-        counter += 1
-    df_100 = pd.DataFrame.from_dict(proteins_100)
-    df_200 = pd.DataFrame.from_dict(proteins_200)
-    df_300 = pd.DataFrame.from_dict(proteins_300)
-    df_100.to_csv("../data/alphafold/confidences_100.csv")
[…]

### 2022-12-29 scripts/simulation.py

@@ -1,266 +1,0 @@
-"""Functions for plotting 2D and 3D simulation amino acid distances."""
-import argparse
-import glob
-import numpy as np
-import pandas as pd
-import scipy.optimize
-import matplotlib.pyplot as plt
-import seaborn as sns
-import theory_functions
-from colour_palette import _COLOUR_PALETTE
-
-
-def get_2d_files():
-    """
-    Get all the files for 2D simulations
-    @return: list containing all simulation files
-    """
-    return glob.glob(f"../data/simulations/2d/2d_sim_no_circle_*")
-
-
-def get_2d_distance_frequencies(files: list) -> np.ndarray:
-    """
-    Get the frequencies from 2D simulation files
-    @param files: 2D simulation files
-    @return: array containing frequencies for amino acid distances
-    """
-    distances_list = []
-    for file in files:
-        distances = np.loadtxt(file)
-        distances_list.append(distances)
-    return np.asarray(distances_list)
-
-
-def get_2d_stats() -> pd.DataFrame:
-    """
-    Compute the mean and confidence interval for 2D simulations
-    @return: None
-    """
-    files = get_2d_files()
-    frequencies = get_2d_distance_frequencies(files)
-    simulation_dataframe = pd.DataFrame(frequencies)
-    melted_dataframe = simulation_dataframe.melt()
-    simulation_stats_dataframe = melted_dataframe.groupby("variable",
-                                                          as_index=False).agg(mean=("value", np.mean),
-                                                                              lower_bound=("value", lambda val:
-                                                                              np.quantile(val, q=0.05)),
-                                                                              upper_bound=("value", lambda val:
-                                                                              np.quantile(val, q=0.95)))
-    simulation_stats_dataframe.to_csv(f"../data/simulations/2d/simulation_stats.csv")
-    return simulation_stats_dataframe
-
-
-def get_2d_plotting_data(simulation_dataframe: pd.DataFrame) -> tuple:
-    """
-    Get 2D simulation data from dataframe
-    @param simulation_dataframe: dataframe containing 2D simulation stats
-    @return: tuple containing the mean frequencies of amino acid distances, distances and the CI bounds
-    """
-    mean_frequencies = simulation_dataframe["mean"].to_numpy()
-    normalised_means = mean_frequencies / np.sum(mean_frequencies)
-    distances = simulation_dataframe["variable"].to_numpy()
[…]

### 2022-12-29 scripts/theory_functions.py

@@ -1,216 +1,0 @@
-"""
-Theoretical functions for plotting.
-"""
-import numpy as np
-from sklearn.metrics import r2_score
-import networkx as nx
-
-
-def harmonic_number(n_numbers: int) -> float:
-    """
-    @param n_numbers: number of natural numbers
-    @return: harmonic: nth harmonic number
-    """
-    n_int = int(n_numbers)
-    harmonic = 1.00
-    for i in range(2, n_int + 1):
-        harmonic += 1 / i
-    return harmonic
-
-
-def f_high(amino_acid_distance: int, chain_length: int, half_n_harmonic_number: float) -> float:
-    """
-    Calculate f for k >> 1
-    @param amino_acid_distance: separation between each amino acid, 2 <= s < N/2
-    @param chain_length: number of C-alphas in a chain
-    @param half_n_harmonic_number: the 'N/2'-th harmonic number
-    @return: amino acid distribution for k >> 1
-    """
-    harmonic = harmonic_number(amino_acid_distance)
-    constant_multiplier = 2 / chain_length
-    s_term = ((half_n_harmonic_number - harmonic + 1) / (half_n_harmonic_number - 1)) * amino_acid_distance
-    constant_term = half_n_harmonic_number / (half_n_harmonic_number - 1)
-    return constant_multiplier * (s_term - constant_term)
-
-
-def f_low(amino_acid_distance: int, chain_length: int) -> float:
-    """
-    Calculate f for k << N/2
-    @param amino_acid_distance: separation between each link, 2 <= s < N/2
-    @param chain_length: number of C-alphas in a chain
-    @return: amino acid distance distribution for k << N/2
-    """
-    s_square_term = (-2 / chain_length ** 2) * amino_acid_distance ** 2
-    s_term = (2 / chain_length + 6 / chain_length ** 2) * amino_acid_distance
-    constant_term = (2 / chain_length) + (4 / chain_length ** 2)
-    return s_square_term + s_term - constant_term
-
-
-def amino_acid_distance_distribution(amino_acid_distances: np.ndarray, chain_length: int, half_n_harmonic_number: float,
-                                     exponent: int, dimensionality: float) -> np.ndarray:
-    """
-
-    @param amino_acid_distances: separation between each amino acid 2 <= s < N/2
-    @param chain_length: number of C-alphas
-    @param half_n_harmonic_number: the 'N/2'-th harmonic number
-    @param exponent: constant a
[…]

### 2022-12-29 scripts/threshold_plots.py

@@ -1,52 +1,0 @@
-""" Supplemental plots for investigating the effect of the threshold value on amino acid distance
-    distributions.
-"""
-
-import functions
-import numpy as np
-import pandas as pd
-import protein_contact_map
-import matplotlib.pyplot as plt
-from colour_palette import _COLOUR_PALETTE
-import seaborn as sns
-
-
-lengths = [100, 200, 300]
-
-for length in lengths:
-    default_threshold = np.load(f"../data/rcsb/histogram_{length}_not_normed.npy", allow_pickle=True)
-    simulation = np.load(f"../data/simulations/3d/histogram_{length}_not_normed.npy", allow_pickle=True)
-    threshold_10 = np.load(f"../data/rcsb/threshold_histograms/histogram_{length}_10.npy", allow_pickle=True)
-    threshold_15 = np.load(f"../data/rcsb/threshold_histograms/histogram_{length}_15.npy", allow_pickle=True)
-    threshold_21 = np.load(f"../data/rcsb/threshold_histograms/histogram_{length}_21.npy", allow_pickle=True)
-    distances = np.linspace(1, 350, 350)[:-1]
-
-    mean_default = np.mean(default_threshold, axis=0)
-    normed_mean_default = mean_default / np.sum(mean_default)
-    mean_10 = np.mean(threshold_10, axis=0)
-    normed_mean_10 = mean_10 / np.sum(mean_10)
-    mean_15 = np.mean(threshold_15, axis=0)
-    normed_mean_15 = mean_15 / np.sum(mean_15)
-    mean_21 = np.mean(threshold_21, axis=0)
-    normed_mean_21 = mean_21 / np.sum(mean_21)
-    simulation_mean = np.mean(simulation, axis=0)
-    normed_simulation_mean = simulation_mean / np.sum(simulation_mean)
-
-    fig = plt.figure()
-    fig.set_size_inches(8, 8)
-    sns.set(context="notebook", palette="colorblind", style="ticks", font_scale=1.88, font="Helvetica")
-    plt.scatter(distances, normed_simulation_mean, label=f"3D Simulation {length}", c="gray")
-    plt.scatter(distances, normed_mean_default, label=f"Threshold 8 Å {length}",
-                c=_COLOUR_PALETTE["PDB_SCATTER"], marker="s")
[…]

### 2022-12-29 scripts/trivial_parallelisation/retrieve_pdbs_parallel.py

@@ -1,36 +1,0 @@
-"""
-Exclusively to be run alongside the embarrassing_parallel.sh script!
-Opens each chunked ID file and retrieves PDB files from RCSB to ../../pdbs/.
-"""
-import numpy as np
-from Bio.PDB.PDBList import PDBList
-import os
-import pandas as pd
-
-
-def concatenate_id_files() -> np.array:
-    """
-    Opens id_file_N.txt as a DF and concatenates them all into one DF
-    @return: pdb_array
-    """
-    frames = []
-    for filename in os.listdir("."):
-        if filename.endswith(".txt"):
-            temp_df = pd.read_csv(filename, header=None)
-            frames.append(temp_df)
-    dataframe = pd.concat(frames).reset_index()
-    dataframe = dataframe.drop(columns='index')
-    pdb_array = dataframe[0].to_numpy()
-    return pdb_array
-
-
-pdblist = PDBList()
-pdb_ids = concatenate_id_files()
-counter = 1
-for pdb in pdb_ids:
-    print(f"At entry {counter}")
-    pdblist.retrieve_pdb_file(pdb_code=pdb,
-                              file_format="pdb",
-                              pdir="../../pdb_files/")
-    counter += 1
-
